# Convert object format {#sec-colocalization}

# Libraries
```{r}
library(Seurat)
library(SeuratWrappers)
library(ggplot2)
library(patchwork)
library(dplyr)
library(wesanderson)
library(ggplotify)
library(cowplot)
library(stringr)
library(reshape2)
library(ggrepel)

source("../source_scripts/slideseq_functions_main.R")

options(Seurat.object.assay.version = 'v5')
```

# Open objects
```{r}
# P102TA
P102TA_10_RNA <- readRDS("../objects/02_Puck10_102TA_TCR_RNA_area_121125.Rds")
P102TA_17_RNA <- readRDS("../objects/02_Puck17_102TA_TCR_RNA_area_121125.Rds")
P102TA_33_RNA <- readRDS("../objects/02_Puck33_102TA_TCR_RNA_area_121125.Rds")
P102TA_39_RNA <- readRDS("../objects/02_Puck39_102TA_TCR_RNA_area_121125.Rds")

# P107TA
P107TA_15_RNA <- readRDS("../objects/02_Puck15_107TA_TCR_RNA_area_121125.Rds")
P107TA_18_RNA <- readRDS("../objects/02_Puck18_107TA_TCR_RNA_area_121125.Rds")
P107TA_30_RNA <- readRDS("../objects/02_Puck20_107TA_TCR_RNA_area_121125.Rds")

# P108TA
P108TA_26_RNA <- readRDS("../objects/02_Puck26_108TA_TCR_RNA_area_121125.Rds")
P108TA_29_RNA <- readRDS("../objects/02_Puck29_108TA_TCR_RNA_area_121125.Rds")
P108TA_30_RNA <- readRDS("../objects/02_Puck30_108TA_TCR_RNA_area_121125.Rds")
P108TA_32_RNA <- readRDS("../objects/02_Puck32_108TA_TCR_RNA_area_121125.Rds")

# P108TB
P108TB_06_RNA <- readRDS("../objects/02_Puck06_108TB_TCR_RNA_area_121125.Rds")
P108TB_14_RNA <- readRDS("../objects/02_Puck14_108TB_TCR_RNA_area_121125.Rds")

# P109TB
P109TB_23_RNA <- readRDS("../objects/02_Puck23_109TB_TCR_RNA_area_121125.Rds")
P109TB_34_RNA <- readRDS("../objects/02_Puck34_109TB_TCR_RNA_area_121125.Rds")

# P110TA
P110TA_28_RNA <- readRDS("../objects/02_Puck28_110TA_TCR_RNA_area_121125.Rds")
P110TA_30_RNA <- readRDS("../objects/02_Puck30_110TA_TCR_RNA_area_121125.Rds")

# P111TA
P111TA_32_RNA <- readRDS("../objects/02_Puck32_111TA_TCR_RNA_area_121125.Rds")
P111TA_37_RNA <- readRDS("../objects/02_Puck37_111TA_TCR_RNA_area_121125.Rds")
```

# Convert all objects to anndata to run LR colocalization script
## sceasy's convertFormat is currently incompatible with Seurat V5 as of March 3rd 2026 as the 'slot' argument of 'GetAssayData()' was deprecated. An alternative is to use the package anndataR instead. It requires R v4.5, so I won't be running it. I'll be using older objects I had previously converted to h5ad. The RNA assay should be identical across my older and newer objects.
```{r, eval = FALSE}
path <- "../objects/"

# P102TA
save_seurat_as_anndata(P102TA_10_RNA, path, "02_Puck10_102TA_TCR_RNA_area_121125")
save_seurat_as_anndata(P102TA_17_RNA, path, "02_Puck17_102TA_TCR_RNA_area_121125")
save_seurat_as_anndata(P102TA_33_RNA, path, "02_Puck33_102TA_TCR_RNA_area_121125")
save_seurat_as_anndata(P102TA_39_RNA, path, "02_Puck39_102TA_TCR_RNA_area_121125")

# P107TA
save_seurat_as_anndata(P107TA_15_RNA, path, "02_Puck15_107TA_TCR_RNA_area_121125")
save_seurat_as_anndata(P107TA_18_RNA, path, "02_Puck18_107TA_TCR_RNA_area_121125")
save_seurat_as_anndata(P107TA_30_RNA, path, "02_Puck20_107TA_TCR_RNA_area_121125")

# P108TA
save_seurat_as_anndata(P108TA_26_RNA, path, "02_Puck26_108TA_TCR_RNA_area_121125")
save_seurat_as_anndata(P108TA_29_RNA, path, "02_Puck29_108TA_TCR_RNA_area_121125")
save_seurat_as_anndata(P108TA_30_RNA, path, "02_Puck30_108TA_TCR_RNA_area_121125")
save_seurat_as_anndata(P108TA_32_RNA, path, "02_Puck32_108TA_TCR_RNA_area_121125")

# P108TB
save_seurat_as_anndata(P108TB_06_RNA, path, "02_Puck06_108TB_TCR_RNA_area_121125")
save_seurat_as_anndata(P108TB_14_RNA, path, "02_Puck14_108TB_TCR_RNA_area_121125")

# P109TB
save_seurat_as_anndata(P109TB_23_RNA, path, "02_Puck23_109TB_TCR_RNA_area_121125")
save_seurat_as_anndata(P109TB_34_RNA, path, "02_Puck34_109TB_TCR_RNA_area_121125")

# P110TA
save_seurat_as_anndata(P110TA_28_RNA, path, "02_Puck28_110TA_TCR_RNA_area_121125")
save_seurat_as_anndata(P110TA_30_RNA, path, "02_Puck30_110TA_TCR_RNA_area_121125")

# P111TA
save_seurat_as_anndata(P111TA_32_RNA, path, "02_Puck32_111TA_TCR_RNA_area_121125")
save_seurat_as_anndata(P111TA_37_RNA, path, "02_Puck37_111TA_TCR_RNA_area_121125")

```

# Get session info
```{r}
sessionInfo()
```


