# Myeloid boundary bead composition {#sec-area-annotation}

# Show increased density of myeloid cells within the myeloid boundary compared to tumor using a linear mixed model

# Libraries
```{r}
library(Seurat)
library(dplyr)
library(ggpointdensity)
library(tidyverse)
library(ComplexHeatmap)
library(stringr)
library(readxl)
library(circlize)
library(RANN)
library(ggridges)
library(ash)
library(ggpubr)
library(rcompanion)
library(patchwork)
library(ggprism)
library(lme4)
library(infer)

source("../source_scripts/slideseq_functions_main.R")

options(Seurat.object.assay.version = 'v5')
```

## Functions
```{r}
get_pct_myeloid_in_bin <- function(obj, bin_length = 30){
  md <- obj@meta.data %>%
    select(plot_x, plot_y, area, assignment1)
  
  # Create bins for x and y
  md$x_bin <- cut(md$plot_x, breaks = bin_length, labels = FALSE) 
  md$y_bin <- cut(md$plot_y, breaks = bin_length, labels = FALSE)
  
  # Don't consider empty areas
  md <- md %>%
    filter(area != "Empty")
  bin_df <- data.frame()

  for(i in c(1:bin_length)){
    for(j in c(1:bin_length)){
      bin <- md %>%
        filter(x_bin == i,
               y_bin == j)
      boundary_count <- bin %>%
        filter(area == "Boundary") %>%
        dplyr::count() %>%
        pull(n)
      tumor_count <- bin %>%
        filter(area == "Tumor") %>%
        dplyr::count() %>%
        pull(n)
      tls_count <- bin %>%
        filter(area %in% c("TLS1", "TLS2", "TLS3", "TLS4", "M2")) %>%
        dplyr::count() %>%
        pull(n)
      total_count <- bin %>%
        dplyr::count() %>%
        pull(n)
      myeloid_count <- bin %>%
        filter(assignment1 == "Myeloid") %>%
        dplyr::count() %>%
        pull(n)
      
      bin_summary <- c(i, j, total_count, boundary_count/total_count, tumor_count/total_count, tls_count/total_count, myeloid_count, myeloid_count/total_count)
      bin_df <- rbind(bin_df, bin_summary)
    }
  }
  colnames(bin_df) <- c("x_bin", "y_bin", "total_beads", "pct_boundary", "pct_tumor", "pct_tls", "ct_myeloid", "pct_myeloid")
  
  return(bin_df)
}

assign_bin_annotation <- function(bin_df){
  bin_df <- bin_df %>%
  mutate(area = case_when(pct_tls > 0.8 ~ "TLS",
                          pct_boundary > 0.8 ~ "Boundary",
                          pct_tumor > 0.8 ~ "Tumor",
                          total_beads == 0 ~ "Empty",
                          T ~ "Mix"))
  return(bin_df)
}

```

# Process spatial data
## Open objects with annotated areas
```{r}
P102TA_10_RNA <- readRDS("../objects/02_Puck10_102TA_TCR_RNA_area_121125.Rds")
P102TA_17_RNA <- readRDS("../objects/02_Puck17_102TA_TCR_RNA_area_121125.Rds")
P102TA_33_RNA <- readRDS("../objects/02_Puck33_102TA_TCR_RNA_area_121125.Rds")

P108TA_26_RNA <- readRDS("../objects/02_Puck26_108TA_TCR_RNA_area_121125.Rds")
P108TA_29_RNA <- readRDS("../objects/02_Puck29_108TA_TCR_RNA_area_121125.Rds")
P108TA_30_RNA <- readRDS("../objects/02_Puck30_108TA_TCR_RNA_area_121125.Rds")
P108TA_32_RNA <- readRDS("../objects/02_Puck32_108TA_TCR_RNA_area_121125.Rds")
```

## Assign celltypes
```{r}
P102TA_10_RNA@meta.data <- P102TA_10_RNA@meta.data %>%
  assign_celltypes() %>%
  order_celltypes()
P102TA_10_RNA@meta.data %>%
  dplyr::count(assignment1)
P102TA_17_RNA@meta.data <- P102TA_17_RNA@meta.data %>%
  assign_celltypes() %>%
  order_celltypes()
P102TA_17_RNA@meta.data %>%
  dplyr::count(assignment1)
P102TA_33_RNA@meta.data <- P102TA_33_RNA@meta.data %>%
  assign_celltypes() %>%
  order_celltypes()
P102TA_33_RNA@meta.data %>%
  dplyr::count(assignment1)

P108TA_26_RNA@meta.data <- P108TA_26_RNA@meta.data %>%
  assign_celltypes() %>%
  order_celltypes()
P108TA_26_RNA@meta.data %>%
  dplyr::count(assignment1)
P108TA_29_RNA@meta.data <- P108TA_29_RNA@meta.data %>%
  assign_celltypes() %>%
  order_celltypes()
P108TA_29_RNA@meta.data %>%
  dplyr::count(assignment1)
P108TA_30_RNA@meta.data <- P108TA_30_RNA@meta.data %>%
  assign_celltypes() %>%
  order_celltypes()
P108TA_30_RNA@meta.data %>%
  dplyr::count(assignment1)
P108TA_32_RNA@meta.data <- P108TA_32_RNA@meta.data %>%
  assign_celltypes() %>%
  order_celltypes()
P108TA_32_RNA@meta.data %>%
  dplyr::count(assignment1)
```

# Bin the beads and calculate percentage of myeloid beads / all beads in each bin
```{r}
P102TA_10_bin <- get_pct_myeloid_in_bin(P102TA_10_RNA)
P102TA_17_bin <- get_pct_myeloid_in_bin(P102TA_17_RNA)
P102TA_33_bin <- get_pct_myeloid_in_bin(P102TA_33_RNA)
P108TA_26_bin <- get_pct_myeloid_in_bin(P108TA_26_RNA)
P108TA_29_bin <- get_pct_myeloid_in_bin(P108TA_29_RNA)
P108TA_30_bin <- get_pct_myeloid_in_bin(P108TA_30_RNA)
P108TA_32_bin <- get_pct_myeloid_in_bin(P108TA_32_RNA)
```

# Filter for bins with 5 beads or more
```{r}
P102TA_10_bin <- P102TA_10_bin %>%
  filter(total_beads >= 5)
P102TA_17_bin <- P102TA_17_bin %>%
  filter(total_beads >= 5)
P102TA_33_bin <- P102TA_33_bin %>%
  filter(total_beads >= 5)

P108TA_26_bin <- P108TA_26_bin %>%
  filter(total_beads >= 5)
P108TA_29_bin <- P108TA_29_bin %>%
  filter(total_beads >= 5)
P108TA_30_bin <- P108TA_30_bin %>%
  filter(total_beads >= 5)
P108TA_32_bin <- P108TA_32_bin %>%
  filter(total_beads >= 5)
```

# Assign an area to each bin 
```{r}
P102TA_10_bin <- P102TA_10_bin %>%
  assign_bin_annotation %>%
  mutate(Patient = "P102TA",
         orig.ident = "P102TA_10")
P102TA_17_bin <- P102TA_17_bin %>%
  assign_bin_annotation %>%
  mutate(Patient = "P102TA",
         orig.ident = "P102TA_17")
P102TA_33_bin <- P102TA_33_bin %>%
  assign_bin_annotation %>%
  mutate(Patient = "P102TA",
         orig.ident = "P102TA_33")

P108TA_26_bin <- P108TA_26_bin %>%
  assign_bin_annotation %>%
  mutate(Patient = "P108TA",
         orig.ident = "P108TA_26")
P108TA_29_bin <- P108TA_29_bin %>%
  assign_bin_annotation %>%
  mutate(Patient = "P108TA",
         orig.ident = "P108TA_29")
P108TA_30_bin <- P108TA_30_bin %>%
  assign_bin_annotation %>%
  mutate(Patient = "P108TA",
         orig.ident = "P108TA_30")
P108TA_32_bin <- P108TA_32_bin %>%
  assign_bin_annotation %>%
  mutate(Patient = "P108TA",
         orig.ident = "P108TA_32")
```

# Plot the bins, colored by area
```{r, fig.width = 6, fig.height = 5}
ggplot(P102TA_10_bin, aes(x = x_bin, y = y_bin, color = area)) +
  geom_point(shape = 15, size = 3) +
  scale_color_manual(values = c("Boundary" = "#ff8f1b", "TLS" = "#8A9A5B", "Tumor" = "#974000", "Mix" = "grey80", "Empty" = "#eeeeee")) +
  xlab("Binned x") +
  ylab("Binned y") +
  theme_bw() +
  coord_fixed(xlim = c(0, 30), ylim = c(0, 30))
ggplot(P102TA_17_bin, aes(x = x_bin, y = y_bin, color = area)) +
  geom_point(shape = 15, size = 3) +
  scale_color_manual(values = c("Boundary" = "#ff8f1b", "TLS" = "#8A9A5B", "Tumor" = "#974000", "Mix" = "grey80", "Empty" = "#eeeeee")) +
  xlab("Binned x") +
  ylab("Binned y") +
  theme_bw() +
  coord_fixed(xlim = c(0, 30), ylim = c(0, 30))
ggplot(P102TA_33_bin, aes(x = x_bin, y = y_bin, color = area)) +
  geom_point(shape = 15, size = 3) +
  scale_color_manual(values = c("Boundary" = "#ff8f1b", "TLS" = "#8A9A5B", "Tumor" = "#974000", "Mix" = "grey80", "Empty" = "#eeeeee")) +
  xlab("Binned x") +
  ylab("Binned y") +
  theme_bw() +
  coord_fixed(xlim = c(0, 30), ylim = c(0, 30))


ggplot(P108TA_26_bin, aes(x = x_bin, y = y_bin, color = area)) +
  geom_point(shape = 15, size = 3) +
  scale_color_manual(values = c("Boundary" = "#ff8f1b", "TLS" = "#8A9A5B", "Tumor" = "#974000", "Mix" = "grey80", "Empty" = "#eeeeee")) +
  xlab("Binned x") +
  ylab("Binned y") +
  theme_bw() +
  coord_fixed(xlim = c(0, 30), ylim = c(0, 30))
ggplot(P108TA_29_bin, aes(x = x_bin, y = y_bin, color = area)) +
  geom_point(shape = 15, size = 3) +
  scale_color_manual(values = c("Boundary" = "#ff8f1b", "TLS" = "#8A9A5B", "Tumor" = "#974000", "Mix" = "grey80", "Empty" = "#eeeeee")) +
  xlab("Binned x") +
  ylab("Binned y") +
  theme_bw() +
  coord_fixed(xlim = c(0, 30), ylim = c(0, 30))
ggplot(P108TA_30_bin, aes(x = x_bin, y = y_bin, color = area)) +
  geom_point(shape = 15, size = 3) +
  scale_color_manual(values = c("Boundary" = "#ff8f1b", "TLS" = "#8A9A5B", "Tumor" = "#974000", "Mix" = "grey80", "Empty" = "#eeeeee")) +
  xlab("Binned x") +
  ylab("Binned y") +
  theme_bw() +
  coord_fixed(xlim = c(0, 30), ylim = c(0, 30))
ggplot(P108TA_32_bin, aes(x = x_bin, y = y_bin, color = area)) +
  geom_point(shape = 15, size = 3) +
  scale_color_manual(values = c("Boundary" = "#ff8f1b", "TLS" = "#8A9A5B", "Tumor" = "#974000", "Mix" = "grey80", "Empty" = "#eeeeee")) +
  xlab("Binned x") +
  ylab("Binned y") +
  theme_bw() +
  coord_fixed(xlim = c(0, 30), ylim = c(0, 30))
```

# Merge data
```{r}
df <- do.call(rbind, list(P102TA_10_bin, P102TA_17_bin, P102TA_33_bin,
                          P108TA_26_bin, P108TA_29_bin, P108TA_30_bin, P108TA_32_bin))

```

## Plot % myeloid beads in tumor and boundary
```{r}
subset_df <- df %>%
  filter(area %in% c("Tumor", "Boundary")) %>%
  mutate(area = factor(area, levels = c("Tumor", "Boundary")))

subset_df %>%
  summarize(mean_total_beads = mean(total_beads))

subset_df %>%
  dplyr::count(area)
```

## Plot proportion of beads
```{r}
subset_df %>%
  ggplot(aes(x = paste0(orig.ident, area), y = pct_myeloid, fill = orig.ident)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

# Try building a linear model
```{r}
model_obs <- lmer(pct_myeloid ~ area + (1 | Patient / orig.ident), data = subset_df, REML = TRUE)
t_val_obs <- coef(summary(model_obs))["areaBoundary","t value"]
```

# Permute data
```{r}
n_perm_safe = 1500

set.seed(123)
# Swap location labels within pucks and clonotypes
df_perm <- subset_df %>%
    rep_sample_n(size = nrow(subset_df), replace = FALSE, reps = n_perm_safe) %>%
    group_by(replicate, orig.ident) %>%
    mutate(perm_area = sample(area, size = n(), replace = FALSE))

```

## Get permuted test statistic
```{r}
# Get permuted test statistic for each gene set
n_perm <- 1000

set.seed(123)

n_run <- 1
n_saved <- 1
non_conv <- 0

t_val_perm.vec <- c()
# Do 1000 converging loops
while(n_saved <= n_perm){
  # Print every saved 100 perms
  if (n_saved %% 100 == 0) { print(n_saved) }
  # Get the permuted replicate
  md_rep <- df_perm %>%
    filter(replicate == n_run)
  # Set runs as good
  bad_run <- FALSE
  # Attempt the model
  model <- withCallingHandlers(
    expr = {
      # Build linear model on permuted data
      lmer(pct_myeloid ~ perm_area + (1 | Patient / orig.ident), data = md_rep, REML = TRUE)},
    warning = function(w) {
    # Note if the model can't converge
    bad_run <<- TRUE
    non_conv <<- non_conv + 1
    n_run <<- n_run + 1
    print("Couldn't converge")
    invokeRestart("muffleWarning")
    }
  )
      
  # Save the t val from the model if it has no errors
  if(!bad_run){
    # Get and save t val
    t_val_perm <- coef(summary(model))["perm_areaBoundary","t value"]
    t_val_perm.vec <- c(t_val_perm.vec, t_val_perm)
    # Count that we've saved a run
    n_saved <- n_saved + 1
    n_run <- n_run + 1
  } else {
    print("Skipping model")
  }
}

model_perm <- t_val_perm.vec
non_convergences <- non_conv
```

## Plot permuted test statstic distribution
```{r, fig.height = 5, fig.width = 8}
ggplot(as.data.frame(model_perm), aes(x = model_perm)) +
  geom_histogram(bins = 100) +
  geom_vline(aes(xintercept = t_val_obs))

```

## Calculate pval
```{r}
# Two sided symmetrical distributions
n_extreme <- length(model_perm[model_perm >= t_val_obs])
pval <- n_extreme/n_perm*2

pval
```

## Get session info
```{r}
sessionInfo()
```


