# 109T Puck23: Area annotation {#sec-area-annotation}

# Integrating RCTD calls and area annotations

## Libraries
```{r}
library(Seurat)
library(SeuratWrappers)
library(ggplot2)
library(patchwork)
library(dplyr)
library(wesanderson)
library(ggplotify)
library(cowplot)
library(stringr)
library(viridis)
library(tidyverse)
library(imager)
library(spacexr)

source("../source_scripts/slideseq_functions_main.R")

options(Seurat.object.assay.version = 'v5')

one_area_colors <- c("#1786f5")
```

## Load object from script 01- where I integrated TCR and RNA data from C26 together
```{r,echo=FALSE}
P109TB_23_RNA <- readRDS("/n/data2/dfci/medonc/cwu/chloe/slideseq_data/tcr_analysis/RCC_neovax_pucks/analysis_110624/objects/01_TCR_RNA_integration/01_Puck23_109TB_TCR_RNA_merged_110724.Rds")
P109TB_23_rctd <- readRDS("/n/data2/dfci/medonc/cwu/chloe/slideseq_data/tcr_analysis/RCC_neovax_pucks/analysis_110624/RCTD/rctd_results_rds_091125/P109TB_23_rctd_results.rds")

P109TB_23_RNA <- UpdateSeuratObject(P109TB_23_RNA)
```
```{r, eval=FALSE}
P109TB_23_RNA <- readRDS("01_Puck23_109TB_TCR_RNA_merged_110724.Rds")
```

## Open rctd objects
```{r,eval=FALSE}
P109TB_23_rctd <- readRDS("P109TB_23_rctd_results.rds")
```

## Remove reject beads and normalize RCTD weights
```{r}
# Remove reject beads
rejects <- P109TB_23_rctd@results[["results_df"]] %>%
  filter(spot_class == "reject") %>%
  rownames()

P109TB_23_weights <- as.data.frame(normalize_weights(P109TB_23_rctd@results$weights)) %>%
  # Remove reject beads
  filter(!(rownames(.) %in% rejects)) %>%
  dplyr::rename("Endothelial" = "ENDO",
                "Tumor" = "RCC",
                "Myeloid" = "MYELOID",
                "Tcell" = "T")
```

## Save normalized weights
```{r, eval = FALSE}
write.csv(P109TB_23_weights, "P109TB_23_normalized_weights.csv")
```

## Add RCTD weights
```{r, echo = FALSE}
P109TB_23_weights <- read.csv("/n/data2/dfci/medonc/cwu/chloe/slideseq_data/tcr_analysis/RCC_neovax_pucks/analysis_110624/RCTD/normalized_weights_091125/P109TB_23_normalized_weights.csv", row.names = 1)
```
```{r, eval = FALSE}
P109TB_23_weights <- read.csv("P109TB_23_normalized_weights.csv", row.names = 1)
```
```{r}
P109TB_23_RNA <- AddMetaData(P109TB_23_RNA, P109TB_23_weights)

P109TB_23_RNA@meta.data <- P109TB_23_RNA@meta.data %>%
  mutate(B = ifelse(is.na(B), 0, B),
         Tcell = ifelse(is.na(Tcell), 0, Tcell),
         Endothelial = ifelse(is.na(Endothelial), 0, Endothelial),
         Myeloid = ifelse(is.na(Myeloid), 0, Myeloid),
         NK = ifelse(is.na(NK), 0, NK),
         Tumor = ifelse(is.na(Tumor), 0, Tumor))
```

## Plot myeloid signature, CD4T, CD8, B cell signature
```{r}
SpatialFeaturePlot(P109TB_23_RNA, features = c("B"), pt.size.factor = 2, alpha = c(0.01,1)) & scale_fill_viridis(direction = -1)
SpatialFeaturePlot(P109TB_23_RNA, features = c("Tcell"), pt.size.factor = 2, alpha = c(0.01,1)) & scale_fill_viridis(direction = -1)
SpatialFeaturePlot(P109TB_23_RNA, features = c("Myeloid"), pt.size.factor = 2, alpha = c(0.01,1)) & scale_fill_viridis(direction = -1)
SpatialFeaturePlot(P109TB_23_RNA, features = c("Endothelial"), pt.size.factor = 2, alpha = c(0.01,1)) & scale_fill_viridis(direction = -1)
SpatialFeaturePlot(P109TB_23_RNA, features = c("Tumor"), pt.size.factor = 2, alpha = c(0.01,1)) & scale_fill_viridis(direction = -1)
```

## Plot feature and counts for empty area annotation
```{r}
SpatialFeaturePlot(P109TB_23_RNA, features = c("log_nFeature_Spatial", "log_nCount_Spatial"))
```

## Areas with less than 25th percentile UMI were removed as empty areas
```{r}
quantile(P109TB_23_RNA$nCount_Spatial)
beads_25th_umi <- Cells(subset(P109TB_23_RNA, nCount_Spatial <= 14))

P109TB_23_RNA %>%
  subset(cells = beads_25th_umi) %>%
  SpatialDimPlot() +
  ggtitle("UMI<25th percentile (14)")

SpatialDimPlot(P109TB_23_RNA, cells.highlight = beads_25th_umi)
```

## Update co-ordinates
```{r}
plot_coords <- P109TB_23_RNA@images$image@coordinates %>%
  mutate(plot_x = y,
         plot_y = -x+5180)

P109TB_23_RNA <- AddMetaData(P109TB_23_RNA, plot_coords)
```

## Check that plotting co-ordinates are perfect
```{r}
P109TB_23_RNA %>%
  subset(tumor_specific == TRUE) %>%
SpatialDimPlot(group.by = "tumor_specific", alpha = 1, stroke = 0, pt.size.factor = 2)

beads_ts <- P109TB_23_RNA@meta.data %>%
  filter(tumor_specific == TRUE)
SpatialDimPlot(P109TB_23_RNA, group.by = "tumor_specific", alpha = 1, stroke = 0, pt.size.factor = 2) +
  geom_point(data = beads_ts, aes(x = plot_x, y = plot_y), size = 0.1)

beads_tas <- P109TB_23_RNA@meta.data %>%
  filter(tumor_antigen_specific == TRUE)
SpatialDimPlot(P109TB_23_RNA,  group.by = "tumor_antigen_specific", alpha = 1, stroke = 0, pt.size.factor = 2) +
  geom_point(data = beads_tas, aes(x = plot_x, y = plot_y), size = 0.1)
```

## Plot the location of TCRs
```{r}
beads_tcr <- P109TB_23_RNA@meta.data %>%
  filter(nFeature_TCR > 0)
SpatialFeaturePlot(P109TB_23_RNA, features = c("Myeloid"), pt.size.factor = 2, alpha = c(0,1)) + scale_fill_viridis(direction = -1) +
  geom_point(data = beads_tcr, aes(x = plot_x, y = plot_y), size = 0.1, color = "blue", alpha = 1) + NoLegend()
```

## Plot the B and myeloid signature of the beads
```{r}
P108TA_26_B <- P109TB_23_RNA@meta.data %>%
  filter(B > 0.3)

SpatialFeaturePlot(P109TB_23_RNA, features = c("B"), pt.size.factor = 2, alpha = c(0,1)) +
  NoLegend() & 
  scale_fill_viridis(direction = -1) 

plot <- SpatialFeaturePlot(P109TB_23_RNA, features = c("Myeloid"), pt.size.factor = 2, alpha = c(0,1)) + scale_fill_viridis(direction = -1) +
  geom_point(data = P108TA_26_B, 
             aes(x = plot_x, y = plot_y, alpha = B), 
             size = 0.5, color = "red") +
  theme(legend.position = "None")

plot
```

## Plot the location of TCRs
```{r}
beads_tcr <- P109TB_23_RNA@meta.data %>%
  filter(nFeature_TCR > 0)
SpatialFeaturePlot(P109TB_23_RNA, features = c("Myeloid"), pt.size.factor = 2, alpha = c(0,1)) + scale_fill_viridis(direction = -1) +
  geom_point(data = beads_tcr, aes(x = plot_x, y = plot_y), size = 0.1, color = "blue", alpha = 1) + NoLegend()
```

## Save images to annotate in Imagescope
```{r}
# No myeloid boundary or empty areas seen, don't bother saving for annotation.
```

## After running python script and creating mask PNGs, convert the PNGs into masks and attach to object metadata
```{r}
# No myeloid boundary or empty areas seen, don't bother saving for annotation.
```

## Skipped adding area annotations (since there are no areas to annotate)
## Add an area mask column of just "Tumor" values
```{r}
P109TB_23_RNA@meta.data <- P109TB_23_RNA@meta.data %>%
  mutate(area = "Tumor")

SpatialDimPlot(P109TB_23_RNA, group.by = "area")
```

## Plot where all TCRs are
```{r}
beads_w_tcr <- P109TB_23_RNA@meta.data %>%
  filter(nFeature_TCR > 0)

legend <- as.ggplot(get_legend(SpatialDimPlot(P109TB_23_RNA, group.by = "area") + 
                                 scale_fill_manual(values = one_area_colors))) 

plot <- SpatialDimPlot(P109TB_23_RNA, group.by = "area", alpha = 0.7, stroke = 0) +
  scale_fill_manual(values = one_area_colors) +
  geom_point(data = beads_w_tcr, aes(x = plot_x, y = plot_y), size = 1) +
  theme(legend.position = "None") 

plot + plot_spacer() + legend + plot_layout(widths = c(3, -1.1 ,3)) + ggtitle("TCR location")
```

## Plot where neoAg TCRs are
```{r}
beads_neoag <- P109TB_23_RNA@meta.data %>%
  filter(tumor_antigen_specific == TRUE)

legend <- as.ggplot(get_legend(SpatialDimPlot(P109TB_23_RNA, group.by = "area") + 
                                 scale_fill_manual(values = one_area_colors))) 

plot <- SpatialDimPlot(P109TB_23_RNA, group.by = "area", alpha = 0.7, stroke = 0) +
  scale_fill_manual(values = one_area_colors) +
  geom_point(data = beads_neoag, aes(x = plot_x, y = plot_y), size = 1) +
  theme(legend.position = "None")

plot + plot_spacer() + legend + plot_layout(widths = c(3, -1.1 , 3)) + ggtitle("NeoAg TCR location")
```

## Plot where EBV reactive TCRs are
```{r}
beads_ebv <- P109TB_23_RNA@meta.data %>%
  filter(virus_antigen == "EBV")

legend <- as.ggplot(get_legend(SpatialDimPlot(P109TB_23_RNA, group.by = "area") + 
                                 scale_fill_manual(values = one_area_colors))) 

plot <- SpatialDimPlot(P109TB_23_RNA, group.by = "area", alpha = 0.7, stroke = 0) +
  scale_fill_manual(values = one_area_colors) +
  geom_point(data = beads_ebv, aes(x = plot_x, y = plot_y), size = 1) +
  theme(legend.position = "None")

plot + plot_spacer() + legend + plot_layout(widths = c(3, -1.1 , 3)) + ggtitle("EBV reactive TCR location")
```

## Save puck
```{r, eval = FALSE}
saveRDS(P109TB_23_RNA, "02_Puck23_109TB_TCR_RNA_area_121125.Rds")
```

## Get session info
```{r}
sessionInfo()
```
