# 110T Puck30: Area annotation {#sec-area-annotation}

# Integrating RCTD calls and area annotations

## Libraries
```{r}
library(Seurat)
library(SeuratWrappers)
library(ggplot2)
library(patchwork)
library(dplyr)
library(tidyverse)
library(wesanderson)
library(ggplotify)
library(cowplot)
library(imager)
library(viridis)
library(spacexr)

source("../source_scripts/slideseq_functions_main.R")

options(Seurat.object.assay.version = 'v5')

four_area_colors <- c("#f2f0f0", "#1786f5", "#ff8f1b", "#8A9A5B", "#8A9A5B")
```

## Load object from script 01
```{r,echo=FALSE}
P110TA_30_RNA <- readRDS("/n/data2/dfci/medonc/cwu/chloe/slideseq_data/tcr_analysis/RCC_neovax_pucks/analysis_110624/objects/01_TCR_RNA_integration/01_Puck30_110TA_TCR_RNA_merged_031925.Rds")
P110TA_30_rctd <- readRDS("/n/data2/dfci/medonc/cwu/chloe/slideseq_data/tcr_analysis/RCC_neovax_pucks/analysis_110624/RCTD/rctd_results_rds_091125/P110TA_30_rctd_results.rds")

P110TA_30_RNA <- UpdateSeuratObject(P110TA_30_RNA)
```
```{r,eval=FALSE}
P110TA_30_RNA <- readRDS("01_Puck30_110TA_TCR_RNA_merged_031925.Rds")
```

## Open rctd objects
```{r,eval=FALSE}
P110TA_30_rctd <- readRDS("P110TA_30_rctd_results.rds")
```

## Remove reject beads and normalize RCTD weights
```{r}
# Remove reject beads
rejects <- P110TA_30_rctd@results[["results_df"]] %>%
  filter(spot_class == "reject") %>%
  rownames()

P110TA_30_weights <- as.data.frame(normalize_weights(P110TA_30_rctd@results$weights)) %>%
  # Remove reject beads
  filter(!(rownames(.) %in% rejects)) %>%
  dplyr::rename("Endothelial" = "ENDO",
                "Tumor" = "RCC",
                "Myeloid" = "MYELOID",
                "Tcell" = "T")
```

## Save normalized weights
```{r, eval = FALSE}
write.csv(P110TA_30_weights, "P110TA_30_normalized_weights.csv")
```

## Add RCTD weights
```{r, echo = FALSE}
P110TA_30_weights <- read.csv("/n/data2/dfci/medonc/cwu/chloe/slideseq_data/tcr_analysis/RCC_neovax_pucks/analysis_110624/RCTD/normalized_weights_091125/P110TA_30_normalized_weights.csv", row.names = 1)
```
```{r, eval = FALSE}
P110TA_30_weights <- read.csv("P110TA_30_normalized_weights.csv", row.names = 1)
```
```{r}
P110TA_30_RNA <- AddMetaData(P110TA_30_RNA, P110TA_30_weights)

P110TA_30_RNA@meta.data <- P110TA_30_RNA@meta.data %>%
  mutate(B = ifelse(is.na(B), 0, B),
         Tcell = ifelse(is.na(Tcell), 0, Tcell),
         Endothelial = ifelse(is.na(Endothelial), 0, Endothelial),
         Myeloid = ifelse(is.na(Myeloid), 0, Myeloid),
         NK = ifelse(is.na(NK), 0, NK),
         Tumor = ifelse(is.na(Tumor), 0, Tumor))
```

## Plot myeloid signature, CD4T, CD8, B cell signature for TLS and myeloid rich tumor annotations
```{r}
SpatialFeaturePlot(P110TA_30_RNA, features = c("B"), pt.size.factor = 2, alpha = c(0,1)) & scale_fill_viridis(direction = -1)
SpatialFeaturePlot(P110TA_30_RNA, features = c("Tcell"), pt.size.factor = 2, alpha = c(0,1)) & scale_fill_viridis(direction = -1)
SpatialFeaturePlot(P110TA_30_RNA, features = c("Myeloid"), pt.size.factor = 2, alpha = c(0,1)) & scale_fill_viridis(direction = -1)
SpatialFeaturePlot(P110TA_30_RNA, features = c("Tumor"), pt.size.factor = 2, alpha = c(0,1)) & scale_fill_viridis(direction = -1)
SpatialFeaturePlot(P110TA_30_RNA, features = c("Endothelial"), pt.size.factor = 2, alpha = c(0,1)) & scale_fill_viridis(direction = -1)
```

## Plot feature and counts for empty area annotation
```{r}
SpatialFeaturePlot(P110TA_30_RNA, features = c("log_nFeature_Spatial", "log_nCount_Spatial"))
```

## Areas with less than 25th percentile UMI were removed as empty areas
```{r}
quantile(P110TA_30_RNA$nCount_Spatial)
beads_25th_umi <- Cells(subset(P110TA_30_RNA, nCount_Spatial <= 108))

P110TA_30_RNA %>%
  subset(cells = beads_25th_umi) %>%
  SpatialDimPlot() +
  ggtitle("UMI<25th percentile (108)")

SpatialDimPlot(P110TA_30_RNA, cells.highlight = beads_25th_umi)
```


## Update co-ordinates
```{r}
plot_coords <- P110TA_30_RNA@images$image@coordinates %>%
  mutate(plot_x = y,
         plot_y = -x+4970)

P110TA_30_RNA <- AddMetaData(P110TA_30_RNA, plot_coords)
```

## Check that plotting co-ordinates are perfect
```{r}
beads_ts <- P110TA_30_RNA@meta.data %>%
  filter(tumor_specific == TRUE)
SpatialDimPlot(P110TA_30_RNA, group.by = "tumor_specific", alpha = 1, stroke = 0, pt.size.factor = 2) +
  geom_point(data = beads_ts, aes(x = plot_x, y = plot_y), size = 0.1)

beads_tas <- P110TA_30_RNA@meta.data %>%
  filter(tumor_antigen_specific == TRUE)
SpatialDimPlot(P110TA_30_RNA,  group.by = "tumor_antigen_specific", alpha = 1, stroke = 0, pt.size.factor = 2) +
  geom_point(data = beads_tas, aes(x = plot_x, y = plot_y), size = 0.1)
```

## Plot the location of TCRs
```{r}
beads_tcr <- P110TA_30_RNA@meta.data %>%
  filter(nFeature_TCR > 0)
SpatialFeaturePlot(P110TA_30_RNA, features = c("Myeloid"), pt.size.factor = 2, alpha = c(0,1)) + scale_fill_viridis(direction = -1) +
  geom_point(data = beads_tcr, aes(x = plot_x, y = plot_y), size = 0.1, color = "blue", alpha = 1) + NoLegend()
```

## Plot the B and myeloid signature of the beads
```{r}
P110TA_30_B <- P110TA_30_RNA@meta.data %>%
  filter(B > 0.3)

plot_b <- SpatialFeaturePlot(P110TA_30_RNA, features = c("B"), pt.size.factor = 2, alpha = c(0,1)) +
  NoLegend() &
  scale_fill_viridis(direction = -1)

plot <- SpatialFeaturePlot(P110TA_30_RNA, features = c("Myeloid"), pt.size.factor = 2, alpha = c(0,1)) + scale_fill_viridis(direction = -1) +
  geom_point(data = P110TA_30_B,
             aes(x = plot_x, y = plot_y, alpha = B),
             size = 0.5, color = "red") +
  theme(legend.position = "None")

plot_b
plot
```

## Save images to annotate in Imagescope
```{r}
# jpeg("/n/data2/dfci/medonc/cwu/chloe/slideseq_data/tcr_analysis/RCC_neovax_pucks/analysis_110624/area_annotation/area_annotation_jpgs_032824/Puck30_110TA_empty.jpeg", width = 7, height = 7, units = "in", res = 300)
  SpatialDimPlot(P110TA_30_RNA, cells.highlight = beads_25th_umi, alpha = 0.5, stroke = 0.1, pt.size.factor = 2) + scale_fill_manual(values = c("darkred", "white")) +
  theme_bw() +
  coord_fixed(ratio = 1, xlim = c(0,5000), ylim = c(0,5000), expand = FALSE, clip = "off") +
  scale_alpha(range = c(0, 1)) +
  theme(legend.position = "None")
# dev.off()

```

## Need anndata object to convert ImageScope XML to PNG in python, so convert seurat object to anndata and save the anndata obj
```{r, eval = FALSE}
save_seurat_as_anndata(P110TA_30_RNA, "/n/data2/dfci/medonc/cwu/chloe/slideseq_data/tcr_analysis/RCC_neovax_pucks/analysis_110624/area_annotation/area_annotation_anndata_032824/", "01_Puck30_110TA_TCR_RNA_area")
```

## After running python script and creating mask PNGs, convert the PNGs into masks and attach to object metadata
```{r}
P110TA_30_RNA <- add_masks_to_obj(P110TA_30_RNA, "/n/data2/dfci/medonc/cwu/chloe/slideseq_data/tcr_analysis/RCC_neovax_pucks/analysis_110624/area_annotation/area_png_masks/Puck30_110TA/", adjust_x = 0, adjust_y = 0)

```

## Make sure the masks look good and are centered correctly, make sure to reference the screenshot of the annotated areas straight from ImageScope
```{r}
suppressMessages(SpatialFeaturePlot(P110TA_30_RNA, features = c("Myeloid"), pt.size.factor = 2, alpha = c(0,1)) + NoLegend() & scale_fill_viridis(direction = -1) )

suppressMessages(P110TA_30_RNA %>%
  subset(cells = beads_25th_umi) %>%
  SpatialDimPlot(stroke = 0) + NoLegend())

SpatialDimPlot(P110TA_30_RNA, group.by = "mask_1") + NoLegend()
SpatialDimPlot(P110TA_30_RNA, group.by = "mask_2") + NoLegend()
```


## Merge layers into one column
```{r}
P110TA_30_RNA@meta.data <- P110TA_30_RNA@meta.data %>%
  mutate(mask_sum = rowSums(select(., mask_1, mask_2)))

SpatialDimPlot(P110TA_30_RNA, group.by = "mask_sum")
```

## Translate numericals into names, make sure to reference the screenshot of the annotated areas taken straight from ImageScope
```{r}
P110TA_30_RNA@meta.data <- P110TA_30_RNA@meta.data %>%
  mutate(area = case_when(mask_sum %in% c(1, 2) ~ "Empty",
                          mask_sum %in% c(3) ~ "Tumor"))


suppressMessages(P110TA_30_RNA %>%
  subset(cells = beads_25th_umi) %>%
  SpatialDimPlot(stroke = 0) +
  NoLegend())

SpatialDimPlot(P110TA_30_RNA, group.by = "area", stroke = 0.1) + scale_fill_manual(values = four_area_colors)
```

## Do we have any straggling beads
```{r}
unique(P110TA_30_RNA$area)
# Plot empty beads only
SpatialDimPlot(subset(P110TA_30_RNA, area == "Empty"), group.by = "area", alpha = 0.7, stroke = 0)
# Plot tumor beads only
SpatialDimPlot(subset(P110TA_30_RNA, area == "Tumor"), group.by = "area", alpha = 0.7, stroke = 0)

```


## Straggling beads have the tumor label when they should be empty
```{r}
# Find affected beads
SpatialDimPlot(P110TA_30_RNA, group.by = "area", alpha = 0.7, stroke = 0) +
  geom_point(data = filter(P110TA_30_RNA@meta.data, plot_x > 4400 & plot_y > 3600), aes(x = plot_x, y = plot_y), size = 1, alpha = 0.05) +
  theme(legend.position = "None") +
  scale_fill_manual(values = four_area_colors)
# Extract straggling beads
beads_to_empty <- P110TA_30_RNA@meta.data %>%
  filter(plot_x > 4400 & plot_y > 3600) %>%
  pull(barcode)
# Cast to empty
P110TA_30_RNA@meta.data <- P110TA_30_RNA@meta.data %>%
  mutate(area = case_when(barcode %in% beads_to_empty ~ "Empty",
                          !(barcode %in% beads_to_empty) ~ area))

```

## Save bead area annotations
```{r, eval = FALSE}
P110TA_30_RNA@meta.data %>%
  select(barcode, area) %>%
  write.csv("Puck30_110TA_bead_area_annot.csv")
```

## Before removing TCRs in the empty areas
## Plot where all TCRs are
```{r}
beads_w_tcr <- P110TA_30_RNA@meta.data %>%
  filter(nFeature_TCR > 0)

legend <- as.ggplot(get_legend(SpatialDimPlot(P110TA_30_RNA, group.by = "area") + 
                                 scale_fill_manual(values = four_area_colors)))

plot <- SpatialDimPlot(P110TA_30_RNA, group.by = "area", alpha = 0.7, stroke = 0) +
  scale_fill_manual(values = four_area_colors) +
  geom_point(data = beads_w_tcr, aes(x = plot_x, y = plot_y), size = 0.7) +
  theme(legend.position = "None") 

plot + plot_spacer() + legend + plot_layout(widths = c(3, -1.1 ,3)) + ggtitle("TCR location")
```

## Plot where neoAg TCRs are
```{r}
beads_neoag <- P110TA_30_RNA@meta.data %>%
  filter(tumor_antigen_specific == TRUE)

legend <- as.ggplot(get_legend(SpatialDimPlot(P110TA_30_RNA, group.by = "area") + 
                                 scale_fill_manual(values = four_area_colors))) 

plot <- SpatialDimPlot(P110TA_30_RNA, group.by = "area", alpha = 0.7, stroke = 0) +
  scale_fill_manual(values = four_area_colors) +
  geom_point(data = beads_neoag, aes(x = plot_x, y = plot_y), size = 1) +
  theme(legend.position = "None")

plot + plot_spacer() + legend + plot_layout(widths = c(3, -1.1 , 3)) + ggtitle("NeoAg TCR location")
```

## Plot where TCRs are AFTER removing TCRs associated with the beads in the "empty" areas

## Remove TCRs, also remove mask columns in metadata 
```{r}
P110TA_30_RNA <- remove_tcrs_in_empty_area(P110TA_30_RNA)

P110TA_30_RNA@meta.data <- P110TA_30_RNA@meta.data %>%
  select(-c(mask_1, mask_2, mask_sum))

```

## Plot where all TCRs are
```{r}
beads_w_tcr <- P110TA_30_RNA@meta.data %>%
  filter(nFeature_TCR > 0)

legend <- as.ggplot(get_legend(SpatialDimPlot(P110TA_30_RNA, group.by = "area") + 
                                 scale_fill_manual(values = four_area_colors))) 

plot <- SpatialDimPlot(P110TA_30_RNA, group.by = "area", alpha = 0.7, stroke = 0) +
  scale_fill_manual(values = four_area_colors) +
  geom_point(data = beads_w_tcr, aes(x = plot_x, y = plot_y), size = 1) +
  theme(legend.position = "None") 

plot + plot_spacer() + legend + plot_layout(widths = c(3, -1.1 ,3)) + ggtitle("TCR location")
```

## Plot where neoAg TCRs are
```{r}
beads_neoag <- P110TA_30_RNA@meta.data %>%
  filter(tumor_antigen_specific == TRUE)

legend <- as.ggplot(get_legend(SpatialDimPlot(P110TA_30_RNA, group.by = "area") + 
                                 scale_fill_manual(values = four_area_colors))) 

plot <- SpatialDimPlot(P110TA_30_RNA, group.by = "area", alpha = 0.7, stroke = 0) +
  scale_fill_manual(values = four_area_colors) +
  geom_point(data = beads_neoag, aes(x = plot_x, y = plot_y), size = 1) +
  theme(legend.position = "None")

plot + plot_spacer() + legend + plot_layout(widths = c(3, -1.1 , 3)) + ggtitle("NeoAg TCR location")
```

## Plot where EBV reactive TCRs are
```{r}
beads_ebv <- P110TA_30_RNA@meta.data %>%
  filter(virus_antigen == "EBV")

legend <- as.ggplot(get_legend(SpatialDimPlot(P110TA_30_RNA, group.by = "area") + 
                                 scale_fill_manual(values = four_area_colors))) 

plot <- SpatialDimPlot(P110TA_30_RNA, group.by = "area", alpha = 0.7, stroke = 0) +
  scale_fill_manual(values = four_area_colors) +
  geom_point(data = beads_ebv, aes(x = plot_x, y = plot_y), size = 1) +
  theme(legend.position = "None")

plot + plot_spacer() + legend + plot_layout(widths = c(3, -1.1 , 3)) + ggtitle("EBV reactive TCR location")
```

## Plot areas
```{r}
SpatialDimPlot(P110TA_30_RNA, group.by = "area", stroke = 0.1) + scale_fill_manual(values = four_area_colors)
```

## Save puck
```{r, eval = FALSE}
saveRDS(P110TA_30_RNA, "02_Puck30_110TA_TCR_RNA_area_121125.Rds")
```

## Get session info
```{r}
sessionInfo()
```
