# Get gene sets {#sec-misc}

# Libraries
```{r}
library(Seurat)
library(SeuratWrappers)
library(ggplot2)
library(patchwork)
library(dplyr)
library(wesanderson)
library(ggplotify)
library(cowplot)
library(stringr)
library(reshape2)
library(ggrepel)
library(tidyverse)
library(infer)
library(RANN)
library(muscat)
library(limma)
library(purrr)
library(SummarizedExperiment)
library(fgsea)
library(msigdbr)
library(ProjecTILs)
library(readxl)
library(UpSetR)

source("../source_scripts/slideseq_functions_main.R")

options(Seurat.object.assay.version = 'v5')

set.seed(123)
```

# Get core T cell gene set
```{r}
core_t <- c("CD3D","CD3E","CD3G","CD247","CD4","CD8A","CD8B","TRAC","TRBC1","TRBC2")
```

# Get gene sets from projecTILs. Downloaded from ProjecTILs' Github.
## CD8
```{r}
cd8_ref <- readRDS("/n/data2/dfci/medonc/cwu/chloe/slideseq_data/tcr_analysis/RCC_neovax_pucks/analysis_110624/objects/gene_set_sources/CD8T_human_ref_v1.rds")

cd8_ref <- NormalizeData(cd8_ref, assay = "RNA")

cd8_deg <- FindAllMarkers(cd8_ref, test.use = "wilcox", group.by = "functional.cluster", assay = "RNA", logfc.threshold = 0.5, min.pct = 0.3) %>%
  filter(str_detect(gene, "^RP", negate = TRUE),
         str_detect(gene, "^MRP", negate = TRUE),
         str_detect(gene, "^MT-", negate = TRUE))

cd8_deg <- split(cd8_deg$gene, f = cd8_deg$cluster)

names(cd8_deg) <- paste0("projecTILs.", names(cd8_deg))
```

## CD4 Downloaded from ProjecTILs' Github.
```{r}
cd4_ref <- readRDS("/n/data2/dfci/medonc/cwu/chloe/slideseq_data/tcr_analysis/RCC_neovax_pucks/analysis_110624/objects/gene_set_sources/CD4T_human_ref_v1.rds")

cd4_ref <- NormalizeData(cd4_ref, assay = "RNA")

cd4_deg <- FindAllMarkers(cd4_ref, test.use = "wilcox", group.by = "functional.cluster", assay = "RNA", logfc.threshold = 0.5, min.pct = 0.3) %>%
  filter(str_detect(gene, "^RP", negate = TRUE),
         str_detect(gene, "^MRP", negate = TRUE),
         str_detect(gene, "^MT-", negate = TRUE))

cd4_deg <- split(cd4_deg$gene, f = cd4_deg$cluster)

names(cd4_deg) <- paste0("projecTILs.", names(cd4_deg))
```

# Get gene sets from "A single-cell tumor immune atlas for precision oncology"
https://pmc.ncbi.nlm.nih.gov/articles/PMC8494216/#abstract1
Supplemental table 4
```{r}
nieto_b <- read_xlsx("/n/data2/dfci/medonc/cwu/chloe/slideseq_data/tcr_analysis/RCC_neovax_pucks/analysis_110624/objects/gene_set_sources/supp_gr.273300.120_Supplemental_Table_S4.xlsx", sheet = 1, skip = 1) %>%
  filter(avg_logFC > 0.5, pct.1 > 0.3, 
         str_detect(gene, "^MRP", negate = TRUE),
         str_detect(gene, "^MT.", negate = TRUE),
         str_detect(gene, "^RP", negate = TRUE)) %>%
  mutate(gene = str_replace_all(gene, fixed("."), "-"))

nieto_b_plasma <- read_xlsx("/n/data2/dfci/medonc/cwu/chloe/slideseq_data/tcr_analysis/RCC_neovax_pucks/analysis_110624/objects/gene_set_sources/supp_gr.273300.120_Supplemental_Table_S4.xlsx", sheet = 3, skip = 1) %>%
  filter(avg_logFC > 0.5, pct.1 > 0.3, 
         str_detect(gene, "^MRP", negate = TRUE),
         str_detect(gene, "^MT.", negate = TRUE),
         str_detect(gene, "^RP", negate = TRUE)) %>%
  mutate(gene = str_replace_all(gene, fixed("."), "-"))

nieto_myeloid_prolif <- read_xlsx("/n/data2/dfci/medonc/cwu/chloe/slideseq_data/tcr_analysis/RCC_neovax_pucks/analysis_110624/objects/gene_set_sources/supp_gr.273300.120_Supplemental_Table_S4.xlsx", sheet = 20, skip = 1) %>%
  filter(avg_logFC > 0.5, pct.1 > 0.3, 
         str_detect(gene, "^MRP", negate = TRUE),
         str_detect(gene, "^MT.", negate = TRUE),
         str_detect(gene, "^RP", negate = TRUE)) %>%
  mutate(gene = str_replace_all(gene, fixed("."), "-"))

nieto_gs <- list(nieto_b = nieto_b$gene, nieto_b_plasma = nieto_b_plasma$gene, nieto_myeloid_prolif = nieto_myeloid_prolif$gene)
```

# Get gene sets from "Single-cell resolution characterization of myeloid-derived cell states with implication in cancer outcome" https://pmc.ncbi.nlm.nih.gov/articles/PMC11228020/
```{r}
guimaraes_m1 <- read_xlsx("/n/data2/dfci/medonc/cwu/chloe/slideseq_data/tcr_analysis/RCC_neovax_pucks/analysis_110624/objects/gene_set_sources/41467_2024_49916_MOESM11_ESM.xlsx", sheet = 1, skip = 2)
guimaraes_m2 <- read_xlsx("/n/data2/dfci/medonc/cwu/chloe/slideseq_data/tcr_analysis/RCC_neovax_pucks/analysis_110624/objects/gene_set_sources/41467_2024_49916_MOESM11_ESM.xlsx", sheet = 2, skip = 2)
guimaraes <- read_xlsx("/n/data2/dfci/medonc/cwu/chloe/slideseq_data/tcr_analysis/RCC_neovax_pucks/analysis_110624/objects/gene_set_sources/41467_2024_49916_MOESM8_ESM (2).xlsx", sheet = 2, skip = 2) %>%
  filter(`Macrophages Subpopulations` %in% c("Mac_AgPres", "Mac_IFN", "RTM_LA"),
         avg_log2FC > 0.5)

guimaraes_gs <- split(guimaraes$gene, f = guimaraes$`Macrophages Subpopulations`)

names(guimaraes_gs) <- paste0("Guimaraes.", c("Mac_AgPres", "Mac_IFN", "RTM_LA"))

guimaraes_gs <- c(guimaraes_gs, list(guimaraes.m1 = guimaraes_m1$`Gene symbol`, guimaraes.m2 = guimaraes_m2$`Gene symbol`))
```

# Compile gene sets
```{r}
gene_sets <- c(cd4_deg, cd8_deg, nieto_gs, guimaraes_gs, coret = list(core_t))
```

# Save gene sets
```{r, eval = FALSE}
saveRDS(gene_sets, "../objects/immune_gene_sets.Rds")
```

# Get session info
```{r}
sessionInfo()
```
