# Shared CD8 TEx TCRs: TPEX phenotype {#sec-shared-tex-tcr-phenotype}

# Phenotype of shared TEx TCRs in the TLS

# Libraries
```{r}
library(Seurat)
library(SeuratWrappers)
library(ggplot2)
library(patchwork)
library(dplyr)
library(wesanderson)
library(ggplotify)
library(cowplot)
library(stringr)
library(reshape2)
library(ggrepel)
library(tidyverse)
library(infer)
library(RANN)
library(muscat)
library(limma)
library(purrr)
library(SummarizedExperiment)
library(fgsea)
library(msigdbr)
library(ProjecTILs)
library(readxl)
library(UpSetR)
library(ggpointdensity)
library(ComplexHeatmap)
library(scales)
library(lme4)
library(escape)

source("../source_scripts/slideseq_functions_main.R")

options(Seurat.object.assay.version = 'v5')

set.seed(123)
```

# Explore the scRNA derived phenotypes of the shared clones
## Open scRNA derived phenotypes
```{r}
scrna_phenotypes <- read.csv("../objects/reformatted_RCC_slideseq_parameters_052925.csv")

scrna_exh_clones <- scrna_phenotypes %>%
  filter(PrimaryCluster == "Tex") %>%
  distinct(sctcr_clone) %>%
  pull(sctcr_clone)
```

# Load objects
```{r}
P102TA_10_RNA <- readRDS("../objects/02_Puck10_102TA_TCR_RNA_area_121125.Rds")
P102TA_33_RNA <- readRDS("../objects/02_Puck33_102TA_TCR_RNA_area_121125.Rds")

P108TA_26_RNA <- readRDS("../objects/02_Puck26_108TA_TCR_RNA_area_121125.Rds")
P108TA_29_RNA <- readRDS("../objects/02_Puck29_108TA_TCR_RNA_area_121125.Rds")
P108TA_30_RNA <- readRDS("../objects/02_Puck30_108TA_TCR_RNA_area_121125.Rds")
P108TA_32_RNA <- readRDS("../objects/02_Puck32_108TA_TCR_RNA_area_121125.Rds")
```

## Add a generalized column "tls_status"
```{r}
add_TLS_column <- function(obj){
  obj <- obj %>%
    mutate(tls_status = case_when(area %in% c("TLS1", "TLS2", "TLS3", "TLS4", "M2") ~ "inTLS",
                              area %in% c("Boundary", "Tumor"
                                          ) ~ "outTLS"))
  
  return(obj)
}

P102TA_10_RNA@meta.data <- P102TA_10_RNA@meta.data %>%
  add_TLS_column()
P102TA_33_RNA@meta.data <- P102TA_33_RNA@meta.data %>%
  add_TLS_column()

P108TA_26_RNA@meta.data <- P108TA_26_RNA@meta.data %>%
  add_TLS_column()
P108TA_29_RNA@meta.data <- P108TA_29_RNA@meta.data %>%
  add_TLS_column()
P108TA_30_RNA@meta.data <- P108TA_30_RNA@meta.data %>%
  add_TLS_column()
P108TA_32_RNA@meta.data <- P108TA_32_RNA@meta.data %>%
  add_TLS_column()
```

## Assign celltypes
```{r}
P102TA_10_RNA@meta.data <- P102TA_10_RNA@meta.data %>%
  assign_celltypes() %>%
  order_celltypes()
P102TA_10_RNA@meta.data %>%
  dplyr::count(assignment1)
P102TA_33_RNA@meta.data <- P102TA_33_RNA@meta.data %>%
  assign_celltypes() %>%
  order_celltypes()
P102TA_33_RNA@meta.data %>%
  dplyr::count(assignment1)

P108TA_26_RNA@meta.data <- P108TA_26_RNA@meta.data %>%
  assign_celltypes() %>%
  order_celltypes()
P108TA_26_RNA@meta.data %>%
  dplyr::count(assignment1)
P108TA_29_RNA@meta.data <- P108TA_29_RNA@meta.data %>%
  assign_celltypes() %>%
  order_celltypes()
P108TA_29_RNA@meta.data %>%
  dplyr::count(assignment1)
P108TA_30_RNA@meta.data <- P108TA_30_RNA@meta.data %>%
  assign_celltypes() %>%
  order_celltypes()
P108TA_30_RNA@meta.data %>%
  dplyr::count(assignment1)
P108TA_32_RNA@meta.data <- P108TA_32_RNA@meta.data %>%
  assign_celltypes() %>%
  order_celltypes()
P108TA_32_RNA@meta.data %>%
  dplyr::count(assignment1)

```

# Remove empty beads
```{r}
P102TA_10_RNA <- subset(P102TA_10_RNA, subset = area == "Empty", invert = TRUE)
P102TA_33_RNA <- subset(P102TA_33_RNA, subset = area == "Empty", invert = TRUE)

P108TA_26_RNA <- subset(P108TA_26_RNA, subset = area == "Empty", invert = TRUE)
P108TA_29_RNA <- subset(P108TA_29_RNA, subset = area == "Empty", invert = TRUE)
P108TA_30_RNA <- subset(P108TA_30_RNA, subset = area == "Empty", invert = TRUE)
P108TA_32_RNA <- subset(P108TA_32_RNA, subset = area == "Empty", invert = TRUE)
```

# Remove extra columns
## P102
```{r}
P102TA_10_RNA@meta.data <- P102TA_10_RNA@meta.data %>%
  select(-c(puck10_count, log_puck10_count, largest_clone_id))
```

# Get shared clonotypes in both the TLS and non-TLS areas per patient
## P102TA
```{r}
P102TA_md <- rbind(P102TA_10_RNA@meta.data,
                   P102TA_33_RNA@meta.data)

P102TA_tls_md <- P102TA_md %>%
  filter(tls_status == "inTLS")

P102TA_nontls_md <- P102TA_md %>%
  filter(tls_status == "outTLS")

P102TA_tls_clones_counts <- count_clones_by_annotated_area(P102TA_tls_md, get_unique_productive_clones(P102TA_tls_md$clone_id)) %>%
  mutate(area = case_when(area %in% c("TLS1", "TLS2", "TLS3", "M2") ~ "TLS",
                          T ~ "wrong")) %>%
  group_by(clone_id, area) %>%
  summarize(sum_count = sum(count)) %>%
  ungroup()
  
P102TA_nontls_clones_counts <- count_clones_by_annotated_area(P102TA_nontls_md, get_unique_productive_clones(P102TA_nontls_md$clone_id)) %>%
  mutate(area = case_when(area %in% c("Tumor", 
    "Boundary") ~ "nonTLS",
                          T ~ "wrong")) %>%
  group_by(clone_id, area) %>%
  summarize(sum_count = sum(count)) %>%
  ungroup()

# We want clones seen in 5 beads or more in each region
P102TA_tls_clones <- P102TA_tls_clones_counts %>%
  filter(sum_count >= 5) %>%
  pull(clone_id)
P102TA_nontls_clones <- P102TA_nontls_clones_counts %>%
  filter(sum_count >= 5) %>%
  pull(clone_id)

table(P102TA_tls_clones %in% P102TA_nontls_clones)
table(P102TA_nontls_clones %in% P102TA_tls_clones)

P102TA_shared_clones <- P102TA_tls_clones[P102TA_tls_clones %in% P102TA_nontls_clones]

# # Get CD8 Tex clones seen in scTCR only
P102TA_shared_clones <- P102TA_shared_clones[P102TA_shared_clones %in% scrna_exh_clones]

length(P102TA_shared_clones)
```

## P108TA
```{r}
P108TA_md <- rbind(P108TA_26_RNA@meta.data, P108TA_29_RNA@meta.data, P108TA_30_RNA@meta.data, P108TA_32_RNA@meta.data)

P108TA_tls_md <- P108TA_md %>%
  filter(tls_status == "inTLS")

P108TA_nontls_md <- P108TA_md %>%
  filter(tls_status == "outTLS")

P108TA_tls_clones_counts <- count_clones_by_annotated_area(P108TA_tls_md, get_unique_productive_clones(P108TA_tls_md$clone_id)) %>%
  mutate(area = case_when(area %in% c("TLS1", "TLS2", "TLS3", "M2") ~ "TLS",
                          T ~ "wrong")) %>%
  group_by(clone_id, area) %>%
  summarize(sum_count = sum(count)) %>%
  ungroup()
  
P108TA_nontls_clones_counts <- count_clones_by_annotated_area(P108TA_nontls_md, get_unique_productive_clones(P108TA_nontls_md$clone_id)) %>%
  mutate(area = case_when(area %in% c("Tumor", 
    "Boundary") ~ "nonTLS",
                          T ~ "wrong")) %>%
  group_by(clone_id, area) %>%
  summarize(sum_count = sum(count)) %>%
  ungroup()

# We want clones seen in 5 beads or more in each region
P108TA_tls_clones <- P108TA_tls_clones_counts %>%
  filter(sum_count >= 5) %>%
  pull(clone_id)
P108TA_nontls_clones <- P108TA_nontls_clones_counts %>%
  filter(sum_count >= 5) %>%
  pull(clone_id)

table(P108TA_tls_clones %in% P108TA_nontls_clones)
table(P108TA_nontls_clones %in% P108TA_tls_clones)

P108TA_shared_clones <- P108TA_tls_clones[P108TA_tls_clones %in% P108TA_nontls_clones]

# # Get CD8 Tex clones seen in scTCR only
P108TA_shared_clones <- P108TA_shared_clones[P108TA_shared_clones %in% scrna_exh_clones]

length(P108TA_shared_clones)
```

#### Which clones are we taking
```{r}
scrna_phenotypes %>% filter(sctcr_clone %in% P102TA_shared_clones) %>% distinct(sctcr_clone, identity, PrimaryCluster)

scrna_phenotypes %>% filter(sctcr_clone %in% P108TA_shared_clones) %>% distinct(sctcr_clone, identity, PrimaryCluster)

```

# Determine shared CD8 Tex clone frequency
```{r}
count_clones_by_annotated_area(P102TA_md, P102TA_shared_clones) %>%
  group_by(clone_id) %>%
  summarize(sum_count = sum(count)) %>%
  arrange(sum_count)

count_clones_by_annotated_area(P108TA_md, P108TA_shared_clones) %>%
  group_by(clone_id) %>%
  summarize(sum_count = sum(count)) %>%
  arrange(sum_count)
```

# Set shared clone order based increasing frequency of clones across all pucks
```{r}
P102TA_shared_clones <- factor(P102TA_shared_clones, levels = c("p102_295", "p102_289", "p102_737", "p102_266", "p102_278", "p102_274", "p102_264", "p102_265"))

P108TA_shared_clones <- factor(P108TA_shared_clones, levels = c("p108_3", "p108_266", "p108_211", "p108_1", "p108_207", "p108_206", "p108_208", "p108_205", "p108_4"))
```

# Cast beads with multiple shared CD8Tex specific clones to the smaller clone
```{r}
cast_shared_cd8s_102 <- function(md){
  md <- md %>%
    mutate(shared_tex_id = case_when(str_detect(sctcr_clone, "p102_295($|,)") ~ "p102_295",
                                  str_detect(sctcr_clone, "p102_289($|,)") ~ "p102_289",
                                  str_detect(sctcr_clone, "p102_737($|,)") ~ "p102_737",
                                  str_detect(sctcr_clone, "p102_266($|,)") ~ "p102_266",
                                  str_detect(sctcr_clone, "p102_278($|,)") ~ "p102_278",
                                  str_detect(sctcr_clone, "p102_274($|,)") ~ "p102_274",
                                  str_detect(sctcr_clone, "p102_264($|,)") ~ "p102_264",
                                  str_detect(sctcr_clone, "p102_265($|,)") ~ "p102_265"))
  return(md)
}

cast_shared_cd8s_108 <- function(md){
  md <- md %>%
    mutate(shared_tex_id = case_when(str_detect(sctcr_clone, "p108_3($|,)") ~ "p108_3",
                                  str_detect(sctcr_clone, "p108_266($|,)") ~ "p108_266",
                                  str_detect(sctcr_clone, "p108_211($|,)") ~ "p108_211",
                                  str_detect(sctcr_clone, "p108_1($|,)") ~ "p108_1",
                                  str_detect(sctcr_clone, "p108_207($|,)") ~ "p108_207",
                                  str_detect(sctcr_clone, "p108_206($|,)") ~ "p108_206",
                                  str_detect(sctcr_clone, "p108_208($|,)") ~ "p108_208",
                                  str_detect(sctcr_clone, "p108_205($|,)") ~ "p108_205",
                                  str_detect(sctcr_clone, "p108_4($|,)") ~ "p108_4"))
  return(md)
}

P102TA_10_RNA@meta.data <- P102TA_10_RNA@meta.data %>%
  cast_shared_cd8s_102
P102TA_33_RNA@meta.data <- P102TA_33_RNA@meta.data %>%
  cast_shared_cd8s_102

P108TA_26_RNA@meta.data <- P108TA_26_RNA@meta.data %>%
  cast_shared_cd8s_108
P108TA_29_RNA@meta.data <- P108TA_29_RNA@meta.data %>%
  cast_shared_cd8s_108
P108TA_30_RNA@meta.data <- P108TA_30_RNA@meta.data %>%
  cast_shared_cd8s_108
P108TA_32_RNA@meta.data <- P108TA_32_RNA@meta.data %>%
  cast_shared_cd8s_108
```

# P102TA
## Puck 10

### Get beads with shared clones in TLS
```{r}
# Get beads with shared clones in TLS
P102TA_10_inTLS <- P102TA_10_RNA@meta.data %>%
  filter(str_detect(clone_id, paste0(paste0(P102TA_shared_clones, "($|,)"), collapse = "|")),
         tls_status == "inTLS") %>%
  pull(barcode)

# Get beads with shared clones outside of TLS
P102TA_10_outTLS <- P102TA_10_RNA@meta.data %>%
  filter(str_detect(clone_id, paste0(paste0(P102TA_shared_clones, "($|,)"), collapse = "|")),
         tls_status == "outTLS") %>%
  pull(barcode)

# How many cells in each category?
length(P102TA_10_inTLS)
length(P102TA_10_outTLS)

```

### Attach assignments to original Seurat object and filter for defined beads
```{r}
P102TA_10_RNA@meta.data <- P102TA_10_RNA@meta.data %>%
  mutate(type = case_when(barcode %in% P102TA_10_inTLS ~ "shared,inTLS",
                          barcode %in% P102TA_10_outTLS ~ "shared,outTLS"),
         muscat = case_when(!is.na(type) ~ "Defined",
                             is.na(type) ~ "Undefined"))

P102TA_10_RNA <- subset(P102TA_10_RNA, subset = muscat == "Defined")
```

### Get beads with shared clones in TLS
```{r}
# Get beads with shared clones in TLS
P102TA_33_inTLS <- P102TA_33_RNA@meta.data %>%
  filter(str_detect(clone_id, paste0(paste0(P102TA_shared_clones, "($|,)"), collapse = "|")),
         tls_status == "inTLS") %>%
  pull(barcode)

# Get beads with shared clones outside of TLS
P102TA_33_outTLS <- P102TA_33_RNA@meta.data %>%
  filter(str_detect(clone_id, paste0(paste0(P102TA_shared_clones, "($|,)"), collapse = "|")),
         tls_status == "outTLS") %>%
  pull(barcode)

# How many cells in each category?
length(P102TA_33_inTLS)
length(P102TA_33_outTLS)

```

### Attach assignments to original Seurat object and filter for defined beads
```{r}
P102TA_33_RNA@meta.data <- P102TA_33_RNA@meta.data %>%
  mutate(type = case_when(barcode %in% P102TA_33_inTLS ~ "shared,inTLS",
                          barcode %in% P102TA_33_outTLS ~ "shared,outTLS"),
         muscat = case_when(!is.na(type) ~ "Defined",
                             is.na(type) ~ "Undefined"))

P102TA_33_RNA <- subset(P102TA_33_RNA, subset = muscat == "Defined")
```

# P108TA
## Puck 26
### Get beads with shared clones in TLS
```{r}
# Get beads with shared clones in TLS
P108TA_26_inTLS <- P108TA_26_RNA@meta.data %>%
  filter(str_detect(clone_id, paste0(paste0(P108TA_shared_clones, "($|,)"), collapse = "|")),
         tls_status == "inTLS") %>%
  pull(barcode)

# Get beads with shared clones outside of TLS
P108TA_26_outTLS <- P108TA_26_RNA@meta.data %>%
  filter(str_detect(clone_id, paste0(paste0(P108TA_shared_clones, "($|,)"), collapse = "|")),
         tls_status == "outTLS") %>%
  pull(barcode)

# How many cells in each category?
length(P108TA_26_inTLS)
length(P108TA_26_outTLS)

```

### Attach assignments to original Seurat object and filter for defined beads
```{r}
P108TA_26_RNA@meta.data <- P108TA_26_RNA@meta.data %>%
  mutate(type = case_when(barcode %in% P108TA_26_inTLS ~ "shared,inTLS",
                          barcode %in% P108TA_26_outTLS ~ "shared,outTLS"),
         muscat = case_when(!is.na(type) ~ "Defined",
                             is.na(type) ~ "Undefined"))

P108TA_26_RNA <- subset(P108TA_26_RNA, subset = muscat == "Defined")
```

## Puck 29
### Get beads with shared clones in TLS
```{r}
# Get beads with shared clones in TLS
P108TA_29_inTLS <- P108TA_29_RNA@meta.data %>%
  filter(str_detect(clone_id, paste0(paste0(P108TA_shared_clones, "($|,)"), collapse = "|")),
         tls_status == "inTLS") %>%
  pull(barcode)

# Get beads with shared clones outside of TLS
P108TA_29_outTLS <- P108TA_29_RNA@meta.data %>%
  filter(str_detect(clone_id, paste0(paste0(P108TA_shared_clones, "($|,)"), collapse = "|")),
         tls_status == "outTLS") %>%
  pull(barcode)

# How many cells in each category?
length(P108TA_29_inTLS)
length(P108TA_29_outTLS)

```

### Attach assignments to original Seurat object and filter for defined beads
```{r}
P108TA_29_RNA@meta.data <- P108TA_29_RNA@meta.data %>%
  mutate(type = case_when(barcode %in% P108TA_29_inTLS ~ "shared,inTLS",
                          barcode %in% P108TA_29_outTLS ~ "shared,outTLS"),
         muscat = case_when(!is.na(type) ~ "Defined",
                             is.na(type) ~ "Undefined"))

P108TA_29_RNA <- subset(P108TA_29_RNA, subset = muscat == "Defined")
```

## Puck 30
### Get beads with shared clones in TLS
```{r}
# Get beads with shared clones in TLS
P108TA_30_inTLS <- P108TA_30_RNA@meta.data %>%
  filter(str_detect(clone_id, paste0(paste0(P108TA_shared_clones, "($|,)"), collapse = "|")),
         tls_status == "inTLS") %>%
  pull(barcode)

# Get beads with shared clones outside of TLS
P108TA_30_outTLS <- P108TA_30_RNA@meta.data %>%
  filter(str_detect(clone_id, paste0(paste0(P108TA_shared_clones, "($|,)"), collapse = "|")),
         tls_status == "outTLS") %>%
  pull(barcode)

# How many cells in each category?
length(P108TA_30_inTLS)
length(P108TA_30_outTLS)

```

### Attach assignments to original Seurat object and filter for defined beads
```{r}
P108TA_30_RNA@meta.data <- P108TA_30_RNA@meta.data %>%
  mutate(type = case_when(barcode %in% P108TA_30_inTLS ~ "shared,inTLS",
                          barcode %in% P108TA_30_outTLS ~ "shared,outTLS"),
         muscat = case_when(!is.na(type) ~ "Defined",
                             is.na(type) ~ "Undefined"))

P108TA_30_RNA <- subset(P108TA_30_RNA, subset = muscat == "Defined")
```

## Puck 32
### Get beads with shared clones in TLS
```{r}
# Get beads with shared clones in TLS
P108TA_32_inTLS <- P108TA_32_RNA@meta.data %>%
  filter(str_detect(clone_id, paste0(paste0(P108TA_shared_clones, "($|,)"), collapse = "|")),
         tls_status == "inTLS") %>%
  pull(barcode)

# Get beads with shared clones outside of TLS
P108TA_32_outTLS <- P108TA_32_RNA@meta.data %>%
  filter(str_detect(clone_id, paste0(paste0(P108TA_shared_clones, "($|,)"), collapse = "|")),
         tls_status == "outTLS") %>%
  pull(barcode)

# How many cells in each category?
length(P108TA_32_inTLS)
length(P108TA_32_outTLS)

```

### Attach assignments to original Seurat object and filter for defined beads
```{r}
P108TA_32_RNA@meta.data <- P108TA_32_RNA@meta.data %>%
  mutate(type = case_when(barcode %in% P108TA_32_inTLS ~ "shared,inTLS",
                          barcode %in% P108TA_32_outTLS ~ "shared,outTLS"),
         muscat = case_when(!is.na(type) ~ "Defined",
                             is.na(type) ~ "Undefined"))

P108TA_32_RNA <- subset(P108TA_32_RNA, subset = muscat == "Defined")
```

# Process RNA data
## Merge RNA objects
```{r}
TLS_pucks <- merge(x = P102TA_10_RNA, y = list(P102TA_33_RNA, P108TA_26_RNA, P108TA_29_RNA, P108TA_30_RNA, P108TA_32_RNA))

# Update metadata
TLS_pucks@meta.data <- TLS_pucks@meta.data %>%
  mutate(tls_status = str_split_i(type, ",", 2),
         Patient = str_split_i(orig.ident, "_", 1),
         area_final = case_when(area %in% c("TLS1", "TLS2", "TLS3", "TLS4", "M2") ~ "TLS",
                                T ~ area))

TLS_pucks[['Spatial']] <- JoinLayers(TLS_pucks[['Spatial']])

TLS_pucks@meta.data %>%
  dplyr::count(tls_status, Patient)
```

## Open the TPEX gene set
```{r}
gene_sets <- readRDS("../objects/immune_gene_sets.Rds")

new_gene_sets <- gene_sets[(names(gene_sets) %in% c("projecTILs.CD8.TPEX", "coret"))]
```

# Run gene set enrichment analysis
```{r}
TLS_pucks <- RenameAssays(TLS_pucks, assay.name = "Spatial", new.assay.name = "RNA")

TLS_pucks <- runEscape(input.data = TLS_pucks, 
                       method = "GSVA",
                       gene.sets = new_gene_sets, 
                       groups = 1000, 
                       min.size = 5, 
                       new.assay.name="escape.GSVA")
```

## Format raw GSVA enrichment scores from escape
```{r}
raw_dat <- as.data.frame(TLS_pucks@assays$escape.GSVA$data) %>%
  t() %>%
  as.data.frame()

# Remove "-"s and remove spaces from data and from gene sets
names(new_gene_sets) <- str_replace_all(names(new_gene_sets), "_", "")

colnames(raw_dat) <- str_replace_all(names(raw_dat), "-", "")

# Get dataframe
raw_md <- raw_dat %>%
  rownames_to_column("barcode") %>%
  full_join(TLS_pucks@meta.data, by = "barcode") %>%
  mutate(tls_status = factor(tls_status))

raw_md$tls_status <- relevel(raw_md$tls_status, ref = "outTLS")

# Select columns to use
raw_md <- raw_md[,c(1:3, 4, 37, 40, 43)]
```

### Plot TCRs

#### Get means, SD and SE
```{r}
sd.df <- raw_md %>%
  # aggregate by puck
  group_by(tls_status, shared_tex_id, orig.ident, Patient) %>% 
  summarize(mean_coret_p = round(mean(coret), 8),
            mean_tpex_p = round(mean(projecTILs.CD8.TPEX), 8)) %>%
  # aggregate across pucks
  group_by(tls_status, shared_tex_id, Patient) %>%
  summarize(n = n(),
            sd_coret = round(sd(mean_coret_p), 8),
            sd_tpex = round(sd(mean_tpex_p), 8),
            se_coret = round(sd_coret/sqrt(n), 8),
            se_tpex = round(sd_tpex/sqrt(n), 8),
            mean_coret = round(mean(mean_coret_p), 8),
            mean_tpex = round(mean(mean_tpex_p), 8)) %>%
  ungroup()
```

#### Pivot for plots
```{r}
long_mean.df <- sd.df %>%
  select(tls_status, shared_tex_id, contains("mean"), Patient) %>%
  pivot_longer(cols = !c(shared_tex_id, tls_status, Patient), names_to = "signature", values_to = "score") %>%
  mutate(signature = str_replace(signature, "mean_", ""))

long_se.df <- sd.df %>%
  select(tls_status, shared_tex_id, contains("se"), Patient) %>%
  pivot_longer(cols = !c(shared_tex_id, tls_status, Patient), names_to = "signature", values_to = "se") %>%
  mutate(signature = str_replace(signature, "se_", ""))

long_merged <- long_mean.df %>%
  full_join(long_se.df, by = c("shared_tex_id", "tls_status", "signature", "Patient")) %>%
  rowwise() %>%
  mutate(upper = score+se,
         lower = score-se,
         tls_status = factor(tls_status, levels = c("inTLS", "outTLS")))
```

#### Plot
Visually both the TPEX phenotype and core T cell phenotype are enriched inTLS
```{r, fig.width = 4, fig.height = 3}
long_merged %>%
  filter(signature == "tpex") %>%
ggplot(aes(x = tls_status, y = score)) +
  geom_line(aes(group = (shared_tex_id)), position=position_dodge(width=0.2)) +
  geom_errorbar(aes(ymin=lower, ymax=upper, group = shared_tex_id), width=.2, position=position_dodge(width=0.2), color = "grey50", linewidth = 0.3) +
  geom_point(aes(group = shared_tex_id, fill = tls_status), color = "black",
             position=position_dodge(width=0.2), size = 3, shape = 22) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  scale_fill_manual(values = c("#b2c4b9", "#d0b39e")) +
  ylab("Raw enrichment (GSVA)") +
  guides(color = "none") +
  facet_wrap(~Patient, scales = "free") +
  ggtitle("TPEX enrichment values")

long_merged %>%
  filter(signature == "coret") %>%
ggplot(aes(x = tls_status, y = score)) +
  geom_line(aes(group = (shared_tex_id)), position=position_dodge(width=0.2)) +
  geom_errorbar(aes(ymin=lower, ymax=upper, group = shared_tex_id), width=.2, position=position_dodge(width=0.2), color = "grey50", linewidth = 0.3) +
  geom_point(aes(group = shared_tex_id, fill = tls_status), color = "black",
             position=position_dodge(width=0.2), size = 3, shape = 22) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  scale_fill_manual(values = c("#b2c4b9", "#d0b39e")) +
  ylab("Raw enrichment (GSVA)") +
  guides(color = "none") +
  facet_wrap(~Patient, scales = "free") +
  ggtitle("Core T enrichment values")
```

## Normalize enrichment values 
```{r}
TLS_pucks <- performNormalization(input.data = TLS_pucks, 
                               assay = "escape.GSVA", 
                               gene.sets = new_gene_sets,
                               scale.factor = TLS_pucks$nFeature_RNA,
                               make.positive = T)
```

### Format results from escape
```{r}
norm_dat <- as.data.frame(TLS_pucks@assays$escape.GSVA_normalized$data) %>%
  t() %>%
  as.data.frame()

# Remove "-"s and remove spaces from data and from gene sets
names(new_gene_sets) <- str_replace_all(names(new_gene_sets), "_", "")

colnames(norm_dat) <- str_replace_all(names(norm_dat), "-", "")

# Get dataframe
md <- norm_dat %>%
  rownames_to_column("barcode") %>%
  full_join(TLS_pucks@meta.data, by = "barcode") %>%
  mutate(tls_status = factor(tls_status))

md$tls_status <- relevel(md$tls_status, ref = "outTLS")

# Select columns to use
md <- md[,c(1:3, 4, 37, 40, 43)]
```

# Run permutation test
#### Hypothesis: Do shared Tex clones found in the TLS have a higher TPEX phenotype than shared clones found outside the TLS

#### Permutation: Randomly swap the location (eg. inTLS) label without replacement, of shared clones both inside and outside the TLS
###### - Swap location label within pucks and TCR IDs (same number of “inTLS” beads in “TCR 1” across all permutations)

#### Test statistic: T value from linear model

# Get observed data
```{r}
# Get observed model of each gene set
model_obs <- vector("list", length(new_gene_sets))
names(model_obs) <- names(new_gene_sets)
# Get observed T statistic for each gene set
t_obs <- vector("list", length(new_gene_sets))
names(model_obs) <- names(new_gene_sets)
names(t_obs) <- names(new_gene_sets)

for (gs in names(new_gene_sets)) {
  print(gs)
  # build formula
  f <- as.formula(paste(gs, "~ tls_status + (1 | Patient / orig.ident)"))
  
  model <- lmer(f, data = md, REML = TRUE)
  t_val <- coef(summary(model))["tls_statusinTLS","t value"]
  
  model_obs[[gs]] <- model
  t_obs[[gs]] <- t_val
}
```

# Permute data
```{r}
n_perm_safe = 1500

set.seed(123)
# Swap location labels within pucks and clonotypes
md_perm <- md %>%
    rep_sample_n(size = nrow(md), replace = FALSE, reps = n_perm_safe) %>%
    group_by(replicate, orig.ident, shared_tex_id
             ) %>%
    mutate(perm_tls_status = sample(tls_status, size = n(), replace = FALSE))

```

## Get permuted test statistic
```{r}
# Get permuted test statistic for each gene set
model_perm <- vector("list", length(new_gene_sets))
names(model_perm) <- names(new_gene_sets)

n_perm <- 1000

# Count non-convergences
non_convergences <- vector("list", length(new_gene_sets))
names(non_convergences) <- names(new_gene_sets)

# Count times tested
tested <- vector("list", length(new_gene_sets))
names(tested) <- names(new_gene_sets)

# Loop across each gene set
for (gs in names(new_gene_sets)) {
  n_run <- 1
  n_saved <- 1
  non_conv <- 0
  
  print(gs)
  t_val_perm.vec <- c()
  # Do 1000 converging loops
  while(n_saved <= n_perm){
    # Print every saved 100 perms
    if (n_saved %% 100 == 0) { print(n_saved) }
    # Get the permuted replicate
    md_rep <- md_perm %>%
      filter(replicate == n_run)
    # Set runs as good
    bad_run <- FALSE
    # Attempt the model
    model <- withCallingHandlers(
      expr = {
        # Build formula with permutated TLs status
        f <- as.formula(paste(gs, "~ perm_tls_status + (1 | Patient/orig.ident)"))
        # Build linear model on permuted data
        lmer(f, data = md_rep, REML = TRUE)},
      warning = function(w) {
      # Note if the model can't converge
      bad_run <<- TRUE
      non_conv <<- non_conv + 1
      n_run <<- n_run + 1
      print("Couldn't converge")
      invokeRestart("muffleWarning")
      }
    )
      
    # Save the t val from the model if it has no warning messages
    if(!bad_run){
      # Get and save t val
      t_val_perm <- coef(summary(model))["perm_tls_statusinTLS","t value"]
      t_val_perm.vec <- c(t_val_perm.vec, t_val_perm)
      # Count that we've saved a run
      n_saved <- n_saved + 1
      n_run <- n_run + 1
    } else {
      print("Skipping model")
    }
  }
  model_perm[[gs]] <- t_val_perm.vec
  non_convergences[[gs]] <- non_conv
}

```

## Format permuted results
```{r}
df_perm <- data.frame(gene_sets = rep(names(model_perm), times = lengths(model_perm)),
                      perm_ts = unlist(model_perm, use.names = FALSE))

df_obs <- data.frame(gene_sets = rep(names(t_obs), times = lengths(t_obs)),
                     obs_ts = unlist(t_obs, use.names = FALSE)) 
```

## Plot permuted test statistic distribution and observed test statistic
```{r, fig.height = 5, fig.width = 8}
df_perm %>%
  ggplot(aes(x = perm_ts)) +
  geom_histogram(bins = 100) +
  facet_wrap(~gene_sets, scales = "free", nrow = 2) +
  geom_vline(data = df_obs, aes(xintercept = obs_ts))
```

## Get one sided pval
```{r}
os_pval.vec <- c()

for(gs in unique(df_obs$gene_sets)){
  obs_val <- df_obs %>%
    filter(gene_sets == gs) %>%
    pull(obs_ts)
  
  perm_val <- df_perm %>%
    filter(gene_sets == gs) %>%
    pull(perm_ts)
  
  # One sided
  n_extreme <- length(perm_val[perm_val >= obs_val])
  pval <- n_extreme/n_perm
  
  os_pval.vec <- c(os_pval.vec, pval)
}

names(os_pval.vec) <- unique(df_obs$gene_sets)

# Replace 0 with small number
os_pval.vec[os_pval.vec == 0] <- 1/(n_perm)

os_pval.vec
```

## Plot observed mean for each puck
```{r}
mean.df <- md %>%
  # aggregate by puck
  group_by(tls_status, orig.ident, Patient) %>% 
  summarize(across(where(is.numeric), ~ mean(.x, na.rm = TRUE)),
            .groups = "drop") %>%
  ungroup()
  
mean.df %>%
  mutate(tls_status = factor(tls_status, levels = c("inTLS", "outTLS"))) %>%
  pivot_longer(cols = !c("tls_status", "orig.ident", "Patient"), values_to = "score", names_to = "signature") %>%
ggplot(aes(x = tls_status, y = score)) +
  geom_boxplot(aes(fill = tls_status)) +
  geom_line(aes(group = orig.ident)) +
  geom_point(aes(group = orig.ident, color = Patient), size = 3, shape = 15) +
  geom_point(aes(group = orig.ident), color = "black", size = 3, shape = 22) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  scale_fill_manual(values = c("#b2c4b9", "#d0b39e")) +
  scale_color_manual(values = c("#FF9300", "#009193")) +
  ylab("Normalized enrichment (GSVA)") +
  xlab("") +
  facet_wrap(~signature, scales = "free_y") +
  guides(fill="none")

```

# Get session info
```{r}
sessionInfo()
```

