# Shared neoantigen-specific TCRs: TPEX phenotype {#sec-shared-neoag-tcr-phenotype}

# Phenotype of shared neoantigen-specific TCRs in the TLS

# Libraries
```{r}
library(Seurat)
library(SeuratWrappers)
library(ggplot2)
library(patchwork)
library(dplyr)
library(wesanderson)
library(ggplotify)
library(cowplot)
library(stringr)
library(reshape2)
library(ggrepel)
library(tidyverse)
library(infer)
library(RANN)
library(muscat)
library(limma)
library(purrr)
library(SummarizedExperiment)
library(fgsea)
library(msigdbr)
library(ProjecTILs)
library(readxl)
library(UpSetR)
library(ggpointdensity)
library(ComplexHeatmap)
library(scales)
library(escape)
library(lme4)

source("../source_scripts/slideseq_functions_main.R")

options(Seurat.object.assay.version = 'v5')

colors <- structure(c("white", paletteer::paletteer_d("khroma::smoothrainbow")), names = as.character(c(0:34)))

set.seed(123)
```

# Set radius to explore
```{r}
rad <- 15
```

# Load objects
```{r}
P108TA_26_RNA <- readRDS("../objects/02_Puck26_108TA_TCR_RNA_area_121125.Rds")
P108TA_29_RNA <- readRDS("../objects/02_Puck29_108TA_TCR_RNA_area_121125.Rds")
P108TA_30_RNA <- readRDS("../objects/02_Puck30_108TA_TCR_RNA_area_121125.Rds")
P108TA_32_RNA <- readRDS("../objects/02_Puck32_108TA_TCR_RNA_area_121125.Rds")

```

## Add a generalized column "tls_status"
```{r}
add_TLS_column <- function(obj){
  obj <- obj %>%
    mutate(tls_status = case_when(area %in% c("TLS1", "TLS2", "TLS3", "TLS4", "M2") ~ "inTLS",
                              area %in% c("Boundary", "Tumor") ~ "outTLS"))
  
  return(obj)
}

P108TA_26_RNA@meta.data <- P108TA_26_RNA@meta.data %>%
  add_TLS_column()
P108TA_29_RNA@meta.data <- P108TA_29_RNA@meta.data %>%
  add_TLS_column()
P108TA_30_RNA@meta.data <- P108TA_30_RNA@meta.data %>%
  add_TLS_column()
P108TA_32_RNA@meta.data <- P108TA_32_RNA@meta.data %>%
  add_TLS_column()
```

## Assign celltypes
```{r}
P108TA_26_RNA@meta.data <- P108TA_26_RNA@meta.data %>%
  assign_celltypes() %>%
  order_celltypes()
P108TA_26_RNA@meta.data %>%
  dplyr::count(assignment1)
P108TA_29_RNA@meta.data <- P108TA_29_RNA@meta.data %>%
  assign_celltypes() %>%
  order_celltypes()
P108TA_29_RNA@meta.data %>%
  dplyr::count(assignment1)
P108TA_30_RNA@meta.data <- P108TA_30_RNA@meta.data %>%
  assign_celltypes() %>%
  order_celltypes()
P108TA_30_RNA@meta.data %>%
  dplyr::count(assignment1)
P108TA_32_RNA@meta.data <- P108TA_32_RNA@meta.data %>%
  assign_celltypes() %>%
  order_celltypes()
P108TA_32_RNA@meta.data %>%
  dplyr::count(assignment1)
```

# Remove empty beads
```{r}
P108TA_26_RNA <- subset(P108TA_26_RNA, subset = area == "Empty", invert = TRUE)
P108TA_29_RNA <- subset(P108TA_29_RNA, subset = area == "Empty", invert = TRUE)
P108TA_30_RNA <- subset(P108TA_30_RNA, subset = area == "Empty", invert = TRUE)
P108TA_32_RNA <- subset(P108TA_32_RNA, subset = area == "Empty", invert = TRUE)
```

# Explore the scRNA derived phenotypes of the shared clones
## Open scRNA derived phenotypes
```{r}
scrna_phenotypes <- read.csv("../objects/reformatted_RCC_slideseq_parameters_052925.csv")

neoantigen_clones <- scrna_phenotypes %>%
  filter(antigen == "NSUN5 p.Q18K") %>%
  distinct(sctcr_clone) %>%
  pull(sctcr_clone)
```

# Get shared clonotypes in both the TLS and non-TLS areas per patient

## P108TA
```{r}
P108TA_md <- rbind(P108TA_26_RNA@meta.data, P108TA_29_RNA@meta.data, P108TA_30_RNA@meta.data, P108TA_32_RNA@meta.data)

P108TA_tls_md <- P108TA_md %>%
  filter(tls_status == "inTLS")

P108TA_nontls_md <- P108TA_md %>%
  filter(tls_status == "outTLS")

P108TA_tls_clones_counts <- count_clones_by_annotated_area(P108TA_tls_md, get_unique_productive_clones(P108TA_tls_md$clone_id)) %>%
  mutate(area = case_when(area %in% c("TLS1", "TLS2", "TLS3", "M2") ~ "TLS",
                          T ~ "wrong")) %>%
  group_by(clone_id, area) %>%
  summarize(sum_count = sum(count)) %>%
  ungroup()
  
P108TA_nontls_clones_counts <- count_clones_by_annotated_area(P108TA_nontls_md, get_unique_productive_clones(P108TA_nontls_md$clone_id)) %>%
  mutate(area = case_when(area %in% c("Tumor", "Boundary") ~ "nonTLS",
                          T ~ "wrong")) %>%
  group_by(clone_id, area) %>%
  summarize(sum_count = sum(count)) %>%
  ungroup()

# We want neoantigen clones seen in 3 beads or more in each region
P108TA_tls_clones <- P108TA_tls_clones_counts %>%
  filter(sum_count >= 3,
         clone_id %in% neoantigen_clones) %>%
  pull(clone_id)
P108TA_nontls_clones <- P108TA_nontls_clones_counts %>%
  filter(sum_count >= 3,
         clone_id %in% neoantigen_clones) %>%
  pull(clone_id)

table(P108TA_tls_clones %in% P108TA_nontls_clones)
table(P108TA_nontls_clones %in% P108TA_tls_clones)

P108TA_shared_clones <- P108TA_tls_clones[P108TA_tls_clones %in% P108TA_nontls_clones]
```

# Set neoantigen clone order based increasing frequency of clones across all pucks
Determined earlier
```{r}
P108TA_shared_clones <- factor(P108TA_shared_clones, levels = c("p108_215", "p108_266", "p108_1", "p108_208", "p108_205"))
```

# Cast beads with multiple neoantigen specific clones to the smaller clone
```{r}
P108TA_26_RNA@meta.data <- P108TA_26_RNA@meta.data %>%
  mutate(neoantigen_id = case_when(str_detect(sctcr_clone, "p108_215($|,)") ~ "p108_215",
                                  str_detect(sctcr_clone, "p108_266($|,)") ~ "p108_266",
                                  str_detect(sctcr_clone, "p108_1($|,)") ~ "p108_1",
                                  str_detect(sctcr_clone, "p108_208($|,)") ~ "p108_208",
                                  str_detect(sctcr_clone, "p108_205($|,)") ~ "p108_205"))
P108TA_29_RNA@meta.data <- P108TA_29_RNA@meta.data %>%
  mutate(neoantigen_id = case_when(str_detect(sctcr_clone, "p108_215($|,)") ~ "p108_215",
                                  str_detect(sctcr_clone, "p108_266($|,)") ~ "p108_266",
                                  str_detect(sctcr_clone, "p108_1($|,)") ~ "p108_1",
                                  str_detect(sctcr_clone, "p108_208($|,)") ~ "p108_208",
                                  str_detect(sctcr_clone, "p108_205($|,)") ~ "p108_205"))
P108TA_30_RNA@meta.data <- P108TA_30_RNA@meta.data %>%
  mutate(neoantigen_id = case_when(str_detect(sctcr_clone, "p108_215($|,)") ~ "p108_215",
                                  str_detect(sctcr_clone, "p108_266($|,)") ~ "p108_266",
                                  str_detect(sctcr_clone, "p108_1($|,)") ~ "p108_1",
                                  str_detect(sctcr_clone, "p108_208($|,)") ~ "p108_208",
                                  str_detect(sctcr_clone, "p108_205($|,)") ~ "p108_205"))
P108TA_32_RNA@meta.data <- P108TA_32_RNA@meta.data %>%
  mutate(neoantigen_id = case_when(str_detect(sctcr_clone, "p108_215($|,)") ~ "p108_215",
                                  str_detect(sctcr_clone, "p108_266($|,)") ~ "p108_266",
                                  str_detect(sctcr_clone, "p108_1($|,)") ~ "p108_1",
                                  str_detect(sctcr_clone, "p108_208($|,)") ~ "p108_208",
                                  str_detect(sctcr_clone, "p108_205($|,)") ~ "p108_205"))
```

# P108TA
## Puck 26

### Get beads near each shared neoantigen clone inside the TLS
```{r}
# Get the beads near shared neoantigen clones
inTLS_list <- list()

for(clone in P108TA_shared_clones){
  result_list <- get_beads_near_specific_neoantigen_TCR2(P108TA_26_RNA, "inTLS", clone, rad)
  inTLS_list <- append(inTLS_list, list(result_list))
}

# Rename lists
names(inTLS_list) <- P108TA_shared_clones
```

### Get beads near each shared neoantigen clone outside the TLS
```{r}
# Get the beads near shared neoantigen clones
outTLS_list <- list()

for(clone in P108TA_shared_clones){
  result_list <- get_beads_near_specific_neoantigen_TCR2(P108TA_26_RNA, "outTLS", clone, rad)
  outTLS_list <- append(outTLS_list, list(result_list))
}

# Rename lists
names(outTLS_list) <- P108TA_shared_clones
```

### Check for overlaps
#### in TLS vs in TLS
```{r}
in_vs_in.df <- data.frame()
for(i in 1:length(P108TA_shared_clones)) {
	for(j in 1:length(P108TA_shared_clones)) {
	  if(i < j){
  		same_beads <- length(intersect(inTLS_list[[i]]$near_beads$barcode, inTLS_list[[j]]$near_beads$barcode))
  		new_row <- c(names(inTLS_list)[i], names(inTLS_list)[j], same_beads)
  		in_vs_in.df <- rbind(in_vs_in.df, new_row)
	  }
	}
}

colnames(in_vs_in.df) <- c("neoantigen1", "neoantigen2", "beads")

# Reformat
in_vs_in.df <- in_vs_in.df %>%
	pivot_wider(names_from=neoantigen2,values_from=beads) %>%
	column_to_rownames("neoantigen1") %>%
	as.matrix()

Heatmap(in_vs_in.df,
        col = colors,
        na_col = "grey80",
        cluster_rows = FALSE,
        cluster_columns = FALSE,
        column_title = "Overlap of nearby beads in TLS vs in TLS",
        name = "No. of\nbeads")
```

#### Outside TLS vs outside TLS
```{r}
out_vs_out.df <- data.frame()
for(i in 1:length(P108TA_shared_clones)) {
	for(j in 1:length(P108TA_shared_clones)) {
	  if(i < j){
  		same_beads <- length(intersect(outTLS_list[[i]]$near_beads$barcode, outTLS_list[[j]]$near_beads$barcode))
  		new_row <- c(names(outTLS_list)[i], names(outTLS_list)[j], same_beads)
  		out_vs_out.df <- rbind(out_vs_out.df, new_row)
	  }
	}
}

colnames(out_vs_out.df) <- c("neoantigen1", "neoantigen2", "beads")

# Reformat
out_vs_out.df <- out_vs_out.df %>%
	pivot_wider(names_from=neoantigen2,values_from=beads) %>%
	column_to_rownames("neoantigen1") %>%
	as.matrix()

Heatmap(out_vs_out.df,
        col = colors,
        na_col = "grey80",
        cluster_rows = FALSE,
        cluster_columns = FALSE,
        column_title = "Overlap of nearby beads outside TLS vs outside TLS",
        name = "No. of\nbeads")
```

##### View overlap of beads in p108_1 and p108_205
```{r}
overlapped_beads <- intersect(outTLS_list$p108_1$near_beads$barcode, outTLS_list$p108_205$near_beads$barcode)

ggplot() +
  geom_point(data = outTLS_list$p108_1$near_beads, aes(x = x, y = y)) +
  geom_point(data = outTLS_list$p108_1$TCRs, aes(x = x, y = y), color = "blue") +
  geom_point(data = filter(P108TA_26_RNA@meta.data, barcode %in% overlapped_beads), aes(x = x, y = y), color = "red") +
  coord_cartesian(xlim = c(0, 5000), ylim = c(0, 5000))

ggplot() +
  geom_point(data = outTLS_list$p108_205$near_beads, aes(x = x, y = y)) +
  geom_point(data = outTLS_list$p108_205$TCRs, aes(x = x, y = y), color = "blue")+
  geom_point(data = filter(P108TA_26_RNA@meta.data, barcode %in% overlapped_beads), aes(x = x, y = y), color = "red") +
  coord_cartesian(xlim = c(0, 5000), ylim = c(0, 5000))
```

##### Reassign overlapping beads
```{r}
# P108_1 and P108_205
# Reassign overlapping beads
P108TA_26_reassigned <- reassign_overlapping_beads(P108TA_26_RNA, tex_beads = outTLS_list$p108_205, tmem_beads = outTLS_list$p108_1)

# Split list back out
outTLS_list$p108_205 <- P108TA_26_reassigned$reassigned_tex_beads
outTLS_list$p108_1 <- P108TA_26_reassigned$reassigned_tmem_beads

# Plot updated beads
ggplot() +
  geom_point(data = outTLS_list$p108_1$near_beads, aes(x = x, y = y)) +
  geom_point(data = outTLS_list$p108_1$TCRs, aes(x = x, y = y), color = "blue") +
  coord_cartesian(xlim = c(0, 5000), ylim = c(0, 5000))

ggplot() +
  geom_point(data = outTLS_list$p108_205$near_beads, aes(x = x, y = y)) +
  geom_point(data = outTLS_list$p108_205$TCRs, aes(x = x, y = y), color = "blue")+
  coord_cartesian(xlim = c(0, 5000), ylim = c(0, 5000))

# P108_208 and P108_205
# Reassign overlapping beads
P108TA_26_reassigned <- reassign_overlapping_beads(P108TA_26_RNA, tex_beads = outTLS_list$p108_205, tmem_beads = outTLS_list$p108_208)

# Split list back out
outTLS_list$p108_205 <- P108TA_26_reassigned$reassigned_tex_beads
outTLS_list$p108_208 <- P108TA_26_reassigned$reassigned_tmem_beads

# P108_215 and P108_205
# Reassign overlapping beads
P108TA_26_reassigned <- reassign_overlapping_beads(P108TA_26_RNA, tex_beads = outTLS_list$p108_205, tmem_beads = outTLS_list$p108_215)

# Split list back out
outTLS_list$p108_205 <- P108TA_26_reassigned$reassigned_tex_beads
outTLS_list$p108_215 <- P108TA_26_reassigned$reassigned_tmem_beads

# P108_266 and P108_205
# Reassign overlapping beads
P108TA_26_reassigned <- reassign_overlapping_beads(P108TA_26_RNA, tex_beads = outTLS_list$p108_205, tmem_beads = outTLS_list$p108_266)

# Split list back out
outTLS_list$p108_205 <- P108TA_26_reassigned$reassigned_tex_beads
outTLS_list$p108_266 <- P108TA_26_reassigned$reassigned_tmem_beads

# P108_266 and P108_208
# Reassign overlapping beads
P108TA_26_reassigned <- reassign_overlapping_beads(P108TA_26_RNA, tex_beads = outTLS_list$p108_266, tmem_beads = outTLS_list$p108_208)

# Split list back out
outTLS_list$p108_266 <- P108TA_26_reassigned$reassigned_tex_beads
outTLS_list$p108_208 <- P108TA_26_reassigned$reassigned_tmem_beads
```

#### Outside TLS vs in TLS
```{r}
out_vs_in.df <- data.frame()
for(i in 1:length(P108TA_shared_clones)) {
	for(j in 1:length(P108TA_shared_clones)) {
	  if(i <= j){
  		same_beads <- length(intersect(outTLS_list[[i]]$near_beads$barcode, inTLS_list[[j]]$near_beads$barcode))
  		new_row <- c(names(outTLS_list)[i], names(inTLS_list)[j], same_beads)
  		out_vs_in.df <- rbind(out_vs_in.df, new_row)
	  }
	}
}

colnames(out_vs_in.df) <- c("neoantigen1", "neoantigen2", "beads")

# Reformat
out_vs_in.df <- out_vs_in.df %>%
	pivot_wider(names_from=neoantigen2,values_from=beads) %>%
	column_to_rownames("neoantigen1") %>%
	as.matrix()

Heatmap(out_vs_in.df,
        col = colors,
        na_col = "grey80",
        cluster_rows = FALSE,
        cluster_columns = FALSE,
        column_title = "Overlap of nearby beads outside TLS vs in TLS",
        name = "No. of\nbeads")
```

### Attach assignments to original Seurat object and filter for defined beads
```{r}
# Attach partition data
partition.df <- data.frame()
for(i in c(1:(length(inTLS_list)))){
  if(nrow(inTLS_list[[i]]$near_beads) > 0){
    near <- inTLS_list[[i]]$near_beads %>%
      select(barcode, partition)
    partition.df <- rbind(partition.df, near)
  }
}
for(i in c(1:(length(outTLS_list)))){
  if(nrow(outTLS_list[[i]]$near_beads) > 0){
    near <- outTLS_list[[i]]$near_beads %>%
      select(barcode, partition)
    partition.df <- rbind(partition.df, near)
  }
}

P108TA_26_RNA@meta.data <- P108TA_26_RNA@meta.data %>%
  left_join(partition.df, by = "barcode") %>%
  column_to_rownames("barcode") %>%
  mutate(barcode = rownames(.))

P108TA_26_RNA@meta.data <- P108TA_26_RNA@meta.data %>%
  mutate(type = case_when(barcode %in% inTLS_list$p108_1$near_beads$barcode ~ "p108_1,inTLS",
                          barcode %in% inTLS_list$p108_205$near_beads$barcode ~ "p108_205,inTLS",
                          barcode %in% inTLS_list$p108_208$near_beads$barcode ~ "p108_208,inTLS",
                          barcode %in% inTLS_list$p108_215$near_beads$barcode ~ "p108_215,inTLS",
                          barcode %in% inTLS_list$p108_266$near_beads$barcode ~ "p108_266,inTLS",
                          barcode %in% outTLS_list$p108_1$near_beads$barcode ~ "p108_1,outTLS",
                          barcode %in% outTLS_list$p108_205$near_beads$barcode ~ "p108_205,outTLS",
                          barcode %in% outTLS_list$p108_208$near_beads$barcode ~ "p108_208,outTLS",
                          barcode %in% outTLS_list$p108_215$near_beads$barcode ~ "p108_215,outTLS",
                          barcode %in% outTLS_list$p108_266$near_beads$barcode ~ "p108_266,outTLS"),
         muscat = case_when(!is.na(type) ~ "Defined",
                             is.na(type) ~ "Undefined"))

P108TA_26_RNA <- subset(P108TA_26_RNA, subset = muscat == "Defined")
```

## Puck 29
### Get beads near each shared neoantigen clone inside the TLS
```{r}
# Get the beads near shared neoantigen clones
inTLS_list <- list()

for(clone in P108TA_shared_clones){
  result_list <- get_beads_near_specific_neoantigen_TCR2(P108TA_29_RNA, "inTLS", clone, rad)
  inTLS_list <- append(inTLS_list, list(result_list))
}

# Rename lists
names(inTLS_list) <- P108TA_shared_clones
```

### Get beads near each shared neoantigen clone outside the TLS
```{r}
# Get the beads near shared neoantigen clones
outTLS_list <- list()

for(clone in P108TA_shared_clones){
  result_list <- get_beads_near_specific_neoantigen_TCR2(P108TA_29_RNA, "outTLS", clone, rad)
  outTLS_list <- append(outTLS_list, list(result_list))
}

# Rename lists
names(outTLS_list) <- P108TA_shared_clones
```

### Check for overlaps
#### in TLS vs in TLS
```{r}
in_vs_in.df <- data.frame()
for(i in 1:length(P108TA_shared_clones)) {
	for(j in 1:length(P108TA_shared_clones)) {
	  if(i < j){
  		same_beads <- length(intersect(inTLS_list[[i]]$near_beads$barcode, inTLS_list[[j]]$near_beads$barcode))
  		new_row <- c(names(inTLS_list)[i], names(inTLS_list)[j], same_beads)
  		in_vs_in.df <- rbind(in_vs_in.df, new_row)
	  }
	}
}

colnames(in_vs_in.df) <- c("neoantigen1", "neoantigen2", "beads")

# Reformat
in_vs_in.df <- in_vs_in.df %>%
	pivot_wider(names_from=neoantigen2,values_from=beads) %>%
	column_to_rownames("neoantigen1") %>%
	as.matrix()

Heatmap(in_vs_in.df,
        col = colors,
        na_col = "grey80",
        cluster_rows = FALSE,
        cluster_columns = FALSE,
        column_title = "Overlap of nearby beads in TLS vs in TLS",
        name = "No. of\nbeads")
```

#### Outside TLS vs outside tls
```{r}
out_vs_out.df <- data.frame()
for(i in 1:length(P108TA_shared_clones)) {
	for(j in 1:length(P108TA_shared_clones)) {
	  if(i < j){
  		same_beads <- length(intersect(outTLS_list[[i]]$near_beads$barcode, outTLS_list[[j]]$near_beads$barcode))
  		new_row <- c(names(outTLS_list)[i], names(outTLS_list)[j], same_beads)
  		out_vs_out.df <- rbind(out_vs_out.df, new_row)
	  }
	}
}

colnames(out_vs_out.df) <- c("neoantigen1", "neoantigen2", "beads")

# Reformat
out_vs_out.df <- out_vs_out.df %>%
	pivot_wider(names_from=neoantigen2,values_from=beads) %>%
	column_to_rownames("neoantigen1") %>%
	as.matrix()

Heatmap(out_vs_out.df,
        col = colors,
        na_col = "grey80",
        cluster_rows = FALSE,
        cluster_columns = FALSE,
        column_title = "Overlap of nearby beads outside TLS vs outside TLS",
        name = "No. of\nbeads")
```

##### Reassign overlapping beads
```{r}
# P108_208 and P108_205
# Reassign overlapping beads
P108TA_29_reassigned <- reassign_overlapping_beads(P108TA_29_RNA, tex_beads = outTLS_list$p108_205, tmem_beads = outTLS_list$p108_208)

# Split list back out
outTLS_list$p108_205 <- P108TA_29_reassigned$reassigned_tex_beads
outTLS_list$p108_208 <- P108TA_29_reassigned$reassigned_tmem_beads

# P108_208 and P108_215
# Reassign overlapping beads
P108TA_29_reassigned <- reassign_overlapping_beads(P108TA_29_RNA, tex_beads = outTLS_list$p108_215, tmem_beads = outTLS_list$p108_208)

# Split list back out
outTLS_list$p108_215 <- P108TA_29_reassigned$reassigned_tex_beads
outTLS_list$p108_208 <- P108TA_29_reassigned$reassigned_tmem_beads
```

#### Outside TLS vs in TLS
```{r}
out_vs_in.df <- data.frame()
for(i in 1:length(P108TA_shared_clones)) {
	for(j in 1:length(P108TA_shared_clones)) {
	  if(i <= j){
  		same_beads <- length(intersect(outTLS_list[[i]]$near_beads$barcode, inTLS_list[[j]]$near_beads$barcode))
  		new_row <- c(names(outTLS_list)[i], names(inTLS_list)[j], same_beads)
  		out_vs_in.df <- rbind(out_vs_in.df, new_row)
	  }
	}
}

colnames(out_vs_in.df) <- c("neoantigen1", "neoantigen2", "beads")

# Reformat
out_vs_in.df <- out_vs_in.df %>%
	pivot_wider(names_from=neoantigen2,values_from=beads) %>%
	column_to_rownames("neoantigen1") %>%
	as.matrix()

Heatmap(out_vs_in.df,
        col = colors,
        na_col = "grey80",
        cluster_rows = FALSE,
        cluster_columns = FALSE,
        column_title = "Overlap of nearby beads outside TLS vs in TLS",
        name = "No. of\nbeads")
```

### Attach assignments to original Seurat object and filter for defined beads
```{r}
# Attach partition data
partition.df <- data.frame()
for(i in c(1:(length(inTLS_list)))){
  if(nrow(inTLS_list[[i]]$near_beads) > 0){
    near <- inTLS_list[[i]]$near_beads %>%
      select(barcode, partition)
    partition.df <- rbind(partition.df, near)
  }
}
for(i in c(1:(length(outTLS_list)))){
  if(nrow(outTLS_list[[i]]$near_beads) > 0){
    near <- outTLS_list[[i]]$near_beads %>%
      select(barcode, partition)
    partition.df <- rbind(partition.df, near)
  }
}

P108TA_29_RNA@meta.data <- P108TA_29_RNA@meta.data %>%
  left_join(partition.df, by = "barcode") %>%
  column_to_rownames("barcode") %>%
  mutate(barcode = rownames(.))


P108TA_29_RNA@meta.data <- P108TA_29_RNA@meta.data %>%
  mutate(type = case_when(barcode %in% inTLS_list$p108_1$near_beads$barcode ~ "p108_1,inTLS",
                          barcode %in% inTLS_list$p108_205$near_beads$barcode ~ "p108_205,inTLS",
                          barcode %in% inTLS_list$p108_208$near_beads$barcode ~ "p108_208,inTLS",
                          barcode %in% inTLS_list$p108_215$near_beads$barcode ~ "p108_215,inTLS",
                          barcode %in% inTLS_list$p108_266$near_beads$barcode ~ "p108_266,inTLS",
                          barcode %in% outTLS_list$p108_1$near_beads$barcode ~ "p108_1,outTLS",
                          barcode %in% outTLS_list$p108_205$near_beads$barcode ~ "p108_205,outTLS",
                          barcode %in% outTLS_list$p108_208$near_beads$barcode ~ "p108_208,outTLS",
                          barcode %in% outTLS_list$p108_215$near_beads$barcode ~ "p108_215,outTLS",
                          barcode %in% outTLS_list$p108_266$near_beads$barcode ~ "p108_266,outTLS"),
         muscat = case_when(!is.na(type) ~ "Defined",
                             is.na(type) ~ "Undefined"))

P108TA_29_RNA <- subset(P108TA_29_RNA, subset = muscat == "Defined")
```

## Puck 30
### Get beads near each shared neoantigen clone inside the TLS
```{r}
# Get the beads near shared neoantigen clones
inTLS_list <- list()

for(clone in P108TA_shared_clones){
  result_list <- get_beads_near_specific_neoantigen_TCR2(P108TA_30_RNA, "inTLS", clone, rad)
  inTLS_list <- append(inTLS_list, list(result_list))
}

# Rename lists
names(inTLS_list) <- P108TA_shared_clones
```

### Get beads near each shared neoantigen clone outside the TLS
```{r}
# Get the beads near shared neoantigen clones
outTLS_list <- list()

for(clone in P108TA_shared_clones){
  result_list <- get_beads_near_specific_neoantigen_TCR2(P108TA_30_RNA, "outTLS", clone, rad)
  outTLS_list <- append(outTLS_list, list(result_list))
}

# Rename lists
names(outTLS_list) <- P108TA_shared_clones
```

### Check for overlaps
#### in TLS vs in TLS
```{r}
in_vs_in.df <- data.frame()
for(i in 1:length(P108TA_shared_clones)) {
	for(j in 1:length(P108TA_shared_clones)) {
	  if(i < j){
  		same_beads <- length(intersect(inTLS_list[[i]]$near_beads$barcode, inTLS_list[[j]]$near_beads$barcode))
  		new_row <- c(names(inTLS_list)[i], names(inTLS_list)[j], same_beads)
  		in_vs_in.df <- rbind(in_vs_in.df, new_row)
	  }
	}
}

colnames(in_vs_in.df) <- c("neoantigen1", "neoantigen2", "beads")

# Reformat
in_vs_in.df <- in_vs_in.df %>%
	pivot_wider(names_from=neoantigen2,values_from=beads) %>%
	column_to_rownames("neoantigen1") %>%
	as.matrix()

Heatmap(in_vs_in.df,
        col = colors,
        na_col = "grey80",
        cluster_rows = FALSE,
        cluster_columns = FALSE,
        column_title = "Overlap of nearby beads in TLS vs in TLS",
        name = "No. of\nbeads")
```

#### Outside TLS vs outside tls
```{r}
out_vs_out.df <- data.frame()
for(i in 1:length(P108TA_shared_clones)) {
	for(j in 1:length(P108TA_shared_clones)) {
	  if(i < j){
  		same_beads <- length(intersect(outTLS_list[[i]]$near_beads$barcode, outTLS_list[[j]]$near_beads$barcode))
  		new_row <- c(names(outTLS_list)[i], names(outTLS_list)[j], same_beads)
  		out_vs_out.df <- rbind(out_vs_out.df, new_row)
	  }
	}
}

colnames(out_vs_out.df) <- c("neoantigen1", "neoantigen2", "beads")

# Reformat
out_vs_out.df <- out_vs_out.df %>%
	pivot_wider(names_from=neoantigen2,values_from=beads) %>%
	column_to_rownames("neoantigen1") %>%
	as.matrix()

Heatmap(out_vs_out.df,
        col = colors,
        na_col = "grey80",
        cluster_rows = FALSE,
        cluster_columns = FALSE,
        column_title = "Overlap of nearby beads outside TLS vs outside TLS",
        name = "No. of\nbeads")
```

#### Outside TLS vs in TLS
```{r}
out_vs_in.df <- data.frame()
for(i in 1:length(P108TA_shared_clones)) {
	for(j in 1:length(P108TA_shared_clones)) {
	  if(i <= j){
  		same_beads <- length(intersect(outTLS_list[[i]]$near_beads$barcode, inTLS_list[[j]]$near_beads$barcode))
  		new_row <- c(names(outTLS_list)[i], names(inTLS_list)[j], same_beads)
  		out_vs_in.df <- rbind(out_vs_in.df, new_row)
	  }
	}
}

colnames(out_vs_in.df) <- c("neoantigen1", "neoantigen2", "beads")

# Reformat
out_vs_in.df <- out_vs_in.df %>%
	pivot_wider(names_from=neoantigen2,values_from=beads) %>%
	column_to_rownames("neoantigen1") %>%
	as.matrix()

Heatmap(out_vs_in.df,
        col = colors,
        na_col = "grey80",
        cluster_rows = FALSE,
        cluster_columns = FALSE,
        column_title = "Overlap of nearby beads outside TLS vs in TLS",
        name = "No. of\nbeads")
```

### Attach assignments to original Seurat object and filter for defined beads
```{r}
# Attach partition data
partition.df <- data.frame()
for(i in c(1:(length(inTLS_list)))){
  if(nrow(inTLS_list[[i]]$near_beads) > 0){
    near <- inTLS_list[[i]]$near_beads %>%
      select(barcode, partition)
    partition.df <- rbind(partition.df, near)
  }
}
for(i in c(1:(length(outTLS_list)))){
  if(nrow(outTLS_list[[i]]$near_beads) > 0){
    near <- outTLS_list[[i]]$near_beads %>%
      select(barcode, partition)
    partition.df <- rbind(partition.df, near)
  }
}

P108TA_30_RNA@meta.data <- P108TA_30_RNA@meta.data %>%
  left_join(partition.df, by = "barcode") %>%
  column_to_rownames("barcode") %>%
  mutate(barcode = rownames(.))


P108TA_30_RNA@meta.data <- P108TA_30_RNA@meta.data %>%
  mutate(type = case_when(barcode %in% inTLS_list$p108_1$near_beads$barcode ~ "p108_1,inTLS",
                          barcode %in% inTLS_list$p108_205$near_beads$barcode ~ "p108_205,inTLS",
                          barcode %in% inTLS_list$p108_208$near_beads$barcode ~ "p108_208,inTLS",
                          barcode %in% inTLS_list$p108_215$near_beads$barcode ~ "p108_215,inTLS",
                          barcode %in% inTLS_list$p108_266$near_beads$barcode ~ "p108_266,inTLS",
                          barcode %in% outTLS_list$p108_1$near_beads$barcode ~ "p108_1,outTLS",
                          barcode %in% outTLS_list$p108_205$near_beads$barcode ~ "p108_205,outTLS",
                          barcode %in% outTLS_list$p108_208$near_beads$barcode ~ "p108_208,outTLS",
                          barcode %in% outTLS_list$p108_215$near_beads$barcode ~ "p108_215,outTLS",
                          barcode %in% outTLS_list$p108_266$near_beads$barcode ~ "p108_266,outTLS"),
         muscat = case_when(!is.na(type) ~ "Defined",
                             is.na(type) ~ "Undefined"))

P108TA_30_RNA <- subset(P108TA_30_RNA, subset = muscat == "Defined")
```

## Puck 32
### Get beads near each shared neoantigen clone inside the TLS
```{r}
# Get the beads near shared neoantigen clones
inTLS_list <- list()

for(clone in P108TA_shared_clones){
  result_list <- get_beads_near_specific_neoantigen_TCR2(P108TA_32_RNA, "inTLS", clone, rad)
  inTLS_list <- append(inTLS_list, list(result_list))
}

# Rename lists
names(inTLS_list) <- P108TA_shared_clones
```

### Get beads near each shared neoantigen clone outside the TLS
```{r}
# Get the beads near shared neoantigen clones
outTLS_list <- list()

for(clone in P108TA_shared_clones){
  result_list <- get_beads_near_specific_neoantigen_TCR2(P108TA_32_RNA, "outTLS", clone, rad)
  outTLS_list <- append(outTLS_list, list(result_list))
}

# Rename lists
names(outTLS_list) <- P108TA_shared_clones
```

### Check for overlaps
#### in TLS vs in TLS
```{r}
in_vs_in.df <- data.frame()
for(i in 1:length(P108TA_shared_clones)) {
	for(j in 1:length(P108TA_shared_clones)) {
	  if(i < j){
  		same_beads <- length(intersect(inTLS_list[[i]]$near_beads$barcode, inTLS_list[[j]]$near_beads$barcode))
  		new_row <- c(names(inTLS_list)[i], names(inTLS_list)[j], same_beads)
  		in_vs_in.df <- rbind(in_vs_in.df, new_row)
	  }
	}
}

colnames(in_vs_in.df) <- c("neoantigen1", "neoantigen2", "beads")

# Reformat
in_vs_in.df <- in_vs_in.df %>%
	pivot_wider(names_from=neoantigen2,values_from=beads) %>%
	column_to_rownames("neoantigen1") %>%
	as.matrix()

Heatmap(in_vs_in.df,
        col = colors,
        na_col = "grey80",
        cluster_rows = FALSE,
        cluster_columns = FALSE,
        column_title = "Overlap of nearby beads in TLS vs in TLS",
        name = "No. of\nbeads")
```

##### Reassign overlapping beads
```{r}
# P108_208 and P108_266
# Reassign overlapping beads
P108TA_32_reassigned <- reassign_overlapping_beads(P108TA_32_RNA, tex_beads = inTLS_list$p108_266, tmem_beads = inTLS_list$p108_208)

# Split list back out
inTLS_list$p108_266 <- P108TA_32_reassigned$reassigned_tex_beads
inTLS_list$p108_208 <- P108TA_32_reassigned$reassigned_tmem_beads

```

#### Outside TLS vs outside tls
```{r}
out_vs_out.df <- data.frame()
for(i in 1:length(P108TA_shared_clones)) {
	for(j in 1:length(P108TA_shared_clones)) {
	  if(i < j){
  		same_beads <- length(intersect(outTLS_list[[i]]$near_beads$barcode, outTLS_list[[j]]$near_beads$barcode))
  		new_row <- c(names(outTLS_list)[i], names(outTLS_list)[j], same_beads)
  		out_vs_out.df <- rbind(out_vs_out.df, new_row)
	  }
	}
}

colnames(out_vs_out.df) <- c("neoantigen1", "neoantigen2", "beads")

# Reformat
out_vs_out.df <- out_vs_out.df %>%
	pivot_wider(names_from=neoantigen2,values_from=beads) %>%
	column_to_rownames("neoantigen1") %>%
	as.matrix()

Heatmap(out_vs_out.df,
        col = colors,
        na_col = "grey80",
        cluster_rows = FALSE,
        cluster_columns = FALSE,
        column_title = "Overlap of nearby beads outside TLS vs outside TLS",
        name = "No. of\nbeads")

```

##### Reassign overlapping beads
```{r}
# P108_205 and P108_1
# Reassign overlapping beads
P108TA_32_reassigned <- reassign_overlapping_beads(P108TA_32_RNA, tex_beads = outTLS_list$p108_205, tmem_beads = outTLS_list$p108_1)

# Split list back out
outTLS_list$p108_205 <- P108TA_32_reassigned$reassigned_tex_beads
outTLS_list$p108_1 <- P108TA_32_reassigned$reassigned_tmem_beads

# P108_266 and P108_208
# Reassign overlapping beads
P108TA_32_reassigned <- reassign_overlapping_beads(P108TA_32_RNA, tex_beads = outTLS_list$p108_266, tmem_beads = outTLS_list$p108_208)

# Split list back out
outTLS_list$p108_266 <- P108TA_32_reassigned$reassigned_tex_beads
outTLS_list$p108_208 <- P108TA_32_reassigned$reassigned_tmem_beads

```

#### Outside TLS vs in TLS
```{r}
out_vs_in.df <- data.frame()
for(i in 1:length(P108TA_shared_clones)) {
	for(j in 1:length(P108TA_shared_clones)) {
	  if(i <= j){
  		same_beads <- length(intersect(outTLS_list[[i]]$near_beads$barcode, inTLS_list[[j]]$near_beads$barcode))
  		new_row <- c(names(outTLS_list)[i], names(inTLS_list)[j], same_beads)
  		out_vs_in.df <- rbind(out_vs_in.df, new_row)
	  }
	}
}

colnames(out_vs_in.df) <- c("neoantigen1", "neoantigen2", "beads")

# Reformat
out_vs_in.df <- out_vs_in.df %>%
	pivot_wider(names_from=neoantigen2,values_from=beads) %>%
	column_to_rownames("neoantigen1") %>%
	as.matrix()

Heatmap(out_vs_in.df,
        col = colors,
        na_col = "grey80",
        cluster_rows = FALSE,
        cluster_columns = FALSE,
        column_title = "Overlap of nearby beads outside TLS vs in TLS",
        name = "No. of\nbeads")
```

### Attach assignments to original Seurat object and filter for defined beads
```{r}
# Attach partition data
partition.df <- data.frame()
for(i in c(1:(length(inTLS_list)))){
  if(nrow(inTLS_list[[i]]$near_beads) > 0){
    near <- inTLS_list[[i]]$near_beads %>%
      select(barcode, partition)
    partition.df <- rbind(partition.df, near)
  }
}
for(i in c(1:(length(outTLS_list)))){
  if(nrow(outTLS_list[[i]]$near_beads) > 0){
    near <- outTLS_list[[i]]$near_beads %>%
      select(barcode, partition)
    partition.df <- rbind(partition.df, near)
  }
}

P108TA_32_RNA@meta.data <- P108TA_32_RNA@meta.data %>%
  left_join(partition.df, by = "barcode") %>%
  column_to_rownames("barcode") %>%
  mutate(barcode = rownames(.))


P108TA_32_RNA@meta.data <- P108TA_32_RNA@meta.data %>%
  mutate(type = case_when(barcode %in% inTLS_list$p108_1$near_beads$barcode ~ "p108_1,inTLS",
                          barcode %in% inTLS_list$p108_205$near_beads$barcode ~ "p108_205,inTLS",
                          barcode %in% inTLS_list$p108_208$near_beads$barcode ~ "p108_208,inTLS",
                          barcode %in% inTLS_list$p108_215$near_beads$barcode ~ "p108_215,inTLS",
                          barcode %in% inTLS_list$p108_266$near_beads$barcode ~ "p108_266,inTLS",
                          barcode %in% outTLS_list$p108_1$near_beads$barcode ~ "p108_1,outTLS",
                          barcode %in% outTLS_list$p108_205$near_beads$barcode ~ "p108_205,outTLS",
                          barcode %in% outTLS_list$p108_208$near_beads$barcode ~ "p108_208,outTLS",
                          barcode %in% outTLS_list$p108_215$near_beads$barcode ~ "p108_215,outTLS",
                          barcode %in% outTLS_list$p108_266$near_beads$barcode ~ "p108_266,outTLS"),
         muscat = case_when(!is.na(type) ~ "Defined",
                             is.na(type) ~ "Undefined"))

P108TA_32_RNA <- subset(P108TA_32_RNA, subset = muscat == "Defined")
```

## Update metadata column
```{r}
P108TA_26_RNA@meta.data <- P108TA_26_RNA@meta.data %>%
  mutate(tls_status_gid = str_split_i(type, ",", 2))
P108TA_29_RNA@meta.data <- P108TA_29_RNA@meta.data %>%
  mutate(tls_status_gid = str_split_i(type, ",", 2))
P108TA_30_RNA@meta.data <- P108TA_30_RNA@meta.data %>%
  mutate(tls_status_gid = str_split_i(type, ",", 2))
P108TA_32_RNA@meta.data <- P108TA_32_RNA@meta.data %>%
  mutate(tls_status_gid = str_split_i(type, ",", 2))
```


## Get beads within 15px of a neoantigen specific TCR to share the same "area" annotation
### Puck 26
```{r}
P108TA_26_RNA@meta.data %>%
  dplyr::count(area, tls_status_gid)

boundary_inTLS <- P108TA_26_RNA@meta.data %>%
  filter(area == "Boundary", str_detect(type, "inTLS"))

# Get beads near the boundary beads
near_beads <- nn2(P108TA_26_RNA@meta.data[,c("x", "y")], boundary_inTLS[,c("x", "y")], radius = 15, k = 10, searchtype = "radius")
# Collapse matrix of row indexes into a vector of unique values
near_beads <- unique(as.vector(near_beads$nn.idx))
# Collect metadata of nearby beads
near_beads_md <- P108TA_26_RNA@meta.data[near_beads,]
# Which areas are those beads from?
near_beads_md %>%
  dplyr::count(area)

# Cast beads to TLS3
P108TA_26_RNA@meta.data <- P108TA_26_RNA@meta.data %>%
  mutate(final_area = case_when(area == "Boundary" & str_detect(type, "inTLS") ~ "TLS3",
                              T ~ area))
```

### Puck 29
```{r}
P108TA_29_RNA@meta.data %>%
  dplyr::count(area, tls_status_gid)

# Cast TLS3 beads to boundary
P108TA_29_RNA@meta.data <- P108TA_29_RNA@meta.data %>%
  mutate(final_area = case_when(area == "TLS3" & str_detect(type, "outTLS") ~ "Boundary",
                              T ~ area))
```

### Puck 30
```{r}
P108TA_30_RNA@meta.data %>%
  dplyr::count(area, tls_status_gid)

P108TA_30_RNA@meta.data <- P108TA_30_RNA@meta.data %>%
  mutate(final_area = area)
```

### Puck 32
```{r}
P108TA_32_RNA@meta.data %>%
  dplyr::count(area, tls_status_gid)

boundary_inTLS <- P108TA_32_RNA@meta.data %>%
  filter(area == "Boundary", str_detect(type, "inTLS"))

# Get beads near the boundary beads
near_beads <- nn2(P108TA_32_RNA@meta.data[,c("x", "y")], boundary_inTLS[1,c("x", "y")], radius = 15, k = 10, searchtype = "radius")
# Collapse matrix of row indexes into a vector of unique values
near_beads <- unique(as.vector(near_beads$nn.idx))
# Collect metadata of nearby beads
near_beads_md <- P108TA_32_RNA@meta.data[near_beads,]
# Which areas are those beads from?
near_beads_md %>%
  dplyr::count(area)
# Which bead is this?
boundary_inTLS[1, "barcode"]

near_beads <- nn2(P108TA_32_RNA@meta.data[,c("x", "y")], boundary_inTLS[2,c("x", "y")], radius = 15, k = 10, searchtype = "radius")
# Collapse matrix of row indexes into a vector of unique values
near_beads <- unique(as.vector(near_beads$nn.idx))
# Collect metadata of nearby beads
near_beads_md <- P108TA_32_RNA@meta.data[near_beads,]
# Which areas are those beads from?
near_beads_md %>%
  dplyr::count(area)
# Which bead is this?
boundary_inTLS[2, "barcode"]

# Cast beads to TLS3
P108TA_32_RNA@meta.data <- P108TA_32_RNA@meta.data %>%
  mutate(final_area = case_when(barcode == "GTCATTACCCACTT-1" ~ "TLS1",
                              barcode == "CCGTGGGTTTTGGG-1" ~ "TLS2",
                              T ~ area))

```

# Merge RNA objects
```{r}
TLS_pucks <- merge(x = P108TA_26_RNA, y = list(P108TA_29_RNA, P108TA_30_RNA, P108TA_32_RNA))

# Update metadata
TLS_pucks@meta.data <- TLS_pucks@meta.data %>%
  mutate(tls_status_gid = str_split_i(type, ",", 2),
         final_neoant_id = str_split_i(type, ",", 1))

TLS_pucks[['Spatial']] <- JoinLayers(TLS_pucks[['Spatial']])

TLS_pucks@meta.data %>%
  dplyr::count(tls_status_gid, final_area)
```

## Open the TPEX gene set
```{r}
gene_sets <- readRDS("../objects/immune_gene_sets.Rds")

new_gene_sets <- gene_sets[(names(gene_sets) %in% c("projecTILs.CD8.TPEX", "coret"))]
```

# Run gene set enrichment analysis
```{r}
TLS_pucks <- RenameAssays(TLS_pucks, assay.name = "Spatial", new.assay.name = "RNA")

TLS_pucks <- runEscape(input.data = TLS_pucks, 
                       method = "GSVA",
                       gene.sets = new_gene_sets, 
                       groups = 1000, 
                       min.size = 5, 
                       new.assay.name="escape.GSVA")

```

## Format raw GSVA enrichment scores from escape
```{r}
raw_dat <- as.data.frame(TLS_pucks@assays$escape.GSVA$data) %>%
  t() %>%
  as.data.frame()

# Remove "-"s and remove spaces from data and from gene sets
# names(new_gene_sets) <- str_replace_all(names(new_gene_sets), "-", "")
names(new_gene_sets) <- str_replace_all(names(new_gene_sets), "_", "")

colnames(raw_dat) <- str_replace_all(names(raw_dat), "-", "")

# Get dataframe
raw_md <- raw_dat %>%
  rownames_to_column("barcode") %>%
  full_join(TLS_pucks@meta.data, by = "barcode") %>%
  mutate(tls_status_gid = factor(tls_status_gid),
         final_neoant_id = factor(final_neoant_id))

raw_md$tls_status_gid <- relevel(raw_md$tls_status_gid, ref = "outTLS")

# # Select columns to use
raw_md <- raw_md[,c(1:4, 41, 44, 46)]
```

### Plot TCRs

#### Get means, SD and SE
```{r}
sd.df <- raw_md %>%
  # aggregate by puck
  group_by(tls_status_gid, final_neoant_id, orig.ident) %>% 
  summarize(mean_coret_p = round(mean(coret), 8),
            mean_tpex_p = round(mean(projecTILs.CD8.TPEX), 8)) %>%
  # aggregate across pucks
  group_by(tls_status_gid, final_neoant_id) %>%
  summarize(n = n(),
            sd_coret = round(sd(mean_coret_p), 8),
            sd_tpex = round(sd(mean_tpex_p), 8),
            se_coret = round(sd_coret/sqrt(n), 8),
            se_tpex = round(sd_tpex/sqrt(n), 8),
            mean_coret = round(mean(mean_coret_p), 8),
            mean_tpex = round(mean(mean_tpex_p), 8)) %>%
  ungroup()
```

#### Pivot for plots
```{r}
long_mean.df <- sd.df %>%
  select(tls_status_gid, final_neoant_id, contains("mean")) %>%
  pivot_longer(cols = !c(final_neoant_id, tls_status_gid), names_to = "signature", values_to = "score") %>%
  mutate(signature = str_replace(signature, "mean_", ""))

long_se.df <- sd.df %>%
  select(tls_status_gid, final_neoant_id, contains("se")) %>%
  pivot_longer(cols = !c(final_neoant_id, tls_status_gid), names_to = "signature", values_to = "se") %>%
  mutate(signature = str_replace(signature, "se_", ""))

long_merged <- long_mean.df %>%
  full_join(long_se.df, by = c("final_neoant_id", "tls_status_gid", "signature")) %>%
  rowwise() %>%
  mutate(upper = score+se,
         lower = score-se,
         tls_status_gid = factor(tls_status_gid, levels = c("inTLS", "outTLS")))
```

#### Plot
Visually both the TPEX phenotype and core T cell phenotype are enriched inTLS
```{r, fig.width = 4, fig.height = 3}
long_merged %>%
  filter(signature == "tpex") %>%
ggplot(aes(x = tls_status_gid, y = score)) +
  geom_line(aes(group = (final_neoant_id)), position=position_dodge(width=0.2)) +
  geom_errorbar(aes(ymin=lower, ymax=upper, group = final_neoant_id), width=.2, position=position_dodge(width=0.2), color = "grey50", linewidth = 0.3) +
  geom_point(aes(group = final_neoant_id, fill = tls_status_gid), color = "black",
             position=position_dodge(width=0.2), size = 3, shape = 22) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  scale_fill_manual(values = c("#b2c4b9", "#d0b39e")) +
  ylab("Raw enrichment (GSVA)") +
  guides(color = "none") +
  ggtitle("TPEX enrichment values")

long_merged %>%
  filter(signature == "coret") %>%
ggplot(aes(x = tls_status_gid, y = score)) +
  geom_line(aes(group = (final_neoant_id)), position=position_dodge(width=0.2)) +
  geom_errorbar(aes(ymin=lower, ymax=upper, group = final_neoant_id), width=.2, position=position_dodge(width=0.2), color = "grey50", linewidth = 0.3) +
  geom_point(aes(group = final_neoant_id, fill = tls_status_gid), color = "black",
             position=position_dodge(width=0.2), size = 3, shape = 22) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  scale_fill_manual(values = c("#b2c4b9", "#d0b39e")) +
  ylab("Raw enrichment (GSVA)") +
  guides(color = "none") +
  ggtitle("Core T enrichment values")
```

## Normalize enrichment values 
```{r}
TLS_pucks <- performNormalization(input.data = TLS_pucks, 
                               assay = "escape.GSVA", 
                               gene.sets = new_gene_sets,
                               scale.factor = TLS_pucks$nFeature_RNA,
                               make.positive = T)
```

### Format results from GSVA
```{r}
norm_dat <- as.data.frame(TLS_pucks@assays$escape.GSVA_normalized$data) %>%
  t() %>%
  as.data.frame()

# Remove "-"s and remove spaces from data and from gene sets
# names(new_gene_sets) <- str_replace_all(names(new_gene_sets), "-", "")
names(new_gene_sets) <- str_replace_all(names(new_gene_sets), "_", "")

colnames(norm_dat) <- str_replace_all(names(norm_dat), "-", "")

# Get dataframe
md <- norm_dat %>%
  rownames_to_column("barcode") %>%
  full_join(TLS_pucks@meta.data, by = "barcode") %>%
  mutate(tls_status_gid = factor(tls_status_gid),
         final_neoant_id = factor(final_neoant_id))

md$tls_status_gid <- relevel(md$tls_status_gid, ref = "outTLS")

# # Select columns to use
md <- md[,c(1:4, 41, 44, 46)]
```

# Run permutation test
#### Hypothesis: Do shared Tex clones found in the TLS have a higher T cell phenotype than shared clones found outside the TLS

#### Permutation: Randomly swap the location (eg. inTLS) label without replacement, of shared clones both inside and outside the TLS
###### - Swap location label within pucks (same number of “inTLS” beads in “puck A” across all permutations)
###### - Swap location label within pucks and TCR IDs (same number of “inTLS” beads in “TCR 1” across all permutations)

#### Test statistic: Median ratio of module score of shared clones in the TLS (in both patients) minus median module score of shared clones outside the TLS (in both patients)

# Get observed data
```{r}
# Get observed model of each gene set
model_obs <- vector("list", length(new_gene_sets))
names(model_obs) <- names(new_gene_sets)
# Get observed T statistic for each gene set
t_obs <- vector("list", length(new_gene_sets))
names(model_obs) <- names(new_gene_sets)
names(t_obs) <- names(new_gene_sets)

for (gs in names(new_gene_sets)) {
  print(gs)
  # build formula
  f <- as.formula(paste(gs, "~ tls_status_gid + (1 | orig.ident)"))
  
  model <- lmer(f, data = md, REML = TRUE)
  t_val <- coef(summary(model))["tls_status_gidinTLS","t value"]
  
  model_obs[[gs]] <- model
  t_obs[[gs]] <- t_val
}

```

# Permute data
```{r}
n_perm_safe = 1500

set.seed(123)
# Swap location labels within pucks and clonotypes and partitions
md_partition <- md

partition_perm <- md_partition %>%
  rep_sample_n(size = nrow(md), replace = FALSE, reps = n_perm_safe) %>%
  distinct(partition, orig.ident, replicate, tls_status_gid) %>%
  mutate(perm_tls_status = sample(tls_status_gid, size = n(), replace = FALSE))

md_perm <- md %>%
  rep_sample_n(size = nrow(md), replace = FALSE, reps = n_perm_safe) %>%
  full_join(partition_perm, by = c("partition", "orig.ident", "replicate", "tls_status_gid"))

```

# Get permuted test statistic
```{r}
# Get permuted test statistic for each gene set
model_perm <- vector("list", length(new_gene_sets))
names(model_perm) <- names(new_gene_sets)

n_perm <- 100

# Count non-convergences
non_convergences <- vector("list", length(new_gene_sets))
names(non_convergences) <- names(new_gene_sets)

# Count times tested
tested <- vector("list", length(new_gene_sets))
names(tested) <- names(new_gene_sets)

set.seed(123)
# Loop across each gene set
for (gs in names(new_gene_sets)) {
  n_run <- 1
  n_saved <- 1
  non_conv <- 0
  
  print(gs)
  t_val_perm.vec <- c()
  # Do 1000 converging loops
  while(n_saved <= n_perm){
    # Print every saved 100 perms
    if (n_saved %% 100 == 0) { print(n_saved) }
    # Get the permuted replicate
    md_rep <- md_perm %>%
      filter(replicate == n_run)
    # Set runs as good
    bad_run <- FALSE
    # Attempt the model
    model <- withCallingHandlers(
      expr = {
        # Build formula with permutated TLs status
        f <- as.formula(paste(gs, "~ perm_tls_status + (1 | orig.ident)"))
        # Build linear model on permuted data
        lmer(f, data = md_rep, REML = TRUE)},
      warning = function(w) {
      # Note if the model can't converge
      bad_run <<- TRUE
      non_conv <<- non_conv + 1
      n_run <<- n_run + 1
      print("Couldn't converge")
      invokeRestart("muffleWarning")
      }
    )
      
    # Save the t val from the model if it has no errors
    if(!bad_run){
      # Get and save t val
      t_val_perm <- coef(summary(model))["perm_tls_statusinTLS","t value"]
      t_val_perm.vec <- c(t_val_perm.vec, t_val_perm)
      # Count that we've saved a run
      n_saved <- n_saved + 1
      n_run <- n_run + 1
    } else {
      print("Skipping model")
    }
  }
  model_perm[[gs]] <- t_val_perm.vec
  non_convergences[[gs]] <- non_conv
}

```

## Format permuted results
```{r}
df_perm <- data.frame(gene_sets = rep(names(model_perm), times = lengths(model_perm)),
                      perm_ts = unlist(model_perm, use.names = FALSE))

df_obs <- data.frame(gene_sets = rep(names(t_obs), times = lengths(t_obs)),
                     obs_ts = unlist(t_obs, use.names = FALSE)) 
```

## Plot permuted test statistic distribution and observed test statistic
```{r, fig.height = 5, fig.width = 8}
df_perm %>%
  ggplot(aes(x = perm_ts)) +
  geom_histogram(bins = 100) +
  facet_wrap(~gene_sets, scales = "free", nrow = 2) +
  geom_vline(data = df_obs, aes(xintercept = obs_ts))

```

## Get one sided pval
```{r}
os_pval.vec <- c()

for(gs in unique(df_obs$gene_sets)){
  obs_val <- df_obs %>%
    filter(gene_sets == gs) %>%
    pull(obs_ts)
  
  perm_val <- df_perm %>%
    filter(gene_sets == gs) %>%
    pull(perm_ts)
  
  # One sided
  n_extreme <- length(perm_val[perm_val >= obs_val])
  pval <- n_extreme/n_perm
  
  os_pval.vec <- c(os_pval.vec, pval)
}

names(os_pval.vec) <- unique(df_obs$gene_sets)

# Replace 0 with small number
os_pval.vec[os_pval.vec == 0] <- 1/(n_perm)

os_pval.vec
```

# Plot observed mean for each puck
```{r}
mean.df <- md %>%
  # aggregate by puck
  group_by(tls_status_gid, orig.ident) %>% 
  summarize(across(where(is.numeric), ~ mean(.x, na.rm = TRUE)),
            .groups = "drop") %>%
  ungroup()
  
mean.df %>%
  mutate(tls_status_gid = factor(tls_status_gid, levels = c("inTLS", "outTLS"))) %>%
  pivot_longer(cols = !c("tls_status_gid", "orig.ident"), values_to = "score", names_to = "signature") %>%
ggplot(aes(x = tls_status_gid, y = score)) +
  geom_boxplot(aes(fill = tls_status_gid)) +
  geom_line(aes(group = orig.ident)) +
  geom_point(aes(group = orig.ident), color = "#009193", size = 3, shape = 15) +
  geom_point(aes(group = orig.ident), color = "black", size = 3, shape = 22) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  scale_fill_manual(values = c("#b2c4b9", "#d0b39e")) +
  ylab("Normalized enrichment (GSVA)") +
  xlab("") +
  facet_wrap(~signature, scales = "free_y") +
  guides(fill="none")

```

# Get session info
```{r}
sessionInfo()
```
