# Plot enrichment {#sec-clone-enrichment}

# Compile and plot enrichment of clonotypes or T cell phenotypes in spatial compartments 

# Libraries
```{r}
library(Seurat)
library(SeuratWrappers)
library(ggplot2)
library(patchwork)
library(dplyr)
library(ggplotify)
library(cowplot)
library(stringr)
library(reshape2)
library(ggrepel)
library(ggbreak)
library(tibble)

source("../source_scripts/slideseq_functions_main.R")
```

###### Joining the enrichment values of clones in the boundary, tumor and TLS across patients for a summarized visualization in volcano plots. Enrichment scores and p-vals (Permutation test) were calculated for clones with at least 5 beads across pucks in a patient. P-vals were FDR corrected using the BH (FDR) method.

###### Note: current_in_all = the number of beads with the clone (or specificity) across all areas in a patient. current_in_given_area = the number of beads with the clone across a given area in a patient.

# Pick significance level
```{r}
sig_level <- 0.05
```

# Open clone objects
```{r}
# P102
P102TA_clone_tls <- readRDS("../objects/P102TA_TLS_clone_list.rds") %>%
  get_vals_df(.) %>%
  mutate(Patient = "P102TA",
         type = "clone",
         pval = case_when(pval == 0 ~ 1/5000,
                          T ~ pval),
         qval = p.adjust(pval, method = "BH"))
P102TA_clone_tumor <- readRDS("../objects/P102TA_tumor_clone_list.rds") %>%
  get_vals_df(.) %>%
  mutate(Patient = "P102TA",
         type = "clone",
         pval = case_when(pval == 0 ~ 1/5000,
                          T ~ pval),
         qval = p.adjust(pval, method = "BH"))
P102TA_clone_boundary <- readRDS("../objects/P102TA_boundary_clone_list.rds") %>%
  get_vals_df(.) %>%
  mutate(Patient = "P102TA",
         type = "clone",
         pval = case_when(pval == 0 ~ 1/5000,
                          T ~ pval),
         qval = p.adjust(pval, method = "BH"))

# P107
P107TA_clone_tls <- readRDS("../objects/P107TA_TLS_clone_list.rds") %>%
  get_vals_df(.) %>%
  mutate(Patient = "P107TA",
         type = "clone",
         pval = case_when(pval == 0 ~ 1/5000,
                          T ~ pval),
         qval = p.adjust(pval, method = "BH"))
P107TA_clone_tumor <- readRDS("../objects/P107TA_tumor_clone_list.rds") %>%
  get_vals_df(.) %>%
  mutate(Patient = "P107TA",
         type = "clone",
         pval = case_when(pval == 0 ~ 1/5000,
                          T ~ pval),
         qval = p.adjust(pval, method = "BH"))

# P108TA
P108TA_clone_tls <- readRDS("../objects/P108TA_TLS_clone_list.rds") %>%
  get_vals_df(.) %>%
  mutate(Patient = "P108TA",
         type = "clone",
         pval = case_when(pval == 0 ~ 1/5000,
                          T ~ pval),
         qval = p.adjust(pval, method = "BH"))
P108TA_clone_tumor <- readRDS("../objects/P108TA_tumor_clone_list.rds") %>%
  get_vals_df(.) %>%
  mutate(Patient = "P108TA",
         type = "clone",
         pval = case_when(pval == 0 ~ 1/5000,
                          T ~ pval),
         qval = p.adjust(pval, method = "BH"))
P108TA_clone_boundary <- readRDS("../objects/P108TA_boundary_clone_list.rds") %>%
  get_vals_df(.) %>%
  mutate(Patient = "P108TA",
         type = "clone",
         pval = case_when(pval == 0 ~ 1/5000,
                          T ~ pval),
         qval = p.adjust(pval, method = "BH"))
```

# Open phenotype objects
```{r}
P102TA_phenotype_tls <- readRDS("../objects/P102TA_TLS_phenotype_list.rds") %>%
  get_vals_df(.) %>%
  mutate(Patient = "P102TA",
         type = "phenotype",
         pval = case_when(pval == 0 ~ 1/5000,
                          T ~ pval),
         qval = p.adjust(pval, method = "BH"))
P102TA_phenotype_boundary <- readRDS("../objects/P102TA_boundary_phenotype_list.rds") %>%
  get_vals_df(.) %>%
  mutate(Patient = "P102TA",
         type = "phenotype",
         pval = case_when(pval == 0 ~ 1/5000,
                          T ~ pval),
         qval = p.adjust(pval, method = "BH"))
P102TA_phenotype_tumor <- readRDS("../objects/P102TA_tumor_phenotype_list.rds") %>%
  get_vals_df(.) %>%
  mutate(Patient = "P102TA",
         type = "phenotype",
         pval = case_when(pval == 0 ~ 1/5000,
                          T ~ pval),
         qval = p.adjust(pval, method = "BH"))

P107TA_phenotype_tls <- readRDS("../objects/P107TA_TLS_phenotype_list.rds") %>%
  get_vals_df(.) %>%
  mutate(Patient = "P107TA",
         type = "phenotype",
         pval = case_when(pval == 0 ~ 1/5000,
                          T ~ pval),
         qval = p.adjust(pval, method = "BH"))
P107TA_phenotype_tumor <- readRDS("../objects/P107TA_tumor_phenotype_list.rds") %>%
  get_vals_df(.) %>%
  mutate(Patient = "P107TA",
         type = "phenotype",
         pval = case_when(pval == 0 ~ 1/5000,
                          T ~ pval),
         qval = p.adjust(pval, method = "BH"))

P108TA_phenotype_tls <- readRDS("../objects/P108TA_TLS_phenotype_list.rds") %>%
  get_vals_df(.) %>%
  mutate(Patient = "P108TA",
         type = "phenotype",
         pval = case_when(pval == 0 ~ 1/5000,
                          T ~ pval),
         qval = p.adjust(pval, method = "BH"))
P108TA_phenotype_tumor <- readRDS("../objects/P108TA_tumor_phenotype_list.rds") %>%
  get_vals_df(.) %>%
  mutate(Patient = "P108TA",
         type = "phenotype",
         pval = case_when(pval == 0 ~ 1/5000,
                          T ~ pval),
         qval = p.adjust(pval, method = "BH"))
P108TA_phenotype_boundary <- readRDS("../objects/P108TA_boundary_phenotype_list.rds") %>%
  get_vals_df(.) %>%
  mutate(Patient = "P108TA",
         type = "phenotype",
         pval = case_when(pval == 0 ~ 1/5000,
                          T ~ pval),
         qval = p.adjust(pval, method = "BH"))
```

# Save results to compare to QC
## Merge clonotype and phenotype dataframes for each patient
```{r}
# P102TA
P102TA_tls <- P102TA_clone_tls %>%
  rbind(P102TA_phenotype_tls) %>%
  mutate(Region = "TLS")
P102TA_boundary <- P102TA_clone_boundary %>%
  rbind(P102TA_phenotype_boundary) %>%
  mutate(Region = "Boundary")
P102TA_tumor <- P102TA_clone_tumor %>%
  rbind(P102TA_phenotype_tumor) %>%
  mutate(Region = "Tumor")

# P107TA
P107TA_tls <- P107TA_clone_tls %>%
  rbind(P107TA_phenotype_tls) %>%
  mutate(Region = "TLS")
P107TA_tumor <- P107TA_clone_tumor %>%
  rbind(P107TA_phenotype_tumor) %>%
  mutate(Region = "Tumor")

# P108TA
P108TA_tls <- P108TA_clone_tls %>%
  rbind(P108TA_phenotype_tls) %>%
  mutate(Region = "TLS")
P108TA_tumor <- P108TA_clone_tumor %>%
  rbind(P108TA_phenotype_tumor) %>%
  mutate(Region = "Tumor")
P108TA_boundary <- P108TA_clone_boundary %>%
  rbind(P108TA_phenotype_boundary) %>%
  mutate(Region = "Boundary")
```

## Merge across patients for each area
```{r}
tls <- rbind(P102TA_tls, P107TA_tls, P108TA_tls)
boundary <- rbind(P102TA_boundary, P108TA_boundary)
tumor <- rbind(P102TA_tumor, P107TA_tumor, P108TA_tumor)
```

## Save results
```{r}
write.csv(tls, "../objects/P102TA_P107TA_P108TA_TLS_aggregated.csv", row.names = FALSE)
write.csv(boundary, "../objects/P102TA_P108TA_boundary_aggregated.csv", row.names = FALSE)
write.csv(tumor, "../objects/P102TA_P107TA_P108TA_tumor_aggregated.csv", row.names = FALSE)
```

# Open scRNA-scTCR derived phenotypes
```{r}
scrna_phenotypes <- read.csv("../objects/reformatted_RCC_slideseq_parameters_052925.csv") %>%
  distinct(sctcr_clone, PrimaryCluster)
```

# Plot TLS
```{r}
TLS_clone <- do.call(rbind, list(P102TA_clone_tls, P107TA_clone_tls, P108TA_clone_tls)) %>%
  mutate(text_size = count_in_area/(12),
         sig_qval = case_when(qval <= sig_level ~ "Significant",
                         qval > sig_level ~ "Non-significant")) %>%
  left_join(scrna_phenotypes, by = c("clone" = "sctcr_clone"))
      
TLS_phenotype <- do.call(rbind, list(P102TA_phenotype_tls, P107TA_phenotype_tls, P108TA_phenotype_tls)) %>%
  mutate(text_size = count_in_area/(12),
         sig_pval = case_when(pval <= sig_level ~ "Significant",
                         pval > sig_level ~ "Non-significant"))
```

## TLS clones
```{r, fig.width = 8, fig.height = 5}
TLS_clone %>% 
  ggplot(aes(x = enrichment, y = -log10(qval), 
             label = ifelse(qval <= sig_level, clone, NA), 
             shape = sig_qval)) +
  geom_hline(color = "grey50", yintercept = -log10(sig_level), linetype = "longdash") +
  geom_vline(color = "black", xintercept = 0) +
  geom_hline(color = "black", yintercept = 0) +
  geom_point(aes(fill = ifelse(qval <= sig_level, PrimaryCluster, ""), 
                 color = ifelse(qval <= sig_level, Patient, ""), 
                 size = count_in_area), shape = 21, stroke = 2) +
  geom_text_repel(alpha = 1, aes(size = 500)) +
  theme_classic() + 
  xlab("Enrichment in TLS") +
  ylab("-log10(adjusted pval)") +
  scale_fill_manual(values = c("white","red", "black")) +
  scale_color_manual(values = c("grey70", "#FF9300","#009193")) +
  guides(color=guide_legend(title="Patient", override.aes = aes(label = "")),
         fill = guide_legend(title = "TCR Clone", override.aes = aes(label = "")),
         size = guide_legend(title = "TCR Beads in Region", override.aes = aes(label = ""))) +
  coord_cartesian(ylim = c(-0.3, 3), xlim = c(-2, 6)) +
  scale_size(breaks=c(1, 31.62, 1000), limits = c(1, 1000))

```

## TLS phenotypes
```{r, fig.width = 8, fig.height = 5}
TLS_phenotype %>%
  ggplot(aes(x = enrichment, y = -log10(pval), label = ifelse(pval <= sig_level, clone, NA))) +
  geom_hline(color = "grey50", yintercept = -log10(sig_level), linetype = "longdash") +
  geom_vline(color = "black", xintercept = 0) +
  geom_hline(color = "black", yintercept = 0) +
  geom_point(aes(fill = ifelse(pval <= sig_level, clone, ""), color = ifelse(pval <= sig_level, Patient, ""), size = count_in_area), shape = 21, stroke = 2) +
  geom_text_repel(alpha = 1, aes(size = 1000)) +
  theme_classic() + 
  xlab("Enrichment in TLS") +
  ylab("-log10(pval)") +
  scale_shape_manual(guide=FALSE, values=c(21, 19)) +
  scale_fill_manual(values = c("white","red")) +
  scale_color_manual(values = c("grey70", "#FF9300","#009193")) +
  guides(color=guide_legend(title="Patient", override.aes = aes(label = "")),
         fill = guide_legend(title = "TCR Phenotype", override.aes = aes(label = "")),
         size = guide_legend(title = "TCR Beads in Region", override.aes = aes(label = ""))) +
  coord_cartesian(ylim = c(-0.3, 4), xlim = c(-1, 2)) +
  scale_size(breaks=c(1, 44.72, 2000), limits = c(1, 2000))
```

# Plot tumor
```{r}
tumor_clone <- do.call(rbind, list(P102TA_clone_tumor, P107TA_clone_tumor, P108TA_clone_tumor)) %>%
  mutate(text_size = count_in_area/(2+40),
         sig_qval = case_when(qval <= sig_level ~ "Significant",
                              qval > sig_level ~ "Non-significant")) %>%
  left_join(scrna_phenotypes, by = c("clone" = "sctcr_clone"))

tumor_phenotype <- do.call(rbind, list(P102TA_phenotype_tumor, P107TA_phenotype_tumor, P108TA_phenotype_tumor)) %>%
  mutate(text_size = count_in_area/(2+40),
         sig_pval = case_when(pval <= sig_level ~ "Significant",
                              pval > sig_level ~ "Non-significant"))
```

## Tumor clones
```{r}
tumor_clone %>% 
  ggplot(aes(x = enrichment, y = -log10(qval), 
             label = ifelse(qval <= sig_level, clone, NA), 
             shape = sig_qval)) +
  geom_hline(color = "grey50", yintercept = -log10(sig_level), linetype = "longdash") +
  geom_vline(color = "black", xintercept = 0) +
  geom_hline(color = "black", yintercept = 0) +
  geom_point(aes(fill = ifelse(qval <= sig_level, PrimaryCluster, ""), 
                 color = ifelse(qval <= sig_level, Patient, ""), 
                 size = count_in_area), shape = 21, stroke = 2) +
  geom_text_repel(alpha = 1, aes(size = 500)) +
  theme_classic() + 
  xlab("Enrichment in Tumor") +
  ylab("-log10(adjusted pval)") +
  scale_fill_manual(values = c("white","red", "black")) +
  scale_color_manual(values = c("grey70", "#FF9300","#009193")) +
  guides(color=guide_legend(title="Patient", override.aes = aes(label = "")),
         fill = guide_legend(title = "TCR Clone", override.aes = aes(label = "")),
         size = guide_legend(title = "TCR Beads in Region", override.aes = aes(label = ""))) +
  coord_cartesian(ylim = c(-0.3, 3), xlim = c(-2, 6)) +
  scale_size(breaks=c(1, 31.62, 1000), limits = c(1, 1000))
```

## Tumor phenotype
```{r, fig.width = 8, fig.height = 5}
tumor_phenotype %>%
  ggplot(aes(x = enrichment, y = -log10(pval), label = ifelse(pval <= sig_level, clone, NA))) +
  geom_hline(color = "grey50", yintercept = -log10(sig_level), linetype = "longdash") +
  geom_vline(color = "black", xintercept = 0) +
  geom_hline(color = "black", yintercept = 0) +
  geom_point(aes(fill = ifelse(pval <= sig_level, clone, ""), color = ifelse(pval <= sig_level, Patient, ""), size = count_in_area), shape = 21, stroke = 2) +
  geom_text_repel(alpha = 1, aes(size = 1000)) +
  theme_classic() + 
  xlab("Enrichment in Tumor") +
  ylab("-log10(pval)") +
  scale_shape_manual(guide=FALSE, values=c(21, 19)) +
  scale_fill_manual(values = c("white","red")) +
  scale_color_manual(values = c("grey70", "#FF9300","#009193")) +
  guides(color=guide_legend(title="Patient", override.aes = aes(label = "")),
         fill = guide_legend(title = "TCR Phenotype", override.aes = aes(label = "")),
         size = guide_legend(title = "TCR Beads in Region", override.aes = aes(label = ""))) +
  coord_cartesian(ylim = c(-0.3, 4), xlim = c(-1, 2)) +
  scale_size(breaks=c(1, 44.72, 2000), limits = c(1, 2000))
```

# Plot boundary
```{r,  fig.width = 8, fig.height = 6}
boundary_clone <- do.call(rbind, list(P102TA_clone_boundary, P108TA_clone_boundary)) %>%
  mutate(text_size = enrichment/(2+100),
         sig_qval = case_when(qval <= sig_level ~ "Significant",
                         qval > sig_level ~ "Non-significant")) %>%
  left_join(scrna_phenotypes, by = c("clone" = "sctcr_clone"))

boundary_phenotype <- do.call(rbind, list(P102TA_phenotype_boundary, P108TA_phenotype_boundary)) %>%
  mutate(text_size = enrichment/(2+100),
         sig_pval = case_when(pval <= sig_level ~ "Significant",
                         pval > sig_level ~ "Non-significant"))
```

## Boundary clones
```{r, fig.width = 8, fig.height = 5}
boundary_clone %>% 
  ggplot(aes(x = enrichment, y = -log10(qval), 
             label = ifelse(qval <= sig_level, clone, NA), 
             shape = sig_qval)) +
  geom_hline(color = "grey50", yintercept = -log10(sig_level), linetype = "longdash") +
  geom_vline(color = "black", xintercept = 0) +
  geom_hline(color = "black", yintercept = 0) +
  geom_point(aes(fill = ifelse(qval <= sig_level, PrimaryCluster, ""), 
                 color = ifelse(qval <= sig_level, Patient, ""), 
                 size = count_in_area), shape = 21, stroke = 2) +
  geom_text_repel(alpha = 1, aes(size = 500)) +
  theme_classic() + 
  xlab("Enrichment in Boundary") +
  ylab("-log10(adjusted pval)") +
  scale_fill_manual(values = c("white","red", "black")) +
  scale_color_manual(values = c("grey70", "#FF9300", "#009193")) +
  guides(color=guide_legend(title="Patient", override.aes = aes(label = "")),
         fill = guide_legend(title = "TCR Clone", override.aes = aes(label = "")),
         size = guide_legend(title = "TCR Beads in Region", override.aes = aes(label = ""))) +
  coord_cartesian(ylim = c(-0.3, 3), xlim = c(-2, 6)) +
  scale_size(breaks=c(1, 31.62, 1000), limits = c(1, 1000)) 
```

## Boundary phenotype
```{r}
boundary_phenotype %>%
  ggplot(aes(x = enrichment, y = -log10(pval), label = ifelse(pval <= sig_level, clone, NA))) +
  geom_hline(color = "grey50", yintercept = -log10(sig_level), linetype = "longdash") +
  geom_vline(color = "black", xintercept = 0) +
  geom_hline(color = "black", yintercept = 0) +
  geom_point(aes(fill = ifelse(pval <= sig_level, clone, ""), color = ifelse(pval <= sig_level, Patient, ""), size = count_in_area), shape = 21, stroke = 2) +
  geom_text_repel(alpha = 1, aes(size = 1000)) +
  theme_classic() + 
  xlab("Enrichment in Boundary") +
  ylab("-log10(pval)") +
  scale_shape_manual(guide=FALSE, values=c(21, 19)) +
  scale_fill_manual(values = c("white","red", "hotpink")) +
  scale_color_manual(values = c("grey70", "#FF9300","#009193")) +
  guides(color=guide_legend(title="Patient", override.aes = aes(label = "")),
         fill = guide_legend(title = "TCR Phenotype", override.aes = aes(label = "")),
         size = guide_legend(title = "TCR Beads in Region", override.aes = aes(label = ""))) +
  coord_cartesian(ylim = c(-0.3, 4), xlim = c(-1, 2)) +
  scale_size(breaks=c(1, 44.72, 2000), limits = c(1, 2000))

```

## Get session info
```{r}
sessionInfo()
```
