# 108TA Boundary: Clone enrichment {#sec-clone-enrichment}

# Calculating enrichment of clonotypes or T cell phenotypes in spatial compartments

## Libraries
```{r}
library(Seurat)
library(SeuratWrappers)
library(ggplot2)
library(patchwork)
library(dplyr)
library(infer)
library(stringr)
library(reshape2)
library(ggrepel)
library(tidyverse)

source("../source_scripts/slideseq_functions_main.R")

options(Seurat.object.assay.version = 'v5')
```

## Set number of permutations and the area to describe clonotype enrichment in
```{r}
n_perm <- 5000
set.seed(123)

given_area = "Boundary"
```


## Load object from script 02- where I integrated ImageScope area annotation for C29 and C29 separately
```{r}
P108TA_26_RNA <- readRDS("../objects/02_Puck26_108TA_TCR_RNA_area_121125.Rds")
P108TA_29_RNA <- readRDS("../objects/02_Puck29_108TA_TCR_RNA_area_121125.Rds")
P108TA_30_RNA <- readRDS("../objects/02_Puck30_108TA_TCR_RNA_area_121125.Rds")
P108TA_32_RNA <- readRDS("../objects/02_Puck32_108TA_TCR_RNA_area_121125.Rds")

unique(P108TA_26_RNA$orig.ident)
unique(P108TA_29_RNA$orig.ident)
unique(P108TA_30_RNA$orig.ident)
unique(P108TA_32_RNA$orig.ident)
```


## Permutation test logic:
Hypothesis:
Our null hypothesis claims that TCR_X is not more enriched in the boundary than any other TCR
The alternative hypothesis claims that TCR_Xs is more enriched in the boundary than any other TCR

Take into account the non-independent variable of puck identity by permuting sample label within the puck. e.g. say Puck30 has 10 TCR_X beads, and Puck32 has 20 TCR_X beads. After permutation, Puck30 will still have 10 TCR_X beads, and Puck32 20 TCR_X beads.

Test statistic:

Enrichment = (number of beads for TCR_X in the structure/total number of beads for TCR_X across all structures)/(number of beads for all other clonotypes in the structure/ total number of beads for all other clonotypes across all structures) - 1

In other words, enrichment = (fraction of TCR_X in the boundary) / (fraction of all other clones in the boundary) - 1

## Get TCRs seen at a count of 5 or more across all pucks
```{r}
# Merge data from all pucks together
P108TA_md <- do.call(rbind, list(P108TA_26_RNA@meta.data, P108TA_29_RNA@meta.data, P108TA_30_RNA@meta.data, P108TA_32_RNA@meta.data))

unique_clones <- get_unique_productive_clones(P108TA_md$clone_id)

P108TA_filtered_clones <- count_clones_by_annotated_area(P108TA_md, unique_clones) %>%
  group_by(clone_id) %>%
  summarize(sum_count = sum(count)) %>%
  filter(sum_count >= 5) %>%
  pull(clone_id)

length(P108TA_filtered_clones)
```

## Filter for beads containing TCRs
```{r}
P108TA_md <- P108TA_md %>%
  filter(!is.na(clone_id))
```

## Permute
```{r, cache = TRUE}
# Save results in a list
boundary_clone_results.list <- list()

for(clone in P108TA_filtered_clones){
  P108TA_TCRX_md <- P108TA_md %>%
    # Select columns of interest
    select(orig.ident, clone_id, barcode, area) %>%
    # Label TCR_X containing beads as "interesting" and label all other beads as "other".
    mutate(tcr_label = case_when(str_detect(clone_id, paste0(clone, "($|,)")) ~ "TCR_of_interest",
                                 T ~ "other_tcr"),
           # Label the boundary as "interesting" and other areas as "other"
           area_label = case_when(area == given_area ~ "area_of_interest",
                                  T ~ "other_area"))
  
  # Find pucks in which TCR_X is not observed
  P108TA_TCRX_pucks <- P108TA_TCRX_md %>%
    filter(tcr_label == "TCR_of_interest") %>%
    distinct(orig.ident) %>%
    pull(orig.ident)
  
  # Remove pucks where TCR_X is not observed
  P108TA_TCRX_md <- P108TA_TCRX_md %>%
    filter(orig.ident %in% P108TA_TCRX_pucks)
  
  # Test enrichment of TCR_X in the boundary
  boundary_results <- enrichment_permutation_test(P108TA_TCRX_md, n_perm = n_perm, clonotype = clone)
  boundary_results$pval
  
  # Print plot only if the clone is significant
  if(boundary_results$pval <= 0.05){
    # Plot the fraction of TCR_X in TLS vs all other TCRs in TLS for each puck.
    # In other words, plot Cc/Call and Ac/Aall
    plot <- P108TA_TCRX_md %>%
      group_by(tcr_label, area_label, orig.ident) %>%
      dplyr::count() %>%
      pivot_wider(names_from = area_label, values_from = "n") %>%
      # Some values are supposed to be 0 but are NA due to dplyr::count
      replace(is.na(.), 0) %>%
      # Calculate Call and Aall
      mutate(all_areas = sum(area_of_interest, other_area),
      # Calculate Cc/Call and Ac/All
             fraction = area_of_interest/all_areas) %>%
      select(tcr_label, fraction, orig.ident) %>%
      ggplot() +
      geom_boxplot(aes(x = tcr_label, y = fraction, fill = tcr_label)) +
      geom_point(aes(x = tcr_label, y = fraction, color = orig.ident)) +
      ggtitle(paste0("TCR of interest: ", clone)) +
      ylab("Fraction in the boundary")
    
    print(plot)
  }
  
  # KEEP THIS ORDER, or the names of the list will be in the wrong order later
  boundary_clone_results.list <- append(boundary_clone_results.list, list(boundary_results))
}


names(boundary_clone_results.list) <- P108TA_filtered_clones
```

## Print pvals
```{r}
boundary_clone_pvals.df <- lapply(boundary_clone_results.list, get_pval) %>%
  as.data.frame() %>%
  t() %>%
  as.data.frame() %>%
  dplyr::rename("pval" = "V1")
```

## Print significant clones (unadjusted)
```{r}
boundary_clone_pvals.df %>%
  filter(pval <= 0.05)
```

## Save object
```{r}
saveRDS(boundary_clone_results.list, "../objects/P108TA_boundary_clone_list.rds")
```

## Get session info
```{r}
sessionInfo()
```
