# Gene expression in myeloid boundary {#sec-gene-exp-plots}

# Boxplot of myeloid cell type signatures in the boundary

# Libraries
```{r}
library(Seurat)
library(dplyr)
library(ggpointdensity)
library(tidyverse)
library(ComplexHeatmap)
library(stringr)
library(readxl)
library(circlize)
library(ggpubr)
library(rcompanion)
library(patchwork)

source("../source_scripts/slideseq_functions_main.R")

options(Seurat.object.assay.version = 'v5')

```

# Process spatial data
## Open objects with annotated areas
```{r}
P102TA_10_RNA <- readRDS("../objects/02_Puck10_102TA_TCR_RNA_area_121125.Rds")
P102TA_17_RNA <- readRDS("../objects/02_Puck17_102TA_TCR_RNA_area_121125.Rds")
P102TA_33_RNA <- readRDS("../objects/02_Puck33_102TA_TCR_RNA_area_121125.Rds")

P108TA_26_RNA <- readRDS("../objects/02_Puck26_108TA_TCR_RNA_area_121125.Rds")
P108TA_29_RNA <- readRDS("../objects/02_Puck29_108TA_TCR_RNA_area_121125.Rds")
P108TA_30_RNA <- readRDS("../objects/02_Puck30_108TA_TCR_RNA_area_121125.Rds")
P108TA_32_RNA <- readRDS("../objects/02_Puck32_108TA_TCR_RNA_area_121125.Rds")
```

## Remove empty areas
```{r}
P102TA_10_RNA <- subset(P102TA_10_RNA, area == "Empty", invert = TRUE)
P102TA_17_RNA <- subset(P102TA_17_RNA, area == "Empty", invert = TRUE)
P102TA_33_RNA <- subset(P102TA_33_RNA, area == "Empty", invert = TRUE)

P108TA_26_RNA <- subset(P108TA_26_RNA, area == "Empty", invert = TRUE)
P108TA_29_RNA <- subset(P108TA_29_RNA, area == "Empty", invert = TRUE)
P108TA_30_RNA <- subset(P108TA_30_RNA, area == "Empty", invert = TRUE)
P108TA_32_RNA <- subset(P108TA_32_RNA, area == "Empty", invert = TRUE)

```

## Normalize data
```{r}
P102TA_10_RNA <- NormalizeData(P102TA_10_RNA, assay = "Spatial")
P102TA_17_RNA <- NormalizeData(P102TA_17_RNA, assay = "Spatial")
P102TA_33_RNA <- NormalizeData(P102TA_33_RNA, assay = "Spatial")

P108TA_26_RNA <- NormalizeData(P108TA_26_RNA, assay = "Spatial")
P108TA_29_RNA <- NormalizeData(P108TA_29_RNA, assay = "Spatial")
P108TA_30_RNA <- NormalizeData(P108TA_30_RNA, assay = "Spatial")
P108TA_32_RNA <- NormalizeData(P108TA_32_RNA, assay = "Spatial")
```

## Load gene sets
```{r} 
ma_gs <- readRDS("../objects/ma_gene_sets.Rds")

angio <- ma_gs$`Angio-TAMs`
la <- ma_gs$`LA-TAMs`
```

# FindAllMarkers
```{r}
myeloid_subclusters <- readRDS( "../objects/myeloid_subclusters_res0.4_cleaned_102_108.rds")

myeloid_subclusters@meta.data <- myeloid_subclusters@meta.data %>%
  mutate(general_myeloid = case_when(RNA_snn_res.0.4 %in% c(0, 8, 3) ~ "Mac",
                                RNA_snn_res.0.4 == 4 ~ "nc_mono",
                                RNA_snn_res.0.4 == 6 ~ "class_mono",
                                RNA_snn_res.0.4 == 7 ~ "cdc"))

Idents(myeloid_subclusters) <- "general_myeloid"
mac_deg <- FindAllMarkers(myeloid_subclusters, only.pos = TRUE, logfc.threshold = 1.5, min.pct = 0.1) 

mac <- mac_deg %>% 
  filter(cluster == "Mac",
         p_val_adj < 0.05) %>% 
  pull(gene)

cdc <- mac_deg %>% 
  filter(cluster == "cdc",
         p_val_adj < 0.05) %>% 
  pull(gene)

mono <- mac_deg %>% 
  filter(cluster == "class_mono",
         p_val_adj < 0.05) %>% 
  pull(gene)

nc_mono <- mac_deg %>% 
  filter(cluster == "nc_mono",
         p_val_adj < 0.05) %>% 
  pull(gene)

```


# Add signature score to spatial objects
```{r, message=FALSE}
P102TA_10_RNA <- P102TA_10_RNA %>%
  AddModuleScore(features = list(mac), name = "mac") %>%
  AddModuleScore(features = list(cdc), name = "cdc") %>%
  AddModuleScore(features = list(mono), name = "mono") %>%
  AddModuleScore(features = list(nc_mono), name = "nc_mono") %>%
  AddModuleScore(features = list(angio), name = "angio") %>%
  AddModuleScore(features = list(la), name = "la")
P102TA_17_RNA <- P102TA_17_RNA %>%
  AddModuleScore(features = list(mac), name = "mac") %>%
  AddModuleScore(features = list(cdc), name = "cdc") %>%
  AddModuleScore(features = list(mono), name = "mono") %>%
  AddModuleScore(features = list(nc_mono), name = "nc_mono") %>%
  AddModuleScore(features = list(angio), name = "angio") %>%
  AddModuleScore(features = list(la), name = "la")
P102TA_33_RNA <- P102TA_33_RNA %>%
  AddModuleScore(features = list(mac), name = "mac") %>%
  AddModuleScore(features = list(cdc), name = "cdc") %>%
  AddModuleScore(features = list(mono), name = "mono") %>%
  AddModuleScore(features = list(nc_mono), name = "nc_mono") %>%
  AddModuleScore(features = list(angio), name = "angio") %>%
  AddModuleScore(features = list(la), name = "la")

P108TA_26_RNA <- P108TA_26_RNA %>%
  AddModuleScore(features = list(mac), name = "mac") %>%
  AddModuleScore(features = list(cdc), name = "cdc") %>%
  AddModuleScore(features = list(mono), name = "mono") %>%
  AddModuleScore(features = list(nc_mono), name = "nc_mono") %>%
  AddModuleScore(features = list(angio), name = "angio") %>%
  AddModuleScore(features = list(la), name = "la")
P108TA_29_RNA <- P108TA_29_RNA %>%
  AddModuleScore(features = list(mac), name = "mac") %>%
  AddModuleScore(features = list(cdc), name = "cdc") %>%
  AddModuleScore(features = list(mono), name = "mono") %>%
  AddModuleScore(features = list(nc_mono), name = "nc_mono")  %>%
  AddModuleScore(features = list(angio), name = "angio") %>%
  AddModuleScore(features = list(la), name = "la")
P108TA_30_RNA <- P108TA_30_RNA %>%
  AddModuleScore(features = list(mac), name = "mac") %>%
  AddModuleScore(features = list(cdc), name = "cdc") %>%
  AddModuleScore(features = list(mono), name = "mono") %>%
  AddModuleScore(features = list(nc_mono), name = "nc_mono") %>%
  AddModuleScore(features = list(angio), name = "angio") %>%
  AddModuleScore(features = list(la), name = "la")
P108TA_32_RNA <- P108TA_32_RNA %>%
  AddModuleScore(features = list(mac), name = "mac") %>%
  AddModuleScore(features = list(cdc), name = "cdc") %>%
  AddModuleScore(features = list(mono), name = "mono") %>%
  AddModuleScore(features = list(nc_mono), name = "nc_mono")  %>%
  AddModuleScore(features = list(angio), name = "angio") %>%
  AddModuleScore(features = list(la), name = "la")

```

# Spatially plot myeloid cluster scores
```{r, message=FALSE}
SpatialFeaturePlot(P102TA_10_RNA, features = c("cdc1", "mono1", "nc_mono1", "mac1"), alpha = c(0.01, 1)) &
  scale_fill_viridis_c(option = "inferno", direction = -1)
SpatialFeaturePlot(P102TA_17_RNA, features = c("cdc1", "mono1", "nc_mono1", "mac1"), alpha = c(0.01, 1)) &
  scale_fill_viridis_c(option = "inferno", direction = -1)
SpatialFeaturePlot(P102TA_33_RNA, features = c("cdc1", "mono1", "nc_mono1", "mac1"), alpha = c(0.01, 1)) &
  scale_fill_viridis_c(option = "inferno", direction = -1)

SpatialFeaturePlot(P108TA_26_RNA, features = c("cdc1", "mono1", "nc_mono1", "mac1"), alpha = c(0.01, 1)) &
  scale_fill_viridis_c(option = "inferno", direction = -1)
SpatialFeaturePlot(P108TA_29_RNA, features = c("cdc1", "mono1", "nc_mono1", "mac1"), alpha = c(0.01, 1)) &
  scale_fill_viridis_c(option = "inferno", direction = -1)
SpatialFeaturePlot(P108TA_30_RNA, features = c("cdc1", "mono1", "nc_mono1", "mac1"), alpha = c(0.01, 1)) &
  scale_fill_viridis_c(option = "inferno", direction = -1)
SpatialFeaturePlot(P108TA_32_RNA, features = c("cdc1", "mono1", "nc_mono1", "mac1"), alpha = c(0.01, 1)) &
  scale_fill_viridis_c(option = "inferno", direction = -1)
```


# Create big dataframe
```{r}
# Remove metadata columns that other pucks don't have
P102TA_17_RNA@meta.data <- P102TA_17_RNA@meta.data %>%
  select(-largest_clone_id, -puck17_count, -log_puck17_count, -Fibroblast)
P102TA_10_RNA@meta.data <- P102TA_10_RNA@meta.data %>%
  select(-largest_clone_id, -puck10_count, -log_puck10_count, -Fibroblast)
P102TA_33_RNA@meta.data <- P102TA_33_RNA@meta.data %>%
  select(-Fibroblast)

P108TA_26_RNA@meta.data <- P108TA_26_RNA@meta.data %>%
  select(-Fibroblast)
P108TA_29_RNA@meta.data <- P108TA_29_RNA@meta.data %>%
  select(-Fibroblast)
P108TA_30_RNA@meta.data <- P108TA_30_RNA@meta.data %>%
  select(-Fibroblast)
P108TA_32_RNA@meta.data <- P108TA_32_RNA@meta.data %>%
  select(-Fibroblast)

md <- do.call(rbind, list(P102TA_10_RNA@meta.data, P102TA_17_RNA@meta.data, P102TA_33_RNA@meta.data,
                                 P108TA_26_RNA@meta.data, P108TA_29_RNA@meta.data, P108TA_30_RNA@meta.data, P108TA_32_RNA@meta.data))

```

# Plot
## Subset for the signature of myeloid cells
```{r}
myeloid_signature_md <- md %>%
  assign_celltypes %>%
  filter(assignment1 == "Myeloid",
         area == "Boundary") %>%
  mutate(Patient = str_split_i(orig.ident, "_", 1),
         area = case_when(area %in% c("TLS1", "TLS2", "TLS3", "M2") ~ "TLS",
                          T ~ area)) %>%
  select(orig.ident, Patient, area, assignment1, mac1, angio1, la1, cdc1, mono1, nc_mono1)
```

## Format data
```{r}
means.short <- myeloid_signature_md %>%
  group_by(orig.ident, Patient, area) %>%
  summarize(mac1 = mean(mac1),
            cdc1 = mean(cdc1),
            angio1 = mean(angio1),
            la1 = mean(la1),
            mono1 = mean(mono1),
            nc_mono1 = mean(nc_mono1)) %>% 
  dplyr::rename("TAM-CD163+" = "mac1",
                "CLASS. MONO" = "mono1",
                "Angio TAMs" = "angio1",
                "LA TAMs" = "la1",
                "NONCLASS. MONO" = "nc_mono1",
                "cDC" = "cdc1")

means <- means.short %>%
  pivot_longer(cols = !c("area", "orig.ident", "Patient"), names_to = "signature", values_to = "score") %>%
  mutate(geneset_ident = case_when(signature %in% c("Angio TAMs", "LA TAMs") ~ "External",
                                   T ~ "Internal"))

points <- myeloid_signature_md %>%
  select(mac1, cdc1, la1, angio1, mono1, nc_mono1, area, orig.ident) %>%
  dplyr::rename("TAM-CD163+" = "mac1",
                "CLASS. MONO" = "mono1",
                "Angio TAMs" = "angio1",
                "LA TAMs" = "la1",
                "NONCLASS. MONO" = "nc_mono1",
                "cDC" = "cdc1") %>%
  pivot_longer(cols = !c("area", "orig.ident"), names_to = "signature", values_to = "score") %>%
  mutate(signature = factor(signature, levels = c("TAM-CD163+", "cDC", "NONCLASS. MONO", "CLASS. MONO", "LA TAMs", "Angio TAMs")),
         geneset_ident = case_when(signature %in% c("Angio TAMs", "LA TAMs") ~ "External",
                                   T ~ "Internal"))
```

# Save as supplementary table
```{r}
supp_table <- myeloid_signature_md %>%
  select(mac1, cdc1, la1, angio1, mono1, nc_mono1) %>%
  mutate(barcode = rownames(.))
  
# write.csv(supp_table, "../Ext Data Fig 9d myeloid signatures.csv", row.names = FALSE)

# write.csv(means.short, "../../../supplementary_tables/Ext Data Fig 9d mean myeloid signatures.csv", row.names = FALSE)
```

## Plot boxplot
```{r, fig.width = 4, fig.height = 6}
plot <- ggplot(points) +
  geom_boxplot(aes(x = signature, y = score, fill = area, color = geneset_ident)) +
  geom_jitter(data = means, aes(y = score, x = signature, fill = Patient), height = 0, width = 0.15, shape = 22) +
  scale_fill_manual(values = c("Boundary" = "white", "P102TA" = "#FF9300", "P108TA" = "#009193")) +
  scale_color_manual(values = c("Internal" = "black", "External" = "grey40")) +
  theme_classic() +
  ylab("") +
  xlab("") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        axis.line = element_blank(),
        panel.border = element_rect(color = "black", fill = NA, linewidth = 0.8)
  ) +
  labs(fill = "Puck-level means")

plot
```

# Save plot
```{r}
# pdf("/n/data2/dfci/medonc/cwu/chloe/slideseq_data/tcr_analysis/RCC_neovax_pucks/final_submission/figures/extended_DEG_boxplot.pdf", width = 5.5, height = 6)
# plot
# dev.off()
```

# Session info
```{r}
sessionInfo()
```

