# Gene expression in myeloid boundary {#sec-gene-exp-plots}

# Boxplot of myeloid cell type signatures in the boundary

# Libraries
```{r}
library(Seurat)
library(dplyr)
library(ggpointdensity)
library(tidyverse)
library(ComplexHeatmap)
library(stringr)
library(readxl)
library(circlize)
library(ggpubr)
library(rcompanion)

source("../source_scripts/slideseq_functions_main.R")

options(Seurat.object.assay.version = 'v5')

```

# Process spatial data
## Open objects with annotated areas
```{r}
P102TA_10_RNA <- readRDS("/n/data2/dfci/medonc/cwu/chloe/slideseq_data/tcr_analysis/RCC_neovax_pucks/analysis_110624/objects/02_area_annotation/02_Puck10_102TA_TCR_RNA_area_121125.Rds")
P102TA_17_RNA <- readRDS("/n/data2/dfci/medonc/cwu/chloe/slideseq_data/tcr_analysis/RCC_neovax_pucks/analysis_110624/objects/02_area_annotation/02_Puck17_102TA_TCR_RNA_area_121125.Rds")
P102TA_33_RNA <- readRDS("/n/data2/dfci/medonc/cwu/chloe/slideseq_data/tcr_analysis/RCC_neovax_pucks/analysis_110624/objects/02_area_annotation/02_Puck33_102TA_TCR_RNA_area_121125.Rds")

P108TA_26_RNA <- readRDS("/n/data2/dfci/medonc/cwu/chloe/slideseq_data/tcr_analysis/RCC_neovax_pucks/analysis_110624/objects/02_area_annotation/02_Puck26_108TA_TCR_RNA_area_121125.Rds")
P108TA_29_RNA <- readRDS("/n/data2/dfci/medonc/cwu/chloe/slideseq_data/tcr_analysis/RCC_neovax_pucks/analysis_110624/objects/02_area_annotation/02_Puck29_108TA_TCR_RNA_area_121125.Rds")
P108TA_30_RNA <- readRDS("/n/data2/dfci/medonc/cwu/chloe/slideseq_data/tcr_analysis/RCC_neovax_pucks/analysis_110624/objects/02_area_annotation/02_Puck30_108TA_TCR_RNA_area_121125.Rds")
P108TA_32_RNA <- readRDS("/n/data2/dfci/medonc/cwu/chloe/slideseq_data/tcr_analysis/RCC_neovax_pucks/analysis_110624/objects/02_area_annotation/02_Puck32_108TA_TCR_RNA_area_121125.Rds")
```

## Remove empty areas
```{r}
P102TA_10_RNA <- subset(P102TA_10_RNA, area == "Empty", invert = TRUE)
P102TA_17_RNA <- subset(P102TA_17_RNA, area == "Empty", invert = TRUE)
P102TA_33_RNA <- subset(P102TA_33_RNA, area == "Empty", invert = TRUE)

P108TA_26_RNA <- subset(P108TA_26_RNA, area == "Empty", invert = TRUE)
P108TA_29_RNA <- subset(P108TA_29_RNA, area == "Empty", invert = TRUE)
P108TA_30_RNA <- subset(P108TA_30_RNA, area == "Empty", invert = TRUE)
P108TA_32_RNA <- subset(P108TA_32_RNA, area == "Empty", invert = TRUE)

```

## Normalize data
```{r}
P102TA_10_RNA <- NormalizeData(P102TA_10_RNA, assay = "Spatial")
P102TA_17_RNA <- NormalizeData(P102TA_17_RNA, assay = "Spatial")
P102TA_33_RNA <- NormalizeData(P102TA_33_RNA, assay = "Spatial")

P108TA_26_RNA <- NormalizeData(P108TA_26_RNA, assay = "Spatial")
P108TA_29_RNA <- NormalizeData(P108TA_29_RNA, assay = "Spatial")
P108TA_30_RNA <- NormalizeData(P108TA_30_RNA, assay = "Spatial")
P108TA_32_RNA <- NormalizeData(P108TA_32_RNA, assay = "Spatial")
```

# Get markers for each myeloid cell type
## Read in the single cell object and create a single cell reference from each patient
```{r}
scrna <- readRDS("/n/data2/dfci/medonc/cwu/alex/objects/merged_RCC.NeoVax.CancerCell.rds")

# Join layers- seurat obj v5 specific code
scrna <- JoinLayers(scrna)
```

## Print clusters
```{r}
DimPlot(scrna, group.by = "unified_cluster") + paletteer::scale_colour_paletteer_d("ggsci::default_igv")
unique(scrna$unified_cluster)
```

## Subset for myeloid clusters
```{r}
myeloid_clusters <- subset(scrna, subset = unified_cluster %in% c("cDC", "MONO.NONCLASS", "MONO.CLASS", "M2.MAC", "MAC", "pDC"))

p102_108_myeloid_clusters <- subset(myeloid_clusters, subset = patient %in% c("16097-102", "16097-108"))

p102_108_myeloid_deg <- FindAllMarkers(p102_108_myeloid_clusters, group.by = "unified_cluster", logfc.threshold = 1.5, min.pct = 0.3, only.pos = TRUE)
```

## Get markers for each myeloid cluster
```{r}
unique(p102_108_myeloid_deg$cluster)

tam_cd163 <- p102_108_myeloid_deg %>% 
  filter(cluster == "M2.MAC",
         p_val_adj < 0.05) %>% 
  pull(gene)

mono <- p102_108_myeloid_deg %>% 
  filter(cluster == "MONO.CLASS",
         p_val_adj < 0.05) %>% 
  pull(gene)

nc_mono <- p102_108_myeloid_deg %>% 
  filter(cluster == "MONO.NONCLASS",
         p_val_adj < 0.05) %>% 
  pull(gene)

cdc <- p102_108_myeloid_deg %>% 
  filter(cluster == "cDC",
         p_val_adj < 0.05) %>% 
  pull(gene)

pdc <- p102_108_myeloid_deg %>% 
  filter(cluster == "pDC",
         p_val_adj < 0.05) %>% 
  pull(gene)

tam2 <- p102_108_myeloid_deg %>% 
  filter(cluster == "MAC",
         p_val_adj < 0.05) %>% 
  pull(gene)

```

# Add signature score to seurat object
```{r}
myeloid_clusters <- myeloid_clusters %>%
  AddModuleScore(features = list(tam_cd163), name = "tam_cd163") %>%
  AddModuleScore(features = list(tam2), name = "tam2") %>%
  AddModuleScore(features = list(pdc), name = "pdc") %>%
  AddModuleScore(features = list(cdc), name = "cdc") %>%
  AddModuleScore(features = list(mono), name = "mono") %>%
  AddModuleScore(features = list(nc_mono), name = "nc_mono")
```

# Check that the gene sets are specific to the cluster
```{r}
DimPlot(myeloid_clusters, group.by = "unified_cluster") + paletteer::scale_colour_paletteer_d("ggsci::default_igv")

FeaturePlot(myeloid_clusters, features = c("tam_cd1631", "tam21", "pdc1", "cdc1", "mono1", "nc_mono1"), max.cutoff = "q95")
```

# Add signature score to spatial objects
```{r, message=FALSE}
P102TA_10_RNA <- P102TA_10_RNA %>%
  AddModuleScore(features = list(tam_cd163), name = "tam_cd163") %>%
  AddModuleScore(features = list(tam2), name = "tam2") %>%
  AddModuleScore(features = list(pdc), name = "pdc") %>%
  AddModuleScore(features = list(cdc), name = "cdc") %>%
  AddModuleScore(features = list(mono), name = "mono") %>%
  AddModuleScore(features = list(nc_mono), name = "nc_mono")
P102TA_17_RNA <- P102TA_17_RNA %>%
  AddModuleScore(features = list(tam_cd163), name = "tam_cd163") %>%
  AddModuleScore(features = list(tam2), name = "tam2") %>%
  AddModuleScore(features = list(pdc), name = "pdc") %>%
  AddModuleScore(features = list(cdc), name = "cdc") %>%
  AddModuleScore(features = list(mono), name = "mono") %>%
  AddModuleScore(features = list(nc_mono), name = "nc_mono")
P102TA_33_RNA <- P102TA_33_RNA %>%
  AddModuleScore(features = list(tam_cd163), name = "tam_cd163") %>%
  AddModuleScore(features = list(tam2), name = "tam2") %>%
  AddModuleScore(features = list(pdc), name = "pdc") %>%
  AddModuleScore(features = list(cdc), name = "cdc") %>%
  AddModuleScore(features = list(mono), name = "mono") %>%
  AddModuleScore(features = list(nc_mono), name = "nc_mono")

P108TA_26_RNA <- P108TA_26_RNA %>%
  AddModuleScore(features = list(tam_cd163), name = "tam_cd163") %>%
  AddModuleScore(features = list(tam2), name = "tam2") %>%
  AddModuleScore(features = list(pdc), name = "pdc") %>%
  AddModuleScore(features = list(cdc), name = "cdc") %>%
  AddModuleScore(features = list(mono), name = "mono") %>%
  AddModuleScore(features = list(nc_mono), name = "nc_mono")
P108TA_29_RNA <- P108TA_29_RNA %>%
  AddModuleScore(features = list(tam_cd163), name = "tam_cd163") %>%
  AddModuleScore(features = list(tam2), name = "tam2") %>%
  AddModuleScore(features = list(pdc), name = "pdc") %>%
  AddModuleScore(features = list(cdc), name = "cdc") %>%
  AddModuleScore(features = list(mono), name = "mono") %>%
  AddModuleScore(features = list(nc_mono), name = "nc_mono")
P108TA_30_RNA <- P108TA_30_RNA %>%
  AddModuleScore(features = list(tam_cd163), name = "tam_cd163") %>%
  AddModuleScore(features = list(tam2), name = "tam2") %>%
  AddModuleScore(features = list(pdc), name = "pdc") %>%
  AddModuleScore(features = list(cdc), name = "cdc") %>%
  AddModuleScore(features = list(mono), name = "mono") %>%
  AddModuleScore(features = list(nc_mono), name = "nc_mono")
P108TA_32_RNA <- P108TA_32_RNA %>%
  AddModuleScore(features = list(tam_cd163), name = "tam_cd163") %>%
  AddModuleScore(features = list(tam2), name = "tam2") %>%
  AddModuleScore(features = list(pdc), name = "pdc") %>%
  AddModuleScore(features = list(cdc), name = "cdc") %>%
  AddModuleScore(features = list(mono), name = "mono") %>%
  AddModuleScore(features = list(nc_mono), name = "nc_mono")

```

# Spatially plot myeloid cluster scores
```{r, message=FALSE}
SpatialFeaturePlot(P102TA_10_RNA, features = c("cdc1", "mono1", "nc_mono1", "tam_cd1631", "tam21"), alpha = c(0.01, 1)) &
  scale_fill_viridis_c(option = "inferno", direction = -1)
SpatialFeaturePlot(P102TA_17_RNA, features = c("cdc1", "mono1", "nc_mono1", "tam_cd1631", "tam21"), alpha = c(0.01, 1)) &
  scale_fill_viridis_c(option = "inferno", direction = -1)
SpatialFeaturePlot(P102TA_33_RNA, features = c("cdc1", "mono1", "nc_mono1", "tam_cd1631", "tam21"), alpha = c(0.01, 1)) &
  scale_fill_viridis_c(option = "inferno", direction = -1)

SpatialFeaturePlot(P108TA_26_RNA, features = c("cdc1", "mono1", "nc_mono1", "tam_cd1631", "tam21"), alpha = c(0.01, 1)) &
  scale_fill_viridis_c(option = "inferno", direction = -1)
SpatialFeaturePlot(P108TA_29_RNA, features = c("cdc1", "mono1", "nc_mono1", "tam_cd1631", "tam21"), alpha = c(0.01, 1)) &
  scale_fill_viridis_c(option = "inferno", direction = -1)
SpatialFeaturePlot(P108TA_30_RNA, features = c("cdc1", "mono1", "nc_mono1", "tam_cd1631", "tam21"), alpha = c(0.01, 1)) &
  scale_fill_viridis_c(option = "inferno", direction = -1)
SpatialFeaturePlot(P108TA_32_RNA, features = c("cdc1", "mono1", "nc_mono1", "tam_cd1631", "tam21"), alpha = c(0.01, 1)) &
  scale_fill_viridis_c(option = "inferno", direction = -1)
```


# Create big dataframe
```{r}
# Remove metadata columns that other pucks don't have
P102TA_17_RNA@meta.data <- P102TA_17_RNA@meta.data %>%
  select(-largest_clone_id, -puck17_count, -log_puck17_count, -Fibroblast)
P102TA_10_RNA@meta.data <- P102TA_10_RNA@meta.data %>%
  select(-largest_clone_id, -puck10_count, -log_puck10_count, -Fibroblast)
P102TA_33_RNA@meta.data <- P102TA_33_RNA@meta.data %>%
  select(-Fibroblast)

P108TA_26_RNA@meta.data <- P108TA_26_RNA@meta.data %>%
  select(-Fibroblast)
P108TA_29_RNA@meta.data <- P108TA_29_RNA@meta.data %>%
  select(-Fibroblast)
P108TA_30_RNA@meta.data <- P108TA_30_RNA@meta.data %>%
  select(-Fibroblast)
P108TA_32_RNA@meta.data <- P108TA_32_RNA@meta.data %>%
  select(-Fibroblast)

md <- do.call(rbind, list(P102TA_10_RNA@meta.data, P102TA_17_RNA@meta.data, P102TA_33_RNA@meta.data,
                                 P108TA_26_RNA@meta.data, P108TA_29_RNA@meta.data, P108TA_30_RNA@meta.data, P108TA_32_RNA@meta.data))

```


# Save as supplementary table
```{r}
saved_md <- md %>%
  assign_celltypes() %>%
  filter(assignment1 == "Myeloid",
         area == "Boundary") %>%
  select(orig.ident, tam_cd1631, tam21, cdc1, mono1, nc_mono1)

```

# Subset for the signature of myeloid cells
```{r}
myeloid_signature_md <- md %>%
  assign_celltypes %>%
  filter(assignment1 == "Myeloid",
         area == "Boundary") %>%
  mutate(Patient = str_split_i(orig.ident, "_", 1),
         area = case_when(area %in% c("TLS1", "TLS2", "TLS3", "M2") ~ "TLS",
                          T ~ area)) %>%
  select(orig.ident, Patient, area, assignment1, tam_cd1631, tam21, pdc1, cdc1, mono1, nc_mono1)
```


# Plot boxplot
```{r, fig.width = 4, fig.height = 4}
means <- myeloid_signature_md %>%
  group_by(orig.ident, Patient, area) %>%
  summarize(tam_cd1631 = mean(tam_cd1631),
            tam21 = mean(tam21),
            cdc1 = mean(cdc1),
            pdc1 = mean(pdc1),
            mono1 = mean(mono1),
            nc_mono1 = mean(nc_mono1)) %>% 
  dplyr::rename("TAM-CD163+" = "tam_cd1631",
                "CLASS. MONO" = "mono1",
                "NONCLASS. MONO" = "nc_mono1",
                "cDC" = "cdc1",
                "pDC" = "pdc1",
                "TAM2" = "tam21") %>%
  pivot_longer(cols = !c("area", "orig.ident", "Patient"), names_to = "signature", values_to = "score")

myeloid_signature_md %>%
  select(tam_cd1631, tam21, cdc1, pdc1, mono1, nc_mono1, area, orig.ident) %>%
  dplyr::rename("TAM-CD163+" = "tam_cd1631",
                "CLASS. MONO" = "mono1",
                "NONCLASS. MONO" = "nc_mono1",
                "cDC" = "cdc1",
                "pDC" = "pdc1",
                "TAM2" = "tam21") %>%
  pivot_longer(cols = !c("area", "orig.ident"), names_to = "signature", values_to = "score") %>%
  mutate(signature = factor(signature, levels = c("TAM-CD163+", "NONCLASS. MONO", "CLASS. MONO", "cDC", "pDC", "TAM2"))) %>%
  ggplot() +
  geom_boxplot(aes(x = signature, y = score, fill = area)) +
  geom_jitter(data = means, aes(y = score, x = signature, fill = Patient), height = 0, width = 0.15, shape = 22) +
  scale_fill_manual(values = c("Boundary" = "white", "P102TA" = "#FF9300", "P108TA" = "#009193")) +
  theme_classic() +
  ylab("Signature score (myeloid cells)") +
  xlab("") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

# Session info
```{r}
sessionInfo()
```

