# Gene expression in shared CD8 Tex TCRs {#sec-gene-exp-plots}

# Dotplot of shared CD8 Tex TCRs in the TLS

# Libraries
```{r}
library(Seurat)
library(SeuratWrappers)
library(ggplot2)
library(patchwork)
library(dplyr)
library(wesanderson)
library(ggplotify)
library(cowplot)
library(stringr)
library(reshape2)
library(ggrepel)
library(tidyverse)
library(infer)
library(RANN)
library(muscat)
library(limma)
library(purrr)
library(SummarizedExperiment)
library(fgsea)
library(msigdbr)
library(ProjecTILs)
library(readxl)
library(UpSetR)
library(ggpointdensity)
library(ComplexHeatmap)
library(scales)

source("../source_scripts/slideseq_functions_main.R")

options(Seurat.object.assay.version = 'v5')

colors <- structure(c("white", paletteer::paletteer_d("khroma::smoothrainbow")), names = as.character(c(0:34)))

set.seed(123)
```

# Explore the scRNA derived phenotypes of the shared clones
## Open scRNA derived phenotypes
```{r}
scrna_phenotypes <- read.csv("../objects/reformatted_RCC_slideseq_parameters_052925.csv")

scrna_exh_clones <- scrna_phenotypes %>%
  filter(PrimaryCluster == "Tex") %>%
  distinct(sctcr_clone) %>%
  pull(sctcr_clone)
```

# Load objects
```{r}
P102TA_10_RNA <- readRDS("../objects/02_Puck10_102TA_TCR_RNA_area_121125.Rds")
P102TA_33_RNA <- readRDS("../objects/02_Puck33_102TA_TCR_RNA_area_121125.Rds")

P108TA_26_RNA <- readRDS("../objects/02_Puck26_108TA_TCR_RNA_area_121125.Rds")
P108TA_29_RNA <- readRDS("../objects/02_Puck29_108TA_TCR_RNA_area_121125.Rds")
P108TA_30_RNA <- readRDS("../objects/02_Puck30_108TA_TCR_RNA_area_121125.Rds")
P108TA_32_RNA <- readRDS("../objects/02_Puck32_108TA_TCR_RNA_area_121125.Rds")
```

## Add a generalized column "tls_status"
```{r}
add_TLS_column <- function(obj){
  obj <- obj %>%
    mutate(tls_status = case_when(area %in% c("TLS1", "TLS2", "TLS3", "TLS4", "M2") ~ "inTLS",
                              area %in% c("Boundary", "Tumor") ~ "outTLS"))
  
  return(obj)
}

P102TA_10_RNA@meta.data <- P102TA_10_RNA@meta.data %>%
  add_TLS_column() %>%
  mutate(area = case_when(area == "M2" ~ "TLS2",
                          T ~ area))
P102TA_33_RNA@meta.data <- P102TA_33_RNA@meta.data %>%
  add_TLS_column()

P108TA_26_RNA@meta.data <- P108TA_26_RNA@meta.data %>%
  add_TLS_column()
P108TA_29_RNA@meta.data <- P108TA_29_RNA@meta.data %>%
  add_TLS_column()
P108TA_30_RNA@meta.data <- P108TA_30_RNA@meta.data %>%
  add_TLS_column()
P108TA_32_RNA@meta.data <- P108TA_32_RNA@meta.data %>%
  add_TLS_column()
```

## Assign celltypes
```{r}
P102TA_10_RNA@meta.data <- P102TA_10_RNA@meta.data %>%
  assign_celltypes() %>%
  order_celltypes()
P102TA_10_RNA@meta.data %>%
  dplyr::count(assignment1)
P102TA_33_RNA@meta.data <- P102TA_33_RNA@meta.data %>%
  assign_celltypes() %>%
  order_celltypes()
P102TA_33_RNA@meta.data %>%
  dplyr::count(assignment1)

P108TA_26_RNA@meta.data <- P108TA_26_RNA@meta.data %>%
  assign_celltypes() %>%
  order_celltypes()
P108TA_26_RNA@meta.data %>%
  dplyr::count(assignment1)
P108TA_29_RNA@meta.data <- P108TA_29_RNA@meta.data %>%
  assign_celltypes() %>%
  order_celltypes()
P108TA_29_RNA@meta.data %>%
  dplyr::count(assignment1)
P108TA_30_RNA@meta.data <- P108TA_30_RNA@meta.data %>%
  assign_celltypes() %>%
  order_celltypes()
P108TA_30_RNA@meta.data %>%
  dplyr::count(assignment1)
P108TA_32_RNA@meta.data <- P108TA_32_RNA@meta.data %>%
  assign_celltypes() %>%
  order_celltypes()
P108TA_32_RNA@meta.data %>%
  dplyr::count(assignment1)

```

# Remove empty beads
```{r}
P102TA_10_RNA <- subset(P102TA_10_RNA, subset = area == "Empty", invert = TRUE)
P102TA_33_RNA <- subset(P102TA_33_RNA, subset = area == "Empty", invert = TRUE)

P108TA_26_RNA <- subset(P108TA_26_RNA, subset = area == "Empty", invert = TRUE)
P108TA_29_RNA <- subset(P108TA_29_RNA, subset = area == "Empty", invert = TRUE)
P108TA_30_RNA <- subset(P108TA_30_RNA, subset = area == "Empty", invert = TRUE)
P108TA_32_RNA <- subset(P108TA_32_RNA, subset = area == "Empty", invert = TRUE)
```

# Remove extra columns
## P102
```{r}
P102TA_10_RNA@meta.data <- P102TA_10_RNA@meta.data %>%
  select(-c(puck10_count, log_puck10_count, largest_clone_id))
```

# Normalize counts
```{r}
P102TA_10_RNA <- NormalizeData(P102TA_10_RNA, assay = "Spatial")
P102TA_33_RNA <- NormalizeData(P102TA_33_RNA, assay = "Spatial")

P108TA_26_RNA <- NormalizeData(P108TA_26_RNA, assay = "Spatial")
P108TA_29_RNA <- NormalizeData(P108TA_29_RNA, assay = "Spatial")
P108TA_30_RNA <- NormalizeData(P108TA_30_RNA, assay = "Spatial")
P108TA_32_RNA <- NormalizeData(P108TA_32_RNA, assay = "Spatial")

```

# Get shared clonotypes in both the TLS and non-TLS areas per patient
## P102TA
```{r}
P102TA_md <- rbind(P102TA_10_RNA@meta.data, P102TA_33_RNA@meta.data)

P102TA_tls_md <- P102TA_md %>%
  filter(tls_status == "inTLS")

P102TA_nontls_md <- P102TA_md %>%
  filter(tls_status == "outTLS")

P102TA_tls_clones_counts <- count_clones_by_annotated_area(P102TA_tls_md, get_unique_productive_clones(P102TA_tls_md$clone_id)) %>%
  mutate(area = case_when(area %in% c("TLS1", "TLS2", "TLS3", "M2") ~ "TLS",
                          T ~ "wrong")) %>%
  group_by(clone_id, area) %>%
  summarize(sum_count = sum(count)) %>%
  ungroup()
  
P102TA_nontls_clones_counts <- count_clones_by_annotated_area(P102TA_nontls_md, get_unique_productive_clones(P102TA_nontls_md$clone_id)) %>%
  mutate(area = case_when(area %in% c("Tumor", "Boundary") ~ "nonTLS",
                          T ~ "wrong")) %>%
  group_by(clone_id, area) %>%
  summarize(sum_count = sum(count)) %>%
  ungroup()

# We want clones seen in 5 beads or more in each region
P102TA_tls_clones <- P102TA_tls_clones_counts %>%
  filter(sum_count >= 5) %>%
  pull(clone_id)
P102TA_nontls_clones <- P102TA_nontls_clones_counts %>%
  filter(sum_count >= 5) %>%
  pull(clone_id)

table(P102TA_tls_clones %in% P102TA_nontls_clones)
table(P102TA_nontls_clones %in% P102TA_tls_clones)

P102TA_shared_clones <- P102TA_tls_clones[P102TA_tls_clones %in% P102TA_nontls_clones]

# # Get CD8 Tex clones seen in scTCR only
P102TA_shared_clones <- P102TA_shared_clones[P102TA_shared_clones %in% scrna_exh_clones]

length(P102TA_shared_clones)
```

## P108TA
```{r}
P108TA_md <- rbind(P108TA_26_RNA@meta.data, P108TA_29_RNA@meta.data, P108TA_30_RNA@meta.data, P108TA_32_RNA@meta.data)

P108TA_tls_md <- P108TA_md %>%
  filter(tls_status == "inTLS")

P108TA_nontls_md <- P108TA_md %>%
  filter(tls_status == "outTLS")

P108TA_tls_clones_counts <- count_clones_by_annotated_area(P108TA_tls_md, get_unique_productive_clones(P108TA_tls_md$clone_id)) %>%
  mutate(area = case_when(area %in% c("TLS1", "TLS2", "TLS3", "M2") ~ "TLS",
                          T ~ "wrong")) %>%
  group_by(clone_id, area) %>%
  summarize(sum_count = sum(count)) %>%
  ungroup()
  
P108TA_nontls_clones_counts <- count_clones_by_annotated_area(P108TA_nontls_md, get_unique_productive_clones(P108TA_nontls_md$clone_id)) %>%
  mutate(area = case_when(area %in% c("Tumor", "Boundary") ~ "nonTLS",
                          T ~ "wrong")) %>%
  group_by(clone_id, area) %>%
  summarize(sum_count = sum(count)) %>%
  ungroup()

# We want clones seen in 5 beads or more in each region
P108TA_tls_clones <- P108TA_tls_clones_counts %>%
  filter(sum_count >= 5) %>%
  pull(clone_id)
P108TA_nontls_clones <- P108TA_nontls_clones_counts %>%
  filter(sum_count >= 5) %>%
  pull(clone_id)

table(P108TA_tls_clones %in% P108TA_nontls_clones)
table(P108TA_nontls_clones %in% P108TA_tls_clones)

P108TA_shared_clones <- P108TA_tls_clones[P108TA_tls_clones %in% P108TA_nontls_clones]

# # Get CD8 Tex clones seen in scTCR only
P108TA_shared_clones <- P108TA_shared_clones[P108TA_shared_clones %in% scrna_exh_clones]

length(P108TA_shared_clones)
```

# Determine shared CD8 Tex clone frequency
```{r}
count_clones_by_annotated_area(P102TA_md, P102TA_shared_clones) %>%
  group_by(clone_id) %>%
  summarize(sum_count = sum(count)) %>%
  arrange(sum_count)

count_clones_by_annotated_area(P108TA_md, P108TA_shared_clones) %>%
  group_by(clone_id) %>%
  summarize(sum_count = sum(count)) %>%
  arrange(sum_count)
```

# Set shared clone order based increasing frequency of clones across all pucks
```{r}
P102TA_shared_clones <- factor(P102TA_shared_clones, levels = c("p102_295", "p102_289", "p102_737", "p102_266", "p102_278", "p102_274", "p102_264", "p102_265"))

P108TA_shared_clones <- factor(P108TA_shared_clones, levels = c("p108_3", "p108_266", "p108_211", "p108_1", "p108_207", "p108_206", "p108_208", "p108_205", "p108_4"))
```

# Cast beads with multiple shared CD8Tex specific clones to the smaller clone
```{r}
cast_shared_cd8s_102 <- function(md){
  md <- md %>%
    mutate(shared_tex_id = case_when(str_detect(sctcr_clone, "p102_295($|,)") ~ "p102_295",
                                  str_detect(sctcr_clone, "p102_289($|,)") ~ "p102_289",
                                  str_detect(sctcr_clone, "p102_737($|,)") ~ "p102_737",
                                  str_detect(sctcr_clone, "p102_266($|,)") ~ "p102_266",
                                  str_detect(sctcr_clone, "p102_278($|,)") ~ "p102_278",
                                  str_detect(sctcr_clone, "p102_274($|,)") ~ "p102_274",
                                  str_detect(sctcr_clone, "p102_264($|,)") ~ "p102_264",
                                  str_detect(sctcr_clone, "p102_265($|,)") ~ "p102_265"))
  return(md)
}

cast_shared_cd8s_108 <- function(md){
  md <- md %>%
    mutate(shared_tex_id = case_when(str_detect(sctcr_clone, "p108_3($|,)") ~ "p108_3",
                                  str_detect(sctcr_clone, "p108_266($|,)") ~ "p108_266",
                                  str_detect(sctcr_clone, "p108_211($|,)") ~ "p108_211",
                                  str_detect(sctcr_clone, "p108_1($|,)") ~ "p108_1",
                                  str_detect(sctcr_clone, "p108_207($|,)") ~ "p108_207",
                                  str_detect(sctcr_clone, "p108_206($|,)") ~ "p108_206",
                                  str_detect(sctcr_clone, "p108_208($|,)") ~ "p108_208",
                                  str_detect(sctcr_clone, "p108_205($|,)") ~ "p108_205",
                                  str_detect(sctcr_clone, "p108_4($|,)") ~ "p108_4"))
  return(md)
}

P102TA_10_RNA@meta.data <- P102TA_10_RNA@meta.data %>%
  cast_shared_cd8s_102
P102TA_33_RNA@meta.data <- P102TA_33_RNA@meta.data %>%
  cast_shared_cd8s_102

P108TA_26_RNA@meta.data <- P108TA_26_RNA@meta.data %>%
  cast_shared_cd8s_108
P108TA_29_RNA@meta.data <- P108TA_29_RNA@meta.data %>%
  cast_shared_cd8s_108
P108TA_30_RNA@meta.data <- P108TA_30_RNA@meta.data %>%
  cast_shared_cd8s_108
P108TA_32_RNA@meta.data <- P108TA_32_RNA@meta.data %>%
  cast_shared_cd8s_108
```


# P102TA
## Puck 10

### Get beads with shared clones in TLS
```{r}
# Get beads with shared clones in TLS
P102TA_10_inTLS <- P102TA_10_RNA@meta.data %>%
  filter(str_detect(clone_id, paste0(paste0(P102TA_shared_clones, "($|,)"), collapse = "|")),
         tls_status == "inTLS") %>%
  pull(barcode)

# Get beads with shared clones outside of TLS
P102TA_10_outTLS <- P102TA_10_RNA@meta.data %>%
  filter(str_detect(clone_id, paste0(paste0(P102TA_shared_clones, "($|,)"), collapse = "|")),
         tls_status == "outTLS") %>%
  pull(barcode)

# How many cells in each category?
length(P102TA_10_inTLS)
length(P102TA_10_outTLS)

```

### Attach assignments to original Seurat object and filter for defined beads
```{r}
P102TA_10_RNA@meta.data <- P102TA_10_RNA@meta.data %>%
  mutate(type = case_when(barcode %in% P102TA_10_inTLS ~ "shared,inTLS",
                          barcode %in% P102TA_10_outTLS ~ "shared,outTLS"),
         muscat = case_when(!is.na(type) ~ "Defined",
                             is.na(type) ~ "Undefined"))

P102TA_10_RNA <- subset(P102TA_10_RNA, subset = muscat == "Defined")
```

## Puck 33
### Get beads with shared clones in TLS
```{r}
# Get beads with shared clones in TLS
P102TA_33_inTLS <- P102TA_33_RNA@meta.data %>%
  filter(str_detect(clone_id, paste0(paste0(P102TA_shared_clones, "($|,)"), collapse = "|")),
         tls_status == "inTLS") %>%
  pull(barcode)

# Get beads with shared clones outside of TLS
P102TA_33_outTLS <- P102TA_33_RNA@meta.data %>%
  filter(str_detect(clone_id, paste0(paste0(P102TA_shared_clones, "($|,)"), collapse = "|")),
         tls_status == "outTLS") %>%
  pull(barcode)

# How many cells in each category?
length(P102TA_33_inTLS)
length(P102TA_33_outTLS)

```

### Attach assignments to original Seurat object and filter for defined beads
```{r}
P102TA_33_RNA@meta.data <- P102TA_33_RNA@meta.data %>%
  mutate(type = case_when(barcode %in% P102TA_33_inTLS ~ "shared,inTLS",
                          barcode %in% P102TA_33_outTLS ~ "shared,outTLS"),
         muscat = case_when(!is.na(type) ~ "Defined",
                             is.na(type) ~ "Undefined"))

P102TA_33_RNA <- subset(P102TA_33_RNA, subset = muscat == "Defined")
```

# P108TA
## Puck 26
### Get beads with shared clones in TLS
```{r}
# Get beads with shared clones in TLS
P108TA_26_inTLS <- P108TA_26_RNA@meta.data %>%
  filter(str_detect(clone_id, paste0(paste0(P108TA_shared_clones, "($|,)"), collapse = "|")),
         tls_status == "inTLS") %>%
  pull(barcode)

# Get beads with shared clones outside of TLS
P108TA_26_outTLS <- P108TA_26_RNA@meta.data %>%
  filter(str_detect(clone_id, paste0(paste0(P108TA_shared_clones, "($|,)"), collapse = "|")),
         tls_status == "outTLS") %>%
  pull(barcode)

# How many cells in each category?
length(P108TA_26_inTLS)
length(P108TA_26_outTLS)

```

### Attach assignments to original Seurat object and filter for defined beads
```{r}
P108TA_26_RNA@meta.data <- P108TA_26_RNA@meta.data %>%
  mutate(type = case_when(barcode %in% P108TA_26_inTLS ~ "shared,inTLS",
                          barcode %in% P108TA_26_outTLS ~ "shared,outTLS"),
         muscat = case_when(!is.na(type) ~ "Defined",
                             is.na(type) ~ "Undefined"))

P108TA_26_RNA <- subset(P108TA_26_RNA, subset = muscat == "Defined")
```

## Puck 29
### Get beads with shared clones in TLS
```{r}
# Get beads with shared clones in TLS
P108TA_29_inTLS <- P108TA_29_RNA@meta.data %>%
  filter(str_detect(clone_id, paste0(paste0(P108TA_shared_clones, "($|,)"), collapse = "|")),
         tls_status == "inTLS") %>%
  pull(barcode)

# Get beads with shared clones outside of TLS
P108TA_29_outTLS <- P108TA_29_RNA@meta.data %>%
  filter(str_detect(clone_id, paste0(paste0(P108TA_shared_clones, "($|,)"), collapse = "|")),
         tls_status == "outTLS") %>%
  pull(barcode)

# How many cells in each category?
length(P108TA_29_inTLS)
length(P108TA_29_outTLS)

```

### Attach assignments to original Seurat object and filter for defined beads
```{r}
P108TA_29_RNA@meta.data <- P108TA_29_RNA@meta.data %>%
  mutate(type = case_when(barcode %in% P108TA_29_inTLS ~ "shared,inTLS",
                          barcode %in% P108TA_29_outTLS ~ "shared,outTLS"),
         muscat = case_when(!is.na(type) ~ "Defined",
                             is.na(type) ~ "Undefined"))

P108TA_29_RNA <- subset(P108TA_29_RNA, subset = muscat == "Defined")
```

## Puck 30
### Get beads with shared clones in TLS
```{r}
# Get beads with shared clones in TLS
P108TA_30_inTLS <- P108TA_30_RNA@meta.data %>%
  filter(str_detect(clone_id, paste0(paste0(P108TA_shared_clones, "($|,)"), collapse = "|")),
         tls_status == "inTLS") %>%
  pull(barcode)

# Get beads with shared clones outside of TLS
P108TA_30_outTLS <- P108TA_30_RNA@meta.data %>%
  filter(str_detect(clone_id, paste0(paste0(P108TA_shared_clones, "($|,)"), collapse = "|")),
         tls_status == "outTLS") %>%
  pull(barcode)

# How many cells in each category?
length(P108TA_30_inTLS)
length(P108TA_30_outTLS)

```

### Attach assignments to original Seurat object and filter for defined beads
```{r}
P108TA_30_RNA@meta.data <- P108TA_30_RNA@meta.data %>%
  mutate(type = case_when(barcode %in% P108TA_30_inTLS ~ "shared,inTLS",
                          barcode %in% P108TA_30_outTLS ~ "shared,outTLS"),
         muscat = case_when(!is.na(type) ~ "Defined",
                             is.na(type) ~ "Undefined"))

P108TA_30_RNA <- subset(P108TA_30_RNA, subset = muscat == "Defined")
```

## Puck 32
### Get beads with shared clones in TLS
```{r}
# Get beads with shared clones in TLS
P108TA_32_inTLS <- P108TA_32_RNA@meta.data %>%
  filter(str_detect(clone_id, paste0(paste0(P108TA_shared_clones, "($|,)"), collapse = "|")),
         tls_status == "inTLS") %>%
  pull(barcode)

# Get beads with shared clones outside of TLS
P108TA_32_outTLS <- P108TA_32_RNA@meta.data %>%
  filter(str_detect(clone_id, paste0(paste0(P108TA_shared_clones, "($|,)"), collapse = "|")),
         tls_status == "outTLS") %>%
  pull(barcode)

# How many cells in each category?
length(P108TA_32_inTLS)
length(P108TA_32_outTLS)

```

### Attach assignments to original Seurat object and filter for defined beads
```{r}
P108TA_32_RNA@meta.data <- P108TA_32_RNA@meta.data %>%
  mutate(type = case_when(barcode %in% P108TA_32_inTLS ~ "shared,inTLS",
                          barcode %in% P108TA_32_outTLS ~ "shared,outTLS"),
         muscat = case_when(!is.na(type) ~ "Defined",
                             is.na(type) ~ "Undefined"))

P108TA_32_RNA <- subset(P108TA_32_RNA, subset = muscat == "Defined")
```


# Genes of interest
```{r}
genes_of_int <- c("PDCD1", "ENTPD1", "LAG3", "HAVCR2", "TOX", "TOX2", "IL7R", "SELL", "CCR7", "TCF7", "CXCL13")
```

# Set up RNA objects

# Merge RNA objects
```{r}
TLS_pucks <- merge(x = P102TA_10_RNA, y = list(P102TA_33_RNA,
                                                   P108TA_26_RNA, P108TA_29_RNA, P108TA_30_RNA, P108TA_32_RNA))

# Update metadata
TLS_pucks@meta.data <- TLS_pucks@meta.data %>%
  mutate(tls_status_gid = str_split_i(type, ",", 2),
         Patient = str_split_i(orig.ident, "_", 1))

TLS_pucks[['Spatial']] <- JoinLayers(TLS_pucks[['Spatial']])

TLS_pucks@meta.data %>%
  dplyr::count(tls_status_gid, Patient)
```


# Get normalized counts and attach metadata
```{r}
TLS_md <- TLS_pucks@meta.data %>%
  select(Patient, orig.ident, area, shared_tex_id, barcode, assignment1, type, tls_status_gid, nCount_Spatial)

counts_of_int <- as.data.frame(LayerData(TLS_pucks, assay = "Spatial", layer = "data")[genes_of_int,]) %>%
  t() %>%
  as.data.frame() %>%
  rownames_to_column("barcode")

TLS_md <- full_join(TLS_md, counts_of_int)

TLS_md <- TLS_md %>%
  mutate(shared_tex_id = factor(shared_tex_id, levels = c("p102_295", "p102_289", "p102_737", "p102_266", "p102_278", "p102_274", "p102_264", "p102_265", "p108_3", "p108_266", "p108_211", "p108_1", "p108_207", "p108_206", "p108_208", "p108_205", "p108_4")))

```

# Explore
## Which shared clones are most expanded?
```{r}
TLS_md %>%
  filter(str_detect(shared_tex_id, "p102")) %>%
  dplyr::count(shared_tex_id, orig.ident) %>%
  dplyr::rename("n_count" = "n") %>%
  pivot_wider(values_from = "n_count", names_from = "orig.ident") %>%
  rowwise() %>%
  mutate(total = rowSums(across(contains("P10")), na.rm = TRUE)) %>%
  arrange(desc(total))

TLS_md %>%
  filter(str_detect(shared_tex_id, "p108")) %>%
  dplyr::count(shared_tex_id, orig.ident) %>%
  dplyr::rename("n_count" = "n") %>%
  pivot_wider(values_from = "n_count", names_from = "orig.ident") %>%
  rowwise() %>%
  mutate(total = rowSums(across(contains("P10")), na.rm = TRUE)) %>%
  arrange(desc(total))
```


## Which shared clones are most expanded in each TLS?
```{r}
TLS_md %>%
  filter(str_detect(shared_tex_id, "p102")) %>%
  dplyr::count(shared_tex_id, tls_status_gid) %>%
  dplyr::rename("n_count" = "n") %>%
  pivot_wider(values_from = "n_count", names_from = "tls_status_gid") %>%
  rowwise() %>%
  mutate(total = rowSums(across(where(is.numeric)), na.rm = TRUE)) %>%
  arrange(desc(total))

TLS_md %>%
  filter(str_detect(shared_tex_id, "p108")) %>%
  dplyr::count(shared_tex_id, tls_status_gid) %>%
  dplyr::rename("n_count" = "n") %>%
  pivot_wider(values_from = "n_count", names_from = "tls_status_gid") %>%
  rowwise() %>%
  mutate(total = rowSums(across(where(is.numeric)), na.rm = TRUE)) %>%
  arrange(desc(total))
```

# Pivot for ggplots
```{r}
TLS_md_long <- TLS_md %>%
  pivot_longer(cols = all_of(genes_of_int), values_to = "counts", names_to = "gene") %>%
mutate(gene = factor(gene, levels = c("TOX", "TOX2", "LAG3", "ENTPD1", "HAVCR2", "PDCD1", "TCF7", "CCR7", "IL7R", "SELL", "CXCL13")))
```

# Plots
## Dotplot split by Patient and area
```{r}
TLS_nonzero <- TLS_md_long %>%
  filter(counts > 0) %>%
  group_by(Patient, tls_status_gid, gene, area) %>%
  summarize(n_cells_nonzero = n())

TLS_md_sum <- TLS_md_long %>%
  group_by(Patient, tls_status_gid, gene, area) %>%
  summarize(mean_counts = mean(counts),
            n_cells = n()) %>%
  full_join(TLS_nonzero, by = c("Patient", "tls_status_gid", "gene", "area")) %>%
  mutate(n_cells_nonzero = replace_na(n_cells_nonzero, 0),
         pct_cells_exp = n_cells_nonzero/n_cells)

TLS_md_sum %>%
  ggplot(aes(x = gene, y = area, fill = mean_counts, size = pct_cells_exp))+
  geom_point(shape = 21, color = "black") +
  theme_bw() +
  facet_wrap(Patient~tls_status_gid, scales = "free_y") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_gradient2(low = "white", mid = "red", high = "red", midpoint = 0.4)

```


## Dotplot split by tcr and tls status
```{r, fig.height = 4, fig.width = 7}
TLS_nonzero <- TLS_md_long %>%
  filter(counts > 0) %>%
  group_by(Patient, tls_status_gid, gene, shared_tex_id) %>%
  summarize(n_cells_nonzero = n())

TLS_md_sum <- TLS_md_long %>%
  group_by(Patient, tls_status_gid, gene, shared_tex_id) %>%
  summarize(mean_counts = mean(counts),
            n_cells = n()) %>%
  full_join(TLS_nonzero, by = c("Patient", "tls_status_gid", "gene", "shared_tex_id")) %>%
  mutate(n_cells_nonzero = replace_na(n_cells_nonzero, 0),
         pct_cells_exp = n_cells_nonzero/n_cells)

TLS_md_sum %>%
  filter(Patient == "P102TA") %>%
  ggplot(aes(x = gene, y = shared_tex_id, fill = mean_counts, size = pct_cells_exp))+
  geom_point(shape = 21, color = "black") +
  theme_bw() +
  facet_wrap(~tls_status_gid) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_gradient2(low = "white", mid = "red", high = "red", midpoint = 1.2) +
  ylab("Shared TEx TCR")

TLS_md_sum %>%
  filter(Patient == "P108TA") %>%
  ggplot(aes(x = gene, y = shared_tex_id, fill = mean_counts, size = pct_cells_exp))+
  geom_point(shape = 21, color = "black") +
  theme_bw() +
  facet_wrap(~tls_status_gid) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_gradient2(low = "white", mid = "red", high = "red", midpoint = 1.2) +
  ylab("Shared TEx TCR")

```

# Get session info
```{r}
sessionInfo()
```

