# Gene expression in shared neoantigen-specific TCRs {#sec-gene-exp-plots}

# Dotplot of shared neoantigen-specific TCRs in the TLS

# Libraries
```{r}
library(Seurat)
library(SeuratWrappers)
library(ggplot2)
library(patchwork)
library(dplyr)
library(wesanderson)
library(ggplotify)
library(cowplot)
library(stringr)
library(reshape2)
library(ggrepel)
library(tidyverse)
library(infer)
library(RANN)
library(muscat)
library(limma)
library(purrr)
library(SummarizedExperiment)
library(fgsea)
library(msigdbr)
library(ProjecTILs)
library(readxl)
library(UpSetR)
library(ggpointdensity)
library(ComplexHeatmap)
library(scales)
library(escape)
library(lme4)

source("../source_scripts/slideseq_functions_main.R")

options(Seurat.object.assay.version = 'v5')

colors <- structure(c("white", paletteer::paletteer_d("khroma::smoothrainbow")), names = as.character(c(0:34)))

set.seed(123)
```

# Set radius to explore
```{r}
rad <- 15
```

# Load objects
```{r}
P108TA_26_RNA <- readRDS("../objects/02_Puck26_108TA_TCR_RNA_area_121125.Rds")
P108TA_29_RNA <- readRDS("../objects/02_Puck29_108TA_TCR_RNA_area_121125.Rds")
P108TA_30_RNA <- readRDS("../objects/02_Puck30_108TA_TCR_RNA_area_121125.Rds")
P108TA_32_RNA <- readRDS("../objects/02_Puck32_108TA_TCR_RNA_area_121125.Rds")

```

## Add a generalized column "tls_status"
```{r}
add_TLS_column <- function(obj){
  obj <- obj %>%
    mutate(tls_status = case_when(area %in% c("TLS1", "TLS2", "TLS3", "TLS4", "M2") ~ "inTLS",
                              area %in% c("Boundary", "Tumor") ~ "outTLS"))
  
  return(obj)
}

P108TA_26_RNA@meta.data <- P108TA_26_RNA@meta.data %>%
  add_TLS_column()
P108TA_29_RNA@meta.data <- P108TA_29_RNA@meta.data %>%
  add_TLS_column()
P108TA_30_RNA@meta.data <- P108TA_30_RNA@meta.data %>%
  add_TLS_column()
P108TA_32_RNA@meta.data <- P108TA_32_RNA@meta.data %>%
  add_TLS_column()
```

## Assign celltypes
```{r}
P108TA_26_RNA@meta.data <- P108TA_26_RNA@meta.data %>%
  assign_celltypes() %>%
  order_celltypes()
P108TA_26_RNA@meta.data %>%
  dplyr::count(assignment1)
P108TA_29_RNA@meta.data <- P108TA_29_RNA@meta.data %>%
  assign_celltypes() %>%
  order_celltypes()
P108TA_29_RNA@meta.data %>%
  dplyr::count(assignment1)
P108TA_30_RNA@meta.data <- P108TA_30_RNA@meta.data %>%
  assign_celltypes() %>%
  order_celltypes()
P108TA_30_RNA@meta.data %>%
  dplyr::count(assignment1)
P108TA_32_RNA@meta.data <- P108TA_32_RNA@meta.data %>%
  assign_celltypes() %>%
  order_celltypes()
P108TA_32_RNA@meta.data %>%
  dplyr::count(assignment1)
```

# Remove empty beads
```{r}
P108TA_26_RNA <- subset(P108TA_26_RNA, subset = area == "Empty", invert = TRUE)
P108TA_29_RNA <- subset(P108TA_29_RNA, subset = area == "Empty", invert = TRUE)
P108TA_30_RNA <- subset(P108TA_30_RNA, subset = area == "Empty", invert = TRUE)
P108TA_32_RNA <- subset(P108TA_32_RNA, subset = area == "Empty", invert = TRUE)
```

# Explore the scRNA derived phenotypes of the shared clones
## Open scRNA derived phenotypes
```{r}
scrna_phenotypes <- read.csv("../objects/reformatted_RCC_slideseq_parameters_052925.csv")

neoantigen_clones <- scrna_phenotypes %>%
  filter(antigen == "NSUN5 p.Q18K") %>%
  distinct(sctcr_clone) %>%
  pull(sctcr_clone)
```

# Get shared clonotypes in both the TLS and non-TLS areas per patient

## P108TA
```{r}
P108TA_md <- rbind(P108TA_26_RNA@meta.data, P108TA_29_RNA@meta.data, P108TA_30_RNA@meta.data, P108TA_32_RNA@meta.data)

P108TA_tls_md <- P108TA_md %>%
  filter(tls_status == "inTLS")

P108TA_nontls_md <- P108TA_md %>%
  filter(tls_status == "outTLS")

P108TA_tls_clones_counts <- count_clones_by_annotated_area(P108TA_tls_md, get_unique_productive_clones(P108TA_tls_md$clone_id)) %>%
  mutate(area = case_when(area %in% c("TLS1", "TLS2", "TLS3", "M2") ~ "TLS",
                          T ~ "wrong")) %>%
  group_by(clone_id, area) %>%
  summarize(sum_count = sum(count)) %>%
  ungroup()
  
P108TA_nontls_clones_counts <- count_clones_by_annotated_area(P108TA_nontls_md, get_unique_productive_clones(P108TA_nontls_md$clone_id)) %>%
  mutate(area = case_when(area %in% c("Tumor", "Boundary") ~ "nonTLS",
                          T ~ "wrong")) %>%
  group_by(clone_id, area) %>%
  summarize(sum_count = sum(count)) %>%
  ungroup()

# We want neoantigen clones seen in 3 beads or more in each region
P108TA_tls_clones <- P108TA_tls_clones_counts %>%
  filter(sum_count >= 3,
         clone_id %in% neoantigen_clones) %>%
  pull(clone_id)
P108TA_nontls_clones <- P108TA_nontls_clones_counts %>%
  filter(sum_count >= 3,
         clone_id %in% neoantigen_clones) %>%
  pull(clone_id)

table(P108TA_tls_clones %in% P108TA_nontls_clones)
table(P108TA_nontls_clones %in% P108TA_tls_clones)

P108TA_shared_clones <- P108TA_tls_clones[P108TA_tls_clones %in% P108TA_nontls_clones]
```

# Set neoantigen clone order based increasing frequency of clones across all pucks
Determined earlier
```{r}
P108TA_shared_clones <- factor(P108TA_shared_clones, levels = c("p108_215", "p108_266", "p108_1", "p108_208", "p108_205"))
```

# Cast beads with multiple neoantigen specific clones to the smaller clone
```{r}
P108TA_26_RNA@meta.data <- P108TA_26_RNA@meta.data %>%
  mutate(neoantigen_id = case_when(str_detect(sctcr_clone, "p108_215($|,)") ~ "p108_215",
                                  str_detect(sctcr_clone, "p108_266($|,)") ~ "p108_266",
                                  str_detect(sctcr_clone, "p108_1($|,)") ~ "p108_1",
                                  str_detect(sctcr_clone, "p108_208($|,)") ~ "p108_208",
                                  str_detect(sctcr_clone, "p108_205($|,)") ~ "p108_205"))
P108TA_29_RNA@meta.data <- P108TA_29_RNA@meta.data %>%
  mutate(neoantigen_id = case_when(str_detect(sctcr_clone, "p108_215($|,)") ~ "p108_215",
                                  str_detect(sctcr_clone, "p108_266($|,)") ~ "p108_266",
                                  str_detect(sctcr_clone, "p108_1($|,)") ~ "p108_1",
                                  str_detect(sctcr_clone, "p108_208($|,)") ~ "p108_208",
                                  str_detect(sctcr_clone, "p108_205($|,)") ~ "p108_205"))
P108TA_30_RNA@meta.data <- P108TA_30_RNA@meta.data %>%
  mutate(neoantigen_id = case_when(str_detect(sctcr_clone, "p108_215($|,)") ~ "p108_215",
                                  str_detect(sctcr_clone, "p108_266($|,)") ~ "p108_266",
                                  str_detect(sctcr_clone, "p108_1($|,)") ~ "p108_1",
                                  str_detect(sctcr_clone, "p108_208($|,)") ~ "p108_208",
                                  str_detect(sctcr_clone, "p108_205($|,)") ~ "p108_205"))
P108TA_32_RNA@meta.data <- P108TA_32_RNA@meta.data %>%
  mutate(neoantigen_id = case_when(str_detect(sctcr_clone, "p108_215($|,)") ~ "p108_215",
                                  str_detect(sctcr_clone, "p108_266($|,)") ~ "p108_266",
                                  str_detect(sctcr_clone, "p108_1($|,)") ~ "p108_1",
                                  str_detect(sctcr_clone, "p108_208($|,)") ~ "p108_208",
                                  str_detect(sctcr_clone, "p108_205($|,)") ~ "p108_205"))
```

# P108TA
## Puck 26

### Get beads near each shared neoantigen clone inside the TLS
```{r}
# Get the beads near shared neoantigen clones
inTLS_list <- list()

for(clone in P108TA_shared_clones){
  result_list <- get_beads_near_specific_neoantigen_TCR2(P108TA_26_RNA, "inTLS", clone, rad)
  inTLS_list <- append(inTLS_list, list(result_list))
}

# Rename lists
names(inTLS_list) <- P108TA_shared_clones
```

### Get beads near each shared neoantigen clone outside the TLS
```{r}
# Get the beads near shared neoantigen clones
outTLS_list <- list()

for(clone in P108TA_shared_clones){
  result_list <- get_beads_near_specific_neoantigen_TCR2(P108TA_26_RNA, "outTLS", clone, rad)
  outTLS_list <- append(outTLS_list, list(result_list))
}

# Rename lists
names(outTLS_list) <- P108TA_shared_clones
```

### Check for overlaps
#### in TLS vs in TLS
```{r}
in_vs_in.df <- data.frame()
for(i in 1:length(P108TA_shared_clones)) {
	for(j in 1:length(P108TA_shared_clones)) {
	  if(i < j){
  		same_beads <- length(intersect(inTLS_list[[i]]$near_beads$barcode, inTLS_list[[j]]$near_beads$barcode))
  		new_row <- c(names(inTLS_list)[i], names(inTLS_list)[j], same_beads)
  		in_vs_in.df <- rbind(in_vs_in.df, new_row)
	  }
	}
}

colnames(in_vs_in.df) <- c("neoantigen1", "neoantigen2", "beads")

# Reformat
in_vs_in.df <- in_vs_in.df %>%
	pivot_wider(names_from=neoantigen2,values_from=beads) %>%
	column_to_rownames("neoantigen1") %>%
	as.matrix()

Heatmap(in_vs_in.df,
        col = colors,
        na_col = "grey80",
        cluster_rows = FALSE,
        cluster_columns = FALSE,
        column_title = "Overlap of nearby beads in TLS vs in TLS",
        name = "No. of\nbeads")
```

#### Outside TLS vs outside TLS
```{r}
out_vs_out.df <- data.frame()
for(i in 1:length(P108TA_shared_clones)) {
	for(j in 1:length(P108TA_shared_clones)) {
	  if(i < j){
  		same_beads <- length(intersect(outTLS_list[[i]]$near_beads$barcode, outTLS_list[[j]]$near_beads$barcode))
  		new_row <- c(names(outTLS_list)[i], names(outTLS_list)[j], same_beads)
  		out_vs_out.df <- rbind(out_vs_out.df, new_row)
	  }
	}
}

colnames(out_vs_out.df) <- c("neoantigen1", "neoantigen2", "beads")

# Reformat
out_vs_out.df <- out_vs_out.df %>%
	pivot_wider(names_from=neoantigen2,values_from=beads) %>%
	column_to_rownames("neoantigen1") %>%
	as.matrix()

Heatmap(out_vs_out.df,
        col = colors,
        na_col = "grey80",
        cluster_rows = FALSE,
        cluster_columns = FALSE,
        column_title = "Overlap of nearby beads outside TLS vs outside TLS",
        name = "No. of\nbeads")
```

##### View overlap of beads in p108_1 and p108_205
```{r}
overlapped_beads <- intersect(outTLS_list$p108_1$near_beads$barcode, outTLS_list$p108_205$near_beads$barcode)

ggplot() +
  geom_point(data = outTLS_list$p108_1$near_beads, aes(x = x, y = y)) +
  geom_point(data = outTLS_list$p108_1$TCRs, aes(x = x, y = y), color = "blue") +
  geom_point(data = filter(P108TA_26_RNA@meta.data, barcode %in% overlapped_beads), aes(x = x, y = y), color = "red") +
  coord_cartesian(xlim = c(0, 5000), ylim = c(0, 5000))

ggplot() +
  geom_point(data = outTLS_list$p108_205$near_beads, aes(x = x, y = y)) +
  geom_point(data = outTLS_list$p108_205$TCRs, aes(x = x, y = y), color = "blue")+
  geom_point(data = filter(P108TA_26_RNA@meta.data, barcode %in% overlapped_beads), aes(x = x, y = y), color = "red") +
  coord_cartesian(xlim = c(0, 5000), ylim = c(0, 5000))
```

##### Reassign overlapping beads
```{r}
# P108_1 and P108_205
# Reassign overlapping beads
P108TA_26_reassigned <- reassign_overlapping_beads(P108TA_26_RNA, tex_beads = outTLS_list$p108_205, tmem_beads = outTLS_list$p108_1)

# Split list back out
outTLS_list$p108_205 <- P108TA_26_reassigned$reassigned_tex_beads
outTLS_list$p108_1 <- P108TA_26_reassigned$reassigned_tmem_beads

# Plot updated beads
ggplot() +
  geom_point(data = outTLS_list$p108_1$near_beads, aes(x = x, y = y)) +
  geom_point(data = outTLS_list$p108_1$TCRs, aes(x = x, y = y), color = "blue") +
  coord_cartesian(xlim = c(0, 5000), ylim = c(0, 5000))

ggplot() +
  geom_point(data = outTLS_list$p108_205$near_beads, aes(x = x, y = y)) +
  geom_point(data = outTLS_list$p108_205$TCRs, aes(x = x, y = y), color = "blue")+
  coord_cartesian(xlim = c(0, 5000), ylim = c(0, 5000))

# P108_208 and P108_205
# Reassign overlapping beads
P108TA_26_reassigned <- reassign_overlapping_beads(P108TA_26_RNA, tex_beads = outTLS_list$p108_205, tmem_beads = outTLS_list$p108_208)

# Split list back out
outTLS_list$p108_205 <- P108TA_26_reassigned$reassigned_tex_beads
outTLS_list$p108_208 <- P108TA_26_reassigned$reassigned_tmem_beads

# P108_215 and P108_205
# Reassign overlapping beads
P108TA_26_reassigned <- reassign_overlapping_beads(P108TA_26_RNA, tex_beads = outTLS_list$p108_205, tmem_beads = outTLS_list$p108_215)

# Split list back out
outTLS_list$p108_205 <- P108TA_26_reassigned$reassigned_tex_beads
outTLS_list$p108_215 <- P108TA_26_reassigned$reassigned_tmem_beads

# P108_266 and P108_205
# Reassign overlapping beads
P108TA_26_reassigned <- reassign_overlapping_beads(P108TA_26_RNA, tex_beads = outTLS_list$p108_205, tmem_beads = outTLS_list$p108_266)

# Split list back out
outTLS_list$p108_205 <- P108TA_26_reassigned$reassigned_tex_beads
outTLS_list$p108_266 <- P108TA_26_reassigned$reassigned_tmem_beads

# P108_266 and P108_208
# Reassign overlapping beads
P108TA_26_reassigned <- reassign_overlapping_beads(P108TA_26_RNA, tex_beads = outTLS_list$p108_266, tmem_beads = outTLS_list$p108_208)

# Split list back out
outTLS_list$p108_266 <- P108TA_26_reassigned$reassigned_tex_beads
outTLS_list$p108_208 <- P108TA_26_reassigned$reassigned_tmem_beads
```

#### Outside TLS vs in TLS
```{r}
out_vs_in.df <- data.frame()
for(i in 1:length(P108TA_shared_clones)) {
	for(j in 1:length(P108TA_shared_clones)) {
	  if(i <= j){
  		same_beads <- length(intersect(outTLS_list[[i]]$near_beads$barcode, inTLS_list[[j]]$near_beads$barcode))
  		new_row <- c(names(outTLS_list)[i], names(inTLS_list)[j], same_beads)
  		out_vs_in.df <- rbind(out_vs_in.df, new_row)
	  }
	}
}

colnames(out_vs_in.df) <- c("neoantigen1", "neoantigen2", "beads")

# Reformat
out_vs_in.df <- out_vs_in.df %>%
	pivot_wider(names_from=neoantigen2,values_from=beads) %>%
	column_to_rownames("neoantigen1") %>%
	as.matrix()

Heatmap(out_vs_in.df,
        col = colors,
        na_col = "grey80",
        cluster_rows = FALSE,
        cluster_columns = FALSE,
        column_title = "Overlap of nearby beads outside TLS vs in TLS",
        name = "No. of\nbeads")
```

### Attach assignments to original Seurat object and filter for defined beads
```{r}
# Attach partition data
partition.df <- data.frame()
for(i in c(1:(length(inTLS_list)))){
  if(nrow(inTLS_list[[i]]$near_beads) > 0){
    near <- inTLS_list[[i]]$near_beads %>%
      select(barcode, partition)
    partition.df <- rbind(partition.df, near)
  }
}
for(i in c(1:(length(outTLS_list)))){
  if(nrow(outTLS_list[[i]]$near_beads) > 0){
    near <- outTLS_list[[i]]$near_beads %>%
      select(barcode, partition)
    partition.df <- rbind(partition.df, near)
  }
}

P108TA_26_RNA@meta.data <- P108TA_26_RNA@meta.data %>%
  left_join(partition.df, by = "barcode") %>%
  column_to_rownames("barcode") %>%
  mutate(barcode = rownames(.))

P108TA_26_RNA@meta.data <- P108TA_26_RNA@meta.data %>%
  mutate(type = case_when(barcode %in% inTLS_list$p108_1$near_beads$barcode ~ "p108_1,inTLS",
                          barcode %in% inTLS_list$p108_205$near_beads$barcode ~ "p108_205,inTLS",
                          barcode %in% inTLS_list$p108_208$near_beads$barcode ~ "p108_208,inTLS",
                          barcode %in% inTLS_list$p108_215$near_beads$barcode ~ "p108_215,inTLS",
                          barcode %in% inTLS_list$p108_266$near_beads$barcode ~ "p108_266,inTLS",
                          barcode %in% outTLS_list$p108_1$near_beads$barcode ~ "p108_1,outTLS",
                          barcode %in% outTLS_list$p108_205$near_beads$barcode ~ "p108_205,outTLS",
                          barcode %in% outTLS_list$p108_208$near_beads$barcode ~ "p108_208,outTLS",
                          barcode %in% outTLS_list$p108_215$near_beads$barcode ~ "p108_215,outTLS",
                          barcode %in% outTLS_list$p108_266$near_beads$barcode ~ "p108_266,outTLS"),
         muscat = case_when(!is.na(type) ~ "Defined",
                             is.na(type) ~ "Undefined"))

P108TA_26_RNA <- subset(P108TA_26_RNA, subset = muscat == "Defined")
```

## Puck 29
### Get beads near each shared neoantigen clone inside the TLS
```{r}
# Get the beads near shared neoantigen clones
inTLS_list <- list()

for(clone in P108TA_shared_clones){
  result_list <- get_beads_near_specific_neoantigen_TCR2(P108TA_29_RNA, "inTLS", clone, rad)
  inTLS_list <- append(inTLS_list, list(result_list))
}

# Rename lists
names(inTLS_list) <- P108TA_shared_clones
```

### Get beads near each shared neoantigen clone outside the TLS
```{r}
# Get the beads near shared neoantigen clones
outTLS_list <- list()

for(clone in P108TA_shared_clones){
  result_list <- get_beads_near_specific_neoantigen_TCR2(P108TA_29_RNA, "outTLS", clone, rad)
  outTLS_list <- append(outTLS_list, list(result_list))
}

# Rename lists
names(outTLS_list) <- P108TA_shared_clones
```

### Check for overlaps
#### in TLS vs in TLS
```{r}
in_vs_in.df <- data.frame()
for(i in 1:length(P108TA_shared_clones)) {
	for(j in 1:length(P108TA_shared_clones)) {
	  if(i < j){
  		same_beads <- length(intersect(inTLS_list[[i]]$near_beads$barcode, inTLS_list[[j]]$near_beads$barcode))
  		new_row <- c(names(inTLS_list)[i], names(inTLS_list)[j], same_beads)
  		in_vs_in.df <- rbind(in_vs_in.df, new_row)
	  }
	}
}

colnames(in_vs_in.df) <- c("neoantigen1", "neoantigen2", "beads")

# Reformat
in_vs_in.df <- in_vs_in.df %>%
	pivot_wider(names_from=neoantigen2,values_from=beads) %>%
	column_to_rownames("neoantigen1") %>%
	as.matrix()

Heatmap(in_vs_in.df,
        col = colors,
        na_col = "grey80",
        cluster_rows = FALSE,
        cluster_columns = FALSE,
        column_title = "Overlap of nearby beads in TLS vs in TLS",
        name = "No. of\nbeads")
```

#### Outside TLS vs outside tls
```{r}
out_vs_out.df <- data.frame()
for(i in 1:length(P108TA_shared_clones)) {
	for(j in 1:length(P108TA_shared_clones)) {
	  if(i < j){
  		same_beads <- length(intersect(outTLS_list[[i]]$near_beads$barcode, outTLS_list[[j]]$near_beads$barcode))
  		new_row <- c(names(outTLS_list)[i], names(outTLS_list)[j], same_beads)
  		out_vs_out.df <- rbind(out_vs_out.df, new_row)
	  }
	}
}

colnames(out_vs_out.df) <- c("neoantigen1", "neoantigen2", "beads")

# Reformat
out_vs_out.df <- out_vs_out.df %>%
	pivot_wider(names_from=neoantigen2,values_from=beads) %>%
	column_to_rownames("neoantigen1") %>%
	as.matrix()

Heatmap(out_vs_out.df,
        col = colors,
        na_col = "grey80",
        cluster_rows = FALSE,
        cluster_columns = FALSE,
        column_title = "Overlap of nearby beads outside TLS vs outside TLS",
        name = "No. of\nbeads")
```

##### Reassign overlapping beads
```{r}
# P108_208 and P108_205
# Reassign overlapping beads
P108TA_29_reassigned <- reassign_overlapping_beads(P108TA_29_RNA, tex_beads = outTLS_list$p108_205, tmem_beads = outTLS_list$p108_208)

# Split list back out
outTLS_list$p108_205 <- P108TA_29_reassigned$reassigned_tex_beads
outTLS_list$p108_208 <- P108TA_29_reassigned$reassigned_tmem_beads

# P108_208 and P108_215
# Reassign overlapping beads
P108TA_29_reassigned <- reassign_overlapping_beads(P108TA_29_RNA, tex_beads = outTLS_list$p108_215, tmem_beads = outTLS_list$p108_208)

# Split list back out
outTLS_list$p108_215 <- P108TA_29_reassigned$reassigned_tex_beads
outTLS_list$p108_208 <- P108TA_29_reassigned$reassigned_tmem_beads
```

#### Outside TLS vs in TLS
```{r}
out_vs_in.df <- data.frame()
for(i in 1:length(P108TA_shared_clones)) {
	for(j in 1:length(P108TA_shared_clones)) {
	  if(i <= j){
  		same_beads <- length(intersect(outTLS_list[[i]]$near_beads$barcode, inTLS_list[[j]]$near_beads$barcode))
  		new_row <- c(names(outTLS_list)[i], names(inTLS_list)[j], same_beads)
  		out_vs_in.df <- rbind(out_vs_in.df, new_row)
	  }
	}
}

colnames(out_vs_in.df) <- c("neoantigen1", "neoantigen2", "beads")

# Reformat
out_vs_in.df <- out_vs_in.df %>%
	pivot_wider(names_from=neoantigen2,values_from=beads) %>%
	column_to_rownames("neoantigen1") %>%
	as.matrix()

Heatmap(out_vs_in.df,
        col = colors,
        na_col = "grey80",
        cluster_rows = FALSE,
        cluster_columns = FALSE,
        column_title = "Overlap of nearby beads outside TLS vs in TLS",
        name = "No. of\nbeads")
```

### Attach assignments to original Seurat object and filter for defined beads
```{r}
# Attach partition data
partition.df <- data.frame()
for(i in c(1:(length(inTLS_list)))){
  if(nrow(inTLS_list[[i]]$near_beads) > 0){
    near <- inTLS_list[[i]]$near_beads %>%
      select(barcode, partition)
    partition.df <- rbind(partition.df, near)
  }
}
for(i in c(1:(length(outTLS_list)))){
  if(nrow(outTLS_list[[i]]$near_beads) > 0){
    near <- outTLS_list[[i]]$near_beads %>%
      select(barcode, partition)
    partition.df <- rbind(partition.df, near)
  }
}

P108TA_29_RNA@meta.data <- P108TA_29_RNA@meta.data %>%
  left_join(partition.df, by = "barcode") %>%
  column_to_rownames("barcode") %>%
  mutate(barcode = rownames(.))


P108TA_29_RNA@meta.data <- P108TA_29_RNA@meta.data %>%
  mutate(type = case_when(barcode %in% inTLS_list$p108_1$near_beads$barcode ~ "p108_1,inTLS",
                          barcode %in% inTLS_list$p108_205$near_beads$barcode ~ "p108_205,inTLS",
                          barcode %in% inTLS_list$p108_208$near_beads$barcode ~ "p108_208,inTLS",
                          barcode %in% inTLS_list$p108_215$near_beads$barcode ~ "p108_215,inTLS",
                          barcode %in% inTLS_list$p108_266$near_beads$barcode ~ "p108_266,inTLS",
                          barcode %in% outTLS_list$p108_1$near_beads$barcode ~ "p108_1,outTLS",
                          barcode %in% outTLS_list$p108_205$near_beads$barcode ~ "p108_205,outTLS",
                          barcode %in% outTLS_list$p108_208$near_beads$barcode ~ "p108_208,outTLS",
                          barcode %in% outTLS_list$p108_215$near_beads$barcode ~ "p108_215,outTLS",
                          barcode %in% outTLS_list$p108_266$near_beads$barcode ~ "p108_266,outTLS"),
         muscat = case_when(!is.na(type) ~ "Defined",
                             is.na(type) ~ "Undefined"))

P108TA_29_RNA <- subset(P108TA_29_RNA, subset = muscat == "Defined")
```

## Puck 30
### Get beads near each shared neoantigen clone inside the TLS
```{r}
# Get the beads near shared neoantigen clones
inTLS_list <- list()

for(clone in P108TA_shared_clones){
  result_list <- get_beads_near_specific_neoantigen_TCR2(P108TA_30_RNA, "inTLS", clone, rad)
  inTLS_list <- append(inTLS_list, list(result_list))
}

# Rename lists
names(inTLS_list) <- P108TA_shared_clones
```

### Get beads near each shared neoantigen clone outside the TLS
```{r}
# Get the beads near shared neoantigen clones
outTLS_list <- list()

for(clone in P108TA_shared_clones){
  result_list <- get_beads_near_specific_neoantigen_TCR2(P108TA_30_RNA, "outTLS", clone, rad)
  outTLS_list <- append(outTLS_list, list(result_list))
}

# Rename lists
names(outTLS_list) <- P108TA_shared_clones
```

### Check for overlaps
#### in TLS vs in TLS
```{r}
in_vs_in.df <- data.frame()
for(i in 1:length(P108TA_shared_clones)) {
	for(j in 1:length(P108TA_shared_clones)) {
	  if(i < j){
  		same_beads <- length(intersect(inTLS_list[[i]]$near_beads$barcode, inTLS_list[[j]]$near_beads$barcode))
  		new_row <- c(names(inTLS_list)[i], names(inTLS_list)[j], same_beads)
  		in_vs_in.df <- rbind(in_vs_in.df, new_row)
	  }
	}
}

colnames(in_vs_in.df) <- c("neoantigen1", "neoantigen2", "beads")

# Reformat
in_vs_in.df <- in_vs_in.df %>%
	pivot_wider(names_from=neoantigen2,values_from=beads) %>%
	column_to_rownames("neoantigen1") %>%
	as.matrix()

Heatmap(in_vs_in.df,
        col = colors,
        na_col = "grey80",
        cluster_rows = FALSE,
        cluster_columns = FALSE,
        column_title = "Overlap of nearby beads in TLS vs in TLS",
        name = "No. of\nbeads")
```

#### Outside TLS vs outside tls
```{r}
out_vs_out.df <- data.frame()
for(i in 1:length(P108TA_shared_clones)) {
	for(j in 1:length(P108TA_shared_clones)) {
	  if(i < j){
  		same_beads <- length(intersect(outTLS_list[[i]]$near_beads$barcode, outTLS_list[[j]]$near_beads$barcode))
  		new_row <- c(names(outTLS_list)[i], names(outTLS_list)[j], same_beads)
  		out_vs_out.df <- rbind(out_vs_out.df, new_row)
	  }
	}
}

colnames(out_vs_out.df) <- c("neoantigen1", "neoantigen2", "beads")

# Reformat
out_vs_out.df <- out_vs_out.df %>%
	pivot_wider(names_from=neoantigen2,values_from=beads) %>%
	column_to_rownames("neoantigen1") %>%
	as.matrix()

Heatmap(out_vs_out.df,
        col = colors,
        na_col = "grey80",
        cluster_rows = FALSE,
        cluster_columns = FALSE,
        column_title = "Overlap of nearby beads outside TLS vs outside TLS",
        name = "No. of\nbeads")
```

#### Outside TLS vs in TLS
```{r}
out_vs_in.df <- data.frame()
for(i in 1:length(P108TA_shared_clones)) {
	for(j in 1:length(P108TA_shared_clones)) {
	  if(i <= j){
  		same_beads <- length(intersect(outTLS_list[[i]]$near_beads$barcode, inTLS_list[[j]]$near_beads$barcode))
  		new_row <- c(names(outTLS_list)[i], names(inTLS_list)[j], same_beads)
  		out_vs_in.df <- rbind(out_vs_in.df, new_row)
	  }
	}
}

colnames(out_vs_in.df) <- c("neoantigen1", "neoantigen2", "beads")

# Reformat
out_vs_in.df <- out_vs_in.df %>%
	pivot_wider(names_from=neoantigen2,values_from=beads) %>%
	column_to_rownames("neoantigen1") %>%
	as.matrix()

Heatmap(out_vs_in.df,
        col = colors,
        na_col = "grey80",
        cluster_rows = FALSE,
        cluster_columns = FALSE,
        column_title = "Overlap of nearby beads outside TLS vs in TLS",
        name = "No. of\nbeads")
```

### Attach assignments to original Seurat object and filter for defined beads
```{r}
# Attach partition data
partition.df <- data.frame()
for(i in c(1:(length(inTLS_list)))){
  if(nrow(inTLS_list[[i]]$near_beads) > 0){
    near <- inTLS_list[[i]]$near_beads %>%
      select(barcode, partition)
    partition.df <- rbind(partition.df, near)
  }
}
for(i in c(1:(length(outTLS_list)))){
  if(nrow(outTLS_list[[i]]$near_beads) > 0){
    near <- outTLS_list[[i]]$near_beads %>%
      select(barcode, partition)
    partition.df <- rbind(partition.df, near)
  }
}

P108TA_30_RNA@meta.data <- P108TA_30_RNA@meta.data %>%
  left_join(partition.df, by = "barcode") %>%
  column_to_rownames("barcode") %>%
  mutate(barcode = rownames(.))


P108TA_30_RNA@meta.data <- P108TA_30_RNA@meta.data %>%
  mutate(type = case_when(barcode %in% inTLS_list$p108_1$near_beads$barcode ~ "p108_1,inTLS",
                          barcode %in% inTLS_list$p108_205$near_beads$barcode ~ "p108_205,inTLS",
                          barcode %in% inTLS_list$p108_208$near_beads$barcode ~ "p108_208,inTLS",
                          barcode %in% inTLS_list$p108_215$near_beads$barcode ~ "p108_215,inTLS",
                          barcode %in% inTLS_list$p108_266$near_beads$barcode ~ "p108_266,inTLS",
                          barcode %in% outTLS_list$p108_1$near_beads$barcode ~ "p108_1,outTLS",
                          barcode %in% outTLS_list$p108_205$near_beads$barcode ~ "p108_205,outTLS",
                          barcode %in% outTLS_list$p108_208$near_beads$barcode ~ "p108_208,outTLS",
                          barcode %in% outTLS_list$p108_215$near_beads$barcode ~ "p108_215,outTLS",
                          barcode %in% outTLS_list$p108_266$near_beads$barcode ~ "p108_266,outTLS"),
         muscat = case_when(!is.na(type) ~ "Defined",
                             is.na(type) ~ "Undefined"))

P108TA_30_RNA <- subset(P108TA_30_RNA, subset = muscat == "Defined")
```

## Puck 32
### Get beads near each shared neoantigen clone inside the TLS
```{r}
# Get the beads near shared neoantigen clones
inTLS_list <- list()

for(clone in P108TA_shared_clones){
  result_list <- get_beads_near_specific_neoantigen_TCR2(P108TA_32_RNA, "inTLS", clone, rad)
  inTLS_list <- append(inTLS_list, list(result_list))
}

# Rename lists
names(inTLS_list) <- P108TA_shared_clones
```

### Get beads near each shared neoantigen clone outside the TLS
```{r}
# Get the beads near shared neoantigen clones
outTLS_list <- list()

for(clone in P108TA_shared_clones){
  result_list <- get_beads_near_specific_neoantigen_TCR2(P108TA_32_RNA, "outTLS", clone, rad)
  outTLS_list <- append(outTLS_list, list(result_list))
}

# Rename lists
names(outTLS_list) <- P108TA_shared_clones
```

### Check for overlaps
#### in TLS vs in TLS
```{r}
in_vs_in.df <- data.frame()
for(i in 1:length(P108TA_shared_clones)) {
	for(j in 1:length(P108TA_shared_clones)) {
	  if(i < j){
  		same_beads <- length(intersect(inTLS_list[[i]]$near_beads$barcode, inTLS_list[[j]]$near_beads$barcode))
  		new_row <- c(names(inTLS_list)[i], names(inTLS_list)[j], same_beads)
  		in_vs_in.df <- rbind(in_vs_in.df, new_row)
	  }
	}
}

colnames(in_vs_in.df) <- c("neoantigen1", "neoantigen2", "beads")

# Reformat
in_vs_in.df <- in_vs_in.df %>%
	pivot_wider(names_from=neoantigen2,values_from=beads) %>%
	column_to_rownames("neoantigen1") %>%
	as.matrix()

Heatmap(in_vs_in.df,
        col = colors,
        na_col = "grey80",
        cluster_rows = FALSE,
        cluster_columns = FALSE,
        column_title = "Overlap of nearby beads in TLS vs in TLS",
        name = "No. of\nbeads")
```

##### Reassign overlapping beads
```{r}
# P108_208 and P108_266
# Reassign overlapping beads
P108TA_32_reassigned <- reassign_overlapping_beads(P108TA_32_RNA, tex_beads = inTLS_list$p108_266, tmem_beads = inTLS_list$p108_208)

# Split list back out
inTLS_list$p108_266 <- P108TA_32_reassigned$reassigned_tex_beads
inTLS_list$p108_208 <- P108TA_32_reassigned$reassigned_tmem_beads

```

#### Outside TLS vs outside tls
```{r}
out_vs_out.df <- data.frame()
for(i in 1:length(P108TA_shared_clones)) {
	for(j in 1:length(P108TA_shared_clones)) {
	  if(i < j){
  		same_beads <- length(intersect(outTLS_list[[i]]$near_beads$barcode, outTLS_list[[j]]$near_beads$barcode))
  		new_row <- c(names(outTLS_list)[i], names(outTLS_list)[j], same_beads)
  		out_vs_out.df <- rbind(out_vs_out.df, new_row)
	  }
	}
}

colnames(out_vs_out.df) <- c("neoantigen1", "neoantigen2", "beads")

# Reformat
out_vs_out.df <- out_vs_out.df %>%
	pivot_wider(names_from=neoantigen2,values_from=beads) %>%
	column_to_rownames("neoantigen1") %>%
	as.matrix()

Heatmap(out_vs_out.df,
        col = colors,
        na_col = "grey80",
        cluster_rows = FALSE,
        cluster_columns = FALSE,
        column_title = "Overlap of nearby beads outside TLS vs outside TLS",
        name = "No. of\nbeads")

```

##### Reassign overlapping beads
```{r}
# P108_205 and P108_1
# Reassign overlapping beads
P108TA_32_reassigned <- reassign_overlapping_beads(P108TA_32_RNA, tex_beads = outTLS_list$p108_205, tmem_beads = outTLS_list$p108_1)

# Split list back out
outTLS_list$p108_205 <- P108TA_32_reassigned$reassigned_tex_beads
outTLS_list$p108_1 <- P108TA_32_reassigned$reassigned_tmem_beads

# P108_266 and P108_208
# Reassign overlapping beads
P108TA_32_reassigned <- reassign_overlapping_beads(P108TA_32_RNA, tex_beads = outTLS_list$p108_266, tmem_beads = outTLS_list$p108_208)

# Split list back out
outTLS_list$p108_266 <- P108TA_32_reassigned$reassigned_tex_beads
outTLS_list$p108_208 <- P108TA_32_reassigned$reassigned_tmem_beads

```

#### Outside TLS vs in TLS
```{r}
out_vs_in.df <- data.frame()
for(i in 1:length(P108TA_shared_clones)) {
	for(j in 1:length(P108TA_shared_clones)) {
	  if(i <= j){
  		same_beads <- length(intersect(outTLS_list[[i]]$near_beads$barcode, inTLS_list[[j]]$near_beads$barcode))
  		new_row <- c(names(outTLS_list)[i], names(inTLS_list)[j], same_beads)
  		out_vs_in.df <- rbind(out_vs_in.df, new_row)
	  }
	}
}

colnames(out_vs_in.df) <- c("neoantigen1", "neoantigen2", "beads")

# Reformat
out_vs_in.df <- out_vs_in.df %>%
	pivot_wider(names_from=neoantigen2,values_from=beads) %>%
	column_to_rownames("neoantigen1") %>%
	as.matrix()

Heatmap(out_vs_in.df,
        col = colors,
        na_col = "grey80",
        cluster_rows = FALSE,
        cluster_columns = FALSE,
        column_title = "Overlap of nearby beads outside TLS vs in TLS",
        name = "No. of\nbeads")
```

### Attach assignments to original Seurat object and filter for defined beads
```{r}
# Attach partition data
partition.df <- data.frame()
for(i in c(1:(length(inTLS_list)))){
  if(nrow(inTLS_list[[i]]$near_beads) > 0){
    near <- inTLS_list[[i]]$near_beads %>%
      select(barcode, partition)
    partition.df <- rbind(partition.df, near)
  }
}
for(i in c(1:(length(outTLS_list)))){
  if(nrow(outTLS_list[[i]]$near_beads) > 0){
    near <- outTLS_list[[i]]$near_beads %>%
      select(barcode, partition)
    partition.df <- rbind(partition.df, near)
  }
}

P108TA_32_RNA@meta.data <- P108TA_32_RNA@meta.data %>%
  left_join(partition.df, by = "barcode") %>%
  column_to_rownames("barcode") %>%
  mutate(barcode = rownames(.))


P108TA_32_RNA@meta.data <- P108TA_32_RNA@meta.data %>%
  mutate(type = case_when(barcode %in% inTLS_list$p108_1$near_beads$barcode ~ "p108_1,inTLS",
                          barcode %in% inTLS_list$p108_205$near_beads$barcode ~ "p108_205,inTLS",
                          barcode %in% inTLS_list$p108_208$near_beads$barcode ~ "p108_208,inTLS",
                          barcode %in% inTLS_list$p108_215$near_beads$barcode ~ "p108_215,inTLS",
                          barcode %in% inTLS_list$p108_266$near_beads$barcode ~ "p108_266,inTLS",
                          barcode %in% outTLS_list$p108_1$near_beads$barcode ~ "p108_1,outTLS",
                          barcode %in% outTLS_list$p108_205$near_beads$barcode ~ "p108_205,outTLS",
                          barcode %in% outTLS_list$p108_208$near_beads$barcode ~ "p108_208,outTLS",
                          barcode %in% outTLS_list$p108_215$near_beads$barcode ~ "p108_215,outTLS",
                          barcode %in% outTLS_list$p108_266$near_beads$barcode ~ "p108_266,outTLS"),
         muscat = case_when(!is.na(type) ~ "Defined",
                             is.na(type) ~ "Undefined"))

P108TA_32_RNA <- subset(P108TA_32_RNA, subset = muscat == "Defined")
```

# Genes of interest
```{r}
genes_of_int <- c("PDCD1", "ENTPD1", "LAG3", "HAVCR2", "TOX", "TOX2", "IL7R", "SELL", "CCR7", "TCF7", "CXCL13")
```

# Set up RNA objects
## Normalize counts
```{r}
P108TA_26_RNA <- NormalizeData(P108TA_26_RNA, assay = "Spatial")
P108TA_29_RNA <- NormalizeData(P108TA_29_RNA, assay = "Spatial")
P108TA_30_RNA <- NormalizeData(P108TA_30_RNA, assay = "Spatial")
P108TA_32_RNA <- NormalizeData(P108TA_32_RNA, assay = "Spatial")

```

## Update metadata column
```{r}
P108TA_26_RNA@meta.data <- P108TA_26_RNA@meta.data %>%
  mutate(tls_status_gid = str_split_i(type, ",", 2))
P108TA_29_RNA@meta.data <- P108TA_29_RNA@meta.data %>%
  mutate(tls_status_gid = str_split_i(type, ",", 2))
P108TA_30_RNA@meta.data <- P108TA_30_RNA@meta.data %>%
  mutate(tls_status_gid = str_split_i(type, ",", 2))
P108TA_32_RNA@meta.data <- P108TA_32_RNA@meta.data %>%
  mutate(tls_status_gid = str_split_i(type, ",", 2))
```


## Get beads within 15px of a neoantigen specific TCR to share the same "area" annotation
### Puck 26
```{r}
P108TA_26_RNA@meta.data %>%
  dplyr::count(area, tls_status_gid)

boundary_inTLS <- P108TA_26_RNA@meta.data %>%
  filter(area == "Boundary", str_detect(type, "inTLS"))

# Get beads near the boundary beads
near_beads <- nn2(P108TA_26_RNA@meta.data[,c("x", "y")], boundary_inTLS[,c("x", "y")], radius = 15, k = 10, searchtype = "radius")
# Collapse matrix of row indexes into a vector of unique values
near_beads <- unique(as.vector(near_beads$nn.idx))
# Collect metadata of nearby beads
near_beads_md <- P108TA_26_RNA@meta.data[near_beads,]
# Which areas are those beads from?
near_beads_md %>%
  dplyr::count(area)

# Cast beads to TLS3
P108TA_26_RNA@meta.data <- P108TA_26_RNA@meta.data %>%
  mutate(final_area = case_when(area == "Boundary" & str_detect(type, "inTLS") ~ "TLS3",
                              T ~ area))
```

### Puck 29
```{r}
P108TA_29_RNA@meta.data %>%
  dplyr::count(area, tls_status_gid)

# Cast TLS3 beads to boundary
P108TA_29_RNA@meta.data <- P108TA_29_RNA@meta.data %>%
  mutate(final_area = case_when(area == "TLS3" & str_detect(type, "outTLS") ~ "Boundary",
                              T ~ area))
```

### Puck 30
```{r}
P108TA_30_RNA@meta.data %>%
  dplyr::count(area, tls_status_gid)

P108TA_30_RNA@meta.data <- P108TA_30_RNA@meta.data %>%
  mutate(final_area = area)
```

### Puck 32
```{r}
P108TA_32_RNA@meta.data %>%
  dplyr::count(area, tls_status_gid)

boundary_inTLS <- P108TA_32_RNA@meta.data %>%
  filter(area == "Boundary", str_detect(type, "inTLS"))

# Get beads near the boundary beads
near_beads <- nn2(P108TA_32_RNA@meta.data[,c("x", "y")], boundary_inTLS[1,c("x", "y")], radius = 15, k = 10, searchtype = "radius")
# Collapse matrix of row indexes into a vector of unique values
near_beads <- unique(as.vector(near_beads$nn.idx))
# Collect metadata of nearby beads
near_beads_md <- P108TA_32_RNA@meta.data[near_beads,]
# Which areas are those beads from?
near_beads_md %>%
  dplyr::count(area)
# Which bead is this?
boundary_inTLS[1, "barcode"]

near_beads <- nn2(P108TA_32_RNA@meta.data[,c("x", "y")], boundary_inTLS[2,c("x", "y")], radius = 15, k = 10, searchtype = "radius")
# Collapse matrix of row indexes into a vector of unique values
near_beads <- unique(as.vector(near_beads$nn.idx))
# Collect metadata of nearby beads
near_beads_md <- P108TA_32_RNA@meta.data[near_beads,]
# Which areas are those beads from?
near_beads_md %>%
  dplyr::count(area)
# Which bead is this?
boundary_inTLS[2, "barcode"]

# Cast beads to TLS3
P108TA_32_RNA@meta.data <- P108TA_32_RNA@meta.data %>%
  mutate(final_area = case_when(barcode == "GTCATTACCCACTT-1" ~ "TLS1",
                              barcode == "CCGTGGGTTTTGGG-1" ~ "TLS2",
                              T ~ area))

```


# Merge RNA objects
```{r}
TLS_pucks <- merge(x = P108TA_26_RNA, y = list(P108TA_29_RNA, P108TA_30_RNA, P108TA_32_RNA))

# Update metadata
TLS_pucks@meta.data <- TLS_pucks@meta.data %>%
  mutate(tls_status_gid = str_split_i(type, ",", 2))

TLS_pucks[['Spatial']] <- JoinLayers(TLS_pucks[['Spatial']])

TLS_pucks@meta.data %>%
  dplyr::count(tls_status_gid, final_area)

TLS_pucks@meta.data %>%
  dplyr::count(tls_status_gid, orig.ident)
```


# Get normalized counts and attach metadata
```{r}
TLS_md <- TLS_pucks@meta.data %>%
  select(orig.ident, clone_id, barcode, final_area, tls_status, assignment1, neoantigen_id, type, tls_status_gid, nCount_Spatial) %>%
  mutate(final_neoant_id = str_split_i(type, ",", 1))

counts_of_int <- as.data.frame(LayerData(TLS_pucks, assay = "Spatial", layer = "data")[genes_of_int,]) %>%
  t() %>%
  as.data.frame() %>%
  rownames_to_column("barcode")

TLS_md <- full_join(TLS_md, counts_of_int)
```

# Explore
## Which neoant clones are most expanded?
```{r}
TLS_md %>%
  dplyr::count(final_neoant_id, orig.ident) %>%
  dplyr::rename("n_count" = "n") %>%
  pivot_wider(values_from = "n_count", names_from = "orig.ident") %>%
  rowwise() %>%
  mutate(total = rowSums(across(contains("P108")), na.rm = TRUE)) %>%
  arrange(desc(total))
```

## Which neoant clones are most expanded per area?
```{r}
TLS_md %>%
  dplyr::count(final_neoant_id, orig.ident, tls_status) %>%
  dplyr::rename("n_count" = "n") %>%
  pivot_wider(values_from = "n_count", names_from = "orig.ident") %>%
  rowwise() %>%
  mutate(total = rowSums(across(contains("P108")), na.rm = TRUE)) %>%
  arrange(desc(final_neoant_id))
```

## Which neoant clones are most expanded in each TLS?
```{r}
TLS_md %>%
  dplyr::count(final_neoant_id, final_area) %>%
  dplyr::rename("n_count" = "n") %>%
  pivot_wider(values_from = "n_count", names_from = "final_area") %>%
  rowwise() %>%
  mutate(total = rowSums(across(where(is.numeric)), na.rm = TRUE)) %>%
  arrange(desc(total))
```


# Pivot for ggplots
```{r}
TLS_md_long <- TLS_md %>%
  pivot_longer(cols = all_of(genes_of_int), values_to = "counts", names_to = "gene") %>%
mutate(gene = factor(gene, levels = c("TOX", "TOX2", "LAG3", "ENTPD1", "HAVCR2", "PDCD1", "TCF7", "CCR7", "IL7R", "SELL", "CXCL13")))
```

# Plots
## Dotplot split by area and tls status
```{r, fig.width = 6, fig.height = 4}
TLS_nonzero <- TLS_md_long %>%
  filter(counts > 0) %>%
  group_by(tls_status_gid, gene, final_area) %>%
  summarize(n_cells_nonzero = n())

TLS_md_sum <- TLS_md_long %>%
  group_by(tls_status_gid, gene, final_area) %>%
  summarize(mean_counts = mean(counts),
            n_cells = n()) %>%
  full_join(TLS_nonzero, by = c("tls_status_gid", "gene", "final_area")) %>%
  mutate(n_cells_nonzero = replace_na(n_cells_nonzero, 0),
         pct_cells_exp = n_cells_nonzero/n_cells)

TLS_md_sum %>%
  ggplot(aes(x = gene, y = final_area, fill = mean_counts, size = pct_cells_exp))+
  geom_point(shape = 21, color = "black") +
  theme_bw() +
  facet_wrap(~tls_status_gid, scales = "free_y", ncol = 1) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.box = "horizontal", 
        legend.position = "right") +
  scale_fill_gradient2(low = "white", mid = "red", high = "red", midpoint = 0.6) +
  ylab("TLS")

```

## Dotplot split by neoantigen and tls status
```{r, fig.width = 7, fig.height = 4}
TLS_nonzero <- TLS_md_long %>%
  filter(counts > 0) %>%
  group_by(tls_status_gid, gene, final_neoant_id) %>%
  summarize(n_cells_nonzero = n())

TLS_md_sum <- TLS_md_long %>%
  group_by(tls_status_gid, gene, final_neoant_id) %>%
  summarize(mean_counts = mean(counts),
            n_cells = n()) %>%
  full_join(TLS_nonzero, by = c("tls_status_gid", "gene", "final_neoant_id")) %>%
  mutate(n_cells_nonzero = replace_na(n_cells_nonzero, 0),
         pct_cells_exp = n_cells_nonzero/n_cells)

TLS_md_sum %>%
  ggplot(aes(x = gene, y = final_neoant_id, fill = mean_counts, size = pct_cells_exp))+
  geom_point(shape = 21, color = "black") +
  theme_bw() +
  facet_wrap(~tls_status_gid) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_gradient2(low = "white", mid = "red", high = "red", midpoint = 0.9) +
  ylab("Neoantigen-specific TCR")

```

# Get session info
```{r}
sessionInfo()
```
