# Gene expression of external macrophage gene sets {#sec-gene-exp-plots}

# Plot the gene expression of external macrophage gene sets in spatial pucks containing a myeloid boundary across all areas

## Libraries
```{r}
library(harmony)
library(Seurat)
library(paletteer)
library(liana)
library(tidyverse)
library(ggpointdensity)
library(rmarkdown)
library(knitr)
library(ComplexHeatmap)
library(clustree)
library(ggridges)

source("../source_scripts/slideseq_functions_main.R")
```


# Process spatial data
## Open objects with annotated areas
```{r}
P102TA_10_RNA <- readRDS("../objects/02_Puck10_102TA_TCR_RNA_area_121125.Rds")
P102TA_17_RNA <- readRDS("../objects/02_Puck17_102TA_TCR_RNA_area_121125.Rds")
P102TA_33_RNA <- readRDS("../objects/02_Puck33_102TA_TCR_RNA_area_121125.Rds")

P108TA_26_RNA <- readRDS("../objects/02_Puck26_108TA_TCR_RNA_area_121125.Rds")
P108TA_29_RNA <- readRDS("../objects/02_Puck29_108TA_TCR_RNA_area_121125.Rds")
P108TA_30_RNA <- readRDS("../objects/02_Puck30_108TA_TCR_RNA_area_121125.Rds")
P108TA_32_RNA <- readRDS("../objects/02_Puck32_108TA_TCR_RNA_area_121125.Rds")
```

## Remove empty areas
```{r}
P102TA_10_RNA <- subset(P102TA_10_RNA, area == "Empty", invert = TRUE)
P102TA_17_RNA <- subset(P102TA_17_RNA, area == "Empty", invert = TRUE)
P102TA_33_RNA <- subset(P102TA_33_RNA, area == "Empty", invert = TRUE)

P108TA_26_RNA <- subset(P108TA_26_RNA, area == "Empty", invert = TRUE)
P108TA_29_RNA <- subset(P108TA_29_RNA, area == "Empty", invert = TRUE)
P108TA_30_RNA <- subset(P108TA_30_RNA, area == "Empty", invert = TRUE)
P108TA_32_RNA <- subset(P108TA_32_RNA, area == "Empty", invert = TRUE)

```

## Normalize data
```{r}
P102TA_10_RNA <- NormalizeData(P102TA_10_RNA, assay = "Spatial")
P102TA_17_RNA <- NormalizeData(P102TA_17_RNA, assay = "Spatial")
P102TA_33_RNA <- NormalizeData(P102TA_33_RNA, assay = "Spatial")

P108TA_26_RNA <- NormalizeData(P108TA_26_RNA, assay = "Spatial")
P108TA_29_RNA <- NormalizeData(P108TA_29_RNA, assay = "Spatial")
P108TA_30_RNA <- NormalizeData(P108TA_30_RNA, assay = "Spatial")
P108TA_32_RNA <- NormalizeData(P108TA_32_RNA, assay = "Spatial")
```


## Load gene sets
```{r} 
genesets <- readRDS("../objects/guimaraes_gene_sets.Rds")
ma_gs <- readRDS("../objects/ma_gene_sets.Rds")
```

# Signature score 
## Score all beads
```{r}
add_signature_score <- function(puck){
  puck <- puck %>%
    AddModuleScore(features = list(ma_gs$`Angio-TAMs`), name = "Ma_Angio-TAM") %>%
    AddModuleScore(features = list(ma_gs$`Classical TIMs`), name = "Ma_Class-TIM") %>%
    AddModuleScore(features = list(ma_gs$`IFN-TAMs`), name = "Ma_IFN-TAM") %>%
    AddModuleScore(features = list(ma_gs$`Inflam-TAMs`), name = "Ma_Inflam-TAM") %>%
    AddModuleScore(features = list(ma_gs$`Intermediate monocytes`), name = "Intermed_mono") %>%
    AddModuleScore(features = list(ma_gs$`LA-TAMs`), name = "Ma_LA-TAM") %>%
    AddModuleScore(features = list(ma_gs$`Nonclassical monocytes`), name = "Ma_Nonclass_mono") %>%
    AddModuleScore(features = list(ma_gs$`Prolif-TAMs`), name = "Ma_Prolif-TAM") %>%
    AddModuleScore(features = list(ma_gs$`Reg-TAMs`), name = "Ma_Reg-TAM") %>%
    AddModuleScore(features = list(ma_gs$`RTM-TAMs_Alveolar-RTM-like`), name = "Ma_RTM_Alv")%>%
    AddModuleScore(features = list(ma_gs$`RTM-TAMs_Interstitial-RTM-like`), name = "Ma_RTM_Int") %>%
    AddModuleScore(features = list(ma_gs$`RTM-TAMs_Kupffer-like`), name = "Ma_RTM_Kupf") %>%
    AddModuleScore(features = list(ma_gs$`RTM-TAMs_Microglia-like`), name = "Ma_RTM_Micro") %>%
    AddModuleScore(features = list(ma_gs$`RTM-TAMs_MT-RTM-like`), name = "Ma_RTM_MT")
  return(puck)
}

P102TA_10_RNA <- P102TA_10_RNA %>% add_signature_score
P102TA_17_RNA <- P102TA_17_RNA %>% add_signature_score
P102TA_33_RNA <- P102TA_33_RNA %>% add_signature_score
P108TA_26_RNA <- P108TA_26_RNA %>% add_signature_score
P108TA_29_RNA <- P108TA_29_RNA %>% add_signature_score
P108TA_30_RNA <- P108TA_30_RNA %>% add_signature_score
P108TA_32_RNA <- P108TA_32_RNA %>% add_signature_score

```

## Spatial plot of all beads
```{r}
SpatialFeaturePlot(P108TA_26_RNA, features = c("Ma_Inflam-TAM1"), alpha = c(0, 1))
SpatialFeaturePlot(P108TA_26_RNA, features = c("Ma_Angio-TAM1"), alpha = c(0, 1))
SpatialFeaturePlot(P108TA_26_RNA, features = c("Ma_LA-TAM1"), alpha = c(0, 1))
SpatialFeaturePlot(P108TA_26_RNA, features = c("Ma_IFN-TAM1"), alpha = c(0, 1))
SpatialFeaturePlot(P108TA_26_RNA, features = c("Ma_Reg-TAM1"), alpha = c(0, 1))
```

# Cast TLS to a singular value
```{r}
format_md <- function(obj){
  obj <- obj %>%
    mutate(final_area = case_when(area %in% c("TLS1", "TLS2", "TLS3", "TLS4", "M2") ~ "TLS",
                                T ~ area),
         Patient = str_sub(orig.ident, start = 1, end = 5),
         tls_bool = case_when(final_area == "TLS" ~ "TLS",
                              T ~ "nonTLS"),
         boundary_bool = case_when(final_area == "Boundary" ~ "Boundary",
                                   T ~ "nonBoundary"),
         tumor_bool = case_when(final_area == "Tumor" ~ "Tumor",
                                   T ~ "nonTumor"),
         Patient_tls = paste0(Patient, "_", tls_bool),
         Patient_boundary = paste0(Patient, "_", boundary_bool),
         Patient_tumor = paste0(Patient, "_", tumor_bool),
         dummy = "dummy")
  
  return(obj)
}

P102TA_10_RNA@meta.data <-  P102TA_10_RNA@meta.data %>%
  format_md()
P102TA_17_RNA@meta.data <-  P102TA_17_RNA@meta.data %>%
  format_md()
P102TA_33_RNA@meta.data <-  P102TA_33_RNA@meta.data %>%
  format_md()

P108TA_26_RNA@meta.data <-  P108TA_26_RNA@meta.data %>%
  format_md()
P108TA_29_RNA@meta.data <-  P108TA_29_RNA@meta.data %>%
  format_md()
P108TA_30_RNA@meta.data <-  P108TA_30_RNA@meta.data %>%
  format_md()
P108TA_32_RNA@meta.data <-  P108TA_32_RNA@meta.data %>%
  format_md()

```

# Create dotplot of genes 

# Make a dotplot of the top macrophage subtypes in spatial for myeloid beads
## Subset all pucks for myeloid cells
```{r}
get_mac_beads <- function(puck){
  puck@meta.data <- puck@meta.data %>% assign_celltypes
  puck <- subset(puck, subset = assignment1 == "Myeloid")
  return(puck)
}

P102TA_10_RNA <- P102TA_10_RNA %>% get_mac_beads
P102TA_17_RNA <- P102TA_17_RNA %>% get_mac_beads
P102TA_33_RNA <- P102TA_33_RNA %>% get_mac_beads
P108TA_26_RNA <- P108TA_26_RNA %>% get_mac_beads
P108TA_29_RNA <- P108TA_29_RNA %>% get_mac_beads
P108TA_30_RNA <- P108TA_30_RNA %>% get_mac_beads
P108TA_32_RNA <- P108TA_32_RNA %>% get_mac_beads
```

## Spatial plot of myeloid beads
```{r}
SpatialFeaturePlot(P108TA_26_RNA, features = c("Ma_Inflam-TAM1"), alpha = c(0, 1))
SpatialFeaturePlot(P108TA_26_RNA, features = c("Ma_Angio-TAM1"), alpha = c(0, 1))
SpatialFeaturePlot(P108TA_26_RNA, features = c("Ma_LA-TAM1"), alpha = c(0, 1))
SpatialFeaturePlot(P108TA_26_RNA, features = c("Ma_IFN-TAM1"), alpha = c(0, 1))
SpatialFeaturePlot(P108TA_26_RNA, features = c("Ma_Reg-TAM1"), alpha = c(0, 1))
```

# Angio and LA gene set
## Get counts of interest
```{r}
ma.subset <- ma_gs[str_detect(names(ma_gs), "Angio|LA")]

get_counts_of_int <- function(puck, genes){
  counts_of_int <- as.data.frame(LayerData(puck, assay = "Spatial", layer = "data", features = unique(unlist(ma.subset)))) %>%
    t() %>%
    as.data.frame() %>%
    rownames_to_column("barcode") %>%
    mutate(orig.ident = unique(Idents(puck)))
  return(counts_of_int)
}

P102TA_10_counts <- P102TA_10_RNA %>% get_counts_of_int
P102TA_17_counts <- P102TA_17_RNA %>% get_counts_of_int
P102TA_33_counts <- P102TA_33_RNA %>% get_counts_of_int
P108TA_26_counts <- P108TA_26_RNA %>% get_counts_of_int
P108TA_29_counts <- P108TA_29_RNA %>% get_counts_of_int
P108TA_30_counts <- P108TA_30_RNA %>% get_counts_of_int
P108TA_32_counts <- P108TA_32_RNA %>% get_counts_of_int

counts_of_int <- P102TA_10_counts %>%
  full_join(P102TA_17_counts)  %>%
  full_join(P102TA_33_counts)  %>%
  full_join(P108TA_26_counts)  %>%
  full_join(P108TA_29_counts)  %>%
  full_join(P108TA_30_counts)  %>%
  full_join(P108TA_32_counts) 
```

# Combine data into one dataframe, noting pucks in column
```{r}
# Remove metadata columns that other pucks don't have
P102TA_17_RNA@meta.data <- P102TA_17_RNA@meta.data %>%
  select(-largest_clone_id, -puck17_count, -log_puck17_count, -Fibroblast)
P102TA_10_RNA@meta.data <- P102TA_10_RNA@meta.data %>%
  select(-largest_clone_id, -puck10_count, -log_puck10_count, -Fibroblast)
P102TA_33_RNA@meta.data <- P102TA_33_RNA@meta.data %>%
  select(-Fibroblast)

P108TA_26_RNA@meta.data <- P108TA_26_RNA@meta.data %>%
  select(-Fibroblast)
P108TA_29_RNA@meta.data <- P108TA_29_RNA@meta.data %>%
  select(-Fibroblast)
P108TA_30_RNA@meta.data <- P108TA_30_RNA@meta.data %>%
  select(-Fibroblast)
P108TA_32_RNA@meta.data <- P108TA_32_RNA@meta.data %>%
  select(-Fibroblast)

puck_md.df <- do.call(rbind, list(P102TA_10_RNA@meta.data, P102TA_17_RNA@meta.data, P102TA_33_RNA@meta.data,
                                 P108TA_26_RNA@meta.data, P108TA_29_RNA@meta.data, P108TA_30_RNA@meta.data, P108TA_32_RNA@meta.data)) %>%
  select(barcode, orig.ident, final_area)

puck_md.df <- full_join(puck_md.df, counts_of_int, by = c("orig.ident", "barcode"))
```

## Pivot for ggplots
```{r}
md.df_long <- puck_md.df %>%
  pivot_longer(cols = !c("barcode", "orig.ident", "final_area"), values_to = "data", names_to = "gene")
```

## Gene annotation df
```{r}
gene_split <- stack(ma.subset)

names(gene_split) <- c("gene", "Annotation")

# Some genes appear in multiple annotations, just take the first instance.
gene_split <- gene_split %>%
  distinct(gene, .keep_all = TRUE)
```

## Plots
### Dotplot split by cluster
```{r, fig.width = 28, fig.height = 6}
nonzero <- md.df_long %>%
  filter(data > 0) %>%
  group_by(orig.ident, gene, final_area) %>%
  summarize(n_cells_nonzero = n())

md_sum <- md.df_long %>%
  group_by(orig.ident, gene, final_area) %>%
  summarize(mean_data = mean(data),
            n_cells = n()) %>%
  full_join(nonzero, by = c("orig.ident", "gene", "final_area")) %>%
  mutate(n_cells_nonzero = replace_na(n_cells_nonzero, 0),
         pct_cells_exp = n_cells_nonzero/n_cells) %>%
  left_join(gene_split, by = "gene")

plot <- md_sum %>%
  ggplot(aes(x = gene, y = orig.ident, fill = mean_data, size = pct_cells_exp))+
  geom_point(shape = 21, color = "black") +
  theme_bw() +
  facet_wrap(Annotation~final_area, scales = "free_x", ncol = 3) +
  scale_fill_gradient(low = "white", high = "red") +
  # scale_fill_gradient2(low = "white", mid = "red", high = "red", midpoint = 4) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.box = "horizontal", 
        legend.position = "right") +
  labs(size = "Pct Cells Expr",
       fill = "Mean norm. counts")

# pdf("../main_myeloid_per_area_highest_ma_genes_dotplot.pdf", width = 28, height = 6)
plot
# dev.off()

```

# IFN and Reg gene sets
## Get counts of interest
```{r}
ma.subset <- ma_gs[str_detect(names(ma_gs), "IFN|Reg")]

get_counts_of_int <- function(puck, genes){
  counts_of_int <- as.data.frame(LayerData(puck, assay = "Spatial", layer = "data", features = unique(unlist(ma.subset)))) %>%
    t() %>%
    as.data.frame() %>%
    rownames_to_column("barcode") %>%
    mutate(orig.ident = unique(Idents(puck)))
  return(counts_of_int)
}

P102TA_10_counts <- P102TA_10_RNA %>% get_counts_of_int
P102TA_17_counts <- P102TA_17_RNA %>% get_counts_of_int
P102TA_33_counts <- P102TA_33_RNA %>% get_counts_of_int
P108TA_26_counts <- P108TA_26_RNA %>% get_counts_of_int
P108TA_29_counts <- P108TA_29_RNA %>% get_counts_of_int
P108TA_30_counts <- P108TA_30_RNA %>% get_counts_of_int
P108TA_32_counts <- P108TA_32_RNA %>% get_counts_of_int

counts_of_int <- P102TA_10_counts %>%
  full_join(P102TA_17_counts)  %>%
  full_join(P102TA_33_counts)  %>%
  full_join(P108TA_26_counts)  %>%
  full_join(P108TA_29_counts)  %>%
  full_join(P108TA_30_counts)  %>%
  full_join(P108TA_32_counts) 
```

# Combine data into one dataframe, noting pucks in column
```{r}
puck_md.df <- do.call(rbind, list(P102TA_10_RNA@meta.data, P102TA_17_RNA@meta.data, P102TA_33_RNA@meta.data,
                                 P108TA_26_RNA@meta.data, P108TA_29_RNA@meta.data, P108TA_30_RNA@meta.data, P108TA_32_RNA@meta.data)) %>%
  select(barcode, orig.ident, final_area)

puck_md.df <- full_join(puck_md.df, counts_of_int, by = c("orig.ident", "barcode"))
```

## Pivot for ggplots
```{r}
md.df_long <- puck_md.df %>%
  pivot_longer(cols = !c("barcode", "orig.ident", "final_area"), values_to = "data", names_to = "gene")
```

## Gene annotation df
```{r}
gene_split <- stack(ma.subset)

names(gene_split) <- c("gene", "Annotation")

# Some genes appear in multiple annotations, just take the first instance.
gene_split <- gene_split %>%
  distinct(gene, .keep_all = TRUE)
```

## Plots
### Dotplot split by cluster
```{r, fig.width = 28, fig.height = 10}
nonzero <- md.df_long %>%
  filter(data > 0) %>%
  group_by(orig.ident, gene, final_area) %>%
  summarize(n_cells_nonzero = n())

md_sum <- md.df_long %>%
  group_by(orig.ident, gene, final_area) %>%
  summarize(mean_data = mean(data),
            n_cells = n()) %>%
  full_join(nonzero, by = c("orig.ident", "gene", "final_area")) %>%
  mutate(n_cells_nonzero = replace_na(n_cells_nonzero, 0),
         pct_cells_exp = n_cells_nonzero/n_cells) %>%
  left_join(gene_split, by = "gene")

plot <- md_sum %>%
  ggplot(aes(x = gene, y = orig.ident, fill = mean_data, size = pct_cells_exp))+
  geom_point(shape = 21, color = "black") +
  theme_bw() +
  facet_wrap(Annotation~final_area, scales = "free_x", ncol = 3) +
  scale_fill_gradient(low = "white", high = "red") +
  # scale_fill_gradient2(low = "white", mid = "red", high = "red", midpoint = 4) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.box = "horizontal", 
        legend.position = "right") +
  labs(size = "Pct Cells Expr",
       fill = "Mean norm. counts")

# pdf("../supp_myeloid_per_area_highest_ma_genes_dotplot.pdf", width = 31, height = 6)
plot
# dev.off()

```

# Get session info
```{r}
sessionInfo()
```


