# Variable and constant TCR gene plots {#sec-gene-exp-plots}

# Libraries
```{r}
library(Seurat)
library(dplyr)
library(ggpointdensity)
library(tidyverse)
library(ComplexHeatmap)
library(stringr)
library(readxl)
library(RANN)
library(ggvenn)
library(eulerr)
library(stringr)
library(plotrix)
library(patchwork)

source("../source_scripts/slideseq_functions_main.R")

options(Seurat.object.assay.version = 'v5')

tls <- c("TLS1", "TLS2", "TLS3", "TLS4", "M2")
```

# Open objects with annotated areas
```{r}
P102TA_10_RNA <- readRDS("../objects/02_Puck10_102TA_TCR_RNA_area_121125.Rds")
P102TA_17_RNA <- readRDS("../objects/02_Puck17_102TA_TCR_RNA_area_121125.Rds")
P102TA_33_RNA <- readRDS("../objects/02_Puck33_102TA_TCR_RNA_area_121125.Rds")
P102TA_39_RNA <- readRDS("../objects/02_Puck39_102TA_TCR_RNA_area_121125.Rds")

P107TA_15_RNA <- readRDS("../objects/02_Puck15_107TA_TCR_RNA_area_121125.Rds")
P107TA_18_RNA <- readRDS("../objects/02_Puck18_107TA_TCR_RNA_area_121125.Rds")
P107TA_20_RNA <- readRDS("../objects/02_Puck20_107TA_TCR_RNA_area_121125.Rds")

P108TA_26_RNA <- readRDS("../objects/02_Puck26_108TA_TCR_RNA_area_121125.Rds")
P108TA_29_RNA <- readRDS("../objects/02_Puck29_108TA_TCR_RNA_area_121125.Rds")
P108TA_30_RNA <- readRDS("../objects/02_Puck30_108TA_TCR_RNA_area_121125.Rds")
P108TA_32_RNA <- readRDS("../objects/02_Puck32_108TA_TCR_RNA_area_121125.Rds")

P108TB_06_RNA <- readRDS("../objects/02_Puck06_108TB_TCR_RNA_area_121125.Rds")
P108TB_14_RNA <- readRDS("../objects/02_Puck14_108TB_TCR_RNA_area_121125.Rds")

P109TB_23_RNA <- readRDS("../objects/02_Puck23_109TB_TCR_RNA_area_121125.Rds")
P109TB_34_RNA <- readRDS("../objects/02_Puck34_109TB_TCR_RNA_area_121125.Rds")

P110TA_28_RNA <- readRDS("../objects/02_Puck28_110TA_TCR_RNA_area_121125.Rds")
P110TA_30_RNA <- readRDS("../objects/02_Puck30_110TA_TCR_RNA_area_121125.Rds")

P111TA_32_RNA <- readRDS("../objects/02_Puck32_111TA_TCR_RNA_area_121125.Rds")
P111TA_37_RNA <- readRDS("../objects/02_Puck37_111TA_TCR_RNA_area_121125.Rds")
```


## Remove empty areas
```{r}
P102TA_10_RNA <- subset(P102TA_10_RNA, area == "Empty", invert = TRUE)
P102TA_17_RNA <- subset(P102TA_17_RNA, area == "Empty", invert = TRUE)
P102TA_33_RNA <- subset(P102TA_33_RNA, area == "Empty", invert = TRUE)
P102TA_39_RNA <- subset(P102TA_39_RNA, area == "Empty", invert = TRUE)

P107TA_15_RNA <- subset(P107TA_15_RNA, area == "Empty", invert = TRUE)
P107TA_18_RNA <- subset(P107TA_18_RNA, area == "Empty", invert = TRUE)
P107TA_20_RNA <- subset(P107TA_20_RNA, area == "Empty", invert = TRUE)

P108TA_26_RNA <- subset(P108TA_26_RNA, area == "Empty", invert = TRUE)
P108TA_29_RNA <- subset(P108TA_29_RNA, area == "Empty", invert = TRUE)
P108TA_30_RNA <- subset(P108TA_30_RNA, area == "Empty", invert = TRUE)
P108TA_32_RNA <- subset(P108TA_32_RNA, area == "Empty", invert = TRUE)

P108TB_06_RNA <- subset(P108TB_06_RNA, area == "Empty", invert = TRUE)
P108TB_14_RNA <- subset(P108TB_14_RNA, area == "Empty", invert = TRUE)

P109TB_23_RNA <- subset(P109TB_23_RNA, area == "Empty", invert = TRUE)
P109TB_34_RNA <- subset(P109TB_34_RNA, area == "Empty", invert = TRUE)

P110TA_28_RNA <- subset(P110TA_28_RNA, area == "Empty", invert = TRUE)
P110TA_30_RNA <- subset(P110TA_30_RNA, area == "Empty", invert = TRUE)

P111TA_32_RNA <- subset(P111TA_32_RNA, area == "Empty", invert = TRUE)
P111TA_37_RNA <- subset(P111TA_37_RNA, area == "Empty", invert = TRUE)
```

# Venn diagram functions
```{r}
compare_tra_constant_variable <- function(obj, puck_name){
  constant <- Cells(subset(x = obj, subset = TRAC > 1))

  variable <- obj@meta.data %>%
    filter(str_detect(chains, "TRA")) %>%
    pull(barcode)

  plot(euler(list(Constant_TRAC = constant, Variable = variable)),
       main = puck_name, 
       quantities = TRUE)
}

compare_trbc1_constant_variable <- function(obj, puck_name){
  constant <- Cells(subset(x = obj, subset = TRBC1 > 1))

  variable <- obj@meta.data %>%
    filter(str_detect(chains, "TRB")) %>%
    pull(barcode)

  plot(euler(list(Constant_TRBC1 = constant, Variable = variable)),
       main = puck_name, 
       quantities = TRUE)
}

compare_trbc2_constant_variable <- function(obj, puck_name){
  constant <- Cells(subset(x = obj, subset = TRBC2 > 1))

  variable <- obj@meta.data %>%
    filter(str_detect(chains, "TRB")) %>%
    pull(barcode)

  plot(euler(list(Constant_TRBC2 = constant, Variable = variable)),
       main = puck_name, 
       quantities = TRUE)
}

compare_troc_constant_variable <- function(obj, puck_name){
  constant <- Cells(subset(x = obj, subset = TRBC1 > 1 | TRAC > 1))

  variable <- obj@meta.data %>%
    filter(!is.na(all_CDR3)) %>%
    pull(barcode)

  plot(euler(list(Constant_TRBC = constant, Variable = variable)),
       main = puck_name, 
       quantities = TRUE)
}

```

# Fraction of beads with variable TCRs functions
```{r}
add_trac <- function(obj){
  trc_gene_exp <- FetchData(object = obj, vars = c("TRAC"), layer = "counts")
  obj <- AddMetaData(obj, trc_gene_exp)
  return(obj@meta.data)
}

calculate_pct_tra_beads_with_tcr <- function(obj){
  md <- add_trac(obj)
  
  tra_df <- data.frame(matrix(nrow = 0, ncol = 6, dimnames = list(rownames = NULL, colnames = c("orig.ident", "n_total_beads", "n_umi", "n_tra_beads", "n_tra_with_tcr_beads", "pct_tra_with_tcr"))))

  expr <- unique(md$TRAC)
  for (i in c(expr)){
    puck <- unique(md$orig.ident)
    num_total_beads <- md %>% nrow()
    # Number of beads with N TRAC counts
    num_beads_tra <- md %>% filter(TRAC == i) %>% nrow()
    # Number of beads with N TRAC counts and a variable TCRA
    num_beads_tra_tcr <- md %>% filter(TRAC == i, str_detect(chains, "TRA")) %>% nrow()
    
    pct_tra_with_tcr <- (num_beads_tra_tcr/num_beads_tra)*100
    
    dat <- data.frame(orig.ident = puck, n_total_beads = num_total_beads, n_umi = i, n_tra_beads = num_beads_tra, n_tra_with_tcr_beads = num_beads_tra_tcr, pct_tra_with_tcr = pct_tra_with_tcr)
    
    tra_df <- rbind(tra_df, dat)
  }
  
  return(tra_df)
}

add_trbc <- function(obj){
  trc_gene_exp <- FetchData(object = obj, vars = c("TRBC2", "TRBC1"), layer = "counts")
  obj <- AddMetaData(obj, trc_gene_exp)
  
  return(obj@meta.data)
}

calculate_pct_trb_beads_with_tcr <- function(obj){
  md <- add_trbc(obj)
  
  md <- md %>%
    rowwise() %>%
    mutate(TRBC = sum(TRBC1, TRBC2)) %>%
    ungroup()
  
  trb_df <- data.frame(matrix(nrow = 0, ncol = 6, dimnames = list(rownames = NULL, colnames = c("orig.ident", "n_total_beads", "n_umi", "n_trb_beads", "n_trb_with_tcr_beads", "pct_trb_with_tcr"))))

  expr <- unique(md$TRBC)
  for (i in c(expr)){
    puck <- unique(md$orig.ident)
    num_total_beads <- md %>% nrow()
    num_beads_trb <- md %>% filter(TRBC == i) %>% nrow()
    num_beads_trb_tcr <- md %>% filter(TRBC == i, !is.na(all_CDR3)) %>% nrow()
    
    pct_trb_with_tcr <- (num_beads_trb_tcr/num_beads_trb)*100
    
    dat <- data.frame(orig.ident = puck, n_total_beads = num_total_beads, n_umi = i, n_trb_beads = num_beads_trb, n_trb_with_tcr_beads = num_beads_trb_tcr, pct_trb_with_tcr = pct_trb_with_tcr)
    
    trb_df <- rbind(trb_df, dat)
  }
  
  return(trb_df)
}


add_trc_exp <- function(obj){
  trc_gene_exp <- FetchData(object = obj, vars = c("TRAC", "TRBC1", "TRBC2"), layer = "counts")
  obj <- AddMetaData(obj, trc_gene_exp)
  obj@meta.data <- obj@meta.data %>%
    rowwise() %>%
    mutate(TRBC = sum(TRBC1, TRBC2))
  
  return(obj)
}


theme_blank <- theme(axis.line=element_blank(),
      axis.text.x=element_blank(),
      axis.text.y=element_blank(),
      axis.ticks=element_blank(),
      axis.title.x=element_blank(),
      axis.title.y=element_blank(),
      legend.position="none",
      panel.background=element_blank(),
      panel.border=element_blank(),
      panel.grid.major=element_blank(),
      panel.grid.minor=element_blank(),
      plot.background=element_blank())

density_fill_14 <- scale_fill_manual(values = c("#4B0055FF", "#44286EFF", "#204C85FF", "#007696FF",  "#009A95FF", "#00B983FF", "#14C673FF", "#6CD05EFF", "#9AD849FF", "#C2DE34FF", "#D4E12DFF", "#ECE32CFF", "#f2ed92", "#fdfdf5"))

density_fill_13 <- scale_fill_manual(values = c("#4B0055FF", "#44286EFF", "#204C85FF", "#007696FF",  "#009A95FF", "#00B983FF", "#14C673FF", "#6CD05EFF", "#9AD849FF", "#D4E12DFF", "#ECE32CFF", "#f2ed92", "#fdfdf5"))

density_fill_12 <- scale_fill_manual(values = c("#4B0055FF", "#44286EFF", "#204C85FF", "#007696FF",  "#009A95FF", "#00B983FF", "#14C673FF", "#9AD849FF", "#D4E12DFF", "#ECE32CFF", "#f2ed92", "#fdfdf5")) 

density_fill_11 <- scale_fill_manual(values = c("#4B0055FF", "#44286EFF", "#204C85FF", "#007696FF",  "#00B983FF", "#14C673FF", "#9AD849FF", "#D4E12DFF", "#ECE32CFF", "#f2ed92", "#fdfdf5")) 

density_fill_10 <- scale_fill_manual(values = c("#4B0055FF",  "#30437FFF", "#007696FF",  "#00B983FF", "#14C673FF", "#9AD849FF", "#D4E12DFF", "#ECE32CFF", "#f2ed92", "#fdfdf5")) 

density_fill_9 <- scale_fill_manual(values = c("#4B0055FF", "#204C85FF", "#008398FF",  "#00B587FF", "#9AD849FF", "#D4E12DFF", "#ECE32CFF", "#f2ed92", "#fdfdf5")) 

density_fill_8 <- scale_fill_manual(values = c("#4B0055FF", "#204C85FF",  "#008197FF", "#7CD357FF", "#D4E12DFF", "#ECE32CFF", "#f2ed92", "#fdfdf5")) 

density_fill_7 <- scale_fill_manual(values = c("#4B0055FF", "#204C85FF",  "#008197FF", "#7CD357FF", "#D4E12DFF", "#f2ed92", "#fdfdf5")) 

density_fill_6 <- scale_fill_manual(values = c("#4B0055FF", "#005F8EFF", "#7CD357FF", "#D4E12DFF", "#f2ed92", "#fdfdf5")) 
```


## Normalize data
```{r}
P102TA_10_RNA <- NormalizeData(P102TA_10_RNA)
P102TA_17_RNA <- NormalizeData(P102TA_17_RNA)
P102TA_33_RNA <- NormalizeData(P102TA_33_RNA)
P102TA_39_RNA <- NormalizeData(P102TA_39_RNA)

P107TA_15_RNA <- NormalizeData(P107TA_15_RNA)
P107TA_18_RNA <- NormalizeData(P107TA_18_RNA)
P107TA_20_RNA <- NormalizeData(P107TA_20_RNA)

P108TA_26_RNA <- NormalizeData(P108TA_26_RNA)
P108TA_29_RNA <- NormalizeData(P108TA_29_RNA)
P108TA_30_RNA <- NormalizeData(P108TA_30_RNA)
P108TA_32_RNA <- NormalizeData(P108TA_32_RNA)

P108TB_06_RNA <- NormalizeData(P108TB_06_RNA)
P108TB_14_RNA <- NormalizeData(P108TB_14_RNA)

P109TB_23_RNA <- NormalizeData(P109TB_23_RNA)
P109TB_34_RNA <- NormalizeData(P109TB_34_RNA)

P110TA_28_RNA <- NormalizeData(P110TA_28_RNA)
P110TA_30_RNA <- NormalizeData(P110TA_30_RNA)

P111TA_32_RNA <- NormalizeData(P111TA_32_RNA)
P111TA_37_RNA <- NormalizeData(P111TA_37_RNA)
```

# Add TRA and TRB counts for supplementary table
```{r}
pucks <- merge(x = P102TA_10_RNA, y = list(P102TA_17_RNA, P102TA_33_RNA, P102TA_39_RNA, 
                                           P107TA_15_RNA, P107TA_18_RNA, P107TA_20_RNA, 
                                           P108TA_26_RNA, P108TA_29_RNA, P108TA_30_RNA, P108TA_32_RNA, 
                                           P108TB_06_RNA, P108TB_14_RNA,
                                           P109TB_23_RNA, P109TB_34_RNA,
                                           P110TA_28_RNA, P110TA_30_RNA,
                                           P111TA_32_RNA, P111TA_37_RNA))

tra_md <- add_trac(pucks) %>%
  select(orig.ident, barcode, TRAC)

trb_md <- add_trbc(pucks) %>%
  select(orig.ident, barcode, TRBC1, TRBC2, chains) %>%
  dplyr::rename("variable_chains" = "chains")

saved_md <- full_join(tra_md, trb_md, by = c("orig.ident", "barcode"))

```

# Calculate fraction of TRA beads with TCR
```{r}
pct_tra <- rbind(calculate_pct_tra_beads_with_tcr(P102TA_10_RNA),
              calculate_pct_tra_beads_with_tcr(P102TA_17_RNA),
              calculate_pct_tra_beads_with_tcr(P102TA_33_RNA),
              calculate_pct_tra_beads_with_tcr(P102TA_39_RNA),
              calculate_pct_tra_beads_with_tcr(P107TA_15_RNA),
              calculate_pct_tra_beads_with_tcr(P107TA_18_RNA),
              calculate_pct_tra_beads_with_tcr(P107TA_20_RNA),
              calculate_pct_tra_beads_with_tcr(P108TA_26_RNA),
              calculate_pct_tra_beads_with_tcr(P108TA_29_RNA),
              calculate_pct_tra_beads_with_tcr(P108TA_30_RNA),
              calculate_pct_tra_beads_with_tcr(P108TA_32_RNA),
              calculate_pct_tra_beads_with_tcr(P108TB_06_RNA),
              calculate_pct_tra_beads_with_tcr(P108TB_14_RNA),
              calculate_pct_tra_beads_with_tcr(P109TB_23_RNA),
              calculate_pct_tra_beads_with_tcr(P109TB_34_RNA),
              calculate_pct_tra_beads_with_tcr(P110TA_28_RNA),
              calculate_pct_tra_beads_with_tcr(P110TA_30_RNA),
              calculate_pct_tra_beads_with_tcr(P111TA_32_RNA),
              calculate_pct_tra_beads_with_tcr(P111TA_37_RNA)) %>%
  mutate(Patient = str_split_i(orig.ident, "_", 1))
```

## Plot after filtering each UMI bin for 3 beads or more
```{r}
pct_tra_sum <- pct_tra %>%
  filter(n_tra_with_tcr_beads > 3,
         n_umi != 0) %>%
  group_by(n_umi) %>%
  summarize(sum_n_tra_beads = sum(n_tra_beads),
            sum_n_tra_beads_with_tcr = sum(n_tra_with_tcr_beads),
            mean_pct_tra_with_tcr = mean(pct_tra_with_tcr),
            std = sd(pct_tra_with_tcr),
            serror = std.error(pct_tra_with_tcr))


tra_plot <- ggplot(pct_tra_sum, aes(x = as.character(n_umi), y = mean_pct_tra_with_tcr)) +
  geom_col(fill = "#B0CBE7FF") +
  geom_text(data = pct_tra_sum, 
              aes(x = as.character(n_umi), 
                  y = 80,
                  angle = 90,
                  label = sum_n_tra_beads),
              colour="black", 
              size=3.5) +
  geom_errorbar(aes(ymin=mean_pct_tra_with_tcr-serror, ymax=mean_pct_tra_with_tcr+serror), width=.2,
                 position=position_dodge(.9)) +
  theme_classic()  +
  theme(axis.ticks.x = element_blank()) +
  xlab("Number of constant TCR UMIs on a bead") +
  ylab("Percentage of beads with variable TCRs")

```

# Calculate fraction of TRB beads with TCR
```{r}
pct_trb <- rbind(calculate_pct_trb_beads_with_tcr(P102TA_10_RNA),
              calculate_pct_trb_beads_with_tcr(P102TA_17_RNA),
              calculate_pct_trb_beads_with_tcr(P102TA_33_RNA),
              calculate_pct_trb_beads_with_tcr(P102TA_39_RNA),
              calculate_pct_trb_beads_with_tcr(P107TA_15_RNA),
              calculate_pct_trb_beads_with_tcr(P107TA_18_RNA),
              calculate_pct_trb_beads_with_tcr(P107TA_20_RNA),
              calculate_pct_trb_beads_with_tcr(P108TA_26_RNA),
              calculate_pct_trb_beads_with_tcr(P108TA_29_RNA),
              calculate_pct_trb_beads_with_tcr(P108TA_30_RNA),
              calculate_pct_trb_beads_with_tcr(P108TA_32_RNA),
              calculate_pct_trb_beads_with_tcr(P108TB_06_RNA),
              calculate_pct_trb_beads_with_tcr(P108TB_14_RNA),
              calculate_pct_trb_beads_with_tcr(P109TB_23_RNA),
              calculate_pct_trb_beads_with_tcr(P109TB_34_RNA),
              calculate_pct_trb_beads_with_tcr(P110TA_28_RNA),
              calculate_pct_trb_beads_with_tcr(P110TA_30_RNA),
              calculate_pct_trb_beads_with_tcr(P111TA_32_RNA),
              calculate_pct_trb_beads_with_tcr(P111TA_37_RNA))
```

## Plot after filtering each UMI bin for 3 TCR-containing beads or more
```{r}
pct_trb_sum <- pct_trb %>%
  filter(n_trb_with_tcr_beads > 3,
         n_umi != 0) %>%
  mutate(new_n_umi = case_when(n_umi %in% c(5, 6) ~ "5+",
                               T ~ as.character(n_umi))) %>%
  group_by(new_n_umi) %>%
  summarize(sum_n_trb_beads = sum(n_trb_beads),
            sum_n_trb_beads_with_tcr = sum(n_trb_with_tcr_beads),
            mean_pct_trb_with_tcr = mean(pct_trb_with_tcr),
            std = sd(pct_trb_with_tcr),
            serror = std.error(pct_trb_with_tcr))

trb_plot <- ggplot(pct_trb_sum, aes(x = new_n_umi, y = mean_pct_trb_with_tcr)) +
  geom_col(fill = "#5D74A5FF") +
  geom_text(data = pct_trb_sum, 
              aes(x = new_n_umi, 
                  y = 80,
                  angle = 90,
                  label = sum_n_trb_beads),
              colour="black", 
              size=3.5) +
  theme_classic() +
  geom_errorbar(aes(ymin=mean_pct_trb_with_tcr-serror, 
                    ymax=mean_pct_trb_with_tcr+serror), 
                width=.2,
                position=position_dodge(.9)) +
  theme(axis.title.y = element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank()) +
  xlab("Number of constant TCR UMIs on a bead")

```

# Plot TRA and TRB together
```{r}
tcr_plot <- tra_plot + plot_spacer() + trb_plot + plot_layout(widths = c(6.4, -1.01, 11), axis_titles = "collect", guides = "collect")

tcr_plot
```

# Add TRC gene expression to metadata
```{r}
P102TA_10_RNA <- add_trc_exp(P102TA_10_RNA)
P102TA_17_RNA <- add_trc_exp(P102TA_17_RNA)
P102TA_33_RNA <- add_trc_exp(P102TA_33_RNA)
P102TA_39_RNA <- add_trc_exp(P102TA_39_RNA)

P108TA_26_RNA <- add_trc_exp(P108TA_26_RNA)
P108TA_29_RNA <- add_trc_exp(P108TA_29_RNA)
P108TA_30_RNA <- add_trc_exp(P108TA_30_RNA)
P108TA_32_RNA <- add_trc_exp(P108TA_32_RNA)
```


# P102TA
## P102TA Puck 10
```{r, fig.width = 7, fig.height = 5}
P102TA_10_TCRA <- P102TA_10_RNA@meta.data %>%
  filter(str_detect(chains, "TRA")) %>%
  ggplot(aes(x = plot_x, y = plot_y)) +
  geom_density_2d_filled(contour_var = "density",
                         adjust = 1/10,
                         show.legend = TRUE) +
  stat_density_2d(contour_var = "density",
                         adjust = 1/2, alpha = 0.2, color = "white") +
  density_fill_8 +
  theme_blank +
  xlim(0,5000) +
  ylim(0,5000) +
  ggtitle("P102TA Puck 10 TCR alpha")
  
P102TA_10_TRAC <- P102TA_10_RNA@meta.data %>%
  filter(TRAC > 0,
         area != "Empty") %>%
  ggplot(aes(x = plot_x, y = plot_y)) +
  geom_density_2d_filled(contour_var = "density",
                         adjust = 1/10,
                         show.legend = TRUE) +
  stat_density_2d(contour_var = "density",
                         adjust = 1/2, alpha = 0.2, color = "white") +
  density_fill_12 +
  theme_blank +
  xlim(0,5000) +
  ylim(0,5000) +
  ggtitle("P102TA Puck 10 TRAC")

P102TA_10_TCRB <- P102TA_10_RNA@meta.data %>%
  filter(str_detect(chains, "TRB")) %>%
  ggplot(aes(x = plot_x, y = plot_y)) +
  geom_density_2d_filled(contour_var = "density",
                         adjust = 1/10,
                         show.legend = TRUE) +
  stat_density_2d(contour_var = "density",
                         adjust = 1/2, alpha = 0.2, color = "white") +
  density_fill_12 +
  theme_blank +
  xlim(0,5000) +
  ylim(0,5000) +
  ggtitle("P102TA Puck 10 TCR beta")
  
P102TA_10_TRBC <- P102TA_10_RNA@meta.data %>%
  filter(TRBC > 0,
         area != "Empty") %>%
  ggplot(aes(x = plot_x, y = plot_y)) +
  geom_density_2d_filled(contour_var = "density",
                         adjust = 1/10,
                         show.legend = TRUE) +
  stat_density_2d(contour_var = "density",
                         adjust = 1/2, alpha = 0.2, color = "white") +
  density_fill_8 +
  theme_blank +
  xlim(0,5000) +
  ylim(0,5000) +
  ggtitle("P102TA Puck 10 TRBC")


P102TA_10_TCRA
P102TA_10_TRAC
P102TA_10_TCRB
P102TA_10_TRBC
```

## P102TA Puck 17
```{r, fig.width = 7, fig.height = 5}
P102TA_17_TCRA <- P102TA_17_RNA@meta.data %>%
  filter(str_detect(chains, "TRA")) %>%
  ggplot(aes(x = plot_x, y = plot_y)) +
  geom_density_2d_filled(contour_var = "density",
                         adjust = 1/10,
                         show.legend = TRUE) +
  stat_density_2d(contour_var = "density",
                         adjust = 1/2, alpha = 0.2, color = "white") +
  density_fill_12 +
  theme_blank +
  xlim(0,5000) +
  ylim(0,5000) +
  ggtitle("P102TA Puck 17 TCR alpha")
  
P102TA_17_TRAC <- P102TA_17_RNA@meta.data %>%
  filter(TRAC > 0,
         area != "Empty") %>%
  ggplot(aes(x = plot_x, y = plot_y)) +
  geom_density_2d_filled(contour_var = "density",
                         adjust = 1/10,
                         show.legend = TRUE) +
  stat_density_2d(contour_var = "density",
                         adjust = 1/2, alpha = 0.2, color = "white") +
  density_fill_12 +
  theme_blank +
  xlim(0,5000) +
  ylim(0,5000) +
  ggtitle("P102TA Puck 17 TRAC")

P102TA_17_TCRB <- P102TA_17_RNA@meta.data %>%
  filter(str_detect(chains, "TRB")) %>%
  ggplot(aes(x = plot_x, y = plot_y)) +
  geom_density_2d_filled(contour_var = "density",
                         adjust = 1/10,
                         show.legend = TRUE) +
  stat_density_2d(contour_var = "density",
                         adjust = 1/2, alpha = 0.2, color = "white") +
  density_fill_13 +
  theme_blank +
  xlim(0,5000) +   ylim(0,5000) +
  ggtitle("P102TA Puck 17 TCR beta")
  
P102TA_17_TRBC <- P102TA_17_RNA@meta.data %>%
  filter(TRBC > 0,
         area != "Empty") %>%
  ggplot(aes(x = plot_x, y = plot_y)) +
  geom_density_2d_filled(contour_var = "density",
                         adjust = 1/10,
                         show.legend = TRUE) +
  stat_density_2d(contour_var = "density",
                         adjust = 1/2, alpha = 0.2, color = "white") +
  density_fill_10 +
  theme_blank +
  xlim(0,5000) +   ylim(0,5000) +
  ggtitle("P102TA Puck 17 TRBC")


P102TA_17_TCRA
P102TA_17_TRAC
P102TA_17_TCRB
P102TA_17_TRBC
```


## P102TA Puck 33
```{r, fig.width = 7, fig.height = 5}
P102TA_33_TCRA <- P102TA_33_RNA@meta.data %>%
  filter(str_detect(chains, "TRA")) %>%
  ggplot(aes(x = plot_x, y = plot_y)) +
  geom_density_2d_filled(contour_var = "density",
                         adjust = 1/10,
                         show.legend = TRUE) +
  stat_density_2d(contour_var = "density",
                         adjust = 1/2, alpha = 0.2, color = "white") +
  density_fill_11 +
  theme_blank +
  xlim(0,5000) +   ylim(0,5000) +
  ggtitle("P102TA Puck 33 TCR alpha")
  
P102TA_33_TRAC <- P102TA_33_RNA@meta.data %>%
  filter(TRAC > 0,
         area != "Empty") %>%
  ggplot(aes(x = plot_x, y = plot_y)) +
  geom_density_2d_filled(contour_var = "density",
                         adjust = 1/10,
                         show.legend = TRUE) +
  stat_density_2d(contour_var = "density",
                         adjust = 1/2, alpha = 0.2, color = "white") +
  density_fill_10 +
  theme_blank +
  xlim(0,5000) +   ylim(0,5000) +
  ggtitle("P102TA Puck 33 TRAC")

P102TA_33_TCRB <- P102TA_33_RNA@meta.data %>%
  filter(str_detect(chains, "TRB")) %>%
  ggplot(aes(x = plot_x, y = plot_y)) +
  geom_density_2d_filled(contour_var = "density",
                         adjust = 1/10,
                         show.legend = TRUE) +
  stat_density_2d(contour_var = "density",
                         adjust = 1/2, alpha = 0.2, color = "white") +
  density_fill_8 +
  theme_blank +
  xlim(0,5000) +   ylim(0,5000) +
  ggtitle("P102TA Puck 33 TCR beta")
  
P102TA_33_TRBC <- P102TA_33_RNA@meta.data %>%
  filter(TRBC > 0,
         area != "Empty") %>%
  ggplot(aes(x = plot_x, y = plot_y)) +
  geom_density_2d_filled(contour_var = "density",
                         adjust = 1/10,
                         show.legend = TRUE) +
  stat_density_2d(contour_var = "density",
                         adjust = 1/2, alpha = 0.2, color = "white") +
  density_fill_14 +
  theme_blank +
  xlim(0,5000) +   ylim(0,5000) +
  ggtitle("P102TA Puck 33 TRBC")

P102TA_33_TCRA
P102TA_33_TRAC
P102TA_33_TCRB
P102TA_33_TRBC
```

## P102TA Puck 39
```{r, fig.width = 7, fig.height = 5}
P102TA_39_TCRA <- P102TA_39_RNA@meta.data %>%
  filter(str_detect(chains, "TRA")) %>%
  ggplot(aes(x = plot_x, y = plot_y)) +
  geom_density_2d_filled(contour_var = "density",
                         adjust = 1/10,
                         show.legend = TRUE) +
  stat_density_2d(contour_var = "density",
                         adjust = 1/2, alpha = 0.2, color = "white") +
  density_fill_14 +
  theme_blank +
  xlim(0,5000) +   ylim(0,5000) +
  ggtitle("P102TA Puck 39 TCR alpha")
  
P102TA_39_TRAC <- P102TA_39_RNA@meta.data %>%
  filter(TRAC > 0,
         area != "Empty") %>%
  ggplot(aes(x = plot_x, y = plot_y)) +
  geom_density_2d_filled(contour_var = "density",
                         adjust = 1/10,
                         show.legend = TRUE) +
  stat_density_2d(contour_var = "density",
                         adjust = 1/2, alpha = 0.2, color = "white") +
  density_fill_12 +
  theme_blank +
  xlim(0,5000) +   ylim(0,5000) +
  ggtitle("P102TA Puck 39 TRAC")

P102TA_39_TCRB <- P102TA_39_RNA@meta.data %>%
  filter(str_detect(chains, "TRB")) %>%
  ggplot(aes(x = plot_x, y = plot_y)) +
  geom_density_2d_filled(contour_var = "density",
                         adjust = 1/10,
                         show.legend = TRUE) +
  stat_density_2d(contour_var = "density",
                         adjust = 1/2, alpha = 0.2, color = "white") +
  density_fill_11 +
  theme_blank +
  xlim(0,5000) +   ylim(0,5000) +
  ggtitle("P102TA Puck 39 TCR beta")
  
P102TA_39_TRBC <- P102TA_39_RNA@meta.data %>%
  filter(TRBC > 0,
         area != "Empty") %>%
  ggplot(aes(x = plot_x, y = plot_y)) +
  geom_density_2d_filled(contour_var = "density",
                         adjust = 1/10,
                         show.legend = TRUE) +
  stat_density_2d(contour_var = "density",
                         adjust = 1/2, alpha = 0.2, color = "white") +
  density_fill_9 +
  theme_blank +
  xlim(0,5000) +   ylim(0,5000) +
  ggtitle("P102TA Puck 39 TRBC")

P102TA_39_TCRA
P102TA_39_TRAC
P102TA_39_TCRB
P102TA_39_TRBC
```

# P108TA
## P108TA Puck 26
```{r, fig.width = 7, fig.height = 5}
P108TA_26_TCRA <- P108TA_26_RNA@meta.data %>%
  filter(str_detect(chains, "TRA")) %>%
  ggplot(aes(x = plot_x, y = plot_y)) +
  geom_density_2d_filled(contour_var = "density",
                         adjust = 1/10,
                         show.legend = TRUE) +
  stat_density_2d(contour_var = "density",
                         adjust = 1/2, alpha = 0.2, color = "white") +
  density_fill_9 +
  theme_blank +
  xlim(0,5000) +   ylim(0,5000) +
  ggtitle("P108TA Puck 26 TCR alpha")
  
P108TA_26_TRAC <- P108TA_26_RNA@meta.data %>%
  filter(TRAC > 0,
         area != "Empty") %>%
  ggplot(aes(x = plot_x, y = plot_y)) +
  geom_density_2d_filled(contour_var = "density",
                         adjust = 1/10,
                         show.legend = TRUE) +
  stat_density_2d(contour_var = "density",
                         adjust = 1/2, alpha = 0.2, color = "white") +
  density_fill_9 +
  theme_blank +
  xlim(0,5000) +   ylim(0,5000) +
  ggtitle("P108TA Puck 26 TRAC")

P108TA_26_TCRB <- P108TA_26_RNA@meta.data %>%
  filter(str_detect(chains, "TRB")) %>%
  ggplot(aes(x = plot_x, y = plot_y)) +
  geom_density_2d_filled(contour_var = "density",
                         adjust = 1/10,
                         show.legend = TRUE) +
  stat_density_2d(contour_var = "density",
                         adjust = 1/2, alpha = 0.2, color = "white") +
  density_fill_9 +
  theme_blank +
  xlim(0,5000) +   ylim(0,5000) +
  ggtitle("P108TA Puck 26 TCR beta")

P108TA_26_TRBC <- P108TA_26_RNA@meta.data %>%
  filter(TRBC > 0,
         area != "Empty") %>%
  ggplot(aes(x = plot_x, y = plot_y)) +
  geom_density_2d_filled(contour_var = "density",
                         adjust = 1/10,
                         show.legend = TRUE) +
  stat_density_2d(contour_var = "density",
                         adjust = 1/2, alpha = 0.2, color = "white") +
  density_fill_12 +
  theme_blank +
  xlim(0,5000) +   ylim(0,5000) +
  ggtitle("P108TA Puck 26 TRBC")

P108TA_26_TCRA
P108TA_26_TRAC
P108TA_26_TCRB
P108TA_26_TRBC
```

## P108TA Puck 29
```{r, fig.width = 7, fig.height = 5}
P108TA_29_TCRA <- P108TA_29_RNA@meta.data %>%
  filter(str_detect(chains, "TRA")) %>%
  ggplot(aes(x = plot_x, y = plot_y)) +
  geom_density_2d_filled(contour_var = "density",
                         adjust = 1/10,
                         show.legend = TRUE) +
  stat_density_2d(contour_var = "density",
                         adjust = 1/2, alpha = 0.2, color = "white") +
  density_fill_8 +
  theme_blank +
  xlim(0,5000) +   ylim(0,5000) +
  ggtitle("P108TA Puck 29 TCR alpha")
  
P108TA_29_TRAC <- P108TA_29_RNA@meta.data %>%
  filter(TRAC > 0,
         area != "Empty") %>%
  ggplot(aes(x = plot_x, y = plot_y)) +
  geom_density_2d_filled(contour_var = "density",
                         adjust = 1/10,
                         show.legend = TRUE) +
  stat_density_2d(contour_var = "density",
                         adjust = 1/2, alpha = 0.2, color = "white") +
  density_fill_12 +
  theme_blank +
  xlim(0,5000) +   ylim(0,5000) +
  ggtitle("P108TA Puck 29 TRAC")

P108TA_29_TCRB <- P108TA_29_RNA@meta.data %>%
  filter(str_detect(chains, "TRB")) %>%
  ggplot(aes(x = plot_x, y = plot_y)) +
  geom_density_2d_filled(contour_var = "density",
                         adjust = 1/10,
                         show.legend = TRUE) +
  stat_density_2d(contour_var = "density",
                         adjust = 1/2, alpha = 0.2, color = "white") +
  density_fill_8 +
  theme_blank +
  xlim(0,5000) +   ylim(0,5000) +
  ggtitle("P108TA Puck 29 TCR beta")
  
P108TA_29_TRBC <- P108TA_29_RNA@meta.data %>%
  filter(TRBC > 0,
         area != "Empty") %>%
  ggplot(aes(x = plot_x, y = plot_y)) +
  geom_density_2d_filled(contour_var = "density",
                         adjust = 1/10,
                         show.legend = TRUE) +
  stat_density_2d(contour_var = "density",
                         adjust = 1/2, alpha = 0.2, color = "white") +
  density_fill_10 +
  theme_blank +
  xlim(0,5000) +   ylim(0,5000) +
  ggtitle("P108TA Puck 29 TRBC")

P108TA_29_TCRA
P108TA_29_TRAC
P108TA_29_TCRB
P108TA_29_TRBC
```

## P108TA Puck 30
```{r, fig.width = 7, fig.height = 5}
P108TA_30_TCRA <- P108TA_30_RNA@meta.data %>%
  filter(str_detect(chains, "TRA")) %>%
  ggplot(aes(x = plot_x, y = plot_y)) +
  geom_density_2d_filled(contour_var = "density",
                         adjust = 1/10,
                         show.legend = TRUE) +
  stat_density_2d(contour_var = "density",
                         adjust = 1/2, alpha = 0.2, color = "white") +
  density_fill_13 +
  theme_blank +
  coord_fixed(ratio = 1, ylim = c(500, 4500)) +
  theme(panel.background = element_rect(fill = '#4B0055FF', colour = "#4B0055FF")) +
  ggtitle("P108TA Puck 30 TCR alpha")
  
P108TA_30_TRAC <- P108TA_30_RNA@meta.data %>%
  filter(TRAC > 0,
         area != "Empty") %>%
  ggplot(aes(x = plot_x, y = plot_y)) +
  geom_density_2d_filled(contour_var = "density",
                         adjust = 1/10,
                         show.legend = TRUE) +
  stat_density_2d(contour_var = "density",
                         adjust = 1/2, alpha = 0.2, color = "white") +
  density_fill_11 +
  theme_blank +
  coord_fixed(ratio = 1, ylim = c(500, 4500)) +
  theme(panel.background = element_rect(fill = '#4B0055FF', colour = "#4B0055FF")) +
  ggtitle("P108TA Puck 30 TRAC")

P108TA_30_TCRB <- P108TA_30_RNA@meta.data %>%
  filter(str_detect(chains, "TRB")) %>%
  ggplot(aes(x = plot_x, y = plot_y)) +
  geom_density_2d_filled(contour_var = "density",
                         adjust = 1/10,
                         show.legend = TRUE) +
  stat_density_2d(contour_var = "density",
                         adjust = 1/2, alpha = 0.2, color = "white") +
  density_fill_13 +
  theme_blank +
  coord_fixed(ratio = 1, ylim = c(500, 4500)) +
  theme(panel.background = element_rect(fill = '#4B0055FF', colour = "#4B0055FF")) +
  ggtitle("P108TA Puck 30 TCR beta")
  
P108TA_30_TRBC <- P108TA_30_RNA@meta.data %>%
  filter(TRBC > 0,
         area != "Empty") %>%
  ggplot(aes(x = plot_x, y = plot_y)) +
  geom_density_2d_filled(contour_var = "density",
                         adjust = 1/10,
                         show.legend = TRUE) +
  stat_density_2d(contour_var = "density",
                         adjust = 1/2, alpha = 0.2, color = "white") +
  density_fill_14 +
  theme_blank +
  coord_fixed(ratio = 1, ylim = c(500, 4500)) +
  theme(panel.background = element_rect(fill = '#4B0055FF', colour = "#4B0055FF")) +
  ggtitle("P108TA Puck 30 TRBC")

P108TA_30_TCRA
P108TA_30_TRAC
P108TA_30_TCRB
P108TA_30_TRBC
```

## P108TA Puck 32
```{r, fig.width = 7, fig.height = 5}
P108TA_32_TCRA <- P108TA_32_RNA@meta.data %>%
  filter(str_detect(chains, "TRA")) %>%
  ggplot(aes(x = plot_x, y = plot_y)) +
  geom_density_2d_filled(contour_var = "density",
                         adjust = 1/10,
                         show.legend = TRUE) +
  stat_density_2d(contour_var = "density",
                         adjust = 1/2, alpha = 0.2, color = "white") +
  density_fill_11 +
  theme_blank +
  xlim(0,5000) +   ylim(0,5000) +
  ggtitle("P108TA Puck 32 TCR alpha")
  
P108TA_32_TRAC <- P108TA_32_RNA@meta.data %>%
  filter(TRAC > 0,
         area != "Empty") %>%
  ggplot(aes(x = plot_x, y = plot_y)) +
  geom_density_2d_filled(contour_var = "density",
                         adjust = 1/10,
                         show.legend = TRUE) +
  stat_density_2d(contour_var = "density",
                         adjust = 1/2, alpha = 0.2, color = "white") +
  density_fill_12 +
  theme_blank +
  xlim(0,5000) +   ylim(0,5000) +
  ggtitle("P108TA Puck 32 TRAC")

P108TA_32_TCRB <- P108TA_32_RNA@meta.data %>%
  filter(str_detect(chains, "TRB")) %>%
  ggplot(aes(x = plot_x, y = plot_y)) +
  geom_density_2d_filled(contour_var = "density",
                         adjust = 1/10,
                         show.legend = TRUE) +
  stat_density_2d(contour_var = "density",
                         adjust = 1/2, alpha = 0.2, color = "white") +
  density_fill_12 +
  theme_blank +
  xlim(0,5000) +   ylim(0,5000) +
  ggtitle("P108TA Puck 32 TCR beta")
  
P108TA_32_TRBC <- P108TA_32_RNA@meta.data %>%
  filter(TRBC > 0,
         area != "Empty") %>%
  ggplot(aes(x = plot_x, y = plot_y)) +
  geom_density_2d_filled(contour_var = "density",
                         adjust = 1/10,
                         show.legend = TRUE) +
  stat_density_2d(contour_var = "density",
                         adjust = 1/2, alpha = 0.2, color = "white") +
  density_fill_10 +
  theme_blank +
  xlim(0,5000) +   ylim(0,5000) +
  ggtitle("P108TA Puck 32 TRBC")

P108TA_32_TCRA
P108TA_32_TRAC
P108TA_32_TCRB
P108TA_32_TRBC
```

# Get session info
```{r}
sessionInfo()
```


