# 111T Puck37: TCR integration {#sec-tcr-integration}

# Integrating spatial RNA and TCR data

## Libraries
```{r}
library(SeuratObject)
library(Seurat)
library(anndata)
library(reticulate)
library(wesanderson)
library(dplyr)
library(stringr)
library(clustree)
library(readxl)
library(reshape2)
library(ggplotify)
library(cowplot)
library(comparator)
library(ggpointdensity)
library(ggvenn)
library(rmarkdown)
library(evaluate)
library(knitr)

source("../source_scripts/slideseq_functions_main.R")

options(Seurat.object.assay.version = 'v5')
```

## Load virtual environment using reticulate
```{r}
use_virtualenv("/home/cht269/my_envs/new_envs/R_anndata_py3.13.1/")
```

## Load RNA puck
```{r}
P111TA_37_RNA <- load_slideseq(gene_exp_dir = "/n/data2/dfci/medonc/cwu/chloe/slideseq_data/raw_data/RCC_pucks/2024-01-12_Puck_231024_37_111TA/gene_exp/", coords_dir = "/n/data2/dfci/medonc/cwu/chloe/slideseq_data/raw_data/RCC_pucks/2024-01-12_Puck_231024_37_111TA/Puck_231024_37_barcode_xy.txt.gz", project_name = "P111TA_37")

dim(P111TA_37_RNA)
```

## Load TCR puck
```{r}
P111TA_37_TCR <- anndata::read_h5ad("/n/data2/dfci/medonc/cwu/chloe/slideseq_data/tcr_analysis/RCC_neovax_pucks/nanoranger_anndata/adata_Puck_231024_37_TCR_111TA.h5ad")
P111TA_37_TCR
```

## Remove non-productive CDR3 sequences
```{r}
all_cdr3 <- P111TA_37_TCR$var_names
all_cdr3 <- all_cdr3[!str_detect(all_cdr3, "_|\\*") & (str_detect(all_cdr3, "^C"))]
P111TA_37_TCR <- P111TA_37_TCR[,all_cdr3]
```

## Extract helpful dataframes from TCR anndata object
```{r}
TCR_seq <- P111TA_37_TCR$var
TCR_xy <- P111TA_37_TCR$obs %>%
  mutate(barcode = paste0(barcode, "-1"),
         x = round(x,1),
         y = round(y,1)) %>%
  dplyr::rename("TCR_x" = "x", 
                "TCR_y" = "y", 
                "TCR_barcode" = "barcode")

rownames(TCR_xy) <- TCR_xy$barcode
```

## Double-check that the coordinate data matches
```{r}
RNA_xy <- P111TA_37_RNA@images$image@coordinates %>%
  mutate(barcode = rownames(.)) %>%
  dplyr::rename("RNA_x" = "x", 
                "RNA_y" = "y", 
                "RNA_barcode" = "barcode")

rownames(RNA_xy) <- RNA_xy$barcode

RNA_xy %>%
  left_join(TCR_xy, by = c("RNA_barcode" = "TCR_barcode")) %>%
  filter(!is.na(TCR_x)) %>%
  mutate(x_match = case_when(RNA_x == TCR_x ~ TRUE,
                             RNA_x != TCR_x ~ FALSE),
         y_match = case_when(RNA_y == TCR_y ~ TRUE,
                             RNA_y != TCR_y ~ FALSE)) %>%
  count(x_match, y_match)
  
```

## How many barcodes gathered from TCR exist in RNA
```{r}
TCR_xy %>%
  mutate(IN_RNA = case_when((TCR_barcode %in% RNA_xy$RNA_barcode) ~ TRUE,
                            !(TCR_barcode %in% RNA_xy$RNA_barcode) ~ FALSE)) %>%
  ggplot(aes(x = 1, fill = IN_RNA)) + 
  geom_bar() +
  scale_fill_manual(values = wes_palette("Darjeeling1")) +
  theme_classic() +
  ggtitle("No. of beads from TCR pipeline in RNA")
```

## Assign all TCR barcodes to an RNA barcode
```{r}
final_xy <- map_tcr_and_rna_barcodes(TCR_xy, RNA_xy)
```

## Add the TCR data to the RNA as an additional assay
```{r}
P111TA_37_RNA <- update_and_add_tcr_assay(P111TA_37_TCR, final_xy, P111TA_37_RNA)
```

## Calculate Hamming distance between TCR barcodes and RNA barcodes
```{r}
hamming_option = Hamming()
final_xy <- final_xy %>%
  mutate(hamming_dist = hamming_option(TCR_barcode, RNA_barcode))
```

## Plot distribution of physical distance and Hamming distance of TCR-unique barcodes
```{r}
final_xy %>%
  filter(TCR_barcode != RNA_barcode) %>%
  ggplot(aes(x = distance, y = hamming_dist)) +
    geom_pointdensity() + 
  theme_classic() +
  xlab("physical distance") +
  ylab("hamming distance")
```

## Plot number of beads w/ clonotype data
```{r, fig.width = 3, fig.height = 3}
P111TA_37_RNA@meta.data %>%
  mutate(TCR = case_when((nFeature_TCR > 0) ~ TRUE,
                         (nFeature_TCR == 0) ~ FALSE)) %>%
  ggplot(aes(x = 1, fill = TCR)) + 
  geom_bar() +
  scale_fill_manual(values = wes_palette("Darjeeling1")) +
  theme_classic() +
  ggtitle("No. of beads w/ TCR")
```

## Plot number of beads w/ multiple TCR chains
```{r, fig.width = 5, fig.height = 3}
P111TA_37_RNA@meta.data %>%
  filter(nFeature_TCR > 0) %>%
  ggplot(aes(x = 1, fill = as.factor(nFeature_TCR))) + 
  geom_bar() +
  scale_fill_manual(values = wes_palette("Darjeeling2")) +
  theme_classic() +
  ggtitle("No. of beads w/ multiple TCR")
```

## Open reactivity and scRNA-scTCR derived phenotypic data
```{r}
reactivities <- read.csv("../objects/reformatted_RCC_slideseq_parameters_052925.csv") %>%
  filter(str_detect(sctcr_clone, "p111"))
``` 

## Attach single cell derived clonotype descriptions to TCR metadata
```{r}
TCR_seq <- TCR_seq %>%
  mutate(CDR3 = rownames(.)) %>%
  left_join(reactivities, by = c("chains", "CDR3"))
```

## Create a summary of clonotypes per bead in object metadata
```{r}
P111TA_37_RNA <- add_CDR3s_to_RNA_md(P111TA_37_RNA)
puck_specific_scTCR_clones <- get_puck_specific_sctcr_clones(TCR_seq, reactivities)
P111TA_37_RNA <- attach_phenotype_specificity_to_RNA_md(puck_specific_scTCR_clones, P111TA_37_RNA)
P111TA_37_RNA <- create_final_clone_ids(P111TA_37_RNA, reactivities)
```

## Plot TCR distribution spatially
```{r}
P111TA_37_RNA$log_nFeature_TCR <- log(P111TA_37_RNA$nFeature_TCR)
P111TA_37_RNA$log_nCount_TCR <- log(P111TA_37_RNA$nCount_TCR)
P111TA_37_RNA$log_nFeature_Spatial <- log(P111TA_37_RNA$nFeature_Spatial)
P111TA_37_RNA$log_nCount_Spatial <- log(P111TA_37_RNA$nCount_Spatial)
```

## Log scale the rna and tcr features and umi
```{r}
SpatialFeaturePlot(P111TA_37_RNA, features = c("log_nFeature_TCR", "log_nCount_TCR"), pt.size = 3, alpha = c(0.5,1))
SpatialFeaturePlot(P111TA_37_RNA, features = c("log_nFeature_Spatial", "log_nCount_Spatial"), pt.size = 3, alpha = c(0.5,1))

```

## Calculate clonal expansion per clonotype
```{r}
# Count up clones
puck37_111TA_clone_id_count <- P111TA_37_RNA@meta.data %>%
  filter(!is.na(clone_id)) %>%
  pull(clone_id) %>%
  paste0(collapse = ",") %>%
  str_split_1(fixed(",")) %>%
  as.data.frame() %>%
  dplyr::rename("clone_id" = ".") %>%
  count(clone_id) %>%
  dplyr::rename("puck37_count" = "n")

clone_counts.df <- data.frame()
  # Remove non-TCR containing beads
  puck_T <- P111TA_37_RNA@meta.data %>%
    filter(nFeature_TCR > 0)
  # Beads can have multiple TCRs- assign the largest count
  for(i in c(1:nrow(puck_T))){
    # Get the clones/CDR3 associated with each barcode
    associated_clones <- puck_T[i,"clone_id"] %>%
      str_split_1(",")
    # Find these clones in the total counts df and pick whichever clone is the largest
    associated_counts <- puck37_111TA_clone_id_count %>%
      filter(clone_id %in% associated_clones) %>%
      filter(puck37_count == max(puck37_count)) %>%
      slice_head(n = 1) %>%
      dplyr::rename("largest_clone_id" = "clone_id") %>%
      select(largest_clone_id, puck37_count) %>%
      mutate(log_puck37_count = log10(puck37_count))
    # Push this barcode, it's associated clone and counts to a dataframe
    new_row <- cbind(associated_counts, barcode = puck_T[i,"barcode"])
    clone_counts.df <- rbind(clone_counts.df, new_row)
  }
  
  # Update metadata of the puck
  P111TA_37_RNA@meta.data <- P111TA_37_RNA@meta.data %>%
    full_join(clone_counts.df, by = "barcode") %>%
    mutate(puck37_count = case_when(is.na(puck37_count) ~ 0,
                                    !is.na(puck37_count) ~ puck37_count),
           log_puck37_count = case_when(is.na(log_puck37_count) ~ 0,
                                      !is.na(log_puck37_count) ~ log_puck37_count))
  
  rownames(P111TA_37_RNA@meta.data) <- P111TA_37_RNA@meta.data$barcode
  
P111TA_37_RNA %>% 
  subset(log_puck37_count < 1 & nFeature_TCR > 0) %>%
SpatialFeaturePlot(features = "log_puck37_count", pt.size = 3, stroke = 0.1, alpha = 0.5) + scale_fill_viridis()
```

## Flip the XY coordinates to match the orientation of the H&E stain
```{r}
P111TA_37_xy <- P111TA_37_RNA@images$image@coordinates

# flip vertically
P111TA_37_xy <- P111TA_37_xy %>%
  dplyr::rename("orig_x" = "x",
                "orig_y" = "y") %>%
  mutate(x = orig_y,
         y = -orig_x+5000) %>%
  select(-orig_x,-orig_y)

P111TA_37_RNA@images$image@coordinates <- P111TA_37_xy

SpatialFeaturePlot(P111TA_37_RNA, features = c("log_nFeature_Spatial"), alpha = c(0,1)) +
  theme_bw() +
  coord_fixed(ratio = 1, xlim = c(-5000,5000), ylim = c(-5000,5000), expand = FALSE, clip = "off") +
  scale_color_gradient(low = "darkgreen", high = "black") +
  scale_alpha(range = c(0, 1)) +
  theme(legend.position = "None")
```

## Save merged object
```{r, eval = FALSE}
saveRDS(P111TA_37_RNA, "01_Puck37_111TA_TCR_RNA_merged_110724.Rds")
```

## Get session info
```{r}
sessionInfo()
```
