# 108T: Spatial bias to TLS1 {#sec-spatial-bias}

# Calculating spatial bias of clonotypes toward particular TLS regions

## Libraries
```{r}
library(Seurat)
library(SeuratWrappers)
library(ggplot2)
library(patchwork)
library(dplyr)
library(wesanderson)
library(ggplotify)
library(cowplot)
library(stringr)
library(reshape2)
library(ggrepel)
library(tidyverse)
library(infer)
library(RANN)

area_fill <- readRDS("../objects/Area_fill.rds")
source("../source_scripts/slideseq_functions_main.R")
options(Seurat.object.assay.version = 'v5')
```

## Following Yaniv Brandvain's book, chapter 17.
https://bookdown.org/ybrandvain/Applied-Biostats/perm1.html#permute-to-generate-a-null-distribution

## Importance of sample sizes
https://www.tandfonline.com/doi/abs/10.1080/00031305.2021.1902856

## Set number of permutations
```{r}
n_perm <- 5000
set.seed(123)
```

## Load object from script 02
```{r}
P108TA_30_RNA <- readRDS("../objects/02_Puck30_108TA_TCR_RNA_area_121125.Rds")
P108TA_32_RNA <- readRDS("../objects/02_Puck32_108TA_TCR_RNA_area_121125.Rds")

```

## Get TCRs seen near TLS1 at a count of 5 or more
```{r}
P108TA_md <- do.call(rbind, list(
  P108TA_30_RNA@meta.data, P108TA_32_RNA@meta.data
  ))

P108TA_clones <- get_unique_productive_clones(P108TA_md$clone_id)

P108TA_filtered_clones <- count_clones_by_annotated_area(P108TA_md, P108TA_clones) %>%
  group_by(clone_id) %>%
  summarize(sum_count = sum(count)) %>%
  filter(sum_count >= 5) %>%
  pull(clone_id)

plot_clones_spatially(P108TA_30_RNA, P108TA_filtered_clones, area_fill)
plot_clones_spatially(P108TA_32_RNA, P108TA_filtered_clones, area_fill)
```

## Permutation test logic:
Hypothesis:
Our null hypothesis claims that TCR_X is not closer to TLS1 compared to other TCRs
The alternative hypothesis claims that TCR_Xs is closer to TLS1 than any other TCR

Take into account the non-independent variable of puck identity by permuting sample label within the puck. e.g. say Puck30 has 10 TCR_X beads, and Puck32 has 20 TCR_X beads. After permutation, Puck30 will still have 10 TCR_X beads, and Puck32 20 beads.

Test statistic:
Median distance from centroid of TLS1 of TCR_X minus Median distance from centroid of TLS1 of all other TCRs
Notes: Median seems superior to mean in this case, since I think outliers would make a big impact
If alternative hypothesis is true: small number - bigger number = negative number

Treatment assignment:
The treatment group are beads or TCR-containing beads assigned TCR_X
The control group are beads or TCR-containing beads not assigned TCR_X

## Get centroid of TLS1
```{r}
tls1_centroid_30 <- P108TA_30_RNA@meta.data %>%
  filter(area == "TLS1") %>%
  summarize(x = mean(x),
            y = mean(y))
tls1_centroid_32 <- P108TA_32_RNA@meta.data %>%
  filter(area == "TLS1") %>%
  summarize(x = mean(x),
            y = mean(y))
```

## Save TLS centroids for plotting purposes
```{r}
tls1_centroid_30_PLOT <- P108TA_30_RNA@meta.data %>%
  filter(area == "TLS1") %>%
  summarize(x = mean(plot_x),
            y = mean(plot_y))
tls1_centroid_32_PLOT <- P108TA_32_RNA@meta.data %>%
  filter(area == "TLS1") %>%
  summarize(x = mean(plot_x),
            y = mean(plot_y))

write.csv(tls1_centroid_30_PLOT, "../objects/P108TA_puck30_tls1_centroid.csv")
write.csv(tls1_centroid_32_PLOT, "../objects/P108TA_puck32_tls1_centroid.csv")
```


## Calculate the distance of all TCRs to the centroid of TLS1 and add label denoting whether the TCR is of interest or not
```{r}
# Calculate the distance of all beads to TLS1
tls1_distances_30 <- calculate_distance_of_beads_to_a_point(P108TA_30_RNA@meta.data, tls1_centroid_30, "TLS1") %>%
  dplyr::rename("distance_to" = "distance_to_TLS1") %>%
  mutate(Puck = "P108TA_Puck30")
tls1_distances_32 <- calculate_distance_of_beads_to_a_point(P108TA_32_RNA@meta.data, tls1_centroid_32, "TLS1") %>%
  dplyr::rename("distance_to" = "distance_to_TLS1") %>%
  mutate(Puck = "P108TA_Puck32")
```

## Run permutation test
```{r}
# Save data
tls1.list <- list()
for(clone in P108TA_filtered_clones){
  # Merge distance distributions across Puck 30 and Puck 32
  tls1_X_distances <- tls1_distances_30 %>%
    rbind(tls1_distances_32) %>%
    filter(!is.na(clone_id)) %>%
    # Label TCR_206 containing beads
    mutate(label = case_when(str_detect(clone_id, paste0(clone, "($|,)")) ~ "TCR_of_interest",
                                    T ~ "Other"))
  
  # Find pucks in which TCR_X is observed
  pucks <- tls1_X_distances %>%
    filter(label == "TCR_of_interest") %>%
    distinct(Puck) %>%
    pull(Puck)
  
  # Filter for pucks where TCR_X is observed
  tls1_X_distances <- tls1_X_distances %>%
    filter(Puck %in% pucks)
  
  # Plot the median distance of TCR_206 vs all other TCRs for each puck
  # Plot descriptive stats of the bead distance to TLS1
  plot <- ggplot(data = tls1_X_distances, aes(x = label, y = distance_to, fill = label)) +
    geom_boxplot() +
    facet_wrap(~Puck) +
    ggtitle(paste0("TCR of interest, ", clone))
  
  print(plot)
  
  # Test TCR_Xs proximity to TLS1
  tls1_X <- distance_permutation_test(tls1_X_distances, n_perm, clone)
  tls1_X$pval
  
  # KEEP THIS ORDER, or the names of the list will be in the wrong order later
  tls1.list <- append(tls1.list, list(tls1_X))
}

names(tls1.list) <- P108TA_filtered_clones
```

## Print pvals
```{r}
get_pval <- function(lis)(
  return(lis[["pval"]])
)

tls1.df <- lapply(tls1.list, get_pval) %>%
  as.data.frame() %>%
  t() %>%
  as.data.frame() %>%
  dplyr::rename("pval" = "V1")
```

## Print significant clones (unadjusted)
```{r}
tls1.df %>%
  filter(pval <= 0.05)
```

## Save object
```{r}
saveRDS(tls1.list, "../objects/P108TA_puck30_32_tls1_list.rds")
```

## Get session info
```{r}
sessionInfo()
```

