# 107T: Plot spatial bias to TLS {#sec-spatial-bias}

# Plot spatial bias of clonotypes toward particular TLS regions

## Libraries
```{r}
library(Seurat)
library(SeuratWrappers)
library(ggplot2)
library(patchwork)
library(dplyr)
library(wesanderson)
library(ggplotify)
library(cowplot)
library(stringr)
library(reshape2)
library(ggrepel)
library(tidyverse)
library(infer)
library(RANN)
library(ComplexHeatmap)
library(circlize)
library(dplyr)

area_fill <- readRDS("../objects/Area_fill.rds")
source("../source_scripts/slideseq_functions_main.R")
options(Seurat.object.assay.version = 'v5')

```

# Open pvals from permutation tests for TLS1, TLS2, TLS3 and TLS4
No results for TLS3 and TLS4- no clones reached n>=5
```{r}
tls1.list <- readRDS("../objects/P107TA_puck20_tls1_list.rds")
tls2.list <- readRDS("../objects/P107TA_puck20_18_tls2_list.rds")
```

# Get pvals from the list
```{r}
get_pval <- function(lis)(
  return(lis[["pval"]])
)

tls1.df <- lapply(tls1.list, get_pval) %>%
  as.data.frame() %>%
  t() 

tls2.df <- lapply(tls2.list, get_pval) %>%
  as.data.frame() %>%
  t() 

```

## Append dataframes together and correct for FDR
```{r}
tls1.df <- tls1.df %>% 
  as.data.frame() %>% 
  rownames_to_column("clone_id") %>% 
  dplyr::rename("TLS1_pval" = "V1") %>%
  mutate(TLS1_pval = case_when(TLS1_pval == 0 ~ 1/5000,
                               T ~ TLS1_pval),
         TLS1_qval = p.adjust(TLS1_pval, method="BH"))
  
tls2.df <- tls2.df %>% as.data.frame() %>% 
  rownames_to_column("clone_id") %>% 
  dplyr::rename("TLS2_pval" = "V1") %>% 
  mutate(TLS2_pval = case_when(TLS2_pval == 0 ~ 1/5000,
                               T ~ TLS2_pval),
         TLS2_qval = p.adjust(TLS2_pval, method="BH"))

tls.df <- tls1.df %>%
  full_join(tls2.df)

```

## Reorder data based on minimum significance
```{r}
reorder.df <- tls.df %>%
  rowwise() %>%
  mutate(min_qval = min(TLS1_qval, TLS2_qval, na.rm = TRUE)) %>%
  arrange(min_qval) %>%
  ungroup()
```

## Plot heatmap without enrichment
```{r, fig.width = 5, fig.height = 3}
reorder.mat <- reorder.df %>%
  arrange(min_qval) %>%
  select(clone_id, TLS1_qval, TLS2_qval) %>%
  column_to_rownames("clone_id") %>%
  as.matrix()

col_fun = colorRamp2(c(0, 0.05, 0.06, 1), c("yellow", "white", "purple", "purple4"))

Heatmap(reorder.mat,
        cluster_columns = FALSE,
        cluster_rows = FALSE,
        row_names_gp = gpar(fontsize = 9),
        col = col_fun,
        na_col = "grey50",
        heatmap_width = unit(6, "cm"),
        column_title = "P107",
        name = "q val in a\nsingle TLS")
```

## Get session info
```{r}
sessionInfo()
```

