# Plot median distances {#sec-spatial-bias}

# Plot spatial bias of clonotypes toward particular TLS regions

## Libraries
```{r}
library(Seurat)
library(SeuratWrappers)
library(ggplot2)
library(patchwork)
library(dplyr)
library(wesanderson)
library(ggplotify)
library(cowplot)
library(stringr)
library(reshape2)
library(ggrepel)
library(tidyverse)
library(infer)
library(ggforce)
library(RANN)

area_fill <- readRDS("../objects/Area_fill.rds")
source("../source_scripts/slideseq_functions_main.R")
options(Seurat.object.assay.version = 'v5')
```

## Load object from script 02
```{r}
P102TA_33_RNA <- readRDS("../objects/02_Puck33_102TA_TCR_RNA_area_121125.Rds")
P102TA_17_RNA <- readRDS("../objects/02_Puck17_102TA_TCR_RNA_area_121125.Rds")
P108TA_32_RNA <- readRDS("../objects/02_Puck32_108TA_TCR_RNA_area_121125.Rds")
```

## Remove empty
```{r}
Idents(P102TA_33_RNA) <- P102TA_33_RNA$area
P102TA_33_RNA <- subset(P102TA_33_RNA, idents = "Empty", invert = TRUE)
Idents(P108TA_32_RNA) <- P108TA_32_RNA$area
P108TA_32_RNA <- subset(P108TA_32_RNA, idents = "Empty", invert = TRUE)
```

## Reorient pucks
```{r}
reorient_pucks <- function(obj){
  obj@images$image@coordinates$x <- obj$plot_x
  obj@images$image@coordinates$y <- -obj$plot_y

  return(obj)
}

P102TA_17_RNA <- reorient_pucks(P102TA_17_RNA)
P102TA_33_RNA <- reorient_pucks(P102TA_33_RNA)
P108TA_32_RNA <- reorient_pucks(P108TA_32_RNA)

```

## Load median distances of clones analyzed
```{r}
p108_tls1 <- readRDS("../objects/P108TA_puck30_32_tls1_list.rds")
p108_tls2 <- readRDS("../objects/P108TA_puck26_29_32_tls2_list.rds")
p108_tls3 <- readRDS("../objects/P108TA_puck26_29_32_tls3_list.rds")

p102_tls3 <- readRDS("../objects/P102TA_puck17_33_tls3_list.rds")
```

## Format median distances
```{r}
p108_tls1_meds <- get_dist_df(p108_tls1) %>%
  mutate(clone = rownames(.),
         TLS = "TLS1") %>%
  column_to_rownames("clone")
p108_tls2_meds <- get_dist_df(p108_tls2) %>%
  mutate(clone = rownames(.),
         TLS = "TLS2") %>%
  column_to_rownames("clone")
p108_tls3_meds <- get_dist_df(p108_tls3) %>%
  mutate(clone = rownames(.),
         TLS = "TLS3") %>%
  column_to_rownames("clone")

p102_tls3_meds <- get_dist_df(p102_tls3) %>%
  mutate(clone = rownames(.),
         TLS = "TLS3") %>%
  column_to_rownames("clone")
```

## Load TLS centroids
```{r}
tls1_centroid_32 <- read.csv("../objects/P108TA_puck32_tls1_centroid.csv", row.names = 1) %>%
  mutate(TLS = "TLS1")
tls2_centroid_32 <- read.csv("../objects/P108TA_puck32_tls2_centroid.csv", row.names = 1) %>%
  mutate(TLS = "TLS2")
tls3_centroid_32 <- read.csv("../objects/P108TA_puck32_tls3_centroid.csv", row.names = 1) %>%
  mutate(TLS = "TLS3")

tls3_centroid_33 <- read.csv("../objects/P102TA_puck33_tls3_centroid.csv", row.names = 1) %>%
  mutate(TLS = "TLS3")
tls3_centroid_17 <- read.csv("../objects/P102TA_puck17_tls3_centroid.csv", row.names = 1) %>%
  mutate(TLS = "TLS3")
```

# P108 plot
## Format TLS centroids
```{r}
p108_centroid <- do.call(rbind, list(tls1_centroid_32, tls2_centroid_32, tls3_centroid_32))
```

## Set clone to investigate
```{r}
clone_of_int <- "p108_206"
```

## Merge data for plot
```{r}
p108_distance_data <- 
  p108_tls1_meds[clone_of_int,] %>%
  rbind(p108_tls2_meds[clone_of_int,]) %>%
  rbind(p108_tls3_meds[clone_of_int,]) %>%
  pivot_longer(cols = c("Other", "TCR_of_interest"), values_to = "radii", names_to = "type") %>%
  left_join(p108_centroid, by = "TLS") %>%
  mutate(radii = round(radii, 3), 
         x = round(x, 3), 
         y = round(y, 3))
```

## Reorient beads
```{r}
# Reorient beads to match the change in SpatialDimPlot
# Pick out the beads with the clone
beads_to_highlight <- P108TA_32_RNA@meta.data %>%
      filter(str_detect(clone_id, paste0(paste0(clone_of_int, "($|,)"), collapse = "|"))) %>%
      mutate(trim = str_extract(clone_id, paste0(clone_of_int, collapse = "|")),
             plot_x = plot_x,
             plot_y = -plot_y)

p108_distance_data <- p108_distance_data %>%
  mutate(x = x,
         y = -y)

tls1_centroid_32 <- tls1_centroid_32 %>%
  mutate(x = x,
         y = -y)
tls2_centroid_32 <- tls2_centroid_32 %>%
  mutate(x = x,
         y = -y)
tls3_centroid_32 <- tls3_centroid_32 %>%
  mutate(x = x,
         y = -y)
```

## Plot clones spatially
```{r}
SpatialDimPlot(P108TA_32_RNA, group.by = "area", alpha = 0.3, stroke = 0) +
  # Add beads of interest
  geom_point(data = beads_to_highlight, aes(x = plot_x, y = plot_y), color = "pink", 
             size = 1) +
    coord_fixed(xlim = c(0, 5000), ylim = c(-5000, 0), ratio = 1) +
  # Add outline to beads of interest
  geom_point(data = beads_to_highlight, aes(x = plot_x, y = plot_y), 
             shape = 1, size = 1.5, alpha = 0.8, stroke = 0.5) +
  # Draw median distances
  geom_circle(data = p108_distance_data, 
              aes(x0 = x, y0 = y, r = radii, fill = NA, color = type, linetype = type), 
              alpha = 0) +
  # Draw centroids
  geom_circle(data = tls1_centroid_32,
  aes(x0 = x, y0 = y, r = 50, fill = "darkgreen"), alpha = 1) +
  geom_circle(data = tls2_centroid_32,
  aes(x0 = x, y0 = y, r = 50, fill = "forestgreen"), alpha = 1) +
  geom_circle(data = tls3_centroid_32, 
              aes(x0 = x, y0 = y, r = 50, fill = "green3"), alpha = 1) +
  # Formatting and color options
  scale_color_manual(values = c("darkgreen", "darkgreen")) +
  scale_linetype_manual(values = c("dashed", "solid")) +
  area_fill +
  ggtitle(clone_of_int) +
  guides(color = "none", fill = "none")
```

# P102 33 plot
## Format TLS centroids
```{r}
p102_centroid <- do.call(rbind, list(tls3_centroid_33))
```

## Set clone to investigate
```{r}
clone_of_int <- "p102_264"
```

## Merge data for plot
```{r}
p102_distance_data <- p102_tls3_meds[clone_of_int,] %>%
  pivot_longer(cols = c("Other", "TCR_of_interest"), values_to = "radii", names_to = "type") %>%
  left_join(p102_centroid, by = "TLS") %>%
  mutate(radii = round(radii, 3), 
         x = round(x, 3), 
         y = round(y, 3))
```

## Reorient beads
```{r}
# Reorient beads to match the change in SpatialDimPlot
# Pick out the beads with the clone
beads_to_highlight <- P102TA_33_RNA@meta.data %>%
      filter(str_detect(clone_id, paste0(paste0(clone_of_int, "($|,)"), collapse = "|"))) %>%
      mutate(trim = str_extract(clone_id, paste0(clone_of_int, collapse = "|")),
             plot_x = plot_x,
             plot_y = -plot_y)

p102_distance_data <- p102_distance_data %>%
  mutate(x = x,
         y = -y)

tls3_centroid_33 <- tls3_centroid_33 %>%
  mutate(x = x,
         y = -y)
```

## Plot clones spatially
```{r}
SpatialDimPlot(P102TA_33_RNA, group.by = "area", alpha = 0.3, stroke = 0) +
  # Add beads of interest
  geom_point(data = beads_to_highlight, aes(x = plot_x, y = plot_y), color = "pink", 
             size = 1) +
    coord_fixed(xlim = c(0, 5000), ylim = c(-5000, 0), ratio = 1) +
  # Add outline to beads of interest
  geom_point(data = beads_to_highlight, aes(x = plot_x, y = plot_y), 
             shape = 1, size = 1.5, alpha = 0.8, stroke = 0.5) +
  # Draw median distances
  geom_circle(data = p102_distance_data, 
              aes(x0 = x, y0 = y, r = radii, fill = NA, color = type, linetype = type), 
              alpha = 0) +
  # Draw centroids
  geom_circle(data = tls3_centroid_33, 
              aes(x0 = x, y0 = y, r = 50, fill = "green3"), alpha = 1) +
  # Formatting and color options
  scale_color_manual(values = c("darkgreen", "darkgreen")) +
  scale_linetype_manual(values = c("dashed", "solid")) +
  area_fill +
  ggtitle(clone_of_int) +
  guides(color = "none", fill = "none")
```


# P102 17 plot
## Format TLS centroids
```{r}
p102_centroid <- do.call(rbind, list(tls3_centroid_17))
```

## Set clone to investigate
```{r}
clone_of_int <- "p102_264"
```

## Merge data for plot
```{r}
p102_distance_data <- p102_tls3_meds[clone_of_int,] %>%
  pivot_longer(cols = c("Other", "TCR_of_interest"), values_to = "radii", names_to = "type") %>%
  left_join(p102_centroid, by = "TLS") %>%
  mutate(radii = round(radii, 3), 
         x = round(x, 3), 
         y = round(y, 3))
```

## Reorient beads
```{r}
# Reorient beads to match the change in SpatialDimPlot
# Pick out the beads with the clone
beads_to_highlight <- P102TA_17_RNA@meta.data %>%
      filter(str_detect(clone_id, paste0(paste0(clone_of_int, "($|,)"), collapse = "|"))) %>%
      mutate(trim = str_extract(clone_id, paste0(clone_of_int, collapse = "|")),
             plot_x = plot_x,
             plot_y = -plot_y)

p102_distance_data <- p102_distance_data %>%
  mutate(x = x,
         y = -y)

tls3_centroid_17 <- tls3_centroid_17 %>%
  mutate(x = x,
         y = -y)
```

## Plot clones spatially
```{r}
SpatialDimPlot(P102TA_17_RNA, group.by = "area", alpha = 0.3, stroke = 0) +
  # Add beads of interest
  geom_point(data = beads_to_highlight, aes(x = plot_x, y = plot_y), color = "pink", 
             size = 1) +
    coord_fixed(xlim = c(0, 5000), ylim = c(-5000, 0), ratio = 1) +
  # Add outline to beads of interest
  geom_point(data = beads_to_highlight, aes(x = plot_x, y = plot_y), 
             shape = 1, size = 1.5, alpha = 0.8, stroke = 0.5) +
  # Draw median distances
  geom_circle(data = p102_distance_data, 
              aes(x0 = x, y0 = y, r = radii, fill = NA, color = type, linetype = type), 
              alpha = 0) +
  # Draw centroids
  geom_circle(data = tls3_centroid_17, 
              aes(x0 = x, y0 = y, r = 50, fill = "green3"), alpha = 1) +
  # Formatting and color options
  scale_color_manual(values = c("darkgreen", "darkgreen")) +
  scale_linetype_manual(values = c("dashed", "solid")) +
  area_fill +
  ggtitle(clone_of_int) +
  guides(color = "none", fill = "none")
```