# 110T: RCTD {#sec-rctd}

# Run RCTD on spatial data using patient-matched scRNAseq data

## Libraries
```{r}
library(spacexr)
library(Seurat)
library(dplyr)

source("../source_scripts/slideseq_functions_main.R")

options(Seurat.object.assay.version = 'v5')
```

# Read in the single cell reference
```{r, echo = FALSE}
ref <- readRDS("/n/data2/dfci/medonc/cwu/chloe/slideseq_data/tcr_analysis/RCC_neovax_pucks/analysis_110624/RCTD/scrna_references_091125/P110_rctd_reference.Rds")
```
```{r, eval = FALSE}
ref <- readRDS("P110_rctd_reference.Rds")
```

# Puck 28
## Read in slide-seq puck
```{r, echo = FALSE}
P110TA_28_RNA <- readRDS("/n/data2/dfci/medonc/cwu/chloe/slideseq_data/tcr_analysis/RCC_neovax_pucks/analysis_110624/objects/01_TCR_RNA_integration/01_Puck28_110TA_TCR_RNA_merged_031925.Rds")
```
```{r, eval = FALSE}
P110TA_28_RNA <- readRDS("01_Puck28_110TA_TCR_RNA_merged_031925.Rds")
```

## Extract RCTD required data
```{r}
P110TA_28_counts <- P110TA_28_RNA[["Spatial"]]$counts
P110TA_28_coords <- GetTissueCoordinates(P110TA_28_RNA)
colnames(P110TA_28_coords) <- c("x", "y")
P110TA_28_coords[is.na(colnames(P110TA_28_coords))] <- NULL
```

## Create SpatialRNA object
```{r}
P110TA_28_query <- SpatialRNA(P110TA_28_coords, P110TA_28_counts, colSums(P110TA_28_counts))

print(dim(P110TA_28_query@counts)) # observe Digital Gene Expression matrix
hist(log(P110TA_28_query@nUMI,2)) # histogram of log_2 nUMI
```

## Run RCTD
```{r, eval = FALSE}
P110TA_28_rctd <- create.RCTD(P110TA_28_query, ref, max_cores = 8)
rm(P110TA_28_RNA, P110TA_28_coords, P110TA_28_query, P110TA_28_counts)
gc()
P110TA_28_rctd <- run.RCTD(P110TA_28_rctd, doublet_mode = 'doublet')
```

## Save RCTD results
```{r, eval = FALSE}
saveRDS(P110TA_28_rctd, "P110TA_28_rctd_results.rds")
```

# Puck 30
## Read in slide-seq puck
```{r, echo = FALSE}
P110TA_30_RNA <- readRDS("/n/data2/dfci/medonc/cwu/chloe/slideseq_data/tcr_analysis/RCC_neovax_pucks/analysis_110624/objects/01_TCR_RNA_integration/01_Puck30_110TA_TCR_RNA_merged_031925.Rds")
```
```{r, eval = FALSE}
P110TA_30_RNA <- readRDS("01_Puck30_110TA_TCR_RNA_merged_031925.Rds")
```

## Extract RCTD required data
```{r}
P110TA_30_counts <- P110TA_30_RNA[["Spatial"]]$counts
P110TA_30_coords <- GetTissueCoordinates(P110TA_30_RNA)
colnames(P110TA_30_coords) <- c("x", "y")
P110TA_30_coords[is.na(colnames(P110TA_30_coords))] <- NULL
```

## Create SpatialRNA object
```{r}
P110TA_30_query <- SpatialRNA(P110TA_30_coords, P110TA_30_counts, colSums(P110TA_30_counts))

print(dim(P110TA_30_query@counts)) # observe Digital Gene Expression matrix
hist(log(P110TA_30_query@nUMI,2)) # histogram of log_2 nUMI
```

## Run RCTD
```{r, eval = FALSE}
P110TA_30_rctd <- create.RCTD(P110TA_30_query, ref, max_cores = 8)
rm(P110TA_30_RNA, P110TA_30_coords, P110TA_30_query, P110TA_30_counts)
gc()
P110TA_30_rctd <- run.RCTD(P110TA_30_rctd, doublet_mode = 'doublet')
```

## Save RCTD results
```{r, eval = FALSE}
saveRDS(P110TA_30_rctd, "P110TA_30_rctd_results.rds")
```

#### Note: RCTD is too resource heavy to render in a quarto file. The original run of RCTD is being loaded and saved

# Get session info
```{r}
sessionInfo()
```

