# 107T: RCTD {#sec-rctd}

# Run RCTD on spatial data using patient-matched scRNAseq data

## Libraries
```{r}
library(spacexr)
library(Seurat)
library(dplyr)

source("../source_scripts/slideseq_functions_main.R")

options(Seurat.object.assay.version = 'v5')
```

# Read in the single cell reference
```{r, echo = FALSE}
ref <- readRDS("/n/data2/dfci/medonc/cwu/chloe/slideseq_data/tcr_analysis/RCC_neovax_pucks/analysis_110624/RCTD/scrna_references_091125/P107_rctd_reference.Rds")
```
```{r, eval = FALSE}
ref <- readRDS("P107_rctd_reference.Rds")
```

# Puck 15
## Read in slide-seq puck
```{r, echo = FALSE}
P107TA_15_RNA <- readRDS("/n/data2/dfci/medonc/cwu/chloe/slideseq_data/tcr_analysis/RCC_neovax_pucks/analysis_110624/objects/01_TCR_RNA_integration/01_Puck15_107TA_TCR_RNA_merged_110724.Rds")
```
```{r, eval = FALSE}
P107TA_15_RNA <- readRDS("01_Puck15_107TA_TCR_RNA_merged_110724.Rds")
```

## Extract RCTD required data
```{r}
P107TA_15_counts <- P107TA_15_RNA[["Spatial"]]$counts
P107TA_15_coords <- GetTissueCoordinates(P107TA_15_RNA)
colnames(P107TA_15_coords) <- c("x", "y")
P107TA_15_coords[is.na(colnames(P107TA_15_coords))] <- NULL
```

## Create SpatialRNA object
```{r}
P107TA_15_query <- SpatialRNA(P107TA_15_coords, P107TA_15_counts, colSums(P107TA_15_counts))

print(dim(P107TA_15_query@counts)) # observe Digital Gene Expression matrix
hist(log(P107TA_15_query@nUMI,2)) # histogram of log_2 nUMI
```

## Run RCTD
```{r, eval = FALSE}
P107TA_15_rctd <- create.RCTD(P107TA_15_query, ref, max_cores = 8)
rm(P107TA_15_RNA, P107TA_15_coords, P107TA_15_query, P107TA_15_counts)
gc()
P107TA_15_rctd <- run.RCTD(P107TA_15_rctd, doublet_mode = 'doublet')
```

## Save RCTD results
```{r, eval = FALSE}
saveRDS(P107TA_15_rctd, "P107TA_15_rctd_results.rds")
```

# Puck 18
## Read in slide-seq puck
```{r, echo = FALSE}
P107TA_18_RNA <- readRDS("/n/data2/dfci/medonc/cwu/chloe/slideseq_data/tcr_analysis/RCC_neovax_pucks/analysis_110624/objects/01_TCR_RNA_integration/01_Puck18_107TA_TCR_RNA_merged_110624.Rds")
```
```{r, eval = FALSE}
P107TA_18_RNA <- readRDS("01_Puck18_107TA_TCR_RNA_merged_110624.Rds")
```

## Extract RCTD required data
```{r}
P107TA_18_counts <- P107TA_18_RNA[["Spatial"]]$counts
P107TA_18_coords <- GetTissueCoordinates(P107TA_18_RNA)
colnames(P107TA_18_coords) <- c("x", "y")
P107TA_18_coords[is.na(colnames(P107TA_18_coords))] <- NULL
```

## Create SpatialRNA object
```{r}
P107TA_18_query <- SpatialRNA(P107TA_18_coords, P107TA_18_counts, colSums(P107TA_18_counts))

print(dim(P107TA_18_query@counts)) # observe Digital Gene Expression matrix
hist(log(P107TA_18_query@nUMI,2)) # histogram of log_2 nUMI
```

## Run RCTD
```{r, eval = FALSE}
P107TA_18_rctd <- create.RCTD(P107TA_18_query, ref, max_cores = 8)
rm(P107TA_18_RNA, P107TA_18_coords, P107TA_18_query, P107TA_18_counts)
gc()
P107TA_18_rctd <- run.RCTD(P107TA_18_rctd, doublet_mode = 'doublet')
```

## Save RCTD results
```{r, eval = FALSE}
saveRDS(P107TA_18_rctd, "P107TA_18_rctd_results.rds")
```

# Puck 20
## Read in slide-seq puck
```{r, echo = FALSE}
P107TA_20_RNA <- readRDS("/n/data2/dfci/medonc/cwu/chloe/slideseq_data/tcr_analysis/RCC_neovax_pucks/analysis_110624/objects/01_TCR_RNA_integration/01_Puck20_107TA_TCR_RNA_merged_110624.Rds")
```
```{r, eval = FALSE}
P107TA_20_RNA <- readRDS("01_Puck20_107TA_TCR_RNA_merged_110624.Rds")
```

## Extract RCTD required data
```{r}
P107TA_20_counts <- P107TA_20_RNA[["Spatial"]]$counts
P107TA_20_coords <- GetTissueCoordinates(P107TA_20_RNA)
colnames(P107TA_20_coords) <- c("x", "y")
P107TA_20_coords[is.na(colnames(P107TA_20_coords))] <- NULL
```

## Create SpatialRNA object
```{r}
P107TA_20_query <- SpatialRNA(P107TA_20_coords, P107TA_20_counts, colSums(P107TA_20_counts))

print(dim(P107TA_20_query@counts)) # observe Digital Gene Expression matrix
hist(log(P107TA_20_query@nUMI,2)) # histogram of log_2 nUMI
```

## Run RCTD
```{r, eval = FALSE}
P107TA_20_rctd <- create.RCTD(P107TA_20_query, ref, max_cores = 8)
rm(P107TA_20_RNA, P107TA_20_coords, P107TA_20_query, P107TA_20_counts)
gc()
P107TA_20_rctd <- run.RCTD(P107TA_20_rctd, doublet_mode = 'doublet')
```

## Save RCTD results
```{r, eval = FALSE}
saveRDS(P107TA_20_rctd, "P107TA_20_rctd_results.rds")
```

#### Note: RCTD is too resource heavy to render in a quarto file. The original run of RCTD is being loaded and saved

# Get session info
```{r}
sessionInfo()
```

