# 108T: RCTD {#sec-rctd}

# Run RCTD on spatial data using patient-matched scRNAseq data

## Libraries
```{r}
library(spacexr)
library(Seurat)
library(dplyr)

source("../source_scripts/slideseq_functions_main.R")

options(Seurat.object.assay.version = 'v5')
```

# Read in the single cell reference
```{r, echo = FALSE}
ref <- readRDS("/n/data2/dfci/medonc/cwu/chloe/slideseq_data/tcr_analysis/RCC_neovax_pucks/analysis_110624/RCTD/scrna_references_091125/P108_rctd_reference.Rds")
```
```{r, eval = FALSE}
ref <- readRDS("P108_rctd_reference.Rds")
```

# Puck 26
## Read in slide-seq puck
```{r, echo = FALSE}
P108TA_26_RNA <- readRDS("/n/data2/dfci/medonc/cwu/chloe/slideseq_data/tcr_analysis/RCC_neovax_pucks/analysis_110624/objects/01_TCR_RNA_integration/01_Puck26_108TA_TCR_RNA_merged_110724.Rds")
```
```{r, eval = FALSE}
P108TA_26_RNA <- readRDS("01_Puck26_108TA_TCR_RNA_merged_110724.Rds")
```

## Extract RCTD required data
```{r}
P108TA_26_counts <- P108TA_26_RNA[["Spatial"]]$counts
P108TA_26_coords <- GetTissueCoordinates(P108TA_26_RNA)
colnames(P108TA_26_coords) <- c("x", "y")
P108TA_26_coords[is.na(colnames(P108TA_26_coords))] <- NULL
```

## Create SpatialRNA object
```{r}
P108TA_26_query <- SpatialRNA(P108TA_26_coords, P108TA_26_counts, colSums(P108TA_26_counts))

print(dim(P108TA_26_query@counts)) # observe Digital Gene Expression matrix
hist(log(P108TA_26_query@nUMI,2)) # histogram of log_2 nUMI
```

## Run RCTD
```{r, eval = FALSE}
P108TA_26_rctd <- create.RCTD(P108TA_26_query, ref, max_cores = 8)
rm(P108TA_26_RNA, P108TA_26_coords, P108TA_26_query, P108TA_26_counts)
gc()
P108TA_26_rctd <- run.RCTD(P108TA_26_rctd, doublet_mode = 'doublet')
```

## Save RCTD results
```{r, eval = FALSE}
saveRDS(P108TA_26_rctd, "P108TA_26_rctd_results.rds")
```

# Puck 29
## Read in slide-seq puck
```{r, echo = FALSE}
P108TA_29_RNA <- readRDS("/n/data2/dfci/medonc/cwu/chloe/slideseq_data/tcr_analysis/RCC_neovax_pucks/analysis_110624/objects/01_TCR_RNA_integration/01_Puck29_108TA_TCR_RNA_merged_110724.Rds")
```
```{r, eval = FALSE}
P108TA_29_RNA <- readRDS("01_Puck29_108TA_TCR_RNA_merged_110724.Rds")
```
## Extract RCTD required data
```{r}
P108TA_29_counts <- P108TA_29_RNA[["Spatial"]]$counts
P108TA_29_coords <- GetTissueCoordinates(P108TA_29_RNA)
colnames(P108TA_29_coords) <- c("x", "y")
P108TA_29_coords[is.na(colnames(P108TA_29_coords))] <- NULL
```

## Create SpatialRNA object
```{r}
P108TA_29_query <- SpatialRNA(P108TA_29_coords, P108TA_29_counts, colSums(P108TA_29_counts))

print(dim(P108TA_29_query@counts)) # observe Digital Gene Expression matrix
hist(log(P108TA_29_query@nUMI,2)) # histogram of log_2 nUMI
```

## Run RCTD
```{r, eval = FALSE}
P108TA_29_rctd <- create.RCTD(P108TA_29_query, ref, max_cores = 8)
rm(P108TA_29_RNA, P108TA_29_coords, P108TA_29_query, P108TA_29_counts)
gc()
P108TA_29_rctd <- run.RCTD(P108TA_29_rctd, doublet_mode = 'doublet')
```

## Save RCTD results
```{r, eval = FALSE}
saveRDS(P108TA_29_rctd, "P108TA_29_rctd_results.rds")
```

# Puck 30
## Read in slide-seq puck
```{r, echo = FALSE}
P108TA_30_RNA <- readRDS("/n/data2/dfci/medonc/cwu/chloe/slideseq_data/tcr_analysis/RCC_neovax_pucks/analysis_110624/objects/01_TCR_RNA_integration/01_Puck30_108TA_TCR_RNA_merged_111124.Rds")
```
```{r, eval = FALSE}
P108TA_30_RNA <- readRDS("01_Puck30_108TA_TCR_RNA_merged_111124.Rds")
```

## Extract RCTD required data
```{r}
P108TA_30_counts <- P108TA_30_RNA[["Spatial"]]$counts
P108TA_30_coords <- GetTissueCoordinates(P108TA_30_RNA)
colnames(P108TA_30_coords) <- c("x", "y")
P108TA_30_coords[is.na(colnames(P108TA_30_coords))] <- NULL
```

## Create SpatialRNA object
```{r}
P108TA_30_query <- SpatialRNA(P108TA_30_coords, P108TA_30_counts, colSums(P108TA_30_counts))

print(dim(P108TA_30_query@counts)) # observe Digital Gene Expression matrix
hist(log(P108TA_30_query@nUMI,2)) # histogram of log_2 nUMI
```

## Run RCTD
```{r, eval = FALSE}
P108TA_30_rctd <- create.RCTD(P108TA_30_query, ref, max_cores = 8)
rm(P108TA_30_RNA, P108TA_30_coords, P108TA_30_query, P108TA_30_counts)
gc()
P108TA_30_rctd <- run.RCTD(P108TA_30_rctd, doublet_mode = 'doublet')
```

## Save RCTD results
```{r, eval = FALSE}
saveRDS(P108TA_30_rctd, "P108TA_30_rctd_results.rds")
```

# Puck 32
## Read in slide-seq puck
```{r, echo = FALSE}
P108TA_32_RNA <- readRDS("/n/data2/dfci/medonc/cwu/chloe/slideseq_data/tcr_analysis/RCC_neovax_pucks/analysis_110624/objects/01_TCR_RNA_integration/01_Puck32_108TA_TCR_RNA_merged_110724.Rds")
```
```{r, eval = FALSE}
P108TA_32_RNA <- readRDS("01_Puck32_108TA_TCR_RNA_merged_110724.Rds")
```

## Extract RCTD required data
```{r}
P108TA_32_counts <- P108TA_32_RNA[["Spatial"]]$counts
P108TA_32_coords <- GetTissueCoordinates(P108TA_32_RNA)
colnames(P108TA_32_coords) <- c("x", "y")
P108TA_32_coords[is.na(colnames(P108TA_32_coords))] <- NULL
```

## Create SpatialRNA object
```{r}
P108TA_32_query <- SpatialRNA(P108TA_32_coords, P108TA_32_counts, colSums(P108TA_32_counts))

print(dim(P108TA_32_query@counts)) # observe Digital Gene Expression matrix
hist(log(P108TA_32_query@nUMI,2)) # histogram of log_2 nUMI
```

## Run RCTD
```{r, eval = FALSE}
P108TA_32_rctd <- create.RCTD(P108TA_32_query, ref, max_cores = 8)
rm(P108TA_32_RNA, P108TA_32_coords, P108TA_32_query, P108TA_32_counts)
gc()
P108TA_32_rctd <- run.RCTD(P108TA_32_rctd, doublet_mode = 'doublet')
```

## Save RCTD results
```{r, eval = FALSE}
saveRDS(P108TA_32_rctd, "P108TA_32_rctd_results.rds")
```

# Puck 06
## Read in slide-seq puck
```{r, echo = FALSE}
P108TB_06_RNA <- readRDS("/n/data2/dfci/medonc/cwu/chloe/slideseq_data/tcr_analysis/RCC_neovax_pucks/analysis_110624/objects/01_TCR_RNA_integration/01_Puck06_108TB_TCR_RNA_merged_110724.Rds")
```
```{r, eval = FALSE}
P108TB_06_RNA <- readRDS("01_Puck06_108TB_TCR_RNA_merged_110724.Rds")
```

## Extract RCTD required data
```{r}
P108TB_06_counts <- P108TB_06_RNA[["Spatial"]]$counts
P108TB_06_coords <- GetTissueCoordinates(P108TB_06_RNA)
colnames(P108TB_06_coords) <- c("x", "y")
P108TB_06_coords[is.na(colnames(P108TB_06_coords))] <- NULL
```

## Create SpatialRNA object
```{r}
P108TB_06_query <- SpatialRNA(P108TB_06_coords, P108TB_06_counts, colSums(P108TB_06_counts))

print(dim(P108TB_06_query@counts)) # observe Digital Gene Expression matrix
hist(log(P108TB_06_query@nUMI,2)) # histogram of log_2 nUMI
```

## Run RCTD
```{r, eval = FALSE}
P108TB_06_rctd <- create.RCTD(P108TB_06_query, ref, max_cores = 8)
rm(P108TB_06_RNA, P108TB_06_coords, P108TB_06_query, P108TB_06_counts)
gc()
P108TB_06_rctd <- run.RCTD(P108TB_06_rctd, doublet_mode = 'doublet')
```

## Save RCTD results
```{r, eval = FALSE}
saveRDS(P108TB_06_rctd, "P108TB_06_rctd_results.rds")
```

# Puck 14
## Read in slide-seq puck
```{r, echo = FALSE}
P108TB_14_RNA <- readRDS("/n/data2/dfci/medonc/cwu/chloe/slideseq_data/tcr_analysis/RCC_neovax_pucks/analysis_110624/objects/01_TCR_RNA_integration/01_Puck14_108TB_TCR_RNA_merged_110724.Rds")
```
```{r, eval = FALSE}
P108TB_14_RNA <- readRDS("01_Puck14_108TB_TCR_RNA_merged_110724.Rds")
```

## Extract RCTD required data
```{r}
P108TB_14_counts <- P108TB_14_RNA[["Spatial"]]$counts
P108TB_14_coords <- GetTissueCoordinates(P108TB_14_RNA)
colnames(P108TB_14_coords) <- c("x", "y")
P108TB_14_coords[is.na(colnames(P108TB_14_coords))] <- NULL
```

## Create SpatialRNA object
```{r}
P108TB_14_query <- SpatialRNA(P108TB_14_coords, P108TB_14_counts, colSums(P108TB_14_counts))

print(dim(P108TB_14_query@counts)) # observe Digital Gene Expression matrix
hist(log(P108TB_14_query@nUMI,2)) # histogram of log_2 nUMI
```

## Run RCTD
```{r, eval = FALSE}
P108TB_14_rctd <- create.RCTD(P108TB_14_query, ref, max_cores = 8)
rm(P108TB_14_RNA, P108TB_14_coords, P108TB_14_query, P108TB_14_counts)
gc()
P108TB_14_rctd <- run.RCTD(P108TB_14_rctd, doublet_mode = 'doublet')
```

## Save RCTD results
```{r, eval = FALSE}
saveRDS(P108TB_14_rctd, "P108TB_14_rctd_results.rds")
```

#### Note: RCTD is too resource heavy to render in a quarto file. The original run of RCTD is being loaded and saved

# Get session info
```{r}
sessionInfo()
```

