# 109T: RCTD {#sec-rctd}

# Run RCTD on spatial data using patient-matched scRNAseq data

## Libraries
```{r}
library(spacexr)
library(Seurat)
library(dplyr)

source("../source_scripts/slideseq_functions_main.R")

options(Seurat.object.assay.version = 'v5')
```

# Read in the single cell reference
```{r, echo = FALSE}
ref <- readRDS("/n/data2/dfci/medonc/cwu/chloe/slideseq_data/tcr_analysis/RCC_neovax_pucks/analysis_110624/RCTD/scrna_references_091125/P109_rctd_reference.Rds")
```
```{r, eval = FALSE}
ref <- readRDS("P109_rctd_reference.Rds")
```

# Puck 23
## Read in slide-seq puck
```{r, echo = FALSE}
P109TB_23_RNA <- readRDS("/n/data2/dfci/medonc/cwu/chloe/slideseq_data/tcr_analysis/RCC_neovax_pucks/analysis_110624/objects/01_TCR_RNA_integration/01_Puck23_109TB_TCR_RNA_merged_110724.Rds")
```
```{r, eval = FALSE}
P109TB_23_RNA <- readRDS("01_Puck23_109TB_TCR_RNA_merged_110724.Rds")
```

## Extract RCTD required data
```{r}
P109TB_23_counts <- P109TB_23_RNA[["Spatial"]]$counts
P109TB_23_coords <- GetTissueCoordinates(P109TB_23_RNA)
colnames(P109TB_23_coords) <- c("x", "y")
P109TB_23_coords[is.na(colnames(P109TB_23_coords))] <- NULL
```

## Create SpatialRNA object
```{r}
P109TB_23_query <- SpatialRNA(P109TB_23_coords, P109TB_23_counts, colSums(P109TB_23_counts))

print(dim(P109TB_23_query@counts)) # observe Digital Gene Expression matrix
hist(log(P109TB_23_query@nUMI,2)) # histogram of log_2 nUMI
```

## Run RCTD
Most beads have UMI ~ 50, so set 20 as minimum instead of 100
```{r, eval = FALSE}
P109TB_23_rctd <- create.RCTD(P109TB_23_query, ref, max_cores = 8, UMI_min = 20)
rm(P109TB_23_RNA, P109TB_23_coords, P109TB_23_query, P109TB_23_counts)
gc()
P109TB_23_rctd <- run.RCTD(P109TB_23_rctd, doublet_mode = 'doublet')
```

## Save RCTD results
```{r, eval = FALSE}
saveRDS(P109TB_23_rctd, "P109TB_23_rctd_results.rds")
```

# Puck 34
## Read in slide-seq puck
```{r, echo = FALSE}
P109TB_34_RNA <- readRDS("/n/data2/dfci/medonc/cwu/chloe/slideseq_data/tcr_analysis/RCC_neovax_pucks/analysis_110624/objects/01_TCR_RNA_integration/01_Puck34_109TB_TCR_RNA_merged_110724.Rds")
```
```{r, eval = FALSE}
P109TB_34_RNA <- readRDS("01_Puck34_109TB_TCR_RNA_merged_110724.Rds")
```

## Extract RCTD required data
```{r}
P109TB_34_counts <- P109TB_34_RNA[["Spatial"]]$counts
P109TB_34_coords <- GetTissueCoordinates(P109TB_34_RNA)
colnames(P109TB_34_coords) <- c("x", "y")
P109TB_34_coords[is.na(colnames(P109TB_34_coords))] <- NULL
```

## Create SpatialRNA object
```{r}
P109TB_34_query <- SpatialRNA(P109TB_34_coords, P109TB_34_counts, colSums(P109TB_34_counts))

print(dim(P109TB_34_query@counts)) # observe Digital Gene Expression matrix
hist(log(P109TB_34_query@nUMI,2)) # histogram of log_2 nUMI
```

## Run RCTD
```{r, eval = FALSE}
P109TB_34_rctd <- create.RCTD(P109TB_34_query, ref, max_cores = 8)
rm(P109TB_34_RNA, P109TB_34_coords, P109TB_34_query, P109TB_34_counts)
gc()
P109TB_34_rctd <- run.RCTD(P109TB_34_rctd, doublet_mode = 'doublet')
```

## Save RCTD results
```{r, eval = FALSE}
saveRDS(P109TB_34_rctd, "P109TB_34_rctd_results.rds")
```

#### Note: RCTD is too resource heavy to render in a quarto file. The original run of RCTD is being loaded and saved

# Get session info
```{r}
sessionInfo()
```

