# 111T: RCTD {#sec-rctd}

# Run RCTD on spatial data using patient-matched scRNAseq data

## Libraries
```{r}
library(spacexr)
library(Seurat)
library(dplyr)

source("../source_scripts/slideseq_functions_main.R")

options(Seurat.object.assay.version = 'v5')
```

# Read in the single cell reference
```{r, echo = FALSE}
ref <- readRDS("/n/data2/dfci/medonc/cwu/chloe/slideseq_data/tcr_analysis/RCC_neovax_pucks/analysis_110624/RCTD/scrna_references_091125/P111_rctd_reference.Rds")
```
```{r, eval = FALSE}
ref <- readRDS("P111_rctd_reference.Rds")
```

# Puck 32
## Read in slide-seq puck
```{r, echo = FALSE}
P111TA_32_RNA <- readRDS("/n/data2/dfci/medonc/cwu/chloe/slideseq_data/tcr_analysis/RCC_neovax_pucks/analysis_110624/objects/01_TCR_RNA_integration/01_Puck32_111TA_TCR_RNA_merged_110724.Rds")
```
```{r, eval = FALSE}
P111TA_32_RNA <- readRDS("01_Puck32_111TA_TCR_RNA_merged_110724.Rds")
```

## Extract RCTD required data
```{r}
P111TA_32_counts <- P111TA_32_RNA[["Spatial"]]$counts
P111TA_32_coords <- GetTissueCoordinates(P111TA_32_RNA)
colnames(P111TA_32_coords) <- c("x", "y")
P111TA_32_coords[is.na(colnames(P111TA_32_coords))] <- NULL
```

## Create SpatialRNA object
```{r}
P111TA_32_query <- SpatialRNA(P111TA_32_coords, P111TA_32_counts, colSums(P111TA_32_counts))

print(dim(P111TA_32_query@counts)) # observe Digital Gene Expression matrix
hist(log(P111TA_32_query@nUMI,2)) # histogram of log_2 nUMI
```

## Run RCTD
```{r, eval = FALSE}
P111TA_32_rctd <- create.RCTD(P111TA_32_query, ref, max_cores = 8)
rm(P111TA_32_RNA, P111TA_32_coords, P111TA_32_query, P111TA_32_counts)
gc()
P111TA_32_rctd <- run.RCTD(P111TA_32_rctd, doublet_mode = 'doublet')
```

## Save RCTD results
```{r, eval = FALSE}
saveRDS(P111TA_32_rctd, "P111TA_32_rctd_results.rds")
```

# Puck 37
## Read in slide-seq puck
```{r, echo = FALSE}
P111TA_37_RNA <- readRDS("/n/data2/dfci/medonc/cwu/chloe/slideseq_data/tcr_analysis/RCC_neovax_pucks/analysis_110624/objects/01_TCR_RNA_integration/01_Puck37_111TA_TCR_RNA_merged_110724.Rds")
```
```{r, eval = FALSE}
P111TA_37_RNA <- readRDS("01_Puck37_111TA_TCR_RNA_merged_110724.Rds")
```

## Extract RCTD required data
```{r}
P111TA_37_counts <- P111TA_37_RNA[["Spatial"]]$counts
P111TA_37_coords <- GetTissueCoordinates(P111TA_37_RNA)
colnames(P111TA_37_coords) <- c("x", "y")
P111TA_37_coords[is.na(colnames(P111TA_37_coords))] <- NULL
```

## Create SpatialRNA object
```{r}
P111TA_37_query <- SpatialRNA(P111TA_37_coords, P111TA_37_counts, colSums(P111TA_37_counts))

print(dim(P111TA_37_query@counts)) # observe Digital Gene Expression matrix
hist(log(P111TA_37_query@nUMI,2)) # histogram of log_2 nUMI
```

## Run RCTD
```{r, eval = FALSE}
P111TA_37_rctd <- create.RCTD(P111TA_37_query, ref, max_cores = 8)
rm(P111TA_37_RNA, P111TA_37_coords, P111TA_37_query, P111TA_37_counts)
gc()
P111TA_37_rctd <- run.RCTD(P111TA_37_rctd, doublet_mode = 'doublet')
```

## Save RCTD results
```{r, eval = FALSE}
saveRDS(P111TA_37_rctd, "P111TA_37_rctd_results.rds")
```

#### Note: RCTD is too resource heavy to render in a quarto file. The original run of RCTD is being loaded and saved

# Get session info
```{r}
sessionInfo()
```

