# Subcluster NeoVax + CancerCell data {#sec-rctd}

# Libraries
```{r}
library(Seurat)
library(paletteer)
library(liana)
library(tidyverse)
library(ggpointdensity)
library(rmarkdown)
library(knitr)
```

# Load NeoVax + Cancer Cell object
```{r}
scrna <- readRDS("/n/data2/dfci/medonc/cwu/alex/objects/merged_RCC.NeoVax.CancerCell.rds")
```

## Print clusters
```{r}
DimPlot(scrna, group.by = "unified_cluster") + paletteer::scale_colour_paletteer_d("ggsci::default_igv")
unique(scrna$unified_cluster)
```

# Create larger cell type groups according to Alex:
Notes on contamination:
B: "B.MEM", "B.PLASMA" 
    alex found some CD3D+ subclusters
Myeloid: "cDC", "MONO.NONCLASS", "MONO.CLASS", "M2.MAC", "MAC", "MAST", "pDC" 
    alex found some CD3D+ CD8+ subclusters
T: "CD8.TEX","CD4.TMEM", "CD8.TRM", "TAP", "TPROL", "CD4.TREG", "CD8.IFN"
    alex found some myeloid contaminant in Tprolif
NK: "NK1", "NK2"
RCC: "RCC1", "RCC2", "RCC3"
    i found some CD68 and PTPRC expression
Endo: "ENDO"
```{r}
scrna@meta.data <- scrna@meta.data %>%
  mutate(generalized_clusters = case_when(unified_cluster %in% c("B.MEM", "B.PLASMA") ~ "B",
                                          unified_cluster %in% c("cDC", "MONO.NONCLASS", "MONO.CLASS", "M2.MAC", "MAC", "MAST", "pDC") ~ "Myeloid",
                                          unified_cluster %in% c("CD8.TEX","CD4.TMEM", "CD8.TRM", "TAP", "TPROL", "CD4.TREG", "CD8.IFN") ~ "T",
                                          unified_cluster %in% c("NK1", "NK2") ~ "NK",
                                          unified_cluster %in% c("RCC1", "RCC2", "RCC3") ~ "RCC",
                                          T ~ unified_cluster))

unique(scrna$generalized_clusters)
```

# B cells ----------------------------------------------------------------------

```{r}
b_subclusters <- readRDS("../objects/b_subclusters.rds")
```

## Subcluster upon the B cell cluster
```{r, eval = FALSE}
Idents(scrna) <- scrna$generalized_clusters
scrna <- FindSubCluster(scrna, cluster = "B", graph.name = "RNA_snn", resolution = 0.5, subcluster.name = "b_subcluster")
```

## Subset for B cell subclusters
```{r, eval = FALSE}
b_subclusters <- subset(scrna, idents = "B")
```

## Plots
```{r}
Idents(b_subclusters) <- b_subclusters$b_subcluster
VlnPlot(b_subclusters, features = c("CD3D", "CD79A", "CD19", "IGHG1"), flip=T,stack=T)
VlnPlot(b_subclusters, features = c("CD3D", "CD79A", "CD19", "IGHG1"), pt.size = 0.1, ncol = 2)

DimPlot(b_subclusters, group.by = "b_subcluster") + paletteer::scale_colour_paletteer_d("ggsci::default_igv")
FeaturePlot(b_subclusters, features = c("CD3D", "CD79A", "CD19", "IGHG1"), order = TRUE) 
```

## Find and plot top variable features per B cluster to see if there's more evidence of contaimination
```{r}
b_markers <- FindAllMarkers(b_subclusters, only.pos = TRUE, nfeatures = 5)

b_markers %>%
  group_by(cluster) %>%
  slice_head(n = 3)
```

## Cell counts
```{r}
b_subclusters@meta.data %>%
  dplyr::count(b_subcluster) %>%
  pivot_wider(names_from = b_subcluster, values_from = n)

b_subclusters@meta.data %>%
  dplyr::count(patient, b_subcluster) %>%
  pivot_wider(names_from = b_subcluster, values_from = n)
```

## B cell markers
```{r}
#16: B-Mem
FeaturePlot(b_subclusters, features = c("CD19", "BANK1"), order = TRUE)
#19: B-PLASMA
FeaturePlot(b_subclusters, features = c("JCHAIN", "DERL3"), order = TRUE)
VlnPlot(b_subclusters, features = c("CD19", "BANK1", "JCHAIN", "DERL3"), flip=T,stack=T)

rm(b_subclusters)
gc()
```


# Myeloid ---------------------------------------------------------------------
## Subcluster upon the myeloid cell cluster
```{r}
myeloid_subclusters <- readRDS("../objects/myeloid_subclusters.rds")
```

```{r, eval = FALSE}
Idents(scrna) <- scrna$generalized_clusters
scrna <- FindSubCluster(scrna, cluster = "Myeloid", graph.name = "RNA_snn", resolution = 0.5, subcluster.name = "myeloid_subcluster")
```

## Subset for myeloid cell subclusters
```{r, eval = FALSE}
myeloid_subclusters <- subset(scrna, idents = "Myeloid")
```

## Plots
```{r}
Idents(myeloid_subclusters) <- myeloid_subclusters$myeloid_subcluster
VlnPlot(myeloid_subclusters, features = c("CD3D", "CD14", "CD68", "NKG7", "CD19"), flip=T,stack=T)
VlnPlot(myeloid_subclusters, features = c("CD3D", "CD14", "CD68", "NKG7"), pt.size = 0.1, ncol = 2)

DimPlot(myeloid_subclusters, group.by = "myeloid_subcluster") + paletteer::scale_colour_paletteer_d("ggsci::default_igv")
FeaturePlot(myeloid_subclusters, features = c("CD3D", "CD14", "CD68", "NKG7"), order = TRUE) 
```

## Find and plot top variable features per myeloid cluster to see if there's more evidence of contaimination
```{r }
myeloid_markers <- FindAllMarkers(myeloid_subclusters, only.pos = TRUE, nfeatures = 5)

myeloid_markers %>%
  group_by(cluster) %>%
  slice_head(n = 3)
```

## Cell counts
```{r}
myeloid_subclusters@meta.data %>%
  dplyr::count(myeloid_subcluster) %>%
  pivot_wider(names_from = myeloid_subcluster, values_from = n)

myeloid_subclusters@meta.data %>%
  dplyr::count(patient, myeloid_subcluster) %>%
  pivot_wider(names_from = myeloid_subcluster, values_from = n)
```

## Myeloid cell markers
```{r}
#1 Macrophage M2
FeaturePlot(myeloid_subclusters, features = c("APOE", "C1QC", "TREM2"), order = TRUE)
#5: Classical Monocytes
FeaturePlot(myeloid_subclusters, features = c("CD14", "S100A8", "FCGR3A"), order = TRUE)
#7: Non classical monocytes
FeaturePlot(myeloid_subclusters, features = c("CD14", "LILRA3", "FCGR3A"), order = TRUE)
#11: cDCs
FeaturePlot(myeloid_subclusters, features = c("CD1E", "CD1C", "FCER1A"), order = TRUE)
#22: pDCs
FeaturePlot(myeloid_subclusters, features = c("CLEC4C", "LILRA4"), order = TRUE)

VlnPlot(myeloid_subclusters, features = c("APOE", "C1QC", "TREM2", "CD14", "S100A8", "FCGR3A", "LILRA3", "CD1E", "CD1C", "FCER1A", "CLEC4C", "LILRA4"), flip=T,stack=T)

rm(myeloid_subclusters)
gc()
```


# T cells ---------------------------------------------------------------------
## Subcluster upon the T cell cluster
```{r}
t_subclusters <- readRDS("../objects/t_subclusters.rds")
```

```{r, eval = FALSE}
Idents(scrna) <- scrna$generalized_clusters
scrna <- FindSubCluster(scrna, cluster = "T", graph.name = "RNA_snn", resolution = 0.5, subcluster.name = "t_subcluster")
```

## Subset for T cell subclusters
```{r, eval = FALSE}
t_subclusters <- subset(scrna, idents = "T")
```

## Plots
```{r}
Idents(t_subclusters) <- t_subclusters$t_subcluster
VlnPlot(t_subclusters, features = c("CD3D", "CD14", "CD68", "CD79A"), flip=T,stack=T)
VlnPlot(t_subclusters, features = c("CD3D", "CD14", "CD68", "CD79A"), pt.size = 0.1, ncol = 2)

DimPlot(t_subclusters, group.by = "t_subcluster") + paletteer::scale_colour_paletteer_d("ggsci::default_igv")
FeaturePlot(t_subclusters, features = c("CD3D", "CD14", "CD68", "CD79A"), order = TRUE) 
```

## Find and plot top variable features per T cluster to see if there's more evidence of contaimination
```{r }
t_markers <- FindAllMarkers(t_subclusters, only.pos = TRUE, nfeatures = 5)

t_markers %>%
  group_by(cluster) %>%
  slice_head(n = 3)
```

## Cell counts
```{r}
t_subclusters@meta.data %>%
  dplyr::count(t_subcluster) %>%
  pivot_wider(names_from = t_subcluster, values_from = n)

t_subclusters@meta.data %>%
  dplyr::count(patient, t_subcluster) %>%
  pivot_wider(names_from = t_subcluster, values_from = n)
```

## T cell markers
```{r}
#0 CD8 TEx, TNFRSF9 TNFRSF9, LAG3, GZMK
FeaturePlot(t_subclusters, features = c("PDCD1", "ENTPD1"), order = TRUE)
#2: CD4 Mem
FeaturePlot(t_subclusters, features = c("CD40LG", "CCR7", "IL7R"), order = TRUE)
#4: CD8 TRM
FeaturePlot(t_subclusters, features = c("CD3E", "CD8A", "CD4"))
#6: TAp - MTRNR2L12
FeaturePlot(t_subclusters, features = c("CD3D"), order = TRUE)
DimPlot(t_subclusters, group.by = "HasClonotype")
#8: Proliferating cells
FeaturePlot(t_subclusters, features = c("UBE2C", "MKI67"), order = TRUE)
#9: NK2
FeaturePlot(t_subclusters, features = c("KRT81", "KRT86", "KLRC1", "ZNF683"), order = TRUE)
#10: CD4-TREG
FeaturePlot(t_subclusters, features = c("FOXP3", "IL2RA", "TNFRSF4"), order = TRUE)
#20: MAST
FeaturePlot(t_subclusters, features = c("MS4A2", "TPSB2"), order = TRUE)
#24: CD8-IFN response/CXCL13

VlnPlot(t_subclusters, features = c("PDCD1", "ENTPD1", "CD40LG", "CCR7", "IL7R", "CD3E", "CD8A", "CD4", "CD3D", "UBE2C", "MKI67", "KRT81", "KRT86", "KLRC1", "ZNF683", "FOXP3", "IL2RA", "TNFRSF4", "MS4A2", "TPSB2"), flip=T,stack=T)

rm(t_subclusters)
gc()
```


# NK cells ---------------------------------------------------------------------
```{r}
nk_subclusters <- readRDS("../objects/nk_subclusters.rds")
```

## Subcluster upon the NK cell cluster
```{r, eval = FALSE}
Idents(scrna) <- scrna$generalized_clusters
scrna <- FindSubCluster(scrna, cluster = "NK", graph.name = "RNA_snn", resolution = 0.5, subcluster.name = "nk_subcluster")
```

## Subset for NK cell subclusters
```{r, eval = FALSE}
nk_subclusters <- subset(scrna, idents = "NK")
```

## Plots
```{r}
Idents(nk_subclusters) <- nk_subclusters$nk_subcluster
VlnPlot(nk_subclusters, features = c("CD3D", "CD8A", "CD68", "CD79A", "NKG7"), flip=T,stack=T)
VlnPlot(nk_subclusters, features = c("CD3D", "CD8A", "CD68", "CD79A", "NKG7"), pt.size = 0.1, ncol = 2)

DimPlot(nk_subclusters, group.by = "nk_subcluster") + paletteer::scale_colour_paletteer_d("ggsci::default_igv")
FeaturePlot(nk_subclusters, features = c("CD3D", "CD8A", "CD68", "CD79A", "NKG7"), order = TRUE) 
```

## Find and plot top variable features per NK cluster to see if there's more evidence of contaimination
```{r }
nk_markers <- FindAllMarkers(nk_subclusters, only.pos = TRUE, selection.method = "vst")

nk_markers %>%
  group_by(cluster) %>%
  slice_head(n = 3) %>%
  print(n = nrow(.))
```

## Cell counts
```{r}
nk_subclusters@meta.data %>%
  dplyr::count(nk_subcluster) %>%
  pivot_wider(names_from = nk_subcluster, values_from = n)

nk_subclusters@meta.data %>%
  dplyr::count(patient, nk_subcluster) %>%
  pivot_wider(names_from = nk_subcluster, values_from = n)
```

## NK cell markers
```{r}
#3: NK1
FeaturePlot(nk_subclusters, features = c("SPON2", "FGFBP2"))
#9: NK2
FeaturePlot(nk_subclusters, features = c("KRT81", "KRT86", "KLRC1", "ZNF683"), order = TRUE)

VlnPlot(nk_subclusters, features = c("SPON2", "FGFBP2", "KRT81", "KRT86", "KLRC1", "ZNF683"), flip=T,stack=T)

rm(nk_subclusters)
gc()
```


# RCC cells ---------------------------------------------------------------------
## Subcluster upon the RCC cell cluster
```{r}
rcc_subclusters <- readRDS("../objects/rcc_subclusters.rds")
```

```{r, eval = FALSE}
Idents(scrna) <- scrna$generalized_clusters
scrna <- FindSubCluster(scrna, cluster = "RCC", graph.name = "RNA_snn", resolution = 0.5, subcluster.name = "rcc_subcluster")
```

## Subset for RCC cell subclusters
```{r, eval = FALSE}
rcc_subclusters <- subset(scrna, idents = "RCC")
```

## Plots
```{r}
Idents(rcc_subclusters) <- rcc_subclusters$rcc_subcluster
VlnPlot(rcc_subclusters, features = c("CD3D", "CD68", "CD79A", "ALB", "SLC22A8", "ALDOB", "CA9"), flip=T,stack=T)
VlnPlot(rcc_subclusters, features = c("CD3D", "CD68", "CD79A", "CA9"), pt.size = 0.1, ncol = 2)

DimPlot(rcc_subclusters, group.by = "rcc_subcluster") + paletteer::scale_colour_paletteer_d("ggsci::default_igv")
FeaturePlot(rcc_subclusters, features = c("CD3D", "CD68", "CD79A", "CA9"), order = TRUE) 
```

## Find and plot top variable features per RCC cluster to see if there's more evidence of contaimination
```{r }
rcc_markers <- FindAllMarkers(rcc_subclusters, only.pos = TRUE, selection.method = "vst")

rcc_markers %>%
  group_by(cluster) %>%
  slice_head(n = 3)
```

## Cell counts
```{r}
rcc_subclusters@meta.data %>%
  dplyr::count(rcc_subcluster) %>%
  pivot_wider(names_from = rcc_subcluster, values_from = n)

rcc_subclusters@meta.data %>%
  dplyr::count(patient, rcc_subcluster) %>%
  pivot_wider(names_from = rcc_subcluster, values_from = n)
```

## RCC cell markers
```{r}
#17: Tumor
FeaturePlot(rcc_subclusters, features = c("CA9", "NDUFA4L2"), order = TRUE)
#25: RCC2 (mes)
FeaturePlot(rcc_subclusters, features = c("LOX", "IGFBP3", "NDUFA4L2"), order = TRUE)
#26: RCC3 (EMT)
FeaturePlot(rcc_subclusters, features = c("MT1A", "ANGPTL4", "NAT8"), order = TRUE, raster = FALSE)

VlnPlot(rcc_subclusters, features = c("CA9", "NDUFA4L2", "LOX", "IGFBP3", "NDUFA4L2", "MT1A", "ANGPTL4", "NAT8"), flip=T,stack=T)

rm(rcc_subclusters)
gc()
```

# Endothelial cells ---------------------------------------------------------------------
## Subcluster upon the endothelial cell cluster
```{r}
endo_subclusters <- readRDS("../objects/endo_subclusters.rds")
```

```{r, eval = FALSE}
Idents(scrna) <- scrna$generalized_clusters
scrna <- FindSubCluster(scrna, cluster = "ENDO", graph.name = "RNA_snn", resolution = 0.5, subcluster.name = "endo_subcluster")
```

## Subset for endothelial cell subclusters
```{r, eval = FALSE}
endo_subclusters <- subset(scrna, idents = "ENDO")
```

## Plots
```{r}
Idents(endo_subclusters) <- endo_subclusters$endo_subcluster
VlnPlot(endo_subclusters, features = c("CD3D", "CD68", "CD79A", "ALB", "SLC22A8", "ALDOB", "CA9"), flip=T,stack=T)
VlnPlot(endo_subclusters, features = c("CD3D", "CD68", "CD79A", "CA9"), pt.size = 0.1, ncol = 2)

DimPlot(endo_subclusters, group.by = "endo_subcluster") + paletteer::scale_colour_paletteer_d("ggsci::default_igv")
FeaturePlot(endo_subclusters, features = c("CD3D", "CD68", "CD79A", "CA9"), order = TRUE) 
```

## Find and plot top variable features per endothelial cluster to see if there's more evidence of contaimination
```{r }
endo_markers <- FindAllMarkers(endo_subclusters, only.pos = TRUE, selection.method = "vst")

endo_markers %>%
  group_by(cluster) %>%
  slice_head(n = 3)
```

## Cell counts
```{r}
endo_subclusters@meta.data %>%
  dplyr::count(endo_subcluster) %>%
  pivot_wider(names_from = endo_subcluster, values_from = n)

endo_subclusters@meta.data %>%
  dplyr::count(patient, endo_subcluster) %>%
  pivot_wider(names_from = endo_subcluster, values_from = n)
```

## Endothelial cell markers
```{r}
#17: Tumor
FeaturePlot(endo_subclusters, features = c("CA9", "NDUFA4L2"), order = TRUE)
#25: RCC2 (mes)
FeaturePlot(endo_subclusters, features = c("LOX", "IGFBP3", "NDUFA4L2"), order = TRUE)
#26: RCC3 (EMT)
FeaturePlot(endo_subclusters, features = c("MT1A", "ANGPTL4", "NAT8"), order = TRUE, raster = FALSE)
#14: Endothelial
FeaturePlot(endo_subclusters, features = c("TEK", "CDH5", "CLEC14A"), order = TRUE)

VlnPlot(endo_subclusters, features = c("CA9", "NDUFA4L2", "LOX", "IGFBP3", "NDUFA4L2", "MT1A", "ANGPTL4", "NAT8", "TEK", "CDH5", "CLEC14A"), flip=T,stack=T)

rm(endo_subclusters)
gc()
```


# Fibroblast cells ---------------------------------------------------------------------
## Subcluster upon the fibroblast cell cluster
```{r}
fib_subclusters <- readRDS("../objects/fib_subclusters.rds")
```

```{r, eval = FALSE}
Idents(scrna) <- scrna$generalized_clusters
scrna <- FindSubCluster(scrna, cluster = "FIBRO", graph.name = "RNA_snn", resolution = 0.5, subcluster.name = "fib_subcluster")
```

## Subset for fibroblast cell subclusters
```{r, eval = FALSE}
fib_subclusters <- subset(scrna, idents = "FIBRO")
```

## Plots
```{r}
Idents(fib_subclusters) <- fib_subclusters$fib_subcluster
VlnPlot(fib_subclusters, features = c("CD3D", "CD68", "PLA2G2A", "COL1A1", "COL1A2", "COL5A1", "LOXL1", "ALB", "SLC22A8", "ALDOB", "CA9"), flip=T,stack=T)
VlnPlot(fib_subclusters, features = c("CD3D", "CD68", "CD79A", "LUM", "FBLN1", "FBLN2"), pt.size = 0.1, ncol = 2)

DimPlot(fib_subclusters, group.by = "fib_subcluster") + paletteer::scale_colour_paletteer_d("ggsci::default_igv")
FeaturePlot(fib_subclusters, features = c("CD3D", "CD68", "CD79A", "CA9", "ALB", "SLC22A8", "ALDOB"), order = TRUE) 
```

## Find and plot top variable features per fibroblast cluster to see if there's more evidence of contaimination
```{r }
fib_markers <- FindAllMarkers(fib_subclusters, only.pos = TRUE, selection.method = "vst")

fib_markers %>%
  group_by(cluster) %>%
  slice_head(n = 3)
```

## Cell counts
```{r}
fib_subclusters@meta.data %>%
  dplyr::count(fib_subcluster) %>%
  pivot_wider(names_from = fib_subcluster, values_from = n)

fib_subclusters@meta.data %>%
  dplyr::count(patient, fib_subcluster) %>%
  pivot_wider(names_from = fib_subcluster, values_from = n)
```

## Fibroblast cell markers
```{r}
#17: Tumor
FeaturePlot(fib_subclusters, features = c("CA9", "NDUFA4L2"), order = TRUE)
#25: RCC2 (mes)
FeaturePlot(fib_subclusters, features = c("LOX", "IGFBP3", "NDUFA4L2"), order = TRUE)
#26: RCC3 (EMT)
FeaturePlot(fib_subclusters, features = c("MT1A", "ANGPTL4", "NAT8"), order = TRUE, raster = FALSE)
FeaturePlot(fib_subclusters, features = c("TEK", "CDH5", "CLEC14A"), order = TRUE)
# Endothelial
VlnPlot(fib_subclusters, features = c("CA9", "NDUFA4L2", "LOX", "IGFBP3", "NDUFA4L2", "MT1A", "ANGPTL4", "NAT8", "TEK", "CDH5", "CLEC14A"), flip=T,stack=T)
# Fibroblast
VlnPlot(fib_subclusters, features = c("PLA2G2A", "COL1A1", "COL1A2", "COL5A1", "LOXL1", "ALB", "SLC22A8", "ALDOB", "CA9"), flip=T,stack=T)

rm(fib_subclusters)
gc()
```

# Save subclustered object
```{r, eval = FALSE}
saveRDS(scrna, "../objects/subclustered_RCC.NeoVax.CancerCell.rds")
```

# Get session info
```{r}
sessionInfo()
```

