---
title: "Puck_231024_32_TCR_analysis"
author: "Chloe R. Tu"
output: html_document
date: "2023-10-26"
---

## Integrate spatial RNA and TCR data

## Libraries
```{r}
library(Seurat)
library(anndata)
library(reticulate)
library(wesanderson)
library(dplyr)
library(stringr)
library(clustree)
library(readxl)
library(reshape2)
library(ggplotify)
library(cowplot)
library(comparator)
library(ggpointdensity)
library(ggvenn)

options(Seurat.object.assay.version = 'v5')
```

## Set up workspace
```{r, message=FALSE}
source("/n/data2/dfci/medonc/cwu/chloe/slideseq_data/tcr_analysis/RCC_neovax_pucks/analysis_110624/scripts/source_scripts/slideseq_functions_main.R")
use_virtualenv("/home/cht269/my_envs/R_anndata_py3.9.14/")
```
```{r, eval=FALSE}
source("slideseq_functions_main.R")
use_virtualenv("R_anndata_py3.9.14/")
```

## Load RNA puck
```{r, message = FALSE}
P108TA_32_RNA <- load_slideseq(gene_exp_dir = "/n/data2/dfci/medonc/cwu/chloe/slideseq_data/raw_data/RCC_pucks/Puck_231024_32_108TA/gene_exp/", coords_dir = "/n/data2/dfci/medonc/cwu/chloe/slideseq_data/raw_data/RCC_pucks/Puck_231024_32_108TA/barcode_matching/Puck_231024_32_barcode_xy.txt.gz", project_name = "P108TA_32")
```
```{r, eval = FALSE}
P108TA_32_RNA <- load_slideseq(gene_exp_dir = "Puck_231024_32_108TA/gene_exp/", coords_dir = "Puck_231024_32_108TA/barcode_matching/Puck_231024_32_barcode_xy.txt.gz", project_name = "P108TA_32")
```

## Load TCR data
```{r, message = FALSE}
P108TA_32_TCR <- anndata::read_h5ad("/n/data2/dfci/medonc/cwu/chloe/slideseq_data/tcr_analysis/RCC_neovax_pucks/mehdi_tcr_analysis/adata_Puck_231024_32_TCR.h5ad")
```
```{r, eval = FALSE}
P108TA_32_TCR <- anndata::read_h5ad("adata_Puck_231024_32_TCR.h5ad")
```

## Filter CDR3 sequences
```{r}
# Extract all CDR3
all_cdr3 <- P108TA_32_TCR$var_names
all_cdr3 <- all_cdr3[!str_detect(all_cdr3, "_|\\*") & (str_detect(all_cdr3, "^C"))]
P108TA_32_TCR <- P108TA_32_TCR[,all_cdr3]
```

## Extract helpful dataframes from TCR anndata object
```{r}
TCR_seq <- P108TA_32_TCR$var
TCR_xy <- P108TA_32_TCR$obs %>%
  mutate(barcode = paste0(barcode, "-1"),
         x = round(x,1),
         y = round(y,1)) %>%
  dplyr::rename("TCR_x" = "x", 
                "TCR_y" = "y", 
                "TCR_barcode" = "barcode")

rownames(TCR_xy) <- TCR_xy$barcode
```

## Double-check that the coordinate data matches
- True, if the TCR data is rounded to 0.1
```{r}
RNA_xy <- P108TA_32_RNA@images$image@coordinates %>%
  mutate(barcode = rownames(.)) %>%
  dplyr::rename("RNA_x" = "x", 
                "RNA_y" = "y", 
                "RNA_barcode" = "barcode")

rownames(RNA_xy) <- RNA_xy$barcode

RNA_xy %>%
  left_join(TCR_xy, by = c("RNA_barcode" = "TCR_barcode")) %>%
  filter(!is.na(TCR_x)) %>%
  mutate(x_match = case_when(RNA_x == TCR_x ~ TRUE,
                             RNA_x != TCR_x ~ FALSE),
         y_match = case_when(RNA_y == TCR_y ~ TRUE,
                             RNA_y != TCR_y ~ FALSE)) %>%
  count(x_match, y_match)
  
```

## Check overlap of TCR barcodes in RNA
```{r, fig.width = 5, fig.height= 4}
TCR_xy %>%
  mutate(IN_RNA = case_when((TCR_barcode %in% RNA_xy$RNA_barcode) ~ TRUE,
                            !(TCR_barcode %in% RNA_xy$RNA_barcode) ~ FALSE)) %>%
  ggplot(aes(x = 1, fill = IN_RNA)) + 
  geom_bar() +
  scale_fill_manual(values = wes_palette("Darjeeling1")) +
  theme_classic() +
  ggtitle("No. of beads from TCR pipeline in RNA")
```

## Assign all TCR barcodes to RNA barcodes
```{r}
final_xy <- map_tcr_and_rna_barcodes(TCR_xy, RNA_xy)
```

## Add the TCR data to the RNA as an additional assay
```{r}
P108TA_32_RNA <- update_and_add_tcr_assay(P108TA_32_TCR, final_xy, P108TA_32_RNA)
```

## Calculate Hamming distance between TCR barcodes and RNA barcodes
```{r}
hamming_option = Hamming()
final_xy <- final_xy %>%
  mutate(hamming_dist = hamming_option(TCR_barcode, RNA_barcode))
```

## Plot distribution of physical distance and Hamming distance of TCR-unique barcodes
```{r}
final_xy %>%
  filter(TCR_barcode != RNA_barcode) %>%
  ggplot(aes(x = distance, y = hamming_dist)) +
    geom_pointdensity() + 
  theme_classic() +
  xlab("physical distance") +
  ylab("hamming distance")
```

## Plot number of beads w/ clonotype data
```{r, fig.width = 3, fig.height = 3}
P108TA_32_RNA@meta.data %>%
  mutate(TCR = case_when((nFeature_TCR > 0) ~ TRUE,
                         (nFeature_TCR == 0) ~ FALSE)) %>%
  ggplot(aes(x = 1, fill = TCR)) + 
  geom_bar() +
  scale_fill_manual(values = wes_palette("Darjeeling1")) +
  theme_classic() +
  ggtitle("No. of beads w/ TCR")
```

## Plot number of beads w/ multiple TCR chains
```{r, fig.width = 5, fig.height = 3}
P108TA_32_RNA@meta.data %>%
  filter(nFeature_TCR > 0) %>%
  ggplot(aes(x = 1, fill = as.factor(nFeature_TCR))) + 
  geom_bar() +
  scale_fill_manual(values = wes_palette("Darjeeling2")) +
  theme_classic() +
  ggtitle("No. of beads w/ multiple TCR")
```

# Open clone phenotype and reactivity data
```{r, message = FALSE}
scrna_reactivities <- read.csv("/n/data2/dfci/medonc/cwu/chloe/slideseq_data/tcr_analysis/RCC_neovax_pucks/analysis_110624/TCR_specificities/reformatted_RCC_slideseq_parameters_110624.csv") %>%
  filter(str_detect(sctcr_clone, "p108"))
```
```{r, eval = FALSE}
scrna_reactivities <- read.csv("reformatted_RCC_slideseq_parameters_052925.csv") %>%
  filter(str_detect(sctcr_clone, "p108"))
```

## Attach scRNA-scTCR derived clone phenotype and reactivities to metadata
```{r}
TCR_seq <- TCR_seq %>%
  mutate(CDR3 = rownames(.)) %>%
  left_join(scrna_reactivities, by = c("chains", "CDR3"))
```

## Add TCR matrix as an assay, and add all CDR3 clones to RNA beads, and their specificities/phenotypes
```{r}
P108TA_32_RNA <- update_and_add_tcr_assay(P108TA_32_TCR, final_xy, P108TA_32_RNA)
P108TA_32_RNA <- add_CDR3s_to_RNA_md(P108TA_32_RNA)
puck_specific_scTCR_clones <- get_puck_specific_sctcr_clones(TCR_seq, scrna_reactivities)
P108TA_32_RNA <- attach_phenotype_specificity_to_RNA_md(puck_specific_scTCR_clones, P108TA_32_RNA)
P108TA_32_RNA <- create_final_clone_ids(P108TA_32_RNA, scrna_reactivities)
```

## Log scale the RNA and TCR features and UMI
```{r}
P108TA_32_RNA$log_nFeature_TCR <- log(P108TA_32_RNA$nFeature_TCR)
P108TA_32_RNA$log_nCount_TCR <- log(P108TA_32_RNA$nCount_TCR)

P108TA_32_RNA$log_nFeature_Spatial <- log(P108TA_32_RNA$nFeature_Spatial)
P108TA_32_RNA$log_nCount_Spatial <- log(P108TA_32_RNA$nCount_Spatial)
```

## Flip the XY coordinates to match the orientation of the H&E stain
```{r}
P108TA_32_xy <- P108TA_32_RNA@images$image@coordinates

# flip vertically
P108TA_32_xy <- P108TA_32_xy %>%
  dplyr::rename("orig_x" = "x",
                "orig_y" = "y") %>%
  mutate(x = orig_y,
         y = orig_x) %>%
  select(-orig_x,-orig_y)

P108TA_32_RNA@images$image@coordinates <- P108TA_32_xy

SpatialFeaturePlot(P108TA_32_RNA, features = c("log_nFeature_Spatial", "log_nCount_TCR"), alpha = c(0,1))
```

## Plot TCR distribution spatially
```{r}
SpatialFeaturePlot(P108TA_32_RNA, features = c("log_nFeature_TCR", "log_nCount_TCR"), pt.size = 2, alpha = c(0.5,1))
```

## Finished joining TCR and RNA data
## Save merged object
```{r, eval = FALSE}
saveRDS(P108TA_32_RNA, "01_Puck32_108TA_TCR_RNA_merged_110724.Rds")
```

## Get session info
```{r}
sessionInfo()
```
