# Plot enrichment {#sec-clone-enrichment-qc}

# Compile and plot enrichment of clonotypes or T cell phenotypes in spatial compartments 

# Libraries
```{r}
library(Seurat)
library(SeuratWrappers)
library(ggplot2)
library(patchwork)
library(dplyr)
library(ggplotify)
library(cowplot)
library(stringr)
library(reshape2)
library(ggrepel)
library(ggbreak)
library(tibble)
library(UpSetR)
library(ggvenn)
library(eulerr)

source("../source_scripts/slideseq_functions_main.R")
```

## Pick significance level
```{r}
sig_level <- 0.05
```

## Open original data
```{r}
orig_tls <- read.csv("../objects/P102TA_P107TA_P108TA_TLS_aggregated.csv")

orig_boundary <- read.csv("../objects/P102TA_P108TA_boundary_aggregated.csv")
```

## Open objects with perimeters removed
```{r}
P102TA_TLS_perimeter_removed_clones <- readRDS("../objects/P102TA_TLS_clone_list_perimeter_removed.rds") %>%
  get_vals_df(.) %>%
  mutate(Patient = "P102TA",
         type = "clone",
         pval = case_when(pval == 0 ~ 1/5000,
                          T ~ pval),
         qval = p.adjust(pval, method = "BH"))
P107TA_TLS_perimeter_removed_clones <- readRDS("../objects/P107TA_TLS_clone_list_perimeter_removed.rds") %>%
  get_vals_df(.) %>%
  mutate(Patient = "P107TA",
         type = "clone",
         pval = case_when(pval == 0 ~ 1/5000,
                          T ~ pval),
         qval = p.adjust(pval, method = "BH"))
P108TA_TLS_perimeter_removed_clones <- readRDS("../objects/P108TA_TLS_clone_list_perimeter_removed.rds") %>%
  get_vals_df(.) %>%
  mutate(Patient = "P108TA",
         type = "clone",
         pval = case_when(pval == 0 ~ 1/5000,
                          T ~ pval),
         qval = p.adjust(pval, method = "BH"))

P102TA_TLS_perimeter_removed_phenotype <- readRDS("../objects/P102TA_TLS_phenotype_list_perimeter_removed.rds") %>%
  get_vals_df(.) %>%
  mutate(Patient = "P102TA",
         type = "phenotype",
         pval = case_when(pval == 0 ~ 1/5000,
                          T ~ pval),
         qval = p.adjust(pval, method = "BH"))
P107TA_TLS_perimeter_removed_phenotype <- readRDS("../objects/P107TA_TLS_phenotype_list_perimeter_removed.rds") %>%
  get_vals_df(.) %>%
  mutate(Patient = "P107TA",
         type = "phenotype",
         pval = case_when(pval == 0 ~ 1/5000,
                          T ~ pval),
         qval = p.adjust(pval, method = "BH"))
P108TA_TLS_perimeter_removed_phenotype <- readRDS("../objects/P108TA_TLS_phenotype_list_perimeter_removed.rds") %>%
  get_vals_df(.) %>%
  mutate(Patient = "P108TA",
         type = "phenotype",
         pval = case_when(pval == 0 ~ 1/5000,
                          T ~ pval),
         qval = p.adjust(pval, method = "BH"))

## Boundary

P102TA_boundary_perimeter_removed_clones <- readRDS("../objects/P102TA_boundary_clone_list_perimeter_removed.rds") %>%
  get_vals_df(.) %>%
  mutate(Patient = "P102TA",
         type = "clone",
         pval = case_when(pval == 0 ~ 1/5000,
                          T ~ pval),
         qval = p.adjust(pval, method = "BH"))
P108TA_boundary_perimeter_removed_clones <- readRDS("../objects/P108TA_boundary_clone_list_perimeter_removed.rds") %>%
  get_vals_df(.) %>%
  mutate(Patient = "P108TA",
         type = "clone",
         pval = case_when(pval == 0 ~ 1/5000,
                          T ~ pval),
         qval = p.adjust(pval, method = "BH"))

P102TA_boundary_perimeter_removed_phenotype <- readRDS("../objects/P102TA_boundary_phenotype_list_perimeter_removed.rds") %>%
  get_vals_df(.) %>%
  mutate(Patient = "P102TA",
         type = "phenotype",
         pval = case_when(pval == 0 ~ 1/5000,
                          T ~ pval),
         qval = p.adjust(pval, method = "BH"))
P108TA_boundary_perimeter_removed_phenotype <- readRDS("../objects/P108TA_boundary_phenotype_list_perimeter_removed.rds") %>%
  get_vals_df(.) %>%
  mutate(Patient = "P108TA",
         type = "phenotype",
         pval = case_when(pval == 0 ~ 1/5000,
                          T ~ pval),
         qval = p.adjust(pval, method = "BH"))
```

## Merge across patients for each area
```{r}
tls_perimeter_removed <- rbind(P102TA_TLS_perimeter_removed_clones, P107TA_TLS_perimeter_removed_clones, P108TA_TLS_perimeter_removed_clones, P102TA_TLS_perimeter_removed_phenotype, P107TA_TLS_perimeter_removed_phenotype, P108TA_TLS_perimeter_removed_phenotype) %>%
  dplyr::rename("pval_perimeter_remove" = "pval",
                "qval_perimeter_remove" = "qval",
                "enrichment_perimeter_remove" = "enrichment",
                "count_in_area_perimeter_remove" = "count_in_area")

boundary_perimeter_removed <- rbind(P102TA_boundary_perimeter_removed_clones, P108TA_boundary_perimeter_removed_clones, P102TA_boundary_perimeter_removed_phenotype, P108TA_boundary_perimeter_removed_phenotype) %>%
  dplyr::rename("pval_perimeter_remove" = "pval",
                "qval_perimeter_remove" = "qval",
                "enrichment_perimeter_remove" = "enrichment",
                "count_in_area_perimeter_remove" = "count_in_area")

```

## Compare to original tls results
```{r}
compare_tls <- orig_tls %>%
  full_join(tls_perimeter_removed) %>%
  select(clone, Patient, type, qval, enrichment, enrichment_perimeter_remove, qval_perimeter_remove)
```

## Compare to original boundary results 
```{r}
compare_boundary <- orig_boundary %>%
  full_join(boundary_perimeter_removed) %>%
  select(clone, Patient, type, qval, enrichment_perimeter_remove, qval_perimeter_remove)
```

## Format data for venn diagram
```{r}
orig_tls_clones_sig <- orig_tls %>%
  filter(type == "clone",
         qval < sig_level)
orig_tls_phenotype_sig <- orig_tls %>%
  filter(type == "phenotype",
         pval < sig_level) %>%
  mutate(clone = paste0(Patient, "_", clone))
remove_perimeter_tls_clones_sig <- tls_perimeter_removed %>%
  filter(type == "clone",
         qval_perimeter_remove < sig_level)
remove_perimeter_tls_phenotype_sig <- tls_perimeter_removed %>%
  filter(type == "phenotype",
         pval_perimeter_remove < sig_level) %>%
  mutate(clone = paste0(Patient, "_", clone))

orig_boundary_clones_sig <- orig_boundary %>%
  filter(type == "clone",
         qval < sig_level)
orig_boundary_phenotype_sig <- orig_boundary %>%
  filter(type == "phenotype",
         pval < sig_level) %>%
  mutate(clone = paste0(Patient, "_", clone))
remove_perimeter_boundary_clones_sig <- boundary_perimeter_removed %>%
  filter(type == "clone",
         qval_perimeter_remove < sig_level)
remove_perimeter_boundary_phenotype_sig <- boundary_perimeter_removed %>%
  filter(type == "phenotype",
         pval_perimeter_remove < sig_level) %>%
  mutate(clone = paste0(Patient, "_", clone))

sig_tls_clones <- list(Original = orig_tls_clones_sig$clone, Reanalysed = remove_perimeter_tls_clones_sig$clone)

sig_tls_phenotypes <- list(Original = orig_tls_phenotype_sig$clone, Reanalysed = remove_perimeter_tls_phenotype_sig$clone)

sig_boundary_clones <- list(Original = orig_boundary_clones_sig$clone, Reanalysed = remove_perimeter_boundary_clones_sig$clone)

sig_boundary_phenotypes <- list(Original = orig_boundary_phenotype_sig$clone, Reanalysed = remove_perimeter_boundary_phenotype_sig$clone)
```


### Plot venn diagram
```{r}
plot(euler(sig_tls_clones), quantities = TRUE, fills = c("#FEF7C7FF", "#B0CBE7FF", "#EBA07EFF"), main = "Significant TLS clones")
plot(euler(sig_tls_phenotypes), quantities = TRUE, fills = c("#FEF7C7FF", "#B0CBE7FF", "#EBA07EFF"), main = "Significant TLS phenotypes")

plot(euler(sig_boundary_clones), quantities = TRUE, fills = c("#B0CBE7FF", "#EBA07EFF", "#FEF7C7FF"), main = "Significant boundary clones")
plot(euler(sig_boundary_phenotypes), quantities = TRUE, fills = c("#FEF7C7FF", "#B0CBE7FF", "#EBA07EFF"), main = "Significant boundary phenotypes")

```

## Plot as enrichment plot
```{r}
TLS_clone_remove <- do.call(rbind, list(P102TA_TLS_perimeter_removed_clones, P107TA_TLS_perimeter_removed_clones, P108TA_TLS_perimeter_removed_clones)) %>%
  mutate(text_size = count_in_area/(2+10),
         sig_qval = case_when(qval <= sig_level ~ "Significant",
                         qval > sig_level ~ "Non-significant"))
      
TLS_phenotype_remove <- do.call(rbind, list(P102TA_TLS_perimeter_removed_phenotype, P107TA_TLS_perimeter_removed_phenotype, P108TA_TLS_perimeter_removed_phenotype)) %>%
  mutate(text_size = count_in_area/(2+10),
         sig_pval = case_when(pval <= sig_level ~ "Significant",
                         pval > sig_level ~ "Non-significant"))

boundary_clone_remove <- do.call(rbind, list(P102TA_boundary_perimeter_removed_clones, P108TA_boundary_perimeter_removed_clones)) %>%
  mutate(text_size = count_in_area/(2+10),
         sig_qval = case_when(qval <= sig_level ~ "Significant",
                         qval > sig_level ~ "Non-significant"))
      
boundary_phenotype_remove <- do.call(rbind, list(P102TA_boundary_perimeter_removed_phenotype, P108TA_boundary_perimeter_removed_phenotype)) %>%
  mutate(text_size = count_in_area/(2+10),
         sig_pval = case_when(pval <= sig_level ~ "Significant",
                         pval > sig_level ~ "Non-significant"))

```

### TLS clones
```{r, fig.width = 8, fig.height = 5}
TLS_clone_original <- orig_tls %>%
  select(-Region) %>%
  mutate(Analysis = "Original",
         text_size = count_in_area/(2+10),
         sig_qval = case_when(qval <= sig_level ~ "Significant",
                         qval > sig_level ~ "Non-significant")) %>%
  filter(type == "clone")

TLS_clone_remove <- TLS_clone_remove %>%
  mutate(Analysis = "remove") %>%
  filter(type == "clone")

compare_tls_clone <- TLS_clone_original %>%
  rbind(TLS_clone_remove) %>%
  mutate(phenotype = case_when(str_detect(clone,"p10") ~ "CD8TEx",
                               T ~ "Spatial"))
  
sig_clones <- compare_tls_clone %>%
  filter(qval < sig_level) %>%
  pull(clone)

sig_compare_tls_clone <- compare_tls_clone %>%
  filter(clone %in% sig_clones)

compare_tls_clone %>%
  ggplot(aes(x = enrichment, y = -log10(qval), label = ifelse(qval <= sig_level & phenotype == "CD8TEx" & Analysis == "Original", clone, NA), shape = sig_qval)) +
  geom_hline(color = "grey50", yintercept = -log10(sig_level), linetype = "longdash") +
  geom_vline(color = "black", xintercept = 0) +
  geom_hline(color = "black", yintercept = 0) +
  geom_point(aes(fill = ifelse(qval <= sig_level, phenotype, ""), color = ifelse(qval <= sig_level, Patient, ""), size = count_in_area), shape = 21, stroke = 2) +
  geom_text_repel(alpha = 1, aes(size = 500)) +
  geom_line(data = sig_compare_tls_clone, aes(group = clone), linewidth = 1, linetype = "dotted") +
  theme_classic() + 
  xlab("Enrichment in TLS") +
  ylab("-log10(adjusted pval)") +
  scale_fill_manual(values = c("white","red", "black")) +
  scale_color_manual(values = c("grey70", "#FF9300","#009193")) +
  guides(color=guide_legend(title="Patient", override.aes = aes(label = "")),
         fill = guide_legend(title = "TCR Phenotype", override.aes = aes(label = "")),
         size = guide_legend(title = "TCR Beads in Region", override.aes = aes(label = ""))) +
  coord_cartesian(ylim = c(-0.3, 3), xlim = c(-2, 6)) +
  scale_size(breaks=c(1, 31.62, 1000), limits = c(1, 1000))

```

### TLS phenotype
```{r, fig.width = 8, fig.height = 5}
TLS_phenotype_original <- orig_tls %>%
  select(-Region) %>%
  mutate(Analysis = "Original",
         text_size = count_in_area/(2+10),
         sig_pval = case_when(pval <= sig_level ~ "Significant",
                         pval > sig_level ~ "Non-significant")) %>%
  filter(type == "phenotype")

TLS_phenotype_remove <- TLS_phenotype_remove %>%
  mutate(Analysis = "remove") %>%
  filter(type == "phenotype")

compare_tls_phenotype <- TLS_phenotype_original %>%
  rbind(TLS_phenotype_remove) %>%
  mutate(phenotype = clone)
  
sig_phenotypes <- compare_tls_phenotype %>%
  filter(pval < sig_level) %>%
  pull(clone)

sig_compare_tls_phenotype <- compare_tls_phenotype %>%
  filter(clone %in% sig_phenotypes)
  
compare_tls_phenotype %>%
  ggplot(aes(x = enrichment, y = -log10(pval), label = ifelse(pval <= sig_level & Analysis == "Original", clone, NA))) +
  geom_hline(color = "grey50", yintercept = -log10(sig_level), linetype = "longdash") +
  geom_vline(color = "black", xintercept = 0) +
  geom_hline(color = "black", yintercept = 0) +
  geom_point(aes(fill = ifelse(pval <= sig_level, phenotype, ""), color = ifelse(pval <= sig_level, Patient, ""), size = count_in_area), shape = 21, stroke = 2) +
  geom_text_repel(alpha = 1, aes(size = 1000)) +
  geom_line(data = sig_compare_tls_phenotype, aes(group = Patient), linewidth = 1, linetype = "dotted") +
  theme_classic() + 
  xlab("Enrichment in TLS") +
  ylab("-log10(pval)") +
  scale_shape_manual(guide=FALSE, values=c(21, 19)) +
  scale_fill_manual(values = c("white","red")) +
  scale_color_manual(values = c("grey70", "#FF9300","#009193")) +
  guides(color=guide_legend(title="Patient", override.aes = aes(label = "")),
         fill = guide_legend(title = "TCR Phenotype", override.aes = aes(label = "")),
         size = guide_legend(title = "TCR Beads in Region", override.aes = aes(label = ""))) +
  coord_cartesian(ylim = c(-0.3, 4), xlim = c(-1, 2)) +
  scale_size(breaks=c(1, 44.72, 2000), limits = c(1, 2000))
```

### Boundary clones
```{r, fig.width = 8, fig.height = 5}
boundary_clone_original <- orig_boundary %>%
  select(-Region) %>%
  mutate(Analysis = "Original",
         text_size = count_in_area/(2+10),
         sig_qval = case_when(qval <= sig_level ~ "Significant",
                         qval > sig_level ~ "Non-significant")) %>%
  filter(type == "clone")

boundary_clone_remove <- boundary_clone_remove %>%
  mutate(Analysis = "remove") %>%
  filter(type == "clone")

compare_boundary_clone <- boundary_clone_original %>%
  rbind(boundary_clone_remove) %>%
  mutate(phenotype = case_when(str_detect(clone,"p10") ~ "CD8TEx",
                               T ~ "Spatial"),
         Analysis = case_when(clone == "p102_286" ~ "Original",
                              T ~ Analysis))
  
sig_clones <- compare_boundary_clone %>%
  filter(qval < sig_level) %>%
  pull(clone)

sig_compare_boundary_clone <- compare_boundary_clone %>%
  filter(clone %in% sig_clones)

compare_boundary_clone %>%
  ggplot(aes(x = enrichment, y = -log10(qval), label = ifelse(qval <= sig_level & phenotype == "CD8TEx" & Analysis == "Original", clone, NA), size = enrichment, shape = sig_qval)) +
  geom_hline(color = "grey50", yintercept = -log10(sig_level), linetype = "longdash") +
  geom_vline(color = "black", xintercept = 0) +
  geom_hline(color = "black", yintercept = 0) +
  geom_point(aes(fill = ifelse(qval <= sig_level, phenotype, ""), color = ifelse(clone %in% sig_clones, Patient, ""), size = count_in_area), shape = 21, stroke = 2) +
  geom_text_repel(alpha = 1, aes(size = 300)) +
  geom_line(data = sig_compare_boundary_clone, aes(group = clone), linewidth = 1, linetype = "dotted") +
  theme_classic() + 
  xlab("Enrichment in boundary") +
  ylab("-log10(adjusted pval)") +
  scale_fill_manual(values = c("white","red", "black")) +
  scale_color_manual(values = c("grey70", "#FF9300","#009193")) +
  guides(color=guide_legend(title="Patient", override.aes = aes(label = "")),
         fill = guide_legend(title = "TCR Phenotype", override.aes = aes(label = "")),
         size = guide_legend(title = "TCR Beads in Region", override.aes = aes(label = ""))) +
  coord_cartesian(ylim = c(-0.3, 3), xlim = c(-2, 6)) +
  scale_size(breaks=c(1, 31.62, 1000), limits = c(1, 1000))
```

### Boundary phenotype
```{r, fig.width = 8, fig.height = 5}
boundary_phenotype_original <- orig_boundary %>%
  select(-Region) %>%
  mutate(Analysis = "Original",
         text_size = count_in_area/(2+10),
         sig_pval = case_when(pval <= sig_level ~ "Significant",
                         pval > sig_level ~ "Non-significant")) %>%
  filter(type == "phenotype")

boundary_phenotype_remove <- boundary_phenotype_remove %>%
  mutate(Analysis = "remove") %>%
  filter(type == "phenotype")

compare_boundary_phenotype <- boundary_phenotype_original %>%
  rbind(boundary_phenotype_remove) %>%
  mutate(phenotype = clone,
         line_group = paste0(phenotype, Patient))
  
sig_phenotypes <- compare_boundary_phenotype %>%
  filter(pval < sig_level) %>%
  pull(clone)

sig_compare_boundary_phenotype <- compare_boundary_phenotype %>%
  filter(clone %in% sig_phenotypes)
  
compare_boundary_phenotype %>%
  ggplot(aes(x = enrichment, y = -log10(pval), label = ifelse(pval <= sig_level & Analysis == "Original", clone, NA))) +
  geom_hline(color = "grey50", yintercept = -log10(sig_level), linetype = "longdash") +
  geom_vline(color = "black", xintercept = 0) +
  geom_hline(color = "black", yintercept = 0) +
  geom_point(aes(fill = ifelse(pval <= sig_level, phenotype, ""), color = ifelse(pval <= sig_level, Patient, ""), size = count_in_area), shape = 21, stroke = 2) +
  geom_text_repel(alpha = 1, aes(size = 1000)) +
  geom_line(data = sig_compare_boundary_phenotype, aes(group = line_group), linewidth = 1, linetype = "dotted") +
  theme_classic() + 
  xlab("Enrichment in boundary") +
  ylab("-log10(pval)") +
  scale_shape_manual(guide=FALSE, values=c(21, 19)) +
  scale_fill_manual(values = c("white","red", "hotpink")) +
  scale_color_manual(values = c("grey70", "#FF9300","#009193")) +
  guides(color=guide_legend(title="Patient", override.aes = aes(label = "")),
         fill = guide_legend(title = "TCR Phenotype", override.aes = aes(label = "")),
         size = guide_legend(title = "TCR Beads in Region", override.aes = aes(label = ""))) +
  coord_cartesian(ylim = c(-0.3, 4), xlim = c(-1, 2)) +
  scale_size(breaks=c(1, 44.72, 2000), limits = c(1, 2000))

```

## Get session info
```{r}
sessionInfo()
```

