---
title: "Reformat_scTCR_and_Reactivity_Data"
author: "Chloe R. Tu"
date: "2024-11-06"
output: html_document
---

## Reformat clonotype information from scRNA, scTCR and reactivity experiments

## Libraries
```{r}
library(ggvenn)
library(stringr)
library(tidyverse)
library(reshape2)
library(readxl)
library(dplyr)
library(arsenal)
```

## Load data
```{r, message = FALSE}
scrna_cd8 <- read_excel("/n/data2/dfci/medonc/cwu/chloe/slideseq_data/tcr_analysis/RCC_neovax_pucks/analysis_110624/TCR_specificities/RCC Slide Seq Analysis Parameters 052925.xlsx", sheet = 1)

scrna_cd4 <- read_excel("/n/data2/dfci/medonc/cwu/chloe/slideseq_data/tcr_analysis/RCC_neovax_pucks/analysis_110624/TCR_specificities/RCC Slide Seq Analysis Parameters 052925.xlsx", sheet = 2)

scrna_cd4_obj <- readRDS("/n/data2/dfci/medonc/cwu/chloe/slideseq_data/tcr_analysis/RCC_neovax_pucks/analysis_110624/scRNA/CD4finalwPrimaryCluster.rds")

scrna_ts <- read_excel("/n/data2/dfci/medonc/cwu/chloe/slideseq_data/tcr_analysis/RCC_neovax_pucks/analysis_110624/TCR_specificities/RCC Slide Seq Analysis Parameters 052925.xlsx", sheet = 4)
scrna_vs <- read_excel("/n/data2/dfci/medonc/cwu/chloe/slideseq_data/tcr_analysis/RCC_neovax_pucks/analysis_110624/TCR_specificities/RCC Slide Seq Analysis Parameters 052925.xlsx", sheet = 5) 
scrna_tas <- read_excel("/n/data2/dfci/medonc/cwu/chloe/slideseq_data/tcr_analysis/RCC_neovax_pucks/analysis_110624/TCR_specificities/RCC Slide Seq Analysis Parameters 052925.xlsx", sheet = 6) 
scrna_ebv_cross_reactive <- read_excel("/n/data2/dfci/medonc/cwu/chloe/slideseq_data/tcr_analysis/RCC_neovax_pucks/analysis_110624/TCR_specificities/RCC Slide Seq Analysis Parameters 052925.xlsx", sheet = 7) %>%
  filter(Antigen == "EBV Cross-reactive")
```

```{r, eval = FALSE}
# CD8 T cell phenotypes
scrna_cd8 <- read_excel("RCC Slide Seq Analysis Parameters 052925.xlsx", sheet = 1)

# CD4 T cell phenotypes
scrna_cd4 <- read_excel("RCC Slide Seq Analysis Parameters 052925.xlsx", sheet = 2)

# CD4 scRNA object
scrna_cd4_obj <- readRDS("CD4finalwPrimaryCluster.rds")

# Reactivities
scrna_ts <- read_excel("RCC Slide Seq Analysis Parameters 052925.xlsx", sheet = 4)
scrna_vs <- read_excel("RCC Slide Seq Analysis Parameters 052925.xlsx", sheet = 5) 
scrna_tas <- read_excel("RCC Slide Seq Analysis Parameters 052925.xlsx", sheet = 6) 
scrna_ebv_cross_reactive <- read_excel("RCC Slide Seq Analysis Parameters 052925.xlsx", sheet = 7) %>%
  filter(Antigen == "EBV Cross-reactive")
```

## Format CD4 and CD8 phenotype data
```{r}
# CD4
colnames(scrna_cd4)[1] <- "TCR.Clone"

# Attach tumor and normal counts for CD4s- open the scTCR-scRNA final CD4 object and attach CD4 clone counts to the metadata file
scrna_cd4_md <- scrna_cd4_obj@meta.data %>%
  select(patient, TCR.Clone, PrimaryCluster, TCRA1, TCRB1, TCRA2, TCRB2, countintuminCD4, countinnorminCD4, countinmetinCD4)

# Attach CD4 clone counts in tumor and normal to reactivity data
scrna_cd4 <- left_join(scrna_cd4, scrna_cd4_md, by = "TCR.Clone") %>%
  distinct()

scrna_cd4 <- scrna_cd4 %>%
  mutate(TRAV1 = str_split_i(TCRA1, fixed(","), 1),
         TRAJ1 = str_split_i(TCRA1, fixed(","), 2),
         CDR3A1 = str_split_i(TCRA1, fixed(","), 3),
         TRBV1 = str_split_i(TCRB1, fixed(","), 1),
         TRBD1 = str_split_i(TCRB1, fixed(","), 2),
         TRBJ1 = str_split_i(TCRB1, fixed(","), 3),
         CDR3B1 = str_split_i(TCRB1, fixed(","), 4),
         TRAV2 = str_split_i(TCRA2, fixed(","), 1),
         TRAJ2 = str_split_i(TCRA2, fixed(","), 2),
         CDR3A2 = str_split_i(TCRA2, fixed(","), 3),
         TRBV2 = str_split_i(TCRB2, fixed(","), 1),
         TRBB2 = str_split_i(TCRB2, fixed(","), 2),
         TRBJ2 = str_split_i(TCRB2, fixed(","), 3),
         CDR3B2 = str_split_i(TCRB2, fixed(","), 4))

# CD8
scrna_cd8 <- scrna_cd8 %>%
  mutate(TRAV1 = str_split_i(TCRA1, fixed(","), 1),
         TRAJ1 = str_split_i(TCRA1, fixed(","), 2),
         CDR3A1 = str_split_i(TCRA1, fixed(","), 3),
         TRBV1 = str_split_i(TCRB1, fixed(","), 1),
         TRBD1 = str_split_i(TCRB1, fixed(","), 2),
         TRBJ1 = str_split_i(TCRB1, fixed(","), 3),
         CDR3B1 = str_split_i(TCRB1, fixed(","), 4),
         TRAV2 = str_split_i(TCRA2, fixed(","), 1),
         TRAJ2 = str_split_i(TCRA2, fixed(","), 2),
         CDR3A2 = str_split_i(TCRA2, fixed(","), 3),
         TRBV2 = str_split_i(TCRB2, fixed(","), 1),
         TRBB2 = str_split_i(TCRB2, fixed(","), 2),
         TRBJ2 = str_split_i(TCRB2, fixed(","), 3),
         CDR3B2 = str_split_i(TCRB2, fixed(","), 4))
```

## How many clones have both CD4 and CD8 cells supporting them?
```{r}
ggvenn(list(cd4_clones = scrna_cd4$TCR.Clone, cd8_clones = scrna_cd8$TCR.Clone))
```

## 471 clones are found in both CD4 and CD8 cells. For such clones, take the CD4 or CD8 identity based on which identity has the most cells in support across all three tissues (normal, tumor and metastatic), if equal, set to "unknown"
```{r}
cd8_summary <- scrna_cd8 %>%
  mutate(CD8_count = countintum + countinnorm + countinmet) %>%
  select(TCR.Clone, CD8_count, PrimaryCluster)

cd4_summary <- scrna_cd4 %>%
  mutate(CD4_count = countintuminCD4 + countinnorminCD4 + countinmetinCD4) %>%
  select(TCR.Clone, CD4_count, PrimaryCluster)

clones_in_both <- inner_join(cd8_summary, cd4_summary, by = "TCR.Clone") %>%
  mutate(identity = case_when(CD8_count > CD4_count ~ "CD8",
                              CD4_count > CD8_count ~ "CD4",
                              CD4_count == CD8_count ~ "unknown"))

table(clones_in_both$identity)
```

## Mark the clones to remove from each set
```{r}
remove_from_cd8 <- clones_in_both %>% 
  filter(identity %in% c("CD4", "unknown"))
remove_from_cd4 <- clones_in_both %>% 
  filter(identity %in% c("CD8", "unknown"))
unknown_clones <- clones_in_both %>% 
  filter(identity == "unknown")   
```

## Attach the clone ID to each CDR3, phenotypic annotation from scRNA, and experimentally confirmed specificity.
```{r}
# CD8
scrna_cd8_reformatted <- scrna_cd8 %>%
  select("TCR.Clone", "PrimaryCluster", "CDR3A1", "CDR3B1", "CDR3A2", "CDR3B2") %>%
  melt(id = c("TCR.Clone", "PrimaryCluster")) %>%
  filter(!is.na(value)) %>%
  mutate(chains = case_when(variable %in% c("CDR3A1", "CDR3A2") ~ "TRA",
                            variable %in% c("CDR3B1", "CDR3B2") ~ "TRB"),
         identity = "CD8") %>%
  select(-variable) %>%
  # Add the total CD8 count
  left_join(cd8_summary[,c("TCR.Clone", "CD8_count")], by = "TCR.Clone") %>%
  dplyr::rename("CDR3" = "value",
                "sctcr_clone" = "TCR.Clone",
                "total_count" = "CD8_count") %>%
  # Remove clones in both the CD8 and CD4 objects that are described as CD4 or unknown in the "clones_in_both" dataframe
  filter(!(sctcr_clone %in% remove_from_cd8$TCR.Clone))
  

# CD4
scrna_cd4_reformatted <- scrna_cd4 %>%
  select("TCR.Clone", "PrimaryCluster", "CDR3A1", "CDR3B1", "CDR3A2", "CDR3B2") %>%
  melt(id = c("TCR.Clone", "PrimaryCluster")) %>%
  filter(!is.na(value)) %>%
  mutate(chains = case_when(variable %in% c("CDR3A1", "CDR3A2") ~ "TRA",
                            variable %in% c("CDR3B1", "CDR3B2") ~ "TRB"),
         identity = "CD4") %>%
  select(-variable) %>%
  # Add the total CD4 count
  left_join(cd4_summary[,c("TCR.Clone", "CD4_count")], by = "TCR.Clone") %>%
  dplyr::rename("CDR3" = "value",
                "sctcr_clone" = "TCR.Clone",
                "total_count" = "CD4_count") %>%
  # Remove clones in both the CD8 and CD4 objects that are described as CD8 or unknown in the "clones_in_both" dataframe
  filter(!(sctcr_clone %in% remove_from_cd4$TCR.Clone))


###### Unknown clones
scrna_unknown_reformatted <- scrna_cd8 %>%
  select("TCR.Clone", "PrimaryCluster", "CDR3A1", "CDR3B1", "CDR3A2", "CDR3B2") %>%
  melt(id = c("TCR.Clone", "PrimaryCluster")) %>%
  filter(!is.na(value)) %>%
  mutate(PrimaryCluster = "unknown",
        chains = case_when(variable %in% c("CDR3A1", "CDR3A2") ~ "TRA",
                            variable %in% c("CDR3B1", "CDR3B2") ~ "TRB"),
         identity = "unknown") %>%
  select(-variable) %>%
  # Add the total CD8 count, since the total counts for these clones are the same in both CD4 and CD8
  left_join(cd8_summary[,c("TCR.Clone", "CD8_count")], by = "TCR.Clone") %>%
  dplyr::rename("CDR3" = "value",
                "sctcr_clone" = "TCR.Clone",
                "total_count" = "CD8_count") %>%
  # Filter for clones in both the CD8 and CD4 objects that are described as CD4 or unknown in the "clones_in_both" dataframe
  filter(sctcr_clone %in% unknown_clones$TCR.Clone)
```

## Reformat phenotypic and reactivity data so that each CDR3 has it's own row
```{r}
scrna_ts_reformatted <- scrna_ts %>%
  select(`Tumor Specific`, "CDR3A", "CDR3B") %>%
  melt(id = c("Tumor Specific")) %>%
  mutate(chains = case_when(variable == "CDR3A" ~ "TRA",
                            variable == "CDR3B" ~ "TRB"),
         tumor_reactive = TRUE) %>%
  select(-variable) %>%
  dplyr::rename("CDR3" = "value", 
                "sctcr_clone" = "Tumor Specific")

scrna_vs_reformatted <- scrna_vs %>%
  select(TCR, Antigen, CDR3A, CDR3B) %>%
  melt(id = c("TCR", "Antigen")) %>%
  mutate(chains = case_when(variable == "CDR3A" ~ "TRA",
                            variable == "CDR3B" ~ "TRB"),
         virus_reactive = TRUE) %>%
  select(-variable) %>%
  dplyr::rename("CDR3" = "value",
                "virus_antigen" = "Antigen",
                "sctcr_clone" = "TCR")

scrna_tas_reformatted <- scrna_tas %>%
  select(TCR, Antigen, CDR3A, CDR3B) %>%
  melt(id = c("TCR", "Antigen")) %>%
  mutate(chains = case_when(variable == "CDR3A" ~ "TRA",
                            variable == "CDR3B" ~ "TRB"),
         tumor_antigen_specific = TRUE) %>%
  select(-variable) %>%
  dplyr::rename("CDR3" = "value",
                "antigen" = "Antigen",
                "sctcr_clone" = "TCR")

scrna_ebv_cross_reactive_reformatted <- scrna_ebv_cross_reactive %>%
  select(TCR, CDR3A, CDR3B) %>%
  melt(id = c("TCR")) %>%
  mutate(chains = case_when(variable == "CDR3A" ~ "TRA",
                            variable == "CDR3B" ~ "TRB"),
         ebv_tumor_reactive = TRUE) %>%
  select(-variable) %>%
  dplyr::rename("CDR3" = "value",
                "sctcr_clone" = "TCR")
```

## Merge data into one sheet
```{r}
scrna_full <- do.call(rbind, list(scrna_cd8_reformatted, scrna_cd4_reformatted, scrna_unknown_reformatted))

scrna_full <- scrna_full %>%
  full_join(scrna_tas_reformatted, by = c("sctcr_clone", "chains", "CDR3")) %>%
  full_join(scrna_ts_reformatted, by = c("sctcr_clone", "chains", "CDR3")) %>%
  full_join(scrna_vs_reformatted, by = c("sctcr_clone", "chains", "CDR3")) %>%
  full_join(scrna_ebv_cross_reactive_reformatted, by = c("sctcr_clone", "chains", "CDR3"))

# Update clones that are "EBV and tumor cross reactive" so that the virus_antigen is EBV, and the virus_reactive and tumor_reactive columns are TRUE
scrna_full <- scrna_full %>%
  mutate(virus_reactive = case_when(sctcr_clone %in% scrna_ebv_cross_reactive_reformatted$sctcr_clone ~ TRUE,
                                    T ~ virus_reactive),
         virus_antigen = case_when(sctcr_clone %in% scrna_ebv_cross_reactive_reformatted$sctcr_clone ~ "EBV",
                                   T ~ virus_antigen),
         tumor_reactive = case_when(sctcr_clone %in% scrna_ebv_cross_reactive_reformatted$sctcr_clone ~ TRUE,
                                    T ~ tumor_reactive)) %>%
  # Remove the row that shows tumor and EBV cross-reactivity
  select(-ebv_tumor_reactive)

# Add another column specifying clones that are truly "tumor_specific"- ie reactive to tumor and tumor alone
scrna_full <- scrna_full %>%
  mutate(tumor_specific = case_when(tumor_reactive == TRUE & is.na(virus_reactive) ~ TRUE,
                                    T ~ NA), .after = "tumor_reactive")
```

## Save sheet
```{r, eval = FALSE}
write.csv(scrna_full, "reformatted_RCC_slideseq_parameters_052925.csv", row.names = FALSE)
```

## Get session info
```{r}
sessionInfo()
```

